### 2015-06-10 content/abstract.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

Continuous Datalog semantics

CoDeR Logical Model (minimal algebra and stream data model) for Continuous Deductive Reasoning

Physical Model and Windowed Rete Algorithm

R4 (\& Squall)

Evaluation (against Sparkwave \cite{sparkwave} and CQELS)

### 2015-06-10 content/conclusions.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Conclusions}{conclusions}
	We are more semantically sound than other RDF Stream Processing offerings and more ``continuous'' than other continuous reasoning offerings, without sacrificing processing efficiency.

%	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}

### 2015-06-10 content/conclusions/placeholder.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Conclusions}{conclusions}
	The problem to be addressed by this work was concerned with how entailments could be continuously inferred, from semantic data sourced from potentially many different data sources, by reasoning according to some potentially complex reasoning task.
	My work thus far has first explored the existing literature concerning each of the primary research areas (database management, Semantic Web technologies, automatic reasoning and continuous stream processing) and their overlap with regard to the research challenges laid out in \refsec{intro-challenges} (reiterated briefly below).
	\begin{description}
		\item[\ref{cha:continuous-data}]\hfill\\
			Data streams can run indefinitely, so the total volume of data to be processed is theoretically infinite.
			As data streams are produced continuously and usually report the state of the world in some way, the distribution of the data on a stream will fluctuate over time.
		\item[\ref{cha:data-rate}]\hfill\\
			Many use cases for data streams involve high streaming data rates, such as traffic-management event-processing systems that track the movements of millions of cars at a granularity of less than a minute.
		\item[\ref{cha:reasoning-expressivity}]\hfill\\
			Solutions for performing more expressive reasoning tend to restrict themselves to processing relatively small, static data sets, while those meant to process larger data sets restrict themselves to relatively inexpressive reasoning.
		\item[\ref{cha:real-time}]\hfill\\
			Because data streams can continue indefinitely, the volume of data produced can become very large, making reasoning and related tasks increasingly expensive as the amount of stored data increases.
			The more expensive the task the lower the throughput of the system.
			If the throughput of a system drops lower than the number of new triples arriving on the stream per second then the system will lose the ability to process data in near or even delayed real-time.
	\end{description}
	Firstly, I determined from the literature that distribution, as discussed in \refsec{background: scalability}, is a common property of systems designed to process either large volumes of data or high rates of input;
	the former case relates to the \refcha{continuous-data}, while the latter relates to the \refcha{data-rate}.
	Secondly, I determined that the querying of semantic data has its foundations in graph pattern matching, which are the same semantics underlying rule-based reasoning, as discussed in \refsec{background: semantic web};
	as the extension of the former with continuous processing techniques has received significant attention in the literature as shown in \refsec{background: streaming}, the application of research in that area to the latter is a potential approach to the \refcha{reasoning-expressivity}.
	Finally, two well known algorithms emerged from the review as being particularly suited to both distribution and pattern matching:
	the Rete pattern matching algorithm discussed in \refsec{background: logic programming,background: continuous queries,background: stream reasoning,background: multi-query optimisation}, proposed in the thesis by \citet{forgy79} for use in production systems, a form of rule-based reasoning;
	and the Eddies algorithm for continuously adaptive query planning, proposed initially by \citet{eddies} in the federated databases community, but being more well known within the data-stream management community as discussed in \refsec{background: adaptive processing}.

	Following the conclusions drawn from the review of the literature, I developed a model for the distribution of continuous processing of streamed semantic data that may be applied to both continuous querying and continuous reasoning to the expressivity of the OWL 2 RL profile \citep{w3cowl2profiles}.
	This model is divided into a logical model and a physical model, presented in \refsec{logical operators: pipelining,physical operators: partitioning} respectively.
	The logical model describes the minimal functional set of share-nothing operators required, the conceptual streams by which they are connected and the nature of the data they act on.
	The physical model, on the other hand, describes means by which processing plans in the logical model may be deployed across a distributed environment.
	My initial investigations into this model, presented in \refsec{physical operators: SPARQLStorm evaluation}, show it to be comparable to state-of-the-art continuous querying engines according to the LSBench continuous querying benchmark \citep{LSBench}, even when using only the most na\"{\i}ve of planning algorithms.

	I also present Tempest in \refsec{tempest}, an implementation of the Eddies algorithm as adapted by the data-stream management community for continuous querying, but intended for continuous reasoning.
	Tempest is founded on the model presented in \refsec{logical operators: pipelining,physical operators: partitioning}, extended with routing policies to carry out the continuously adaptive planning characteristic to Eddy-based systems.
	However, a greedy incremental approach to data routing, with semantic data represented natively in RDF graphs from which its \emph{lineage} \citep{CACQ} may be inferred, is shown to be unable to meet the \refcha{data-rate} in \refsec{tempest-experiments-results}.
	As such, Tempest does not constitute a contribution, but provides scope for the work to be conducted during the remainder of my PhD, as discussed in \refsec{future-work}.

	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}

### 2015-06-10 content/contribution/CoDeR.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
	Stream-based Data Model

	Minimal Algebra

%	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}

### 2015-06-10 content/contribution/R4.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{R4: Rule-based Reasoning using Rete\dots Roar}{implementation}
	Data

	Operators

%	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}

### 2015-06-10 content/contribution/continuous-datalog.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Continuous Datalog Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
	\begin{description}
		\item[Stream]\label{def:continuous datalog: stream}\hfill\\
			A stream is a \emph{sequence} of elements ordered by some temporal characteristic of each element, e.g. its creation time, entailment time, or arrival time.
			Each element is uniquely identifiable by its contents (which may duplicate that of another element in the stream) and its position in the sequence.
		\item[Window]\label{def:continuous datalog: window}\hfill\\
			A window of range $X$ over any stream $S$ at any given time ${t \in \mathbb{N}}$ is the \emph{set} of elements that occur in $S$ during the interval ${(t-X,t]}$, and is denoted ${W^{S,X}_{t}}$.
	\end{description}
	\begin{enumerate}\newcounter{continuousDatalogAxioms}
		\item\label{axiom:continuous datalog: window range leq 0}
			A window of range ${{\leq} 0}$ over any stream $S$ at any time $t$ is always empty.
			\begin{equation*}
				\forall S \, \forall t \in \mathbb{N} \, \forall X \leq 0 
					\left( W^{S,X}_{t} = \varnothing \right)
			\end{equation*}
		\item\label{axiom:continuous datalog: window composition}
			A window of range $X$ over any stream $S$ at any given time $t$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at time $t$.
			\begin{equation*}
				\forall S \, \forall t \in \mathbb{N} 
					\left( W^{S,X}_t = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{t-i} \right)
			\end{equation*}
		\item\label{axiom:continuous datalog: window disjointness}
			A window of range $X$ over stream $S$ at time $t$ is disjoint from a second window of any range $Y$ over $S$ at any time after time ${t - X}$.
			\begin{equation*}
				\forall S \, \forall t,X,Y \in \mathbb{N} \, \forall X' \geq X 
					\left( W^{S,X}_t \cap W^{S,Y}_{t-X'} = \varnothing \right)
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}
	\begin{description}
		\item[Continuous Datalog Program]\label{def:continuous datalog: CDP}\hfill\\
			A Continuous Datalog Program (${CDP}$) is composed of a set of Datalog axioms $A$, the union of a set of streams $S$ of Datalog facts and the range $R$ of a window over $S$.
			Such a continuous program entails a triple composed of the axioms $A$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
			The truth of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program $CDP$ and a time $t$ as arguments.
			\begin{equation*}
				CDP = \{A,S,R\} \vDash \{A,P,N\}
			\end{equation*}
		\item[State of a \emph{CDP}]\label{def:continuous datalog: CDPt}\hfill\\
			The state of a \emph{CDP} at a given time $t$ (${CDP_t}$) is the union of the axioms $A$ and the window $W^{S,R}_{t}$.
			As both $A$ and $W^{S,R}_{t}$ are sets of Datalog axioms, a ${CDP_t}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the instantaneous truth of the ${CDP}$ at time $t$ by $T^{CD}$.
			The instantaneous entailments of a $CDP_{t}$ are the union of the axioms $A$ and the set of entailments $E_{t}$ that are justified in part by a subset of $W^{S,R}_{t}$.
			\begin{multline*}
				\forall CDP = \{A,S,R\} \, \forall t \in \mathbb{N} \\
					\left( CDP_t = A \cup W^{S,R}_t \vDash A \cup E_t \right)
			\end{multline*}
			\begin{equation*}
				\forall t \in \mathbb{N} \left( T^{D} \left( CDP_t \right) = T^{CD} \left( CDP, t \right) \right)
			\end{equation*}
	\end{description}
	Both streams entailed by a ${CDP}$ are defined as sequences of the windows of range 1 over themselves, with those windows being defined as the difference between the sets of entailments of the ${CDP}$ between two consecutive instances, as in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: positive window increment}
			The window of range 1 at any given time $t$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at time $t$ that were not entailed by the state of ${CDP}$ at the previous time instance ${t-1}$.
			\begin{multline*}
				\forall CDP \vDash \{A,P,N\} \, \forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \\
					\left( W^{P,1}_{t} = E_{t} \setminus E_{t-1} \right)
			\end{multline*}
		\item\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given time $t$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous time instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at time $t$.
			\begin{multline*}
				\forall CDP \vDash \{A,P,N\} \, \forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \\
					\left( W^{N,1}_{t} = E_{t-1} \setminus E_{t} \right)
			\end{multline*}
		\item\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated (i.e. appear in the stream $N$ at time $t$) after they have been entailed (i.e. appeared in the stream $P$ at some previous time ${t-X}$ where ${X > 0}$).
			\begin{multline*}
				\forall CDP = \{A,S,R\} \vDash \{A,P,N\} \, \forall t,X \in \mathbb{N} \, \forall Y \geq R \\
					\left( W^{P,R}_{t} \cap W^{N,X}_{t-Y} = \varnothing \right)
			\end{multline*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: disjoint entailments}
			The instantaneous set of stream-justified entailments $E_t$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{t-X}$ of the same ${CDP}$, where ${X \geq R}$.
			\begin{multline*}
				\forall CDP = \{A,S,R\} \, \forall CDP_t \vDash A \cup E_t \, \forall t \in \mathbb{N} \, \forall X \geq R \\
					\left( E_t \cap E_{t-X} = \varnothing \right)
			\end{multline*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	\begin{hyp}\labelhyp{continuous datalog: entailment}
		For any ${CDP = \{A,S,R\} \vDash \{A,P,N\}}$, where ${CDP_t \vDash A \cup E_t}$, the set of facts in a window of range ${X \geq R}$ over the stream $P$ at time $t$ and not in a window of range ${Y \geq X}$ over the stream $N$ at $t$ is equal to the set of facts entailed by the state of the ${CDP}$ at time $t$, for any ${t \in \mathbb{N}}$:
		\begin{multline*}
			\forall CDP = \{A,S,R\} \vDash \{A,P,N\} \, \forall X \geq R \, \forall Y \geq X \\
				\forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \left( W^{P,X}_t \setminus W^{N,Y}_t = E_t \right)
		\end{multline*}
	\end{hyp}
	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams are of the same Range as that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary Range.
	\begin{lem}\label{hyp:continuous datalog: entailment lemma}
		For any ${CDP \vDash \{A,P,N\}}$, where ${CDP_t \vDash A \cup E_t}$%, the set of facts in the window of range $X$ over the stream $P$ and not in the window also of range $X$ over the stream $N$ is equal to the set of facts entailed by the state of the ${CDP}$ at time $t$ but not entailed by all states of the ${CDP}$ in the preceding $X$ time instances
		:
		\begin{multline*}
			\forall CDP \vDash \{A,P,N\} \, \forall CDP_t \vDash A \cup E_t \, \forall t,X \in \mathbb{N} \, \forall Y \geq X \\
				\left( W^{P,X}_t \setminus W^{N,Y}_t = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i} \right)
		\end{multline*}
	\end{lem}

	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]
		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every time ${t \in \mathbb{N}}$:
		\begin{align*}
			E_t & = W^{P,X}_t \setminus W^{N,Y}_t \\
			\text{By Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i} \\
			& = \mathop{\cup}^{X}_{i=1} \left( E_t \setminus E_{t-i} \right) \\
			& = \left( E_t \setminus E_{t-R} \right) \cup \mathop{\cup}^{R-1}_{i=1} \left( E_t \setminus E_{t-i} \right) \cup \mathop{\cup}^{X}_{i=R+1} \left( E_t \setminus E_{t-i} \right) \\
			\text{By Axiom~\ref{axiom:continuous datalog: disjoint entailments}} & = E_t \cup \mathop{\cup}^{R-1}_{i=1} \left( E_t \setminus E_{t-i} \right) \cup \mathop{\cup}^{X}_{i=R+1} \left( E_t \setminus E_{t-i} \right) \\
			& = E_t \cup \left( E_t \setminus \left( \mathop{\cap}^{R-1}_{i=1} E_{t-i} \cap \mathop{\cap}^{X}_{i=R+1} E_{t-i} \right) \right) \\
			& = E_t
		\end{align*}
		Thus showing that the RHS is equal to the LHS.
	\end{proof}

	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${X \in \mathbb{N}}$, regardless of the Range of of the ${CDP}$ entailing $P$ and $N$.

	\begin{proof}[Proof of Lemma~\ref{hyp:continuous datalog: entailment lemma} by Induction]
		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${X = k}$ and ${Y \geq X}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${X = k + 1}$, where ${k \in \mathbb{N}}$.

		Firstly, the \textbf{general case} can be defined in terms of sub-windows of range $1$ by Axiom~\ref{axiom:continuous datalog: window composition} over the entailed streams $P$ and $N$, for which Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment} provide an equivalence to sets of entailments.
		\begin{align*}
			W^{P,X}_t \setminus W^{P,Y}_t & = \\
			\text{By Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{X-1}_{i=0}W^{P,1}_{t-i} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{t-j} \\
			& = \mathop{\cup}^{X-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{t-j} \right) \\
			& = \mathop{\cup}^{X-1}_{i=0}\mathop{\cap}^{Y-1}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
			\text{By Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } Y \geq X & = \mathop{\cup}^{X-1}_{i=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right)
		\end{align*}
		The general case also shows that any negative entailments within the window of range $Y$ over $N$ not in the window of range $X$ over $N$ do not influence the composition of the set of values ${W^{P,X}_t \setminus W^{P,Y}_t}$.

		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${X = 1}$ (therefore ${Y \geq 1}$).
		\begin{align*}
			E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-1} & = W^{P,1}_t \setminus W^{N,Y}_t \\
			\text{By the \textbf{general case}} & = W^{P,1}_t \setminus W^{N,1}_t \\
			\text{By Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E_t \setminus E_{t-1} \right) \setminus \left( E_{t-1} \setminus E_t \right) \\
			& = E_t \setminus \left( E_{t-1} \setminus \left( E_{t-1} \setminus E_t \right) \right) \\
			& = E_t \setminus \left( E_{t-1} \cap E_t \right) \\
			& = \left( E_t \setminus E_{t-1} \right) \cup \left( E_t \setminus E_t \right) \\
			& = E_t \setminus E_{t-1} \\
			& = E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-i}
		\end{align*}

		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
		\begin{figure}[p]
			\centering
			\caption[Proof of Entailment Lemma for Continuous Datalog]{Proof of Inductive Case of the Entailment Lemma for Continuous Datalog.}
			\labelfig{continuous datalog: entailment lemma inductive case}
			\begin{align*}
				E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} & = W^{P,k+1}_{t+1} \setminus W^{P,k+1}_{t+1}
				\text{By the \textbf{general case}} & = \mathop{\cup}^{k-1}_{i=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{t-j} \right) \\
				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \left( W^{N,1}_{t+1} \cup \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \right) \right) \\
				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \setminus W^{N,1}_{t+1} \right) \\
				& = \left( W^{P,1}_{t+1} \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \right) \right) \setminus W^{N,1}_{t+1} \\
				& = \left( W^{P,1}_{t+1} \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \right) \setminus W^{N,1}_{t+1} \\
				& = \left( W^{P,1}_{t+1} \cup \left( W^{P,k}_{t} \setminus W^{N,k}_{t} \right) \right) \setminus W^{N,1}_{t+1} \\
				\text{\textbf{Inductive Step}:}& = \left( W^{P,1}_{t+1} \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \setminus W^{N,1}_{t+1} \\
				\text{By Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( \left( E_{t+1} \setminus E_{t} \right) \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \setminus \left( E_{t} \setminus E_{t+1} \right) \\
				& = \left( E_{t+1} \setminus E_{t} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \\
				& = \left( E_{t+1} \setminus E_{t} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \cup \left( E_t \setminus \left( E_{t} \setminus E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
				& = \left( E_{t+1} \setminus \left( E_{t} \cap E_{t+1} \right) \right) \cup \left( \left( E_t \cap E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
				& = \left( E_{t+1} \setminus E_{t} \right) \cup \left( \left( E_t \cap E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
				& = \left( E_{t+1} \setminus E_{t} \right) \cup \left( E_{t+1} \cap \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
				& = \left( E_{t+1} \cap E^C_{t} \right) \cup \left( E_{t+1} \cap \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right)\\
				& = E_{t+1} \cap \left( E^C_{t} \cup \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
				& = E_{t+1} \setminus \left( E_{t} \setminus \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
				& = E_{t+1} \setminus \left( E_{t} \cap \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
				& = E_{t+1} \setminus \mathop{\cap}^{k}_{i=0} E_{t-i} \\
				& = E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i}
			\end{align*}
		\end{figure}
	\end{proof}
\end{nestedsection}

### 2015-06-10 content/evaluation.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Evaluation}{evaluation}
	Ontology

	Data Stream and Window

	Performance comparison of R4 against Sparkwave and CQELS

%	\subimport{conclusions/}{future-work.tex}
\end{nestedsection}

### 2015-06-10 content/introduction.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Introduction}{intro}
	Lack of consensus on stream processing semantics within RDF stream processing community, even more dramatic between RSP and continuous reasoning communities.

	Motivating use case (Glacier stuff, to match evaluation?)
\end{nestedsection}

### 2015-06-10 content/introduction/placeholder.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Introduction}{intro}
	% As computer systems become more prolific and better connected with the ever increasing popularity of the World Wide Web, data is being produced and consumed at an ever increasing rate and in an ever increasing number of contexts.
	The phenomenon of large volume data production on the scale now exhibited across the World Wide Web poses problems in the field of data management, and is the focus of a research area known as Big Data.
	Much of the new data being produced is monitor data, be it from sensor networks (monitoring the physical world), stock market data (monitoring the financial world), or social networking feeds (monitoring the perception of the world) \citep{linkedstreamproc,massivelyConnectedWorld}.
	It can be relatively short lived, becoming less relevant (``stale'') to some tasks quickly either due to the arrival of more up-to-date data or the nature of the task.
	An example of the latter would be Tweets posted on Twitter expressing information about or prompted by ongoing events, which become ``stale'' to a task such as mapping events in real time once the event in question ends.

	While the data can be stored and used as historical data for later analysis through DataBase Management Systems (DBMSs), a lot of potential lies in the ability to perform real-time processing over the most recent data.
	Existing use cases include traffic management systems for automatically mitigating the effects of traffic jams and/or incidents, and decision support systems for environmental disaster response.
	Less time critical cases may nonetheless be time sensitive, such as advertisement targeting algorithms based on recent browsing patterns.
	Use cases such as these have given rise to the concepts of data streams and Data Stream Management Systems (DSMSs), where high volumes of data are produced, processed and stored until they are stale according to some specified task, then discarded.

	Another problem posed by the phenomenon of large volume data production is the increasing demand for information gleaned from multiple disparate data sources, e.g. planning a route across a major city incorporating public transport data, weather forecasts, and social network data indicating disruptive events such as protests or flash mobs.
	The application of two distinct but closely linked concepts concerning the presentation of data accessible via the World Wide Web is required for such cross-source querying:
	The first is the concept of Open Data, in which organisations expose their non-sensitive, non-private data for free consumption via the Web (as a primarily social/legal concept, Open Data is not considered further in this thesis);
	The second is the concept of a standard data structure used to represent all data exposed via the Web in a meaningful way, such that applications can draw on data from different sources without explicit prior knowledge of each source's proprietary Web Application Programming Interface (API).

	Standards relating to the World Wide Web are the purview of the World Wide Web Consortium (W3C), which is ``a community of organisations, full-time staff and private individuals dedicated to leading the Web to its full potential''\footnote{\url{http://www.w3.org/Consortium/}}.
	In order to meet the requirements of providing meaning to data in a Web-wide standard data structure, the W3C has developed the Resource Description Framework (RDF).
	This represents data in a structure whose basic building blocks are known as ``triples'', so named because each building block of the structure holds three pieces of data, a ``Subject'', a ``Predicate'' and an ``Object''.
	RDF also provides an explicit meaning, or ``model theory'', for any data stored in this way:
	The \emph{Subject} is related to the \emph{Object} by the \emph{Predicate}.
	As such, data in this format is known as \emph{semantic data}, because it conveys meaning as well as data.

	The representation of the meaning of data allows the application of reasoning to the data in addition to the querying performed in DBMSs and DSMSs.
	Reasoning is a form of processing that differs from the querying of traditional databases, just as semantic data differs from relational data, in that any data produced through reasoning also subscribes to the same standardised model theory as the original data.
	This allows reasoning processes to interact without explicit instruction, as the output of one process may constitute semantically valid input for others (unlike database queries, which are inherently stand-alone).
	% A group of reasoning processes that define a model according to some model theory is referred to in this work as a ``TBox'', which is the term used in the branch of reasoning called Description Logics (or DLs).
	% Also in DL terminology, a set of data that is compatible with a specific logical model is referred to as an ``ABox''.
	% The combination of a TBox with a compatible ABox is called an ``ontology'' in DL terminology, but is referred to by the more general term ``reasoning task'' in this work.
	\subimport{introduction/}{problem.tex}
	\subimport{introduction/}{challenges.tex}
	\subimport{introduction/}{questions.tex}
	\subimport{introduction/}{contributions.tex}
	\subimport{introduction/}{overview.tex}
\end{nestedsection}

### 2015-06-10 content/literature-report/placeholder.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\begin{nestedsection}{Related Work}{background}
	The research challenges stated in \refsec{intro-challenges} of this thesis span a number of different research areas, including the Semantic Web, deductive reasoning, multiple-query optimisation, stream processing, continuous querying and distributed systems.
	In every case, the challenges relate to drawing together knowledge from two or more of these areas and investigating the overlap therein.
	In this section I review the work present in the literature where it relates to my research challenges.

	The first point I discuss in \refsec{background-queries-vs-reasoning} is the implicit question that underpins the rest of the thesis: What are the similarities and differences between querying and reasoning?
	This defines to what extent I can apply research into forms of querying to this work on reasoning (as per \refdef{deductive-reasoning}).
	It also highlights the differing characteristics of querying and reasoning tasks, which informs my analysis of the literature in the later sections.

	I then explore the notion of stream processing as an answer to the \refcha{real-time}, and an answer to the ``continuous'' part of the \refcha{continuous-data}, in \refsec{background-native-stream-processing}.
	I focus my investigation on what characterises an algorithm that could be considered to perform reasoning over streamed data.
	One of the prime characteristics I identify in stream processing systems is a notion of ``forgetting'' older data in favour of that more recently received, epitomised by a data structure known as a ``sliding window''.
	The other important characteristic of such systems is an intrinsically incremental algorithm, which precludes the direct application of many techniques found in more powerful reasoning algorithms.
	This, in turn, highlights the \refcha{reasoning-expressivity} in conjunction with the \refcha{real-time}.

	Having elicited the key features of stream processing systems, I next detail the various methods for processing large amounts of data employed in the state of the art in \refsec{background-high-rate-volume}, focusing again on how those methods can be applied to stream processing in the face of the \refcha{data-rate}.
	Of particular interest are the ``divide and conquer'' technique and multi-query optimisation.
	The former refers to the practise of partitioning a problem and distributing it across many processing units, then combining the results into a complete solution.
	The latter involves detecting common tasks, known as ``common sub-expressions'', shared between a set of concurrent queries, then using that knowledge to eliminate duplicate task processing across the given query set.

	Finally, in \refsec{background-cont-changing-data} I explore the solutions proposed in the literature to the \refcha{continuous-data}.
	These solutions diverge from the concept of fixed plans explored in \refsec{background-queries-vs-reasoning}, exploring methods of adapting query plans dynamically in response to fluctuating data and operator statistics during processing.
	For a query plan to continuously adapt, the query planning algorithms used must be applied during stream processing;
	as such these planning algorithms must also display the characteristics common to stream processing algorithms as identified in \refsec{background-native-stream-processing}.

	In the concluding \refsec{background-conclusions}, I present a summary of the conclusions that I have drawn from literature.

	% \begin{table}[h]
	% 	\centering
	% 	\begin{tabular}{c | r | *{3}{c}}
	% 		\multicolumn{2}{c |}{} & \rot{Hello} & \rot{there} & \rot{World}\\\hline
	% 		& Hello & * & & \\
	% 		& there & & * & \\
	% 		\rot{\rlap{Greeting!}}
	% 		& World & & & *
	% 	\end{tabular}
	% 	\caption{Hello there, World!}
	% 	\labeltab{}
	% \end{table}

%	These elements can be divided into three primary concerns:
%	those of the Semantic Web as discussed in \refsec{background-semweb};
%	those of reasoning by logic programming and semantic networks as discusse in \refsec{background-reasoning};
%	and those of streaming as discussed in \refsec{background-streaming}, including both the processing of streamed data sources and the streaming of data between components of a distributed system.

%	In addition to these distinct concerns, the current state of the art concerning the overlap of these areas is discussed in \refsec{background-streaming-semantic-data}, and the diverse methods used to measure the performance of solutions within these research areas are discussed in \refsec{background-benchmarking}.

	\subimport{literature-report/}{queries-vs-reasoning.tex}
	\subimport{literature-report/}{native-stream-processing.tex}
	\subimport{literature-report/}{high-rate-volume.tex}
	\subimport{literature-report/}{cont-changing-data.tex}

%	\subimport{literature-report/}{semweb.tex}
%	\subimport{literature-report/}{reasoning.tex}
%	\subimport{literature-report/}{streaming.tex}
%	\subimport{literature-report/}{streaming-semantic-data.tex}
%	\subimport{literature-report/}{benchmarking.tex}

	\begin{nestedsection}{Preliminary Conclusions}{background-conclusions}

	\end{nestedsection}
\end{nestedsection}

### 2015-06-10 template/part_locators.tex
BUGGY: Constructed basic template of the paper, and began testing the Proof of Continuous Datalog section.
PROBLEMS:
- Inductive case is too wide for the columns
- I don't know Rehab's surname or email address!

\graphicspath{{../content/images/}}   % Location of your graphics files
\newcommand{\inputgraphics}[3][width=0.9\textwidth]{\IfFileExists{../content/images/#2}{\includegraphics[#1]{../content/images/#2}}{}} % Define a new command to input text from other files stored in the content directory.
\newcommand{\inputdoc}[2][content]{\IfFileExists{../#1/#2}{\input{../#1/#2}}{}} % Define a new command to input text from other files stored in the content directory.
\newcommand{\includechap}[2][content]{\IfFileExists{../#1/#2}{\include{../#1/#2}}{}} % Define a new command to include tex from other files stored in the content directory.

### 2015-06-10 latexbuild.sh

@@ -1,0 +1,84 @@
+#!/bin/bash
+cd template
+docs=""
+count=0
+while [ $# != 0 ]; do
+	tflag=false
+	dflag=false
+	t=""
+	if [[ $1 ==  "-t"* ]]; then
+		tflag=true
+		t=`echo "$1" | sed -e 's/-t//'`
+		if [ -z $t ]; then
+			shift
+			t="$1"
+		fi
+		t=${t}"-wrapper"
+		shift
+		cat > "$t.tex" << EOF
+\\input{preamble-thesis.tex}
+\\begin{document}
+	\\begin{nestedsection}{$t}{main}
+EOF
+		count=$((count+1))
+	elif [[ $1 == "-d" ]]; then
+		dflag=true
+		t=`echo "$2" | sed -e 's/.*\///'`
+		t=${t}"-wrapper"
+		cat > "$t.tex" << EOF
+\\input{preamble-thesis.tex}
+\\begin{document}
+EOF
+		count=$((count+1))
+	fi
+
+	while [[ $1 ==  "-d" ]]; do
+		shift
+		cat >> "$t.tex" << EOF
+		\\subimport{../content/}{$1.tex}
+EOF
+		shift
+	done;
+	
+	if $tflag ; then
+		cat >> "$t.tex" << EOF
+	\\end{nestedsection}
+	\\bibliographystyle{apalike}
+	\\bibliography{../bibliography}
+\\end{document}
+EOF
+	elif $dflag ; then
+		cat >> "$t.tex" << EOF
+	\\input{bibliography}
+\\end{document}
+EOF
+	fi
+	
+	if [[ -z $t ]]; then
+		t="$1"
+		count=$((count+1))
+		shift
+	fi
+
+	docs="$docs $t"
+done
+
+echo \*\*\* Processing $count documents: $docs \*\*\*
+
+for t in $docs ; do
+	echo \*\*\* "$t" about to be processed \*\*\*
+	make -f ../makefile DOC="$t"
+	echo \*\*\* "$t" processed \*\*\*
+
+	tFinal=`echo "$t" | sed -e 's/-wrapper$//'`
+	mv "$t.pdf" "../$tFinal.pdf"
+	open "../$tFinal.pdf"
+done
+wrappers=`ls -1 *-wrapper* 2>/dev/null`
+rm -f $wrappers
+if [[ -n $wrappers ]] ; then
+	echo \*\*\* Template files removed: $wrappers \*\*\*
+fi
[…]

### 2015-06-10 content/personal-details-ieee.tex
Added the IOS article style (I think this is the right one, though the call for papers section doesn't mention the call I thought we were working towards...), and migrated preambles and personal-details files to be for specific article styles

% Personal Details
\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
\begin{IEEEkeywords}
      Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation
\end{IEEEkeywords}
%\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
\author    {\IEEEauthorblockN{David L. Monks, Rehab ???, Dr. Nicholas M. Gibbins}
            \IEEEauthorblockA{
                              Web and Internet Science\\
                              Electronics and Computer Science\\
                              Physical Sciences and Engineering\\
                              University of Southampton\\
                              \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
                              }
            }
% TODO fix Rehab's details
\date{\today}

### 2015-06-10 content/personal-details-ios.tex
Added the IOS article style (I think this is the right one, though the call for papers section doesn't mention the call I thought we were working towards...), and migrated preambles and personal-details files to be for specific article styles

% Personal Details
\begin{frontmatter}
	\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
	\runningtitle{CoDeR Model and R4 Algorithm: Continuous Deductive Reasoning over Streamed RDF Data}

	\review{}{}{}

	\author{\fnms{David L.} \snm{Monks}}
	\sep
	\author{\fnms{Rehab} \snm{?}}
	and
	\author{\fnms{Nicholas M.} \snm{Gibbins}}
	\runningauthor{D.L. Monks et al.}
	\address{
	          Department of Electronics and Computer Science,\\
	          Faculty of Physical Sciences and Engineering,\\
	          University of Southampton,\\
	          University Rd., Southampton, SO17 1BJ,\\
	          United Kingdom\\
	          E-mail: \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
	}
	% TODO fix Rehab's details

	\begin{abstract}
		\input{abstract.tex}
	\end{abstract}

	\begin{keyword}
	      Deductive Reasoning\sep Streamed Data\sep Semantic Data\sep Continuous Processing\sep Plan Optimisation
	\end{keyword}

	\date{\today}

\end{frontmatter}

### 2015-06-10 content/contribution/continuous-datalog.tex
moved away from \forall symbols etc to avoid cluttering the axioms, instead leaving the text to specify such quantifiers

@@
 		\item\label{axiom:continuous datalog: window range leq 0}
-			A window of range ${{\leq} 0}$ over any stream $S$ at any time $t$ is always empty.
+			A window of range ${X \leq 0}$ over any stream $S$ at any time ${t \in \mathbb{N}}$ is always empty.
 			\begin{equation*}
-				\forall S \, \forall t \in \mathbb{N} \, \forall X \leq 0 
-					\left( W^{S,X}_{t} = \varnothing \right)
+				W^{S,X}_{t} = \varnothing
 			\end{equation*}
 		\item\label{axiom:continuous datalog: window composition}
-			A window of range $X$ over any stream $S$ at any given time $t$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at time $t$.
+			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given time ${t \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at time $t$.
 			\begin{equation*}
-				\forall S \, \forall t \in \mathbb{N} 
-					\left( W^{S,X}_t = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{t-i} \right)
+				W^{S,X}_t = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{t-i}
 			\end{equation*}
 		\item\label{axiom:continuous datalog: window disjointness}
-			A window of range $X$ over stream $S$ at time $t$ is disjoint from a second window of any range $Y$ over $S$ at any time after time ${t - X}$.
+			A window of any range ${X \in \mathbb{N}}$ over stream $S$ at time ${t \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any time after time ${t - X}$.
 			\begin{equation*}
-				\forall S \, \forall t,X,Y \in \mathbb{N} \, \forall X' \geq X 
-					\left( W^{S,X}_t \cap W^{S,Y}_{t-X'} = \varnothing \right)
+				W^{S,X}_t \cap W^{S,Y}_{t-X'} = \varnothing
 			\end{equation*}
@@
 		\item[Continuous Datalog Program]\label{def:continuous datalog: CDP}\hfill\\
-			A Continuous Datalog Program (${CDP}$) is composed of a set of Datalog axioms $A$, the union of a set of streams $S$ of Datalog facts and the range $R$ of a window over $S$.
-			Such a continuous program entails a triple composed of the axioms $A$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
-			The truth of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program $CDP$ and a time $t$ as arguments.
+			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, the union of a set of streams $S$ of Datalog facts and the range ${R \in \mathbb{N}}$ of a window over $S$.
+			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
+			The truth of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and a time ${t \in \mathbb{N}}$ as arguments.
 			\begin{equation*}
-				CDP = \{A,S,R\} \vDash \{A,P,N\}
+				CDP = \{A,S,R\} \vDash \{A',P,N\}
 			\end{equation*}
-		\item[State of a \emph{CDP}]\label{def:continuous datalog: CDPt}\hfill\\
-			The state of a \emph{CDP} at a given time $t$ (${CDP_t}$) is the union of the axioms $A$ and the window $W^{S,R}_{t}$.
+			\begin{equation*}
+				T^{CD} \left( CDP, t \right)
+			\end{equation*}
+		\item[State of a ${CDP}$]\label{def:continuous datalog: CDPt}\hfill\\
+			The state of a ${CDP}$ at a given time ${t \in \mathbb{N}}$ (${CDP_t}$) is the union of the axioms $A$ and the window $W^{S,R}_{t}$.
 			As both $A$ and $W^{S,R}_{t}$ are sets of Datalog axioms, a ${CDP_t}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the instantaneous truth of the ${CDP}$ at time $t$ by $T^{CD}$.
-			The instantaneous entailments of a $CDP_{t}$ are the union of the axioms $A$ and the set of entailments $E_{t}$ that are justified in part by a subset of $W^{S,R}_{t}$.
-			\begin{multline*}
-				\forall CDP = \{A,S,R\} \, \forall t \in \mathbb{N} \\
-					\left( CDP_t = A \cup W^{S,R}_t \vDash A \cup E_t \right)
-			\end{multline*}
+			The instantaneous entailments of a $CDP_{t}$ are the union of the persistent entailments of ${A \vDash A'}$ and the set of transient entailments $E_{t}$ that are justified at least in part by some subset of $W^{S,R}_{t}$.
 			\begin{equation*}
-				\forall t \in \mathbb{N} \left( T^{D} \left( CDP_t \right) = T^{CD} \left( CDP, t \right) \right)
+				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_t \right)
+			\end{equation*}
+			\begin{equation*}
+				CDP_t = A \cup W^{S,R}_t \vDash A' \cup E_t
 			\end{equation*}
@@
 		\item\label{axiom:continuous datalog: positive window increment}
-			The window of range 1 at any given time $t$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at time $t$ that were not entailed by the state of ${CDP}$ at the previous time instance ${t-1}$.
-			\begin{multline*}
-				\forall CDP \vDash \{A,P,N\} \, \forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \\
-					\left( W^{P,1}_{t} = E_{t} \setminus E_{t-1} \right)
-			\end{multline*}
+			The window of range 1 at any given time ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at time $t$ that were not entailed by the state of ${CDP}$ at the previous time instance ${t-1}$.
+			\begin{equation*}
+				W^{P,1}_{t} = \left( A' \cup E_{t} \right) \setminus \left( A' \cup E_{t-1} \right) = E_{t} \setminus E_{t-1}
+			\end{equation*}
 		\item\label{axiom:continuous datalog: negative window increment}
-			The window of range 1 at any given time $t$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous time instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at time $t$.
-			\begin{multline*}
-				\forall CDP \vDash \{A,P,N\} \, \forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \\
-					\left( W^{N,1}_{t} = E_{t-1} \setminus E_{t} \right)
-			\end{multline*}
+			The window of range 1 at any given time ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous time instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at time $t$.
+			\begin{equation*}
+				W^{N,1}_{t} = \left( A' \cup E_{t-1} \right) \setminus \left( A' \cup E_{t} \right) = E_{t-1} \setminus E_{t}
+			\end{equation*}
 		\item\label{axiom:continuous datalog: entailment precedes negation}
-			Facts may only be negated (i.e. appear in the stream $N$ at time $t$) after they have been entailed (i.e. appeared in the stream $P$ at some previous time ${t-X}$ where ${X > 0}$).
-			\begin{multline*}
-				\forall CDP = \{A,S,R\} \vDash \{A,P,N\} \, \forall t,X \in \mathbb{N} \, \forall Y \geq R \\
-					\left( W^{P,R}_{t} \cap W^{N,X}_{t-Y} = \varnothing \right)
-			\end{multline*}
+			Facts may only be negated (i.e.
+appear in the stream $N$ entailed by some ${CDP}$ at time ${t \in \mathbb{N}}$) after they have been entailed (i.e.
+appeared in the stream $P$ entailed by the same ${CDP}$ at some time ${t-X}$ where ${X \geq 0}$).
+			\begin{equation*}
+				W^{P,1}_{t} \cap W^{N,1}_{t-X} = \varnothing
+			\end{equation*}
 		\setcounter{continuousDatalogAxioms}{\theenumi}
@@
 		\item\label{axiom:continuous datalog: disjoint entailments}
-			The instantaneous set of stream-justified entailments $E_t$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{t-X}$ of the same ${CDP}$, where ${X \geq R}$.
-			\begin{multline*}
-				\forall CDP = \{A,S,R\} \, \forall CDP_t \vDash A \cup E_t \, \forall t \in \mathbb{N} \, \forall X \geq R \\
-					\left( E_t \cap E_{t-X} = \varnothing \right)
-			\end{multline*}
+			The instantaneous set of stream-justified entailments $E_t$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{t-X}$ of the same ${CDP}$ when ${X \geq R}$.
+			\begin{equation*}
+				E_t \cap E_{t-X} = \varnothing
+			\end{equation*}
 		\setcounter{continuousDatalogAxioms}{\theenumi}
@@
 	\begin{hyp}\labelhyp{continuous datalog: entailment}
-		For any ${CDP = \{A,S,R\} \vDash \{A,P,N\}}$, where ${CDP_t \vDash A \cup E_t}$, the set of facts in a window of range ${X \geq R}$ over the stream $P$ at time $t$ and not in a window of range ${Y \geq X}$ over the stream $N$ at $t$ is equal to the set of facts entailed by the state of the ${CDP}$ at time $t$, for any ${t \in \mathbb{N}}$:
-		\begin{multline*}
-			\forall CDP = \{A,S,R\} \vDash \{A,P,N\} \, \forall X \geq R \, \forall Y \geq X \\
-				\forall t \in \mathbb{N} \, \forall CDP_t \vDash A \cup E_t \left( W^{P,X}_t \setminus W^{N,Y}_t = E_t \right)
-		\end{multline*}
+		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_t \vDash A' \cup E_t}$ for any ${t \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given time ${t}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same time $t$ is equal to the set of facts $E_t$ transitively entailed by the ${CDP}$ at that time $t$:
+		\begin{equation*}
+			W^{P,X}_t \setminus W^{N,Y}_t = E_t
+		\end{equation*}
 	\end{hyp}
-	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams are of the same Range as that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary Range.
+	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
 	\begin{lem}\label{hyp:continuous datalog: entailment lemma}
-		For any ${CDP \vDash \{A,P,N\}}$, where ${CDP_t \vDash A \cup E_t}$%, the set of facts in the window of range $X$ over the stream $P$ and not in the window also of range $X$ over the stream $N$ is equal to the set of facts entailed by the state of the ${CDP}$ at time $t$ but not entailed by all states of the ${CDP}$ in the preceding $X$ time instances
-		:
-		\begin{multline*}
-			\forall CDP \vDash \{A,P,N\} \, \forall CDP_t \vDash A \cup E_t \, \forall t,X \in \mathbb{N} \, \forall Y \geq X \\
-				\left( W^{P,X}_t \setminus W^{N,Y}_t = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i} \right)
-		\end{multline*}
+		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_t \vDash A' \cup E_t}$ for any ${t \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given time ${t}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same time $t$ is equal to the set of transient entailments at that time $t$ less the intersection of all previous sets of transient entailments from time ${t-1}$ up to and including ${t-X}$:
+		\begin{equation*}
+			W^{P,X}_t \setminus W^{N,Y}_t = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i}
+		\end{equation*}
 	\end{lem}
 
-	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]
+	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
 		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every time ${t \in \mathbb{N}}$:
@@
 			& = E_t \cup \left( E_t \setminus \left( \mathop{\cap}^{R-1}_{i=1} E_{t-i} \cap \mathop{\cap}^{X}_{i=R+1} E_{t-i} \right) \right) \\
-			& = E_t
+			& = E_t \qedhere
 		\end{align*}
-		Thus showing that the RHS is equal to the LHS.
 	\end{proof}
@@
 			& = E_t \setminus E_{t-1} \\
-			& = E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-i}
+			& = E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-i} \qedhere
 		\end{align*}
@@
 			\begin{align*}
-				E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} & = W^{P,k+1}_{t+1} \setminus W^{P,k+1}_{t+1}
+				E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} & = W^{P,k+1}_{t+1} \setminus W^{P,k+1}_{t+1} \\
 				\text{By the \textbf{general case}} & = \mathop{\cup}^{k-1}_{i=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
@@
 				& = E_{t+1} \setminus \mathop{\cap}^{k}_{i=0} E_{t-i} \\
-				& = E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i}
+				& = E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} \qedhere
 			\end{align*}

### 2015-06-12 content/personal-details-els.tex
Added the CORRECT article class, and set up the template correctly apart from the problem listed below.  Also fixed the details for Rehab in all the personal-details-*.tex
PROBLEMS
- inductive step still overflows both column and page
- abstract or the rest of the frontmatter does not format properly in twocolumn mode

--- 2015-06-10 content/personal-details-ios.tex
@@
 	\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
-	\runningtitle{CoDeR Model and R4 Algorithm: Continuous Deductive Reasoning over Streamed RDF Data}
 
-	\review{}{}{}
-
-	\author{\fnms{David L.} \snm{Monks}}
-	\sep
-	\author{\fnms{Rehab} \snm{?}}
-	and
-	\author{\fnms{Nicholas M.} \snm{Gibbins}}
-	\runningauthor{D.L. Monks et al.}
+	\author{David L. Monks}
+	\ead{dm11g08@ecs.soton.ac.uk}
+	\author{Rehab Albeladi}
+	\ead{raab1g09@ecs.soton.ac.uk}
+	\author{Nicholas M. Gibbins}
+	\ead{nmg@ecs.soton.ac.uk}
+	% \ead{\{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk}
 	\address{
@@
 	          University Rd., Southampton, SO17 1BJ,\\
-	          United Kingdom\\
-	          E-mail: \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
+	          United Kingdom
 	}

### 2015-06-12 latexbuild.sh
added a "start" command to support automatic opening of PDF documents after build on windows machines, as "open" allows on Mac.

@@ -73,6 +73,7 @@ for t in $docs ; do
 	tFinal=`echo "$t" | sed -e 's/-wrapper$//'`
 	mv "$t.pdf" "../$tFinal.pdf"
 	open "../$tFinal.pdf"
+	start "../$tFinal.pdf"
 done
 wrappers=`ls -1 *-wrapper* 2>/dev/null`
 rm -f $wrappers

### 2015-06-12 content/contribution/continuous-datalog.tex
started to expand on Continuous Datalog... seems way too long at them moment so I'll need to discuss what can be cut with nmg

@@
-\begin{nestedsection}{Continuous Datalog Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
+\begin{nestedsection}{Continuous Datalog}{semantics}
+	The first requirement of stream reasoning is a well-defined semantics.
+	This must take into account the continuous nature of a subset of the base axioms of a stream reasoning problem, as well as its instantaneous truth and entailments, in line with the work of the reasoning community.
+	However, it must also provide a continuous interpretation of those instantaneous entailments in order to integrate with other stream-processing technologies.
+	More specifically, an appropriate semantics for stream reasoning on the semantic web must allow the expression of OWL 2 ontologies to some degree of expressivity, not sacrifice the precise semantics thereof whilst incorporating the semantics of stream processing, and allow for the continuous derivation of entailments in a concise manner.
+
+	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
+	This comprises the extension of Datalog programs to include sources of continuously generated/reported facts, and the application of sliding window semantics over those sources, thereby providing for the interpretation of Continuous Datalog programs as sequences of instantaneous states that are each valid positive Datalog programs.
+	As such, these states each inherit the semantics of positive Datalog.
+	Furthermore, it involves the continuous derivation of entailments from this sequence of instantaneous states as a set of streams of facts;
+	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
+	In addition, these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community, as proven in \refsec{semantics: proof for entailments}.
+
+	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
 	\begin{description}
-		\item[Stream]\label{def:continuous datalog: stream}\hfill\\
-			A stream is a \emph{sequence} of elements ordered by some temporal characteristic of each element, e.g. its creation time, entailment time, or arrival time.
-			Each element is uniquely identifiable by its contents (which may duplicate that of another element in the stream) and its position in the sequence.
-		\item[Window]\label{def:continuous datalog: window}\hfill\\
-			A window of range $X$ over any stream $S$ at any given time ${t \in \mathbb{N}}$ is the \emph{set} of elements that occur in $S$ during the interval ${(t-X,t]}$, and is denoted ${W^{S,X}_{t}}$.
+		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
+			An instant is a discrete point in time.
+			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
+			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
+		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
+			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g.
+the time at which each was entailment by some process at the source of the stream.
+			Each instance of a fact is considered to be uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
+			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
+			In practice, if two instances of the same fact also share an entailment time, they are considered to be the same instance.
+		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
+			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
+			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
 	\end{description}
+	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog.
 	\begin{enumerate}\newcounter{continuousDatalogAxioms}
 		\item\label{axiom:continuous datalog: window range leq 0}
-			A window of range ${X \leq 0}$ over any stream $S$ at any time ${t \in \mathbb{N}}$ is always empty.
+			A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
 			\begin{equation*}
-				W^{S,X}_{t} = \varnothing
+				W^{S,X}_{i} = \varnothing
 			\end{equation*}
 		\item\label{axiom:continuous datalog: window composition}
-			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given time ${t \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at time $t$.
+			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
 			\begin{equation*}
-				W^{S,X}_t = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{t-i}
+				W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
 			\end{equation*}
 		\item\label{axiom:continuous datalog: window disjointness}
-			A window of any range ${X \in \mathbb{N}}$ over stream $S$ at time ${t \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any time after time ${t - X}$.
+			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
 			\begin{equation*}
-				W^{S,X}_t \cap W^{S,Y}_{t-X'} = \varnothing
+				\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
 			\end{equation*}
+		% \item\label{axiom:continuous datalog: stream disjointness}
+		% 	The contents of any two streams are disjoint, thus a window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over any other stream ${Z \neq S}$ at any instant ${i \in \mathbb{N}}$.
+		% 	\begin{equation*}
+		% 		W^{S,X}_{i} \cap W^{Z,Y}_{i} = \varnothing
+		% 	\end{equation*}
 		\setcounter{continuousDatalogAxioms}{\theenumi}
 	\end{enumerate}
+
+	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
 	\begin{description}
-		\item[Continuous Datalog Program]\label{def:continuous datalog: CDP}\hfill\\
-			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, the union of a set of streams $S$ of Datalog facts and the range ${R \in \mathbb{N}}$ of a window over $S$.
-			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
-			The truth of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and a time ${t \in \mathbb{N}}$ as arguments.
+		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
+			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
+			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
+			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.
 			\begin{equation*}
@@
 			\end{equation*}
+		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
+			The state of a ${CDP}$ at a given instant ${t \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
+			As both $A$ and $W^{S,R}_{i}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $t$ by $T^{CD}$.
+			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$.
 			\begin{equation*}
-				T^{CD} \left( CDP, t \right)
-			\end{equation*}
-		\item[State of a ${CDP}$]\label{def:continuous datalog: CDPt}\hfill\\
-			The state of a ${CDP}$ at a given time ${t \in \mathbb{N}}$ (${CDP_t}$) is the union of the axioms $A$ and the window $W^{S,R}_{t}$.
-			As both $A$ and $W^{S,R}_{t}$ are sets of Datalog axioms, a ${CDP_t}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the instantaneous truth of the ${CDP}$ at time $t$ by $T^{CD}$.
-			The instantaneous entailments of a $CDP_{t}$ are the union of the persistent entailments of ${A \vDash A'}$ and the set of transient entailments $E_{t}$ that are justified at least in part by some subset of $W^{S,R}_{t}$.
-			\begin{equation*}
-				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_t \right)
+				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_{i} \right)
 			\end{equation*}
 			\begin{equation*}
-				CDP_t = A \cup W^{S,R}_t \vDash A' \cup E_t
+				CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
 			\end{equation*}
 	\end{description}
-	Both streams entailed by a ${CDP}$ are defined as sequences of the windows of range 1 over themselves, with those windows being defined as the difference between the sets of entailments of the ${CDP}$ between two consecutive instances, as in Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}.
+	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
 	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
 		\item\label{axiom:continuous datalog: positive window increment}
-			The window of range 1 at any given time ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at time $t$ that were not entailed by the state of ${CDP}$ at the previous time instance ${t-1}$.
+			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $t$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${t-1}$.
 			\begin{equation*}
-				W^{P,1}_{t} = \left( A' \cup E_{t} \right) \setminus \left( A' \cup E_{t-1} \right) = E_{t} \setminus E_{t-1}
+				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
 			\end{equation*}
 		\item\label{axiom:continuous datalog: negative window increment}
-			The window of range 1 at any given time ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous time instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at time $t$.
+			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at instant $t$.
 			\begin{equation*}
-				W^{N,1}_{t} = \left( A' \cup E_{t-1} \right) \setminus \left( A' \cup E_{t} \right) = E_{t-1} \setminus E_{t}
+				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
 			\end{equation*}
 		\item\label{axiom:continuous datalog: entailment precedes negation}
-			Facts may only be negated (i.e.
-appear in the stream $N$ entailed by some ${CDP}$ at time ${t \in \mathbb{N}}$) after they have been entailed (i.e.
-appeared in the stream $P$ entailed by the same ${CDP}$ at some time ${t-X}$ where ${X \geq 0}$).
+			Facts may only be negated after they have been entailed.
+			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 
 			\begin{equation*}
-				W^{P,1}_{t} \cap W^{N,1}_{t-X} = \varnothing
+				\forall Y \geq 1 \left( W^{P,1}_{i} \cap W^{N,1}_{i-Y} = \varnothing \right)
 			\end{equation*}
@@
 
-	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
-		\item\label{axiom:continuous datalog: disjoint entailments}
-			The instantaneous set of stream-justified entailments $E_t$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{t-X}$ of the same ${CDP}$ when ${X \geq R}$.
-			\begin{equation*}
-				E_t \cap E_{t-X} = \varnothing
-			\end{equation*}
-		\setcounter{continuousDatalogAxioms}{\theenumi}
-	\end{enumerate}
+	From this, we can see that each instance $e$ of an entailment in some $E_{i}$ may take the same form as any other fact in a stream of facts, being ${e = p(s,o):t}$, where $t$ is the real time associated with the instant in which $e$ is first entailed by the ${CDP}$ (i.e.
+it appears in $E_{i}$ but not $E_{i-1}$, and so is part of $W^{P,1}_{i}$).
+	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
+	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e.
+they appear in $E_{i-1}$ but not $E_{i}$, and so is part of $W^{N,1}_{i}$).
 
-	\begin{hyp}\labelhyp{continuous datalog: entailment}
-		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_t \vDash A' \cup E_t}$ for any ${t \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given time ${t}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same time $t$ is equal to the set of facts $E_t$ transitively entailed by the ${CDP}$ at that time $t$:
-		\begin{equation*}
-			W^{P,X}_t \setminus W^{N,Y}_t = E_t
-		\end{equation*}
-	\end{hyp}
-	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
-	\begin{lem}\label{hyp:continuous datalog: entailment lemma}
-		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_t \vDash A' \cup E_t}$ for any ${t \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given time ${t}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same time $t$ is equal to the set of transient entailments at that time $t$ less the intersection of all previous sets of transient entailments from time ${t-1}$ up to and including ${t-X}$:
-		\begin{equation*}
-			W^{P,X}_t \setminus W^{N,Y}_t = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i}
-		\end{equation*}
-	\end{lem}
+	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
+	However, it can be shown from these axioms that it is also possible to reconstruct the instantaneous set of entailments of a ${CDP}$ at a given instant from the contents of windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
 
-	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
-		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every time ${t \in \mathbb{N}}$:
-		\begin{align*}
-			E_t & = W^{P,X}_t \setminus W^{N,Y}_t \\
-			\text{By Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_t \setminus \mathop{\cap}^{X}_{i=1} E_{t-i} \\
-			& = \mathop{\cup}^{X}_{i=1} \left( E_t \setminus E_{t-i} \right) \\
-			& = \left( E_t \setminus E_{t-R} \right) \cup \mathop{\cup}^{R-1}_{i=1} \left( E_t \setminus E_{t-i} \right) \cup \mathop{\cup}^{X}_{i=R+1} \left( E_t \setminus E_{t-i} \right) \\
-			\text{By Axiom~\ref{axiom:continuous datalog: disjoint entailments}} & = E_t \cup \mathop{\cup}^{R-1}_{i=1} \left( E_t \setminus E_{t-i} \right) \cup \mathop{\cup}^{X}_{i=R+1} \left( E_t \setminus E_{t-i} \right) \\
-			& = E_t \cup \left( E_t \setminus \left( \mathop{\cap}^{R-1}_{i=1} E_{t-i} \cap \mathop{\cap}^{X}_{i=R+1} E_{t-i} \right) \right) \\
-			& = E_t \qedhere
-		\end{align*}
-	\end{proof}
-
-	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${X \in \mathbb{N}}$, regardless of the Range of of the ${CDP}$ entailing $P$ and $N$.
-
-	\begin{proof}[Proof of Lemma~\ref{hyp:continuous datalog: entailment lemma} by Induction]
-		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${X = k}$ and ${Y \geq X}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${X = k + 1}$, where ${k \in \mathbb{N}}$.
-
-		Firstly, the \textbf{general case} can be defined in terms of sub-windows of range $1$ by Axiom~\ref{axiom:continuous datalog: window composition} over the entailed streams $P$ and $N$, for which Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment} provide an equivalence to sets of entailments.
-		\begin{align*}
-			W^{P,X}_t \setminus W^{P,Y}_t & = \\
-			\text{By Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{X-1}_{i=0}W^{P,1}_{t-i} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{t-j} \\
-			& = \mathop{\cup}^{X-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{t-j} \right) \\
-			& = \mathop{\cup}^{X-1}_{i=0}\mathop{\cap}^{Y-1}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
-			\text{By Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } Y \geq X & = \mathop{\cup}^{X-1}_{i=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right)
-		\end{align*}
-		The general case also shows that any negative entailments within the window of range $Y$ over $N$ not in the window of range $X$ over $N$ do not influence the composition of the set of values ${W^{P,X}_t \setminus W^{P,Y}_t}$.
-
-		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${X = 1}$ (therefore ${Y \geq 1}$).
-		\begin{align*}
-			E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-1} & = W^{P,1}_t \setminus W^{N,Y}_t \\
-			\text{By the \textbf{general case}} & = W^{P,1}_t \setminus W^{N,1}_t \\
-			\text{By Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E_t \setminus E_{t-1} \right) \setminus \left( E_{t-1} \setminus E_t \right) \\
-			& = E_t \setminus \left( E_{t-1} \setminus \left( E_{t-1} \setminus E_t \right) \right) \\
-			& = E_t \setminus \left( E_{t-1} \cap E_t \right) \\
-			& = \left( E_t \setminus E_{t-1} \right) \cup \left( E_t \setminus E_t \right) \\
-			& = E_t \setminus E_{t-1} \\
-			& = E_t \setminus \mathop{\cap}^{1}_{i=1} E_{t-i} \qedhere
-		\end{align*}
-
-		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
-		\begin{figure}[p]
-			\centering
-			\caption[Proof of Entailment Lemma for Continuous Datalog]{Proof of Inductive Case of the Entailment Lemma for Continuous Datalog.}
-			\labelfig{continuous datalog: entailment lemma inductive case}
-			\begin{align*}
-				E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} & = W^{P,k+1}_{t+1} \setminus W^{P,k+1}_{t+1} \\
-				\text{By the \textbf{general case}} & = \mathop{\cup}^{k-1}_{i=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
-				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
-				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \\
-				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{t-j} \right) \\
-				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \left( W^{N,1}_{t+1} \cup \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \right) \right) \\
-				& = \left( W^{P,1}_{t+1} \setminus W^{N,1}_{t+1} \right) \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \setminus W^{N,1}_{t+1} \right) \\
-				& = \left( W^{P,1}_{t+1} \cup \mathop{\cup}^{k-1}_{i=0} \left( W^{P,1}_{t-i} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{t-j} \right) \right) \setminus W^{N,1}_{t+1} \\
-				& = \left( W^{P,1}_{t+1} \cup \mathop{\cup}^{k-1}_{i=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{t-i} \setminus W^{N,1}_{t-j} \right) \right) \setminus W^{N,1}_{t+1} \\
-				& = \left( W^{P,1}_{t+1} \cup \left( W^{P,k}_{t} \setminus W^{N,k}_{t} \right) \right) \setminus W^{N,1}_{t+1} \\
-				\text{\textbf{Inductive Step}:}& = \left( W^{P,1}_{t+1} \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \setminus W^{N,1}_{t+1} \\
-				\text{By Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( \left( E_{t+1} \setminus E_{t} \right) \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \setminus \left( E_{t} \setminus E_{t+1} \right) \\
-				& = \left( E_{t+1} \setminus E_{t} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \cup \left( E_t \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \\
-				& = \left( E_{t+1} \setminus E_{t} \setminus \left( E_{t} \setminus E_{t+1} \right) \right) \cup \left( E_t \setminus \left( E_{t} \setminus E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
-				& = \left( E_{t+1} \setminus \left( E_{t} \cap E_{t+1} \right) \right) \cup \left( \left( E_t \cap E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
-				& = \left( E_{t+1} \setminus E_{t} \right) \cup \left( \left( E_t \cap E_{t+1} \right) \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
-				& = \left( E_{t+1} \setminus E_{t} \right) \cup \left( E_{t+1} \cap \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
-				& = \left( E_{t+1} \cap E^C_{t} \right) \cup \left( E_{t+1} \cap \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right)\\
-				& = E_{t+1} \cap \left( E^C_{t} \cup \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
-				& = E_{t+1} \setminus \left( E_{t} \setminus \left( E_{t} \setminus \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \right) \\
-				& = E_{t+1} \setminus \left( E_{t} \cap \mathop{\cap}^{k}_{i=1} E_{t-i} \right) \\
-				& = E_{t+1} \setminus \mathop{\cap}^{k}_{i=0} E_{t-i} \\
-				& = E_{t+1} \setminus \mathop{\cap}^{k+1}_{i=1} E_{t+1-i} \qedhere
-			\end{align*}
-		\end{figure}
-	\end{proof}
+	\subimport{continuous-datalog/}{proof-of-entailment-semantics.tex}
 \end{nestedsection}

### 2015-06-12 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
started to expand on Continuous Datalog... seems way too long at them moment so I'll need to discuss what can be cut with nmg

\begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}

	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: disjoint entailments}
			The instantaneous set of stream-justified entailments $E_{i}$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{i-X}$ of the same ${CDP}$ when ${X \geq R}$.
			\begin{equation*}
				E_{i} \cap E_{i-X} = \varnothing
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	\begin{hyp}\labelhyp{continuous datalog: entailment}
		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i}
		\end{equation*}
	\end{hyp}
	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
	\begin{lem}\label{hyp:continuous datalog: entailment lemma}
		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d}
		\end{equation*}
	\end{lem}

	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$:
		\begin{align*}
			E_{i} & = W^{P,X}_{i} \setminus W^{N,Y}_{i} \\
			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d} \\
			& = \mathop{\cup}^{X}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			\text{Axiom~\ref{axiom:continuous datalog: disjoint entailments}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\
			& = E_{i} && \qedhere
		\end{align*}
	\end{proof}

	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${X \in \mathbb{N}}$, regardless of the Range of of the ${CDP}$ entailing $P$ and $N$.

	\begin{proof}[Proof of Lemma~\ref{hyp:continuous datalog: entailment lemma} by Induction]
		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${X = k}$ and ${Y \geq X}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${X = k + 1}$, where ${k \in \mathbb{N}}$.

		Firstly, the \textbf{general case} can be defined in terms of sub-windows of range $1$ by Axiom~\ref{axiom:continuous datalog: window composition} over the entailed streams $P$ and $N$, for which Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment} provide an equivalence to sets of entailments.
		\begin{align*}
			W^{P,X}_{i} \setminus W^{P,Y}_{i} & = \\
			\text{Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{X-1}_{d=0}W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \\
			& = \mathop{\cup}^{X-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \right) \\
			& = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{Y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
			\text{Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } Y \geq X & = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right)
		\end{align*}
		The general case also shows that any negative entailments within the window of range $Y$ over $N$ not in the window of range $X$ over $N$ do not influence the composition of the set of values ${W^{P,X}_{i} \setminus W^{P,Y}_{i}}$.

		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${X = 1}$ (therefore ${Y \geq 1}$).
		\begin{align*}
			E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-1} & = W^{P,1}_{i} \setminus W^{N,Y}_{i} \\
			\text{\textbf{General Case}} & = W^{P,1}_{i} \setminus W^{N,1}_{i} \\
			\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E_{i} \setminus E_{i-1} \right) \setminus \left( E_{i-1} \setminus E_{i} \right) \\
			& = E_{i} \setminus \left( E_{i-1} \setminus \left( E_{i-1} \setminus E_{i} \right) \right) \\
			& = E_{i} \setminus \left( E_{i-1} \cap E_{i} \right) \\
			& = \left( E_{i} \setminus E_{i-1} \right) \cup \left( E_{i} \setminus E_{i} \right) \\
			& = E_{i} \setminus E_{i-1} \\
			& = E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-d} && \qedhere
		\end{align*}

		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
		\begin{figure}[p]
			\centering
			\caption[Proof of Entailment Lemma for Continuous Datalog]{Proof of Inductive Case of the Entailment Lemma for Continuous Datalog.}
			\labelfig{continuous datalog: entailment lemma inductive case}
			\begin{align*}
				E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
				\text{\textbf{General Case}} & = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{i-j} \right) \\
				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \left( W^{N,1}_{i+1} \cup \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \right) \right) \\
				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
				& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
				\text{\textbf{Inductive Step}:}& = \left( W^{P,1}_{i+1} \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus W^{N,1}_{i+1} \\
				\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus \left( E_{i} \setminus E_{i+1} \right) \\
				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \\
				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
				& = \left( E_{i+1} \setminus \left( E_{i} \cap E_{i+1} \right) \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
				& = \left( E_{i+1} \cap E^C_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right)\\
				& = E_{i+1} \cap \left( E^C_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
				& = E_{i+1} \setminus \left( E_{i} \setminus \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
				& = E_{i+1} \setminus \left( E_{i} \cap \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
				& = E_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E_{i-d} \\
				& = E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} \qedhere
			\end{align*}
		\end{figure}
	\end{proof}
\end{nestedsection}

### 2015-06-16 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
moved Hypothesis and Lemma titles to their own lines
forced inductive step to span both columns in 2column layout mode

@@
 
-	\begin{hyp}\labelhyp{continuous datalog: entailment}
+	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
 		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
@@
 	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
-	\begin{lem}\label{hyp:continuous datalog: entailment lemma}
+	\begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
 		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
@@
 		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
-		\begin{figure}[p]
+		\begin{figure*}[p]
 			\centering
@@
 			\end{align*}
-		\end{figure}
+		\end{figure*}
 	\end{proof}

### 2015-06-16 content/contribution/continuous-datalog.tex
concluded section on continuous datalog by bringing up the concept of "valid windows" as a generalisation of "sliding windows" (i.e. windows of data valid by some definition as opposed to windows of data valid by the specific definition of a duration from receipt).

@@
 			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
+			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
 			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
@@
 	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e.
-they appear in $E_{i-1}$ but not $E_{i}$, and so is part of $W^{N,1}_{i}$).
+they appear in $E_{i-1}$ but not $E_{i}$, and so are part of $W^{N,1}_{i}$).
 
 	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
-	However, it can be shown from these axioms that it is also possible to reconstruct the instantaneous set of entailments of a ${CDP}$ at a given instant from the contents of windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
-
-	\subimport{continuous-datalog/}{proof-of-entailment-semantics.tex}
+	However, it is proven in \refapp{semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
+	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
+	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
+	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
+	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
 \end{nestedsection}

### 2015-06-16 content/contribution/continuous-datalog.tex
minor modifications to cite IStream/DStream, and explain "entailment precedes negation" in terms of justification from windows over S

@@
 			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
+			Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
 			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.
@@
 		\item\label{axiom:continuous datalog: entailment precedes negation}
-			Facts may only be negated after they have been entailed.
+			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
 			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 

### 2015-06-16 content/contribution/CoDeR.tex
started work on CoDeR data model

@@
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-	Stream-based Data Model
+	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog.
+	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$s, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
+	% The primary difference between the CoDeR model and the stream-to-stream algebras of existing continuous semantic query engines \citep{CQELS,EP-SPARQL} is the nature of the windows maintained by operators, which are \emph{valid windows} rather than \emph{sliding windows}.
 
-	Minimal Algebra
+	% Stream-based Data Model
+	The data model of CoDeR differs from that of many RDF stream processing engines \citep{CQELS,EP-SPARQL} in that the streams each express both entailment \emph{and} negation streams, e.g.
+streams $P$ and $N$ of a ${CDP}$.
+	Due to Axiom~\ref{axiom:continuous datalog: entailment precedes negation} of Continuous Datalog, and the fact that the instant at which data will leave $W^{S,R}_{i}$ can be calculated in advance from the instant of its arrival and the value of $R$, the instant of a fact instance's negation may be determined at the time of its entailment:
+	the negation instant $i_{n}$ of a fact instance is the earliest instant ${i + 1}$ in which any of the base fact instance's in $W^{S,R}_{i}$ that justify it will no longer be found in $W^{S,R}_{i+1}$.
+	This allows a logical negation stream to be expressed as \emph{valid-time} annotations \citep{TA-SPARQL,sparkwave} on the instances of entailments in the corresponding entailment stream, making each entailment instance on a CoDeR output stream a quad composed of the fact ${(s,p,o)}$ and its negation instant $i_{n}$: ${\langle (s,p,o),i_{n} \rangle}$
+
+	% Minimal Algebra
 

### 2015-06-17 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
merged some less important and more obvious steps of the proof of the inductive step together so that it fits on a single page.

@@
 				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
 				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \left( W^{N,1}_{i+1} \cup \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \right) \right) \\
 				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
-				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
 				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
 				& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
-				\text{\textbf{Inductive Step}:}& = \left( W^{P,1}_{i+1} \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus W^{N,1}_{i+1} \\
-				\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus \left( E_{i} \setminus E_{i+1} \right) \\
+				\text{\textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}:}& = \left( \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus \left( E_{i} \setminus E_{i+1} \right) \\
 				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \\
@@
 				& = E_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E_{i-d} \\
-				& = E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} \qedhere
+				& = E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} && \qedhere
 			\end{align*}

### 2015-06-17 content/contribution/CoDeR.tex
made some major changes and extensions to what I had originally regarding the CoDeR streamed data model. Unfortunately, it is a horrible mess when reading it back, with monstrous sentences and so much clarification sentence by sentence that it is near impossible to understand overall...

--- 2015-06-10 content/contribution/CoDeR.tex
@@
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-	Stream-based Data Model
+	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog.
+	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
+	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
+	% The primary difference between the CoDeR model and the stream-to-stream algebras of existing continuous semantic query engines \citep{CQELS,EP-SPARQL} is the nature of the windows maintained by operators, which are \emph{valid windows} rather than \emph{sliding windows}.
 
-	Minimal Algebra
+	% Stream-based Data Model
+	The data model of CoDeR differs from that of many RDF stream processing engines \citep{CQELS,EP-SPARQL} in that each stream entailed within the model expresses both the entailment \emph{and} negation streams of the semantics, e.g.
+the streams $P$ and $N$ of a ${CDP}$.
+	Due to Axiom~\ref{axiom:continuous datalog: entailment precedes negation} of Continuous Datalog, and the fact that the instant at which a datum will leave $W^{S,R}_{i}$ can be calculated in advance from the instant of the datum's arrival and the value of $R$, the instant of a fact instance's negation may be determined at the time of its entailment:
+	the negation instant $i_{n}$ of a fact instance is the earliest instant $i_{n}$ in which any of the base fact instances in $W^{S,R}_{i_n - 1}$ that justify it will no longer be found in $W^{S,R}_{i_n}$.
+	This allows a logical negation stream to be expressed as \emph{valid-time} annotations \citep{SemanticStreamingManagement,sparkwave} on the instances of entailments in the corresponding entailment stream, making each entailment instance on a CoDeR output stream $O$ a triple composed of the fact ${f = (s,p,o)}$, its entailment instant $i_{e}$ and its negation instant $i_{n}$: ${\langle f,i_{e},i_{n} \rangle}$
+
+	As the algebra is composed of stream-to-stream operators for pattern-matching, each inter-operator stream is a sequence of instances of ``matches'' to a specific sub-graph-pattern of the body clause of one of the axiomatic rules $A$.
+	Each inter-operator stream can, therefore, be considered to be ``entailed'' by the axioms $A$, the window of range $R$ over the base streamed data $S$ and, crucially, the body-clause-fragment expressed by the sequence of operators culminating in the stream's source operator;
+	as such, each match instance in the stream must be annotated with its valid-time interval as described in relation to output streams, in order to express the semantic streams of entailments and negations for the specific body-clause-fragment.
+	Furthermore, the nature of a match instance as a match for some RDF graph pattern makes it an instance of an RDF graph, rather than a set of instances of RDF triples.
+	This means that a match instance is a triple composed of an RDF graph ${g = (f_1,\dots,f_k)}$, its entailment instant $i_{e}$ and its negation instant $i_{n}$: ${\langle g,i_{e},i_{n} \rangle}$
+
+	While the input streams $S$ may be composed of valid-time annotated RDF triples, if published by systems that share the entailment semantics of Continuous Datalog, they may alternately be composed of unannotated RDF triples to allow compatibility with the results of existing RDF stream processing engines with distinct semantics.
+	These unannotated triples must be annotated upon arrival at a CoDeR-based system with valid-times based on the instant of their arrival and the range $R$ of the ${CDP}$ expressed by the system, thereby imposing the entailment semantics of Continuous Datalog on the data.
+	In complement, the output stream $O$ of a CoDeR-based system expressing a ${CDP \vDash \{A',P,N\}}$ may have the valid-time annotations removed, making it equivalent to the stream $P$ in isolation.
+	While this sacrifices the correctness of the entailment semantics of Continuous Datalog, it allows entailments from a CoDeR-based system to be used as input for existing stream processing systems that lack such strict semantics.
+
+	% Minimal Algebra
 

### 2015-06-17 content/contribution/continuous-datalog.tex
fixed a reference we will probably remove

@@
 	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
-	However, it is proven in \refapp{semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
+	However, it is proven in \ref{sec:semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
 	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]

### 2015-06-18 content/contribution/continuous-datalog.tex
slight modifications to make the nature of entailment instances consistent

--- 2015-06-16 content/contribution/continuous-datalog.tex
@@
 the time at which each was entailment by some process at the source of the stream.
-			Each instance of a fact is considered to be uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
+			Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
 			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
-			In practice, if two instances of the same fact also share an entailment time, they are considered to be the same instance.
 		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
@@
 
-	From this, we can see that each instance $e$ of an entailment in some $E_{i}$ may take the same form as any other fact in a stream of facts, being ${e = p(s,o):t}$, where $t$ is the real time associated with the instant in which $e$ is first entailed by the ${CDP}$ (i.e.
-it appears in $E_{i}$ but not $E_{i-1}$, and so is part of $W^{P,1}_{i}$).
+	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes entailed by the ${CDP}$
+	(i.e. when it becomes newly justified by some subset of instances in $W^{S,R}_{i_{e}}$).
 	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
-	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e.
-they appear in $E_{i-1}$ but not $E_{i}$, and so are part of $W^{N,1}_{i}$).
+	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
+	(i.e.
+the first instant ${i_{n} \geq i_{e}}$ in which $W^{S,R}_{i_{n}}$ no longer contains the subset of fact instances first present in $W^{S,R}_{i_{e}}$ that justified the entailment instance ${p(s,o):t}$).
 
 	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
-	However, it is proven in \refapp{semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
+	However, it is proven in \ref{sec:semantics: proof for entailments}, from these axioms, that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
 	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]

### 2015-06-18 content/contribution/CoDeR.tex
moved first draft of CoDeR description to backup file, started again in used file

--- 2015-06-10 content/contribution/CoDeR.tex
@@
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-	Stream-based Data Model
+	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog.
+	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
+	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
 
-	Minimal Algebra
+	% Stream-based Data Model
+
+	% Minimal Algebra
 

### 2015-06-18 content/contribution/CoDeR.tex
started fleshing out the data model: starting from the form of the model and working back to the semantics expressed, rather than starting with the semantics and working forward to the form of the model (see firstDraft copy)

@@
 	% Stream-based Data Model
+	The data model of CoDeR is composed of two classes of streams.
+	The first are \emph{input}/\emph{output} streams, and provide the semantics of streams $S$, $P$ and $N$ for any ${CDP}$ expressed by a CoDeR-based system;
+	they are sequences of instances of RDF triples, where an instance of an RDF triple is an RDF triple annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle (s,p,o),i_{e},i_{n} \rangle}$.
+	This interval for any entailed fact instance is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}:
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=0.45\textwidth]{intersected-intervals}
+		\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
+		\labelfig{intersected-intervals}
+	\end{figure}
+	if ${A = \{c\,\text{:-}\,a\,,\,b\,.\}}$ and ${W^{S,3}_{2} = \{ \langle a,1,3 \rangle , \langle b,2,4 \rangle \}}$, then ${CDP_{2} = A \cup W^{S,3}_{2} \vDash \{ \langle a,1,3 \rangle , \langle b,2,4 \rangle , \langle c,2,3 \rangle \}}$, where $a$, $b$ and $c$ are RDF triples.
+	It is in the base case of this derivation that the semantic window range $R$ of a ${CDP}$ is expressed:
+	each fact instance arriving on an \emph{input} stream has a valid interval from the time of its arrival at the system for the duration of the semantic window expressed by that range $R$.
+
+	Given that the ordering of instances in the \emph{output} stream by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives all streams of CoDeR a dual semantics with respect to Continuous Datalog.
 

### 2015-06-18 content/contribution/CoDeR.tex
finished describing the streamed data model of CoDeR in a more coherent manner.  still a few long sentences, but not as bad now :-)

@@
 	they are sequences of instances of RDF triples, where an instance of an RDF triple is an RDF triple annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle (s,p,o),i_{e},i_{n} \rangle}$.
-	This interval for any entailed fact instance is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}:
+	This interval, for any entailed fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}:
 	\begin{figure}[t]
@@
 	It is in the base case of this derivation that the semantic window range $R$ of a ${CDP}$ is expressed:
-	each fact instance arriving on an \emph{input} stream has a valid interval from the time of its arrival at the system for the duration of the semantic window expressed by that range $R$.
+	each fact instance arriving on an input stream has a valid interval from the time of its arrival at the system for a duration equal to the range $R$.
 
-	Given that the ordering of instances in the \emph{output} stream by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives all streams of CoDeR a dual semantics with respect to Continuous Datalog.
+	The second class of streams the make up the CoDeR data model are \emph{inter-operator} streams, which are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
+	These streams are sequences of the partial results of processing between one step in the production of entailments and the next.
+	As the production of entailments in CoDeR is cast as a pattern matching problem, all partial results of this process will take the form of instances of RDF \emph{graphs};
+	for a given inter-operator stream, each such RDF graph will match the same graph pattern, that being a sub-pattern of the body of one of the axiomatic rules $A$.
+	As with instances of RDF triples, instances of RDF graphs are RDF graphs that are annotated with the time interval for which they are ``valid'': ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
+	As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same semantics of validity may be applied to these partial matches as to the true entailments of the model, and their valid-time intervals may derived from their justifying instances in the same manner.
+	It follows that inter-operator streams are, functionally, a general case of input/output streams, as the latter are simply sequences of instances of singleton RDF graphs matching the pattern ${\{(?s,?p,?o)\}}$.
+
+	Given that the ordering of instances in the output stream of a CoDeR-based system by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives the output streams of CoDeR a dual semantics with respect to Continuous Datalog.
+	Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by a specific pipeline of CoDeR operators applied to some set of continuous RDF data sources.
 

### 2015-06-19 content/contribution/R4.tex
migrated Rehab's draft into the latex, including using a new package in the template for including "listings"

@@
 \begin{nestedsection}{R4: Rule-based Reasoning using Rete\dots Roar}{implementation}
-	Data
+	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
+	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
+	The system architecture is shown in \reffig{R4-architecture}.
+	\begin{figure}[t]
+		\centering
+		\includegraphics[width=0.45\textwidth]{R4-architecture}
+		\caption{The system architecture of R4.}
+		\labelfig{R4-architecture}
+	\end{figure}
 
-	Operators
+	A rule document containing any number of rules is submitted to the system.
+	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
+	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
+	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in Listing~\ref{lst-S-RIF-Core}.
+	\begin{figure*}
+		\centering
+		\begin{lstlisting}[caption=Streaming RIF-Core Grammar,label=lst-S-RIF-Core]
+Document       ::= IRIMETA? 'Document' '(' Base? Prefix* Import* Group?')'
+Base           ::= 'Base' '(' ANGLEBRACKIRI ')'
+Prefix         ::= 'Prefix' '(' Name ANGLEBRACKIRI ')'
+Import         ::= IRIMETA? 'Import' '(' LOCATOR PROFILE? ')'
+Group          ::= IRIMETA? 'Group' '(' (RULE | Group)*')' 
+RULE           ::= ((IRIMETA? 'Forall' Var+ '(' CLAUSE')') | CLAUSE)
+CLAUSE         ::= Implies | ATOMIC
+Implies        ::= IRIMETA? (ATOMIC | 'And' '(' ATOMIC* ')') ':-' FORMULA
+LOCATOR        ::= ANGLEBRACKIRI
+PROFILE        ::= ANGLEBRACKIRI Window?
+Window         ::= Number TimeUnit
+TimeUnit       ::= 'ms' | 's' | 'm' | 'h'
+		\end{lstlisting}
+	\end{figure*}
 
-%	\subimport{conclusions/}{future-work.tex}
+	These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete networks.
+	The optimizer employs simple known heuristics to generate a good plan.
+	These heuristics include sharing nodes between rules where possible, avoiding Cartesian products as much as possible by joining patterns that have common variables, pushing more selective patterns (the ones with less variables) earlier in the network to minimize intermediate results.
+
+	The optimizer instantiates a Rete network for background reasoning if required and another network for generic rules.
+	The first network feeds into the second one and is also re-entrant to enable iterative inference of results, e.g. the calculation of transitive closure.
+
+	These networks read TBox input data, and receive streaming data in the form of RDF streams and operate directly on them.
+	We chose this RDF-native approach as opposed to reusing existing technologies as in \citep{C-SPARQL,streaming-sparql} to allow full control over the low-level operators.
+	This can offer maximum optimization opportunities such as adaptively optimizing the network topologies, which is our future work.
+
+	\begin{nestedsection}{Continuous Reasoning}{implementation: continuous reasoning}
+		The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, can be well fit into the stream reasoning model as it operates incrementally.
+		The RETE algorithm can process large data sets efficiently because it avoids iterating over both data elements (facts or working memory) and over the production rules.
+		To avoid iteration over data elements, the RETE algorithm stores with each condition (or pattern), a list of the data elements that it matches.
+		These lists are updated when the working memory changes, in a forward chaining process.
+		To avoid iteration over the production rules, the RETE algorithm creates a dataflow tree-structured network to represent the rules.
+
+		However, the Rete algorithm is memory-intensive as it stores all partial matches, trading memory for speed.
+		In a streaming context, this is infeasible as streams can grow without bounds.
+		Therefore, we place time constraints on the memories using a window join operator instead of the traditional joins.
+
+		Talk about operators and link them to CoDeR.
+
+ 		Explain an example Rete network.
+
+ 		\begin{figure*}
+ 			\centering
+ 			\includegraphics[width=0.9\textwidth]{example-rete-network}
+ 			\caption{An example Rete network with corresponding (RDFS) rules.}
+ 			\labelfig{example-rete-network}
+ 		\end{figure*}
 	\end{nestedsection}
+\end{nestedsection}

### 2015-06-19 content/contribution/CoDeR.tex
moved the data model to a subsection of the description of CoDeR, and modified the expression of the example state of a CDP to first present the CDP, then the state of it whose transitive data validity is expressed in the figure (which needs editing)

--- 2015-06-18 content/contribution/CoDeR.tex
@@
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog.
+	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWl 2 RL.
 	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
@@
 	% Stream-based Data Model
+	\begin{nestedsection}{Data Model}{model: data}
+		The data model of CoDeR is composed of two classes of streams: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
+
+		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any ${CDP}$ expressed by a CoDeR-based system;
+		they are sequences of instances of RDF triples, where an instance of an RDF triple is an RDF triple annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle (s,p,o),i_{e},i_{n} \rangle}$.
+		This interval, for any entailed fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}, being the graphical representation of valid times for transitive data entailed by the state of the ${CDP = \{\{c\,\text{:-}\,a\,,\,b\,.\},S,3\}}$ at instant ${i = 2}$:
+		\begin{figure}[t]
+			\centering
+			\includegraphics[width=0.45\textwidth]{intersected-intervals}
+			\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
+			\labelfig{intersected-intervals}
+		\end{figure}
+		\begin{multline*}
+			CDP_{2} = \{c\,\text{:-}\,a\,,\,b\,.\} \cup \{ \langle a,1,4 \rangle . \langle b,2,5 \rangle \} \vDash \\
+			\{ c\,\text{:-}\,a\,,\,b\,.\,\langle a,1,4 \rangle . \langle b,2,5 \rangle . \langle c,2,4 \rangle \}
+		\end{multline*}
+		It is in the base case of this derivation that the semantic window range $R$ of a ${CDP}$ is expressed:
+		each fact instance arriving on an input stream has a valid interval from the time of its arrival at the system for a duration equal to the range $R$.
+
+		In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
+		These streams are sequences of the intermediate results of processing between one step in the production of entailments and the next.
+		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will take the form of instances of RDF \emph{graphs};
+		for a given inter-operator stream, each such RDF graph will match the same graph pattern, that being a sub-pattern of the body of one of the axiomatic rules $A$.
+		As with instances of RDF triples, instances of RDF graphs are RDF graphs that are annotated with the time interval for which they are ``valid'': ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
+		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same semantics of validity may be applied to these intermediate matches as to the true entailments of the model, and their valid-time intervals may derived from their justifying instances in the same manner.
+		It follows that inter-operator streams are, functionally, a general case of input/output streams, as the latter are simply sequences of instances of singleton RDF graphs matching the pattern ${\{(?s,?p,?o)\}}$.
+
+		Given that the ordering of instances in the output stream of a CoDeR-based system by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives the output streams of CoDeR a dual semantics with respect to Continuous Datalog.
+		Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by a specific pipeline of CoDeR operators applied to some set of continuous RDF data sources.
+	\end{nestedsection}
 

### 2015-06-19 content/contribution/R4.tex
corrected R4 section title

@@
-\begin{nestedsection}{R4: Rule-based Reasoning using Rete\dots Roar}{implementation}
+\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
 	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.

### 2015-06-22 content/contribution/CoDeR.tex
Updated the data model to accommodate streams of graphs as input/output streams

--- 2015-06-18 content/contribution/CoDeR.tex
@@
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog.
+	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
 	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
@@
 	% Stream-based Data Model
-	The data model of CoDeR is composed of two classes of streams.
-	The first are \emph{input}/\emph{output} streams, and provide the semantics of streams $S$, $P$ and $N$ for any ${CDP}$ expressed by a CoDeR-based system;
-	they are sequences of instances of RDF triples, where an instance of an RDF triple is an RDF triple annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle (s,p,o),i_{e},i_{n} \rangle}$.
-	This interval for any entailed fact instance is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}:
+	\begin{nestedsection}{Data Model}{model: data}
+		The data model of CoDeR is composed of two classes of streams: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
+
+		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any ${CDP}$ expressed by a CoDeR-based system.
+		These streams are sequences of instances of RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
+		an instance of an RDF graph is an RDF graph annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
+		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the streams $S$ must be translated into sequences of fact instances using the graph-to-triple operator described in \refsec{model: algebra}, where each fact instance inherits the valid time interval of the graph instance in which it originated.
+		This interval, for any entailed graph/fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}, being the graphical representation of valid times for transitive data continuously entailed by the ${CDP = \{\{c\,\text{:-}\,a,b\,.\},S,3\}}$ over instants ${i = 1 \dots 4}$:
 		\begin{figure}[t]
@@
 		\end{figure}
-	if ${A = \{c\,\text{:-}\,a\,,\,b\,.\}}$ and ${W^{S,3}_{2} = \{ \langle a,1,3 \rangle , \langle b,2,4 \rangle \}}$, then ${CDP_{2} = A \cup W^{S,3}_{2} \vDash \{ \langle a,1,3 \rangle , \langle b,2,4 \rangle , \langle c,2,3 \rangle \}}$, where $a$, $b$ and $c$ are RDF triples.
+		\begin{align*}
+			CDP_{1} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \} \\
+			CDP_{2} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \, b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \, b \, . \, c \, . \} \\
+			CDP_{3} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \, b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \, b \, . \, c \, . \} \\
+			CDP_{4} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, b \, . \}
+		\end{align*}
 		It is in the base case of this derivation that the semantic window range $R$ of a ${CDP}$ is expressed:
-	each fact instance arriving on an \emph{input} stream has a valid interval from the time of its arrival at the system for the duration of the semantic window expressed by that range $R$.
+		each graph instance arriving on an input stream has a valid interval from the time of its arrival at the system for a duration equal to the range $R$.
 
-	Given that the ordering of instances in the \emph{output} stream by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives all streams of CoDeR a dual semantics with respect to Continuous Datalog.
+		In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
+		These streams are sequences of the intermediate results of processing between steps in the production of entailments.
+		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
+		for a given inter-operator stream, each such RDF graph will match the same graph pattern, that being a sub-pattern of the body of one of the axiomatic rules $A$.
+		In addition, the continuous partial results for each specific graph pattern will occur in exactly one logical stream within a system, meaning that some logical streams may have multiple sources and/or multiple sinks within the system.
+		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same semantics of validity may be applied to these intermediate matches as to the true entailments of the model found on output streams, and their valid-time intervals may derived from their justifying instances in the same manner.
+		It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
+		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are the entailments of either the system or its predecessor in the stream pipeline.
+
+		Given that the ordering of instances in the output stream of a CoDeR-based system by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives the output streams of CoDeR a dual semantics with respect to Continuous Datalog.
+		Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by a specific pipeline of CoDeR operators applied to some set of continuous RDF data sources.
+	\end{nestedsection}
 

### 2015-06-22 content/contribution/CoDeR.tex
started work on the algebra, starting with a summary then moving to detailed descriptions of each operatior

@@
 	% Minimal Algebra
+	\begin{nestedsection}{Minimal Algebra}{model: algebra}
+		The minimal algebra of CoDeR is based on that of Positive Datalog, that being ${\{\bot, \text{:-}, \wedge\}}$.
+		Of these, $\text{:-}$ and $\wedge$ may be expressed as stream-to-stream operators, the former being trivially incremental and the latter being the window-join \citep{ReteDBMS}, in order to support forward-chaining processing of fixed rules;
+		$\bot$ is not an operator, but a special class of value represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}).
+		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
+		In addition to $\text{:-}$ and $\wedge$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and the datatype functions of RIF-Core \citep{w3crifcore}.
+		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community to the streams of triples on which $\sigma_{triple}$ operate.
+		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}, {}_b\Join_p, \sigma_{triple}, \sigma_{external}, {}_g\pi_t\}}$.
+
+		\begin{description}
+			\item[$\sigma_{triple}$] \hfill \\
+				The \emph{basic pattern match} of SPARQL is one-pass, and so is trivially interpreted as the stream-to-stream operator $\sigma_{(s,p,o)}$.
+				It takes a stream of triple instances as input and produces a stream of triple instances that contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the basic pattern ${(s,p,o)}$.
+			\item[$\sigma_{external}$] \hfill \\
+				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where each ${a_1,\dots,a_n}$ is either a variable or a constant value.
+				It filters instances from its single input stream to a stream of instances that contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				This filtering is based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern, thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern};
+				as such, the contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
+		\end{description}
+	\end{nestedsection}
 

### 2015-06-24 content/contribution/CoDeR.tex
further modification of data model and expansion of algebra

--- 2015-06-22 content/contribution/CoDeR.tex
@@
 		an instance of an RDF graph is an RDF graph annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
-		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the streams $S$ must be translated into sequences of fact instances using the graph-to-triple operator described in \refsec{model: algebra}, where each fact instance inherits the valid time interval of the graph instance in which it originated.
-		This interval, for any entailed graph/fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}, being the graphical representation of valid times for transitive data continuously entailed by the ${CDP = \{\{c\,\text{:-}\,a,b\,.\},S,3\}}$ over instants ${i = 1 \dots 4}$:
+		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the streams $S$ must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the valid time interval of the graph instance in which it originated.
+		This interval, for any entailed graph/fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}, being the graphical representation of valid times for transitive data continuously entailed by the ${CDP = \{\{c\,\text{:-}\,a,b\,.\},(\{a\}\text{:}1,\{b\}\text{:}2),3\}}$ over instants ${i = 1 \dots 4}$:
 		\begin{figure}[t]
@@
 		\begin{align*}
-			CDP_{1} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \} \\
-			CDP_{2} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \, b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \, b \, . \, c \, . \} \\
-			CDP_{3} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ a \, . \, b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, a \, . \, b \, . \, c \, . \} \\
-			CDP_{4} & = \{c\,\text{:-}\,a,b\,.\} \cup \{ b \, . \} & \vDash & \{ c\,\text{:-}\,a,b\,. \, b \, . \}
+			CDP_{1} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \} \\
+			CDP_{2} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \, b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \, b\text{:}2. \, c\text{:}2. \} \\
+			CDP_{3} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \, b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \, b\text{:}2. \, c\text{:}2. \} \\
+			CDP_{4} & = \{c\,\text{:-}\,a,b.\} \cup \{ b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, b\text{:}2. \}
 		\end{align*}
@@
 		for a given inter-operator stream, each such RDF graph will match the same graph pattern, that being a sub-pattern of the body of one of the axiomatic rules $A$.
-		In addition, the continuous partial results for each specific graph pattern will occur in exactly one logical stream within a system, meaning that some logical streams may have multiple sources and/or multiple sinks within the system.
 		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same semantics of validity may be applied to these intermediate matches as to the true entailments of the model found on output streams, and their valid-time intervals may derived from their justifying instances in the same manner.
@@
 		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are the entailments of either the system or its predecessor in the stream pipeline.
+
+		In addition, the continuous partial results for each specific graph pattern will occur in exactly one inter-operator stream within a system, meaning that some streams may have multiple sources and/or multiple sinks within the system.
+		The most obvious example of this is the ``generic triple stream'', whose graph pattern is ${(?s,?p,?o)}$, generated by the graph-to-triples operator ${{}_g\pi_t}$, which is also contributed to by all instance implication operators $\text{:-}_I$ and input to all basic pattern match operators $\sigma_{triple}$.
 
@@
 	% Minimal Algebra
+	\begin{nestedsection}{Minimal Algebra}{model: algebra}
+		The minimal algebra of CoDeR is based on that of Positive Datalog, that being ${\{\bot, \text{:-}, \wedge\}}$.
+		Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_I$ and the latter being the window-join \citep{ReteDBMS}, in order to support forward-chaining processing of fixed rules;
+		$\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}).
+		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
+		In addition to $\text{:-}_I$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and the datatype functions of RIF-Core \citep{w3crifcore}.
+		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community to the streams of triples on which $\sigma_{triple}$ operate.
+		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_I, {}_b\Join_p, \sigma_{triple}, \sigma_{external}, {}_g\pi_t\}}$.
+
+		\begin{description}
+			\item[$\sigma_{triple}$] \hfill \\
+				The \emph{basic pattern match} of SPARQL is one-pass, and so is trivially interpreted as the stream-to-stream operator $\sigma_{(s,p,o)}$.
+				It takes an inter-operator stream of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the basic pattern ${(s,p,o)}$.
+			\item[$\sigma_{external}$] \hfill \\
+				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value.
+				It filters elements of its input stream based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
+				thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern}, as specified in SPARQL.
+				As such, it takes an inter-operator stream of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				The contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
+			\item[${{}_g\pi_t}$] \hfill \\
+				The graph-to-triples operator ${{}_g\pi_t}$ is a one-pass operator that takes a heterogeneous (\emph{input/output}) stream of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of instances of the RDF triples contained in the graphs instantiated in the input stream.
+				Each triple instance produced is assigned the same valid-time interval as the graph instance in whose graph the instantiated triple was contained.
+				Consider, for example, the input/output stream containing an instance of a graph containing the RDF triples $a$ and $b$ and valid from instant $t_e$ to instant $t_n$:
+				\begin{multline*}
+					{}_g\pi_t (\dots,\langle \{a,b\},t_e,t_n \rangle,\dots) = \\
+						(\dots,\langle a,t_e,t_n \rangle,\langle b,t_e,t_n \rangle,\dots)
+				\end{multline*}
+			\item[$\text{:-}_I$] \hfill \\
+				Instance implication is a one-pass operator that represents 
+		\end{description}
+	\end{nestedsection}
 

### 2015-06-25 content/contribution/continuous-datalog.tex
added a few paragraphs and an axiom regarding instance entailment, allowing the proof to work without proving from the previous axioms that E_i \cap E_{i-R} |= {}

@@
 	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
-	In addition, these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community, as proven in \refsec{semantics: proof for entailments}.
+	In addition, in can be shown that these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community.
 
@@
 
+	Traditionally, a set of axioms $A$ entails another set of axioms $E$ in which the existence of one fact in $E$ may be justified by many distinct subsets of $A$.
+	When considering streams of facts, in which one fact may occur multiple times and be considered distinct based on its entailment time, temporal information would be lost if sets of axioms including fact instances (denoted $A_{inst}$) were to entail sets of axioms without.
+	As such, where a fact would be entailed by some set of facts $A$ that are instantiated on some stream, an instance of that fact should be entailed by the set of instances of those facts present on that stream.
+	Furthermore, each such justification in $A_{inst}$ for a fact entails a unique instance of that fact, such that each entailed instance has a single justification within the original set of axioms, and many instances of the entailed fact may occur in the entailed set of axioms $E_{inst}$.
+	This form of entailment shall be henceforth referred to as ``instance entailment'', and shall be the norm when discussing the entailments of sets of axioms that include instances of facts.
+
+	It should be noted that removing all temporal information from a set of axioms including fact instances $A_{inst}$ produces the set of traditional Datalog axioms $A$ instantiated in $A_{inst}$.
+	This is be represented as the function ${trad\left( A_{inst} \right) = A}$.
+	In addition, applying the ${trad()}$ function to both the original axiom instances $A_{inst}$ and the entailed axiom instances $E_{inst}$ maintains the relationship of entailment between the two sets:
+	\[ \left( A_{inst} \vDash E_{inst} \right) \rightarrow \left( A \vDash E \right) \]
+
+	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.
+	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
+		\item\label{axiom:continuous datalog: entailment disjointness}
+			If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
+			\begin{multline*}
+				\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
+					\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
+			\end{multline*}
+		\setcounter{continuousDatalogAxioms}{\theenumi}
+	\end{enumerate}
+
 	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
@@
 			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
-			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
+			Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
 			Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
-			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.
+			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.
 			\begin{equation*}
@@
 		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
-			The state of a ${CDP}$ at a given instant ${t \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
-			As both $A$ and $W^{S,R}_{i}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $t$ by $T^{CD}$.
+			The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
+			As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.
 			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$.
 			\begin{equation*}
-				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_{i} \right)
+				T^{CD} \left( CDP, i \right) = T^{D} \left( CDP_{i} \right)
 			\end{equation*}
@@
 		\item\label{axiom:continuous datalog: positive window increment}
-			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $t$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${t-1}$.
+			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
 			\begin{equation*}
@@
 		\item\label{axiom:continuous datalog: negative window increment}
-			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at instant $t$.
+			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
 			\begin{equation*}
@@
 			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
-			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 
+			As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
 			\begin{equation*}
-				\forall Y \geq 1 \left( W^{P,1}_{i} \cap W^{N,1}_{i-Y} = \varnothing \right)
+				\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
 			\end{equation*}
@@
 
-	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes entailed by the ${CDP}$
-	(i.e. when it becomes newly justified by some subset of instances in $W^{S,R}_{i_{e}}$).
-	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
+	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes newly justified by the ${CDP}$
+	(more specifically, by some subset of the instances in $W^{S,R}_{i_{e}}$).
+	As such, the stream of entailment instances $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
 	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
@@
 	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
-	However, it is proven in \ref{sec:semantics: proof for entailments}, from these axioms, that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
+	However, it can be shown from these axioms,
+	as in \ref{sec:semantics: proof for entailments},
+	that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
 	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]

### 2015-06-25 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
added a few paragraphs and an axiom regarding instance entailment, allowing the proof to work without proving from the previous axioms that E_i \cap E_{i-R} |= {}

@@
 \begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
-
-	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
-		\item\label{axiom:continuous datalog: disjoint entailments}
-			The instantaneous set of stream-justified entailments $E_{i}$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{i-X}$ of the same ${CDP}$ when ${X \geq R}$.
-			\begin{equation*}
-				E_{i} \cap E_{i-X} = \varnothing
-			\end{equation*}
-		\setcounter{continuousDatalogAxioms}{\theenumi}
-	\end{enumerate}
-
 	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
@@
 	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
-		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$:
+		Assume that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$.
+		Remember that ${W^{S,R}_{i} \cap W^{S,R}_{i-R} = \varnothing}$, ${W^{S,R}_{i} \vDash E_{i}}$ and ${W^{S,R}_{i-R} \vDash E_{i-R}}$.
+		Therefore:
 		\begin{align*}
@@
 			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			\text{Axiom~\ref{axiom:continuous datalog: disjoint entailments}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
+			\text{Axiom~\ref{axiom:continuous datalog: entailment disjointness}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
 			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\

### 2015-06-25 content/contribution/CoDeR.tex
added description of stream union through multiple sources, a description of the fact that operators applied to streams do themselves describe streams, the definition of the SteM operator, and the summisation that CoDeR supports correct and complete processing with respect to the semantics of continuous datalog.
Also reworded a lot of the existing operators to read better, and more directly relate basic pattern matches and SteMs to Rete.

@@
 
-		In addition, the continuous partial results for each specific graph pattern will occur in exactly one inter-operator stream within a system, meaning that some streams may have multiple sources and/or multiple sinks within the system.
-		The most obvious example of this is the ``generic triple stream'', whose graph pattern is ${(?s,?p,?o)}$, generated by the graph-to-triples operator ${{}_g\pi_t}$, which is also contributed to by all instance implication operators $\text{:-}_I$ and input to all basic pattern match operators $\sigma_{triple}$.
+		In addition, each stream in the model is semantically distinct, with each input/output stream being from a distinct source and uniquely identifiable by the source's IRI, and each inter-operator stream containing graphs matching a specific graph pattern and uniquely identifiable by that pattern.
+		As the same graph pattern may be a sub-pattern of many rule bodies, and so appear in multiple places in a processing plan, one inter-operator stream may be the input to multiple operators.
+		Furthermore, two or more operators may contribute to the same stream within the model, where the pattern of the stream in question is the union of the patterns of the streams produced by each of the operators in isolation.
+		In this case, each operator contributes results to the common stream as they are produced, thereby interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment time $i_e$ of the constituent instances.
+		As this is the intended behaviour of a stream union operator, such an operator is not needed in CoDeR-based systems.
+		However, within the algebra of CoDeR, such a stream is still denoted ${op_{application} \cup op_{application}}$.
 
 		Given that the ordering of instances in the output stream of a CoDeR-based system by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives the output streams of CoDeR a dual semantics with respect to Continuous Datalog.
-		Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by a specific pipeline of CoDeR operators applied to some set of continuous RDF data sources.
+		Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by specific pipelines of CoDeR operators applied to some set of continuous RDF data sources.
 	\end{nestedsection}
@@
 		The minimal algebra of CoDeR is based on that of Positive Datalog, that being ${\{\bot, \text{:-}, \wedge\}}$.
-		Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_I$ and the latter being the window-join \citep{ReteDBMS}, in order to support forward-chaining processing of fixed rules;
+		Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_{head}$ and the latter being the window-join $\Join$ \citep{niagaraCQ}, in order to support forward-chaining processing of fixed rules;
 		$\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}).
 		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
-		In addition to $\text{:-}_I$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and the datatype functions of RIF-Core \citep{w3crifcore}.
+		In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
 		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community to the streams of triples on which $\sigma_{triple}$ operate.
-		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_I, {}_b\Join_p, \sigma_{triple}, \sigma_{external}, {}_g\pi_t\}}$.
+		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
+
+		As all operators are stream-to-stream in nature, they each take one or two streams as input and produce a single stream (though they may contribute to any number of union streams).
+		As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
+		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${CDP = \{ \{b\,\text{:-}\,a.\}, S, R \}}$, where $a$ and $b$ are RDF triples.
+		\begin{equation}\labeleqn{basic-example-plan}
+			\text{:-}_b \left( \sigma_a \left( {}_g{\pi_t}\,S \right) \right)
+		\end{equation}
 
 		\begin{description}
-			\item[$\sigma_{triple}$] \hfill \\
-				The \emph{basic pattern match} of SPARQL is one-pass, and so is trivially interpreted as the stream-to-stream operator $\sigma_{(s,p,o)}$.
-				It takes an inter-operator stream of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+			\item[$\sigma_{triple}\,S$] \hfill \\
+				The triple-select operator identifies instances of triples in a stream $S$ that match some \emph{basic pattern} \citep{w3csparql}.
+				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm applied to semantic data.
+				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{(s,p,o)}$, where ${(s,p,o)}$ is the basic pattern to be matched.
+				It takes an inter-operator stream $S$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
 				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the basic pattern ${(s,p,o)}$.
-			\item[$\sigma_{external}$] \hfill \\
-				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value.
-				It filters elements of its input stream based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
+			\item[$\sigma_{external}\,S$] \hfill \\
+				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
+				It filters elements of its input stream $S$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
 				thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern}, as specified in SPARQL.
-				As such, it takes an inter-operator stream of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				As such, it takes an inter-operator stream $S$ of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
 				The contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
-			\item[${{}_g\pi_t}$] \hfill \\
-				The graph-to-triples operator ${{}_g\pi_t}$ is a one-pass operator that takes a heterogeneous (\emph{input/output}) stream of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of instances of the RDF triples contained in the graphs instantiated in the input stream.
+			\item[${{}_g{\pi_t}\,S}$] \hfill \\
+				The graph-to-triples operator ${{}_g{\pi_t}}$ is a unary operator that deconstructs graph instances into sequences of triple instances, the union of the triples instantiated therein being equal to the contents of the original graph.
+				It takes a heterogeneous (\emph{input/output}) stream $S$ of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of RDF triple instances.
 				Each triple instance produced is assigned the same valid-time interval as the graph instance in whose graph the instantiated triple was contained.
@@
 				\begin{multline*}
-					{}_g\pi_t (\dots,\langle \{a,b\},t_e,t_n \rangle,\dots) = \\
+					{}_g{\pi_t} (\dots,\langle \{a,b\},t_e,t_n \rangle,\dots) = \\
 						(\dots,\langle a,t_e,t_n \rangle,\langle b,t_e,t_n \rangle,\dots)
 				\end{multline*}
-			\item[$\text{:-}_I$] \hfill \\
-				Instance implication is a one-pass operator that represents 
+			\item[$\text{:-}_{head}\,S$] \hfill \\
+				Instance implication is a one-pass operator that takes as input an inter-operator stream $S$ whose graph pattern is the body of one of the axiomatic rules $A$ of the expressed ${CDP}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
+				Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
+				In addition, each produced instance inherits its valid-time interval from the graph instance that implied it.
+			\item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
+				State Modules (\emph{SteM}s) are asymmetrical binary operators that each perform the dual function of maintaining a window of graph instances and ``joining'' each graph instance on a stream against those within the maintained window (known as ``probing'' the window).
+				These are the same two functions performed for each window in a symmetric window-join, extended from Rete \emph{beta nodes} in \citep{ReteDBMS}, with results of each probe being produced to the same result stream as they occur.
+				Window-joins are, therefore, functionally identical to the union of two complementary SteMs:
+				${A \Join B = \left( A\,{}_b{\Join_p}\,B \right) \cup \left( B\,{}_b{\Join_p}\,A \right)}$, where $A$ and $B$ are inter-operator streams.
+				In addition, it should be noted that SteMs express the semantics of the sequential intersection operator ${seq}$ proposed in EP-SPARQL \citep{EP-SPARQL}, in that they join new data from one stream with the prior (valid) contents of another, thereby supporting rudimentary event processing.
+
+				The implementations of SteMs in the literature show a great variety of behaviour regarding window maintenance \citep{SteMs}, in the most extreme case ceasing to be a wholly stream-to-stream operator and supporting the incorporation of ``static'' data into stream processing systems.
+				In the minimal algebra of CoDeR, however, SteMs simply support the building of a \emph{valid window} (as defined in \refsec{semantics}) from its first inter-operator stream $S_1$, and the probing of that window with data from its second inter-operator stream $S_2$ to produce an inter-operator stream of the homogeneously joined graph instances.
+				Furthermore, the nature of probing in CoDeR is strictly limited to that of the natural join between the variables in the patterns for streams $S_1$ and $S_2$.
+				Finally, in contrast to the direct inheritance of valid-times in unary operators of CoDeR, the valid-time interval assigned to each graph produced by a SteM is the intersection of the valid-times of the two joined graphs.
 		\end{description}
+
+		As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
+		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $R$ over the system input streams $S$ defined in a given ${CDP = \{A,S,R\}}$.
+		It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
+		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $A$ of a ${CDP}$, all facts that are logically entailed by that ${CDP}$ at a given instant will be persisted within that system at that instant,
+		i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
 	\end{nestedsection}

### 2015-06-26 content/contribution/R4.tex

@@
+\include{ebnf}
 \begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
@@
 	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in Listing~\ref{lst-S-RIF-Core}.
+
 \begin{figure*}
-		\centering
-		\begin{lstlisting}[caption=Streaming RIF-Core Grammar,label=lst-S-RIF-Core]
-Document       ::= IRIMETA? 'Document' '(' Base? Prefix* Import* Group?')'
-Base           ::= 'Base' '(' ANGLEBRACKIRI ')'
-Prefix         ::= 'Prefix' '(' Name ANGLEBRACKIRI ')'
-Import         ::= IRIMETA? 'Import' '(' LOCATOR PROFILE? ')'
-Group          ::= IRIMETA? 'Group' '(' (RULE | Group)*')' 
-RULE           ::= ((IRIMETA? 'Forall' Var+ '(' CLAUSE')') | CLAUSE)
-CLAUSE         ::= Implies | ATOMIC
-Implies        ::= IRIMETA? (ATOMIC | 'And' '(' ATOMIC* ')') ':-' FORMULA
-LOCATOR        ::= ANGLEBRACKIRI
-PROFILE        ::= ANGLEBRACKIRI Window?
-Window         ::= Number TimeUnit
-TimeUnit       ::= 'ms' | 's' | 'm' | 'h'
-		\end{lstlisting}
+\begin{ebnf}
+\Rule{Document}{
+        \Optional{\NTerm{IRIMETA}} 
+        \Token{Document} 
+        \Token{(} 
+        \Optional{\NTerm{Base}} 
+        \ZeroOrMore{\NTerm{Prefix}} 
+        \ZeroOrMore{\NTerm{Import}} 
+        \Optional{\NTerm{Group}} 
+        \Token{)}
+}
+\Rule{Base}{
+        \Token{Base} 
+        \Token{(} 
+        \NTerm{ANGLEBRACKIRI} 
+        \Token{)}
+}
+\Rule{Prefix}{
+        \Token{Prefix} 
+        \Token{(} 
+        \NTerm{Name} 
+        \NTerm{ANGLEBRACKIRI} \Token{)}
+}
+\Rule{Import}{
+        \Optional{\NTerm{IRIMETA}} 
+        \Token{Import} 
+        \Token{(} 
+        \NTerm{LOCATOR} 
+        \Optional{\NTerm{PROFILE}} 
+        \Token{)}
+}
+\Rule{Group}{
+        \Optional{\NTerm{IRIMETA}} 
+        \Token{Group} 
+        \Token{(} 
+        \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
+        \Token{)}
+}
+\Rule{RULE}{
+        \Optional{\NTerm{IRIMETA}}
+        \Token{Forall}
+        \OneOrMore{\NTerm{Variable}}
+        \Token{(}
+        \NTerm{CLAUSE}
+        \Token{)}
+}
+\OrRule{
+        \NTerm{CLAUSE}
+}
+\Rule{CLAUSE}{
+        \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
+}
+\Rule{Implies}{
+        \Optional{\NTerm{IRIMETA}}
+        \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
+        \Token{:-}
+        \NTerm{FORMULA}
+}
+\Rule{LOCATOR}{
+        \NTerm{ANGLEBRACKURI}
+}
+\Rule{PROFILE}{
+        \NTerm{ANGLEBRACKURI} 
+        \Optional{\NTerm{Window}}
+}
+\Rule{Window}{
+        \NTerm{Number}
+        \NTerm{TimeUnit}
+}
+\Rule{TimeUnit}{
+        \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
+}
+\end{ebnf}
+\caption{Streaming RIF-Core Grammar}
+\label{lst-S-RIF-Core}
 \end{figure*}

### 2015-06-26 content/contribution/continuous-datalog.tex
Converted to axiom environment

@@
 	\end{description}
-	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog.
-	\begin{enumerate}\newcounter{continuousDatalogAxioms}
-		\item\label{axiom:continuous datalog: window range leq 0}
-			A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
+
+The following axioms define the semantics of streams and windows over
+those streams in Continuous Datalog:
+
+\begin{axiom}
+\label{axiom:continuous datalog: window range leq 0}
+A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i
+  \in \mathbb{N}}$ is always empty.
 \begin{equation*}
@@
 \end{equation*}
-		\item\label{axiom:continuous datalog: window composition}
-			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
+\end{axiom}
+
+\begin{axiom}
+\label{axiom:continuous datalog: window composition}
+A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
+given instant ${i \in \mathbb{N}}$ is equivalent to the union of the
+$X$ consecutive windows of range 1 over the stream $S$ preceding and
+including that at instant $i$.
 \begin{equation*}
@@
 \end{equation*}
+\end{axiom}
+
+\begin{axiom}
 \item\label{axiom:continuous datalog: window disjointness}
-			A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
+A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
+instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
+range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
 \begin{equation*}
@@
 \end{equation*}
-		% \item\label{axiom:continuous datalog: stream disjointness}
-		% 	The contents of any two streams are disjoint, thus a window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over any other stream ${Z \neq S}$ at any instant ${i \in \mathbb{N}}$.
-		% 	\begin{equation*}
-		% 		W^{S,X}_{i} \cap W^{Z,Y}_{i} = \varnothing
-		% 	\end{equation*}
-		\setcounter{continuousDatalogAxioms}{\theenumi}
-	\end{enumerate}
+\end{axiom}
 
@@
 	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.
-	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
-		\item\label{axiom:continuous datalog: entailment disjointness}
-			If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
+
+\begin{axiom}
+\label{axiom:continuous datalog: entailment disjointness}
+If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
+disjoint, then the any set $E_{inst}$ of axiom instances each
+justified at least in part by $A_{inst}$ is necessarily disjoint from
+any set $F_{inst}$ of axiom instances each justified at least in part
+by $B_{inst}$, because entailed fact instances are distinct with
+regards to the fact instances that justify them.
 \begin{multline*}
@@
 \end{multline*}
-		\setcounter{continuousDatalogAxioms}{\theenumi}
-	\end{enumerate}
+\end{axiom}
 
@@
 	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
-	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
-		\item\label{axiom:continuous datalog: positive window increment}
-			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
+
+\begin{axiom}
+\label{axiom:continuous datalog: positive window increment}
+The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
+the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
+that are entailed by the state of the ${CDP}$ at instant $i$ that were
+not entailed by the state of ${CDP}$ at the previous instant instance
+${i-1}$.
 			\begin{equation*}
@@
 			\end{equation*}
-		\item\label{axiom:continuous datalog: negative window increment}
+\end{axiom}
+
+\begin{axiom}
+\label{axiom:continuous datalog: negative window increment}
 			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
@@
 			\end{equation*}
-		\item\label{axiom:continuous datalog: entailment precedes negation}
+\end{axiom}
+
+\begin{axiom}
+\label{axiom:continuous datalog: entailment precedes negation}
 			Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
@@
 			\end{equation*}
-		\setcounter{continuousDatalogAxioms}{\theenumi}
-	\end{enumerate}
+\end{axiom}
 

### 2015-06-26 content/contribution/continuous-datalog.tex
Fixed axioms and definitions

@@
 	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
-	\begin{description}
-		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
-			An instant is a discrete point in time.
-			The continuous distance in time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
-			In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
-		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
-			A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g.
-the time at which each was entailment by some process at the source of the stream.
-			Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
-			This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
-		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
-			A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
-			The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
-	\end{description}
+
+\begin{definition}[Instant]
+\label{def:continuous datalog: instant}
+An instant is a discrete point in time.  The continuous distance in
+time between these discrete points may be theoretically infinitesimal,
+but will in practise have a lower bound set by hardware limitations,
+or will be defined as part of some specific implementation.  In these
+cases, anything that occurs between discrete instants is considered to
+have occurred at the next instant.
+\end{definition}
+
+\begin{definition}[Stream of Facts]
+\label{def:continuous datalog: stream}
+A stream of facts is a \emph{sequence} of Datalog facts ordered by
+some temporal characteristic of each instance of each fact, e.g. the
+time at which each was entailment by some process at the source of the
+stream.  Each instance of a fact is uniquely identifiable by the
+combination of its signature (its predicate $p$, subject $s$ and
+object $o$) and its entailment time $t$, so an instance is denoted
+${p(s,o):t}$.  This distinction between instances of facts within a
+stream is represented in Axiom~\ref{axiom:continuous datalog: window
+  disjointness}.
+\end{definition}
+
+\begin{definition}[Sliding Window]
+\label{def:continuous datalog: window}
+A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at
+any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances
+of facts that occur in $S$ during the interval ${(i-X,i]}$, and is
+denoted ${W^{S,X}_{i}}$.  The concept of a sliding window within the
+semantics of Continuous Datalog are further defined by
+Axioms~\ref{axiom:continuous datalog: window range leq 0}
+and~\ref{axiom:continuous datalog: window composition}.
+\end{definition}
 
@@
 \begin{axiom}
-\item\label{axiom:continuous datalog: window disjointness}
+\label{axiom:continuous datalog: window disjointness}
 A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
@@
 	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
-	\begin{description}
-		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
-			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
-			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
-			Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
-			Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
-			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.
+
+\begin{definition}[Continuous Datalog Program]
+\label{def:continuous datalog: CDP}
+
+A Continuous Datalog Program (${CDP}$) is composed of a set of
+Positive Datalog axioms $A$, a set of streams of facts $S$ and the
+range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
+This specifies that the streamed data from $S$ is to be considered
+valid from its arrival for the duration represented by $R$.  Such a
+continuous program entails a triple composed of the traditional
+entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and
+a stream of instance negations $N$, whose contents are defined in
+Axioms~\ref{axiom:continuous datalog: positive window
+  increment},~\ref{axiom:continuous datalog: negative window
+  increment} and~\ref{axiom:continuous datalog: entailment precedes
+  negation}.  Streams $P$ and $N$ are therefore the continuous
+reasoning equivalents of the streams produced by the \emph{IStream}
+and \emph{DStream} operators of the \emph{CQL} continuous query
+language \citep{CQL} respectively.  The truth value of a ${CDP}$ is
+instantaneous, with the the truth function of Continuous Datalog
+$T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in
+  \mathbb{N}}$ as arguments.
 \begin{equation*}
@@
 \end{equation*}
-		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
-			The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
-			As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.
-			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$.
+\end{definition}
+
+\begin{definition}[State of a ${CDP}$]
+\label{def:continuous datalog: CDPt}
+
+The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$
+(${CDP_{i}}$) is the union of the axioms $A$ and the window
+$W^{S,R}_{i}$.  As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are
+sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog
+program that may be evaluated by the truth function of Datalog
+$T^{D}$, the result of which is the definition for the instantaneous
+truth of the ${CDP}$ at instant $i$ by $T^{CD}$.  The instantaneous
+entailments of a $CDP_{i}$ are the union of the persistent entailments
+of ${A \vDash A'}$ (which are guaranteed to be entailed by every
+state, as Positive Datalog is monotonic) and the set $E_{i}$ of
+instances of the transient entailments that are justified at least in
+part by some subset of $W^{S,R}_{i}$.
+
 \begin{equation*}
@@
 \end{equation*}
-	\end{description}
+\end{definition}
+
 	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.

### 2015-06-26 content/contribution/CoDeR.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

@@
 				The triple-select operator identifies instances of triples in a stream $S$ that match some \emph{basic pattern} \citep{w3csparql}.
-				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm applied to semantic data.
-				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{(s,p,o)}$, where ${(s,p,o)}$ is the basic pattern to be matched.
+				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to semantic data.
+				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the basic pattern to be matched.
 				It takes an inter-operator stream $S$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
-				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the basic pattern ${(s,p,o)}$.
+				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
+				\begin{figure}[t]
+					\centering
+					\includegraphics[width=0.45\textwidth]{basic-pattern-match}
+					\caption{The triple-select operator.}
+					\labelfig{continuous datalog: basic pattern match}
+				\end{figure}
 			\item[$\sigma_{external}\,S$] \hfill \\
-				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{f(a_1,\dots,a_n)}$, where $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
+				\begin{figure}[b]
+					\centering
+					\includegraphics[width=0.45\textwidth]{datatype-function}
+					\caption{The operator encompassing all datatype functions of RIF-Core.}
+					\labelfig{continuous datalog: datatype function}
+				\end{figure}
+				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
 				It filters elements of its input stream $S$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
@@
 			\item[${{}_g{\pi_t}\,S}$] \hfill \\
+				\begin{figure}[t]
+					\centering
+					\includegraphics[width=0.45\textwidth]{graph-to-triple}
+					\caption{The graph-to-triple operator.}
+					\labelfig{continuous datalog: graph-to-triple}
+				\end{figure}
 				The graph-to-triples operator ${{}_g{\pi_t}}$ is a unary operator that deconstructs graph instances into sequences of triple instances, the union of the triples instantiated therein being equal to the contents of the original graph.
@@
 			\item[$\text{:-}_{head}\,S$] \hfill \\
+				\begin{figure}[b]
+					\centering
+					\includegraphics[width=0.45\textwidth]{instance-implication}
+					\caption{The instance implication operator.}
+					\labelfig{continuous datalog: instance implication}
+				\end{figure}
 				Instance implication is a one-pass operator that takes as input an inter-operator stream $S$ whose graph pattern is the body of one of the axiomatic rules $A$ of the expressed ${CDP}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
 				Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
+				This is similar to the CONSTRUCT clause of SPARQL, or the final step of \emph{conflict set resolution} within the Rete algorithm \citep{forgy79}.
 				In addition, each produced instance inherits its valid-time interval from the graph instance that implied it.
 			\item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
+				\begin{figure}[t]
+					\centering
+					\subfloat[The SteM operator.]{\includegraphics[width=0.45\textwidth]{SteM}}\\
+					\subfloat[A typical window-join.]{\includegraphics[width=0.45\textwidth]{window-join}}
+					\caption{Exploded views of two-pass joining operators.}
+					\labelfig{continuous datalog: SteM}
+				\end{figure}
 				State Modules (\emph{SteM}s) are asymmetrical binary operators that each perform the dual function of maintaining a window of graph instances and ``joining'' each graph instance on a stream against those within the maintained window (known as ``probing'' the window).

### 2015-06-26 content/contribution/R4.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

@@
-\include{ebnf}
 \begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
@@
 	The system architecture is shown in \reffig{R4-architecture}.
-	\begin{figure}[t]
+	\begin{figure}[b]
 		\centering
@@
 	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
-	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in Listing~\ref{lst-S-RIF-Core}.
+	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.
 
-\begin{figure*}
+	\begin{figure*}[t]
+		\centering
 		\begin{ebnf}
@@
 		\caption{Streaming RIF-Core Grammar}
-\label{lst-S-RIF-Core}
+		\labelfig{lst-S-RIF-Core}
 	\end{figure*}
@@
 
- 		\begin{figure*}
+ 		\begin{figure*}[t]
  			\centering

### 2015-06-26 content/contribution/continuous-datalog.tex
added figures for operators, fixed EBNF templating (moved EBNF environment definition to the template folder, moved include to the article template to avoid the pagebreak), added the Axiom environment missing from Nick's last commit

--- 2015-06-26 content/contribution/continuous-datalog.tex
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: window range leq 0}
-A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i
-  \in \mathbb{N}}$ is always empty.
+	\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
+		A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
 		\begin{equation*}
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: window composition}
-A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
-given instant ${i \in \mathbb{N}}$ is equivalent to the union of the
-$X$ consecutive windows of range 1 over the stream $S$ preceding and
-including that at instant $i$.
+	\begin{axiom}\label{axiom:continuous datalog: window composition}
+		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
 		\begin{equation*}
@@
 
-\begin{axiom}
-\item\label{axiom:continuous datalog: window disjointness}
-A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
-instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
-range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
+	\begin{axiom}\label{axiom:continuous datalog: window disjointness}
+		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
 		\begin{equation*}
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: entailment disjointness}
+	\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
 		If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: positive window increment}
-The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
-the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
-that are entailed by the state of the ${CDP}$ at instant $i$ that were
-not entailed by the state of ${CDP}$ at the previous instant instance
-${i-1}$.
+	\begin{axiom}\label{axiom:continuous datalog: positive window increment}
+		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
 		\begin{equation*}
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: negative window increment}
+	\begin{axiom}\label{axiom:continuous datalog: negative window increment}
 		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: entailment precedes negation}
+	\begin{axiom}\label{axiom:continuous datalog: entailment precedes negation}
 		Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.

### 2015-06-26 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
changed the Hypothesis environment to a theorem environment

@@
 \begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
-	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
+	\begin{thrm}\labelhyp{continuous datalog: entailment} \hfill \\
 		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
@@
 		\end{equation*}
-	\end{hyp}
+	\end{thrm}
 	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.

### 2015-06-26 content/contribution/continuous-datalog.tex

@@
 	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
-	\begin{description}
-		\item[Instant\label{def:continuous datalog: instant}]\hfill\\
+
+	\begin{definition}[Instant]\label{def:continuous datalog: instant}
 		An instant is a discrete point in time.
@@
 		In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
-		\item[Stream of Facts\label{def:continuous datalog: stream}]\hfill\\
+	\end{definition}
+
+	\begin{definition}[Stream of Facts]\label{def:continuous datalog: stream}
 		A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g.
@@
 This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
-		\item[Sliding Window\label{def:continuous datalog: window}]\hfill\\
+	\end{definition}
+
+	\begin{definition}[Sliding Window]\label{def:continuous datalog: window}
 		A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
 		The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
-	\end{description}
+	\end{definition}
 
-	The following axioms define the semantics of streams and windows over
-	those streams in Continuous Datalog:
+	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog:
 
@@
 	\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-		If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are
-		disjoint, then the any set $E_{inst}$ of axiom instances each
-		justified at least in part by $A_{inst}$ is necessarily disjoint from
-		any set $F_{inst}$ of axiom instances each justified at least in part
-		by $B_{inst}$, because entailed fact instances are distinct with
-		regards to the fact instances that justify them.
+		If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
 		\begin{multline*}
@@
 	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
-	\begin{description}
-		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
+
+	\begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
 		A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
@@
 		\end{equation*}
-		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
+	\end{definition}
+
+	\begin{definition}[State of a ${CDP}$]\label{def:continuous datalog: CDPt}
 		The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
@@
 		\end{equation*}
-	\end{description}
+	\end{definition}
+
 	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.

### 2015-06-26 content/contribution/R4.tex
fixed the graphics inclusion in R4 to explicitly include 'png's
added the new text from Rehab, and added references to the operators of CoDeR and semantics of Continuous Datalog, where appropriate.  Also modified the EBNF to reflect the windowing as actually implemented, based on the description of the implementation.

@@
 		\centering
-		\includegraphics[width=0.45\textwidth]{R4-architecture}
+		\includegraphics[width=0.45\textwidth]{R4-architecture.png}
 		\caption{The system architecture of R4.}
@@
 	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.
+	Further, we propose the extension of the set of PROFILEs that describe the IMPORTs of a rule set to include those definitions of RDF streams recognised by the W3C.
 
@@
 			        \ZeroOrMore{\NTerm{Import}}
+			        \Optional{\NTerm{Window}}
 			        \Optional{\NTerm{Group}} 
@@
 			        \NTerm{ANGLEBRACKURI} 
-			        \Optional{\NTerm{Window}}
 			}
@@
 		In a streaming context, this is infeasible as streams can grow without bounds.
-		Therefore, we place time constraints on the memories using a window join operator instead of the traditional joins.
+		Therefore, we place time constraints on the memories using a window-join operator instead of the traditional joins.
 
-		Talk about operators and link them to CoDeR.
+		Data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
+		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
+		Source nodes also time stamps the incoming tuples with system timestamp if they are not already stamped from their origins.
+		They propagate the data to the successor nodes, which are the alpha nodes.
+
+		Alpha nodes are single-entry nodes that form a discrimination network.
+		In R4, alpha nodes can receive input from any number of nodes through its single entry point.
+		Each input triple is matched against some conditions, and is either dropped if there is no match, or propagated downstream to the successor nodes in the case of a match.
+		As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
+		The first alpha node in the network propagates its output to a special node called the left input adaptor, while the other alpha nodes propagate their output to join nodes.
+		The left input adaptor node is responsible of preparing the incoming triples to enter the beta network from the left input of the first join node in a given join-tree.
+		It creates a new token i.e. a partial match, defined as a list of triples, for each incoming triple and adds this triple as the first item in the list.
+		Tokens are then sent to the first join node in the beta network.
+
+		Beta nodes are two-entry nodes that are responsible for joining the branches of the alpha network.
+		In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
+		Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing triples received from alpha nodes.
+		Join nodes are beta nodes that match tuples from both sides according to some conditions, e.g. a shared variable binding.
+		As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
+		In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
+		These windows are implemented as priorities queues and their items are ordered according to their timestamps.
+
+		When a join node is left activated, i.e.
+it receives a token through its left input, it first adds the new token to its left window, then it iterates over its right window elements trying to find a match.
+		When a match is found, a new token is created by adding the right triple to the triples list of the left token.
+		The new token is stamped with the earliest timestamp of its triples, and then propagated to the next join node or to the terminal node if it is the last join.
+		At this point, the join node performs a garbage collection over its left window, ensuring that items with timestamps that are smaller that current time minus the semantic window size are deleted.
+		Conversely, when the join node is right activated by receiving a triple from an alpha node, it adds the new triple to the right window, matches it against the elements of the left window, and performs a garbage collection over the right window.
+
+		This method of annotating tokens with the earliest timestamp of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
+		However, the ordering of tokens in their streams and their propagation through the system expresses those annotation semantics implicitly.
+		This leaves the timestamp of a token, in combination with the size of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances in CoDeR.
+		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
+
+ 		Finally, terminal nodes receive tokens from the last node in the network and are responsible of producing the new entailed triples.
 
@@
  			\centering
- 			\includegraphics[width=0.9\textwidth]{example-rete-network}
+ 			\includegraphics[width=0.9\textwidth]{example-rete-network.png}
  			\caption{An example Rete network with corresponding (RDFS) rules.}

### 2015-06-29 content/introduction.tex
started introduction, with description of "why stream reasoning"

@@
 \begin{nestedsection}{Introduction}{intro}
-	Lack of consensus on stream processing semantics within RDF stream processing community, even more dramatic between RSP and continuous reasoning communities.
+	% general framing of the problem (why reason over streams)
+	As the Semantic Web becomes more extensive, with more people providing and describing their data in the semantic formats recommended by the W3C, the variety in the nature of the data expressed continues to approach that of previous, more established data formats.
+	This includes the growing number of use cases for the continuous processing of streams of semantic data, in which large volumes of time-sensitive data must be processed in near-real time, with the aim of minimising the volume of data transported, duplicated and stored across the Web whilst maximising the variety of problems that can be solved.
+	With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
+	This may be characterised more precisely as continuous deductive reasoning over streamed RDF data.
 
-	Motivating use case (Glacier stuff, to match evaluation?)
+	The value of such processing is the same as that of reasoning over comparatively static data, that being that the division of knowledge into particular knowledge and general knowledge allows for the expression of a greater volume of information in a lesser volume of data.
+	In addition, this division allows the application of a single domain of general knowledge to a variety of distinct sets of particular knowledge, and the application of many distinct domains of general knowledge to a single set of particular knowledge.
+	This is a particularly useful characteristic in the context of semantic stream processing, in which all current use cases consider the application of some static domain of knowledge to some stream of particular knowledge;
+	thus, with stream reasoning the volume of streamed data may be dramatically reduced by extracting the static domain of general knowledge that underpins the data and providing that domain as a static ontology, then streaming only the minimal particular knowledge that would be necessary to deduce the full data set.
+
+	% *brief* summary of state of the art (to be expanded in related work)
+	% 	RDF stream processing and reasoning
+	% 	issues with existing art: no explicit semantics for reasoning
+	% Rete as non-streaming implementationexample (application to dataflow networks)
+	% motivating scenario
 \end{nestedsection}

### 2015-06-29 content/personal-details-ios.tex
Removed middle initial

@@
 	\sep
-	\author{\fnms{Rehab} \snm{?}}
+	\author{\fnms{Rehab} \snm{Albeladi}}
 	and
-	\author{\fnms{Nicholas M.} \snm{Gibbins}}
+	\author{\fnms{Nicholas} \snm{Gibbins}}
 	\runningauthor{D.L. Monks et al.}
@@
 	          United Kingdom\\
-	          E-mail: \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
+	          E-mail: \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
 	}

### 2015-06-29 content/contribution/R4.tex
added that terminal nodes implement instance entailment

@@
 
- 		Finally, terminal nodes receive tokens from the last node in the network and are responsible of producing the new entailed triples.
+ 		Finally, terminal nodes receive tokens from the last node in the network and are responsible of producing the new entailed triples, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
 

### 2015-06-29 content/evaluation.tex
PROBLEM: missing good graph of results.
Otherwise integrated that draft of the evaluation (no conclusions)

@@
 \begin{nestedsection}{Evaluation}{evaluation}
-	Ontology
+	To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning on semantic streaming data.
+	These include Etalis \citep{EP-SPARQL}, Sparkwave \citep{sparkwave}, Streaming knowledge bases \citep{walavalkar08streamingkb}, and the incremental reasoner presented in \citep{inc-reasoning-background-knowledge}.
+	To the best of our knowledge, the latter two implementations were never made public, so we compare R4 against Etalis and Sparkwave only.
+	The experiments are performed on an Intel Core i5 computer with 3.2 GHz and 8 GB memory.
+	We set up Jtalis (the Java wrapper of Etalis) over SWI-Prolog v7.2.1 and installed Sparkwave v0.5.1.
 
-	Data Stream and Window
+	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that cont ains weather observations over a period of ten days.
+	Inspired by a query in \citep{SRBench}, we formulated a rule that asserts the sensors’ information that observe a wind speed higher than a known hurricane (\reffig{SRBench-rule}).
+	\begin{figure}
+		\centering
+		\begin{lstlisting}
+Forall ?ob ?s ?hur (
+    If And( ?ob [ssn:observedProperty -> wind_speed]
+            ?ob [ssn:observedBy -> ?s]
+            ?ob [ssn:observationResult -> ?result]
+            ?result [ssn:hasValue -> ?value]
+            ?value [ssnExt:hasQuantityValue -> ?v]
+            ?hur [rdf:type -> yago:Hurricane111467018]
+            ?hur [dbpprop:1MinWinds -> ?hurWind]
+            External (pred:numeric-greater-than(?v ?hurWind)))
+    Then Do (Print “Wind speed at: ” ?s “ is higher than hurricane: ” ?hur))
+		\end{lstlisting}
+		\caption{}
+		\labelfig{SRBench-rule}
+	\end{figure}
+	This rule involves dealing with static and dynamic data and also needs to perform background ontology reasoning to find compare against all the hurricanes that are subclasses of the specified class of hurricanes.
+	For the historical hurricane data, we prepared a small dataset taken from dbpedia, containing information about a number of hurricanes.
+	We push the whole stream and measure the time needed to process all the data while varying the window size. We ran this rule successfully on R4 and Jtalis getting correct results.
+	However, Sparkwave was unable to read the dataset correctly as it does not parse blank node because it builds hashtables based on the URIs of the incoming data.
 
-	Performance comparison of R4 against Sparkwave and CQELS
+	In order to compare against Sparkwave, we ran the experiment again reusing the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+	We generated a 1.1 million triples dataset containing information about 100,000 offers.
+	We used a small schema that has 329 product types arranged in a 4-levels hierarchy.
+	The rule, in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
+	\begin{figure}
+		\centering
+		\begin{lstlisting}
+Forall ?offer ?product ?vendor ?price ?from ?to(
+    If And( ?offer [rdf:type -> bsbm_voc:Offer]
+            ?offer [bsbm_voc:product -> ?product]
+            ?offer [bsbm_voc:vendor -> ?vendor]
+            ?offer [bsbm_voc:price -> ?price]
+            ?offer [bsbm_voc:validFrom -> ?from]
+            ?offer [bsbm_voc:validTo -> ?to]
+            ?offer [bsbm_voc:deliveryDays -> ?delivery]
+            ?offer [bsbm_voc:offerWebpage -> ?webpage]
+            ?offer [dc:publisher -> ?publisher]
+            ?offer [dcc:date -> ?date])
+    Then ?product [bsbm_voc:offerPrice -> ?price])
+		\end{lstlisting}
+		\caption{}
+		\labelfig{BSBM-rule}
+	\end{figure}.
+	Before we move on to the results, we noticed that for Etalis, when we write the rule with ‘and’ conjunction between the patterns, the system runs very slow (3 hours for the 1.1 million dataset with 1 second time window).
+	However, when we changed the `and' into `seq', the system runs 20x times faster (7.5 minutes for the same test).
+	This is because Etalis is optimized for the \emph{seq} operator.
+	In order to present results that provide a meaningful comparison between the efficiency of each system with regards to changing window sizes, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
+	We recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, but is sufficient to contrast the effect of window size on the efficiency of the three systems.
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.45\textwidth]{all-systems-varying-windows.png}
+		\caption{}
+		\labelfig{all-systems-varying-windows}
+	\end{figure}
+	The X axis shows the time windows in milliseconds while the Y axis shows the processing time in milliseconds.
+	Generally, R4 is faster than Etalis but slower than Sparkwave.
+	The increasing difference between R4 and Sparkwave while increasing the window size is due to the fact that Sparkwave uses a hash join algorithm while the current implementation of R4 uses sequential joins, so as the beta memories (i.e time windows) grow bigger, the join needs significantly longer time to iterate over the windows in the matching process.
+	More efficient join algorithms are considered in the mean time.
+
+	Talk about Sparkwave limited exprisseviness?
 

### 2015-06-29 content/introduction.tex
finished first draft of introduction less the portion on rete, which I am struggling to integrate

@@
 
+	% motivating scenario
+	Take, for example, the case of extensive sensor networks with many and varied sensors, such as those served by the SemsorGrid4Env project\footnote{\url{http://www.semsorgrid4env.eu/}}.
+	Any stand-alone semantic representation of this data may contain classification information for each sensor, according to what they measure, as well as how frequently and how accurately, as well as other factors.
+	It may also contain explicit data concerning events that are reported by no one sensor, and are instead extrapolated from the base sensor data.
+	In both of these cases, the classification criteria and extrapolation process for macro-events may be expressed as general knowledge that is applicable to all sensor readings.
+	As such, a consumer of the streams produced using SemsorGrid4Env could download the appropriate ontologies once, then use them to reason over a reduced set of streamed data containing only the base readings from the sensor network, thus obtaining the classification and macro-event information desired without having to stream it explicitly.
+
 	% *brief* summary of state of the art (to be expanded in related work)
+	Recent research has approached the task of stream reasoning from two different directions.
 	% 	RDF stream processing and reasoning
+	The RDF stream processing community has proposed extensions to their semantic query languages to support the iterative application of queries to the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}.
 	% 	issues with existing art: no explicit semantics for reasoning
+	While this approach is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, the query-by-query windowing semantics of the base systems do not translate to a coherent semantics for multi-rule stream reasoning.
+	On the other hand, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
+	While these algorithms do support a coherent semantics for continuous reasoning, this semantics casts the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams, and so preclude the further processing of entailments in a stream-processing manner.
+	As such, stream reasoning in its current form either lacks a coherent semantics for continuous truth and entailment, or is necessarily a terminal rather than intermediate process in any streamed data pipeline.
+
 	% Rete as non-streaming implementationexample (application to dataflow networks)
-	% motivating scenario
+
+	% overview
+	In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Continuous Datalog.
+	As positive Datalog is also the semantic equivalent of RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, this continuous semantics may be applied to the continuous application of OWL 2 ontologies to streamed RDF data.
+	Next, we present the abstract processing model CoDeR that provides for the expression of processing plans that satisfy the semantics of Continuous Datalog.
+	We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
+	Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-querying benchmarks to contrast the performance of R4 against that of contemporary systems EP-SPARQL and Sparkwave.
+	Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Continuous Datalog.
 \end{nestedsection}

### 2015-06-29 content/introduction.tex
finished first draft of introduction

@@
 	% Rete as non-streaming implementationexample (application to dataflow networks)
+	Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
+	These systems send data from operator to operator, with `memories' compiled for two-pass operators in which to retain the specific data received, in much the same way as sliding windows are used in stream processing.
+	This makes them functionally similar to many stream-to-stream processing systems, and have even been a foundation for some such as Sparkwave \citep{sparkwave}.
+	However, as \citep{forgy79} does not consider the semantics of stream processing, and those semantics are not clearly defined elsewhere for reasoning, such a semantics must still be defined before any stream reasoning system based on Rete may claim to adhere to them.
 
@@
 	We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
-	Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-querying benchmarks to contrast the performance of R4 against that of contemporary systems EP-SPARQL and Sparkwave.
+	Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems EP-SPARQL and Sparkwave.
 	Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Continuous Datalog.

### 2015-06-29 content/contribution/CoDeR.tex
migrated away from annotated join symbols to the new streamjoin symbols (only rstreamjoin is correct at the moment, but it should be trivial to create an lstreamjoin alos)

@@
 		$\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}).
-		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
-		In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
+		% Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
+		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by $\rstreamjoin$.
+		% In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
+		In addition to $\text{:-}_{head}$ and $\rstreamjoin$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
 		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community to the streams of triples on which $\sigma_{triple}$ operate.
-		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
+		% In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
+		In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, \rstreamjoin, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
 
@@
 				In addition, each produced instance inherits its valid-time interval from the graph instance that implied it.
-			\item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
+			% \item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
+			\item[${S_1\,\rstreamjoin\,S_2}$] \hfill \\
 				\begin{figure}[t]
@@
 				Window-joins are, therefore, functionally identical to the union of two complementary SteMs:
-				${A \Join B = \left( A\,{}_b{\Join_p}\,B \right) \cup \left( B\,{}_b{\Join_p}\,A \right)}$, where $A$ and $B$ are inter-operator streams.
+				% ${A \Join B = \left( A\,{}_b{\Join_p}\,B \right) \cup \left( B\,{}_b{\Join_p}\,A \right)}$, where $A$ and $B$ are inter-operator streams.
+				${A \Join B = \left( A\,\rstreamjoin\,B \right) \cup \left( B\,\rstreamjoin\,A \right)}$, where $A$ and $B$ are inter-operator streams.
 				In addition, it should be noted that SteMs express the semantics of the sequential intersection operator ${seq}$ proposed in EP-SPARQL \citep{EP-SPARQL}, in that they join new data from one stream with the prior (valid) contents of another, thereby supporting rudimentary event processing.
@@
 
-		As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+		% As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+		As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
 		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;

### 2015-06-29 content/contribution/R4.tex
migrated away from annotated join symbols to the new streamjoin symbols (only rstreamjoin is correct at the moment, but it should be trivial to create an lstreamjoin alos)

@@
 		This method of annotating tokens with the earliest timestamp of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
-		However, the ordering of tokens in their streams and their propagation through the system expresses those annotation semantics implicitly.
+		However, the ordering of tokens in their streams and their propagation through the synchronous system expresses those annotation semantics implicitly.
 		This leaves the timestamp of a token, in combination with the size of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances in CoDeR.
-		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
+		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
 

### 2015-06-30 content/contribution/continuous-datalog.tex
Minor changes to §2

@@
 The first requirement of stream reasoning is a well-defined semantics.
-	This must take into account the continuous nature of a subset of the base axioms of a stream reasoning problem, as well as its instantaneous truth and entailments, in line with the work of the reasoning community.
-	However, it must also provide a continuous interpretation of those instantaneous entailments in order to integrate with other stream-processing technologies.
-	More specifically, an appropriate semantics for stream reasoning on the semantic web must allow the expression of OWL 2 ontologies to some degree of expressivity, not sacrifice the precise semantics thereof whilst incorporating the semantics of stream processing, and allow for the continuous derivation of entailments in a concise manner.
+This must take into account the continuous nature of a subset of the
+base axioms of a stream reasoning problem, as well as its
+instantaneous truth and entailments, in line with the work of the
+reasoning community.  However, it must also provide a continuous
+interpretation of those instantaneous entailments in order to
+integrate with other stream-processing technologies.  More
+specifically, an appropriate semantics for stream reasoning on the
+Semantic Web must: allow the expression of OWL 2 ontologies to some
+degree of expressivity; not sacrifice the precise semantics thereof
+whilst incorporating the semantics of stream processing; and allow for
+the continuous derivation of entailments in a concise manner.
 
@@
 	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
-	In addition, in can be shown that these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community.
+
+In addition, it can be shown that these streams may be trivially
+manipulated to retrieve the instantaneous sets of entailments produced
+by the more semantically well-defined offerings from the reasoning
+community.
 
@@
 \begin{definition}[Instant]\label{def:continuous datalog: instant}
-		An instant is a discrete point in time.
-The continuous distance in	time between these discrete points may be theoretically infinitesimal, but will in practise have a lower bound set by hardware limitations, or will be defined as part of some specific implementation.
-		In these cases, anything that occurs between discrete instants is considered to have occurred at the next instant.
+An instant is a discrete point in time.  The continuous distance in
+time between these discrete points may be theoretically infinitesimal,
+but will in practise have a lower bound set by hardware limitations,
+or will be defined as part of some specific implementation.  In these
+cases, anything that occurs between discrete instants is considered to
+have occurred at the next instant.
+\end{definition}
+
+\begin{definition}[Fact Instance]\label{def:continuous datalog: fact instance}
+A fact instance is a combination of the signature of a fact (its
+predicate $p$, subject $s$ and object $o$) and its entailment time
+$t$; a fact instance is denoted ${p(s,o):t}$, and is unique.
 \end{definition}
@@
 \begin{definition}[Stream of Facts]\label{def:continuous datalog: stream}
-		A stream of facts is a \emph{sequence} of Datalog facts ordered by some temporal characteristic of each instance of each fact, e.g.
-the time at which each was entailment by some process at the source of the stream.
-		Each instance of a fact is uniquely identifiable by the combination of its signature (its predicate $p$, subject $s$ and object $o$) and its entailment time $t$, so an instance is denoted ${p(s,o):t}$.
-This distinction between instances of facts within a stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
+A stream of facts is a \emph{sequence} of Datalog facts ordered by
+some temporal characteristic of each instance of each fact, e.g. the
+time at which each was entailment by some process at the source of the
+stream.  The distinction between instances of facts within a
+stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
 \end{definition}
@@
 \begin{definition}[Sliding Window]\label{def:continuous datalog: window}
-		A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances of facts that occur in $S$ during the interval ${(i-X,i]}$, and is denoted ${W^{S,X}_{i}}$.
-		The concept of a sliding window within the semantics of Continuous Datalog are further defined by Axioms~\ref{axiom:continuous datalog: window range leq 0} and~\ref{axiom:continuous datalog: window composition}.
+A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at
+any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances
+of facts that occur in $S$ during the interval ${(i-X,i]}$, and is
+  denoted ${W^{S,X}_{i}}$.
 \end{definition}
 
-	The following axioms define the semantics of streams and windows over those streams in Continuous Datalog:
+The following axioms define the semantics of streams and windows over
+those streams in Continuous Datalog:
 
 \begin{axiom}\label{axiom:continuous datalog: window range leq 0}
-		A window of range ${X \leq 0}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is always empty.
+A window of range $0$ over any stream $S$ at any instant 
+${i \in \mathbb{N}}$ is always empty.
 \begin{equation*}
-			W^{S,X}_{i} = \varnothing
+W^{S,0}_{i} = \varnothing
 \end{equation*}
@@
 \begin{axiom}\label{axiom:continuous datalog: window composition}
-		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any given instant ${i \in \mathbb{N}}$ is equivalent to the union of the $X$ consecutive windows of range 1 over the stream $S$ preceding and including that at instant $i$.
+A window of any range ${X \ge 0}$ over any stream $S$ at any
+given instant ${i \in \mathbb{N}}$ is equal to the union of the
+$X$ consecutive windows of range 1 over the stream $S$ preceding and
+including that at instant $i$.
 \begin{equation*}
@@
 \begin{axiom}\label{axiom:continuous datalog: window disjointness}
-		A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any instant ${i \in \mathbb{N}}$ is disjoint from a second window of any range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
+A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
+instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
+range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
 \begin{equation*}

### 2015-06-30 content/contribution/continuous-datalog.tex
separated the discussion of continuous time from the main definition of an 'instant'

@@
 \begin{definition}[Instant]\label{def:continuous datalog: instant}
-An instant is a discrete point in time.  The continuous distance in
-time between these discrete points may be theoretically infinitesimal,
-but will in practise have a lower bound set by hardware limitations,
-or will be defined as part of some specific implementation.  In these
-cases, anything that occurs between discrete instants is considered to
-have occurred at the next instant.
+An instant is a discrete point in time, such that an instant ${i \in \mathbb{N}}$.
+
+It should be noted that time is a continuous concept, with infinitely many divisions possible between any two discrete instants.
+However, the resolution with which such divisions can be recognised will, in practise, have a lower bound set by hardware limitations, or may be defined as part of some specific implementation.
+In these cases, anything that occurs between discrete instants is considered to have occurred at the next recognised instant.
 \end{definition}

### 2015-06-30 content/evaluation.tex
minor typesetting corrections to evaluation section, and added captions

@@
 
-	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that cont ains weather observations over a period of ten days.
+	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
 	Inspired by a query in \citep{SRBench}, we formulated a rule that asserts the sensors’ information that observe a wind speed higher than a known hurricane (\reffig{SRBench-rule}).
-	\begin{figure}
+	\begin{figure*}
 		\centering
-		\begin{lstlisting}
+		\begin{verbatim}
 Forall ?ob ?s ?hur (
@@
     Then Do (Print “Wind speed at: ” ?s “ is higher than hurricane: ” ?hur))
-		\end{lstlisting}
-		\caption{}
+		\end{verbatim}
+		\caption{The RIF-Core rule inspired by SRBench}
 		\labelfig{SRBench-rule}
-	\end{figure}
+	\end{figure*}
 	This rule involves dealing with static and dynamic data and also needs to perform background ontology reasoning to find compare against all the hurricanes that are subclasses of the specified class of hurricanes.
@@
 	We used a small schema that has 329 product types arranged in a 4-levels hierarchy.
-	The rule, in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
-	\begin{figure}
+	The rule, shown in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
+	\begin{figure*}
 		\centering
-		\begin{lstlisting}
+		\begin{verbatim}
 Forall ?offer ?product ?vendor ?price ?from ?to(
@@
     Then ?product [bsbm_voc:offerPrice -> ?price])
-		\end{lstlisting}
-		\caption{}
+		\end{verbatim}
+		\caption{The RIF-Core rule inspired by the Berlin SPARQL benchmark}
 		\labelfig{BSBM-rule}
-	\end{figure}.
+	\end{figure*}
 	Before we move on to the results, we noticed that for Etalis, when we write the rule with ‘and’ conjunction between the patterns, the system runs very slow (3 hours for the 1.1 million dataset with 1 second time window).
@@
 		\includegraphics[width=0.45\textwidth]{all-systems-varying-windows.png}
-		\caption{}
+		\caption{A comparison of the times taken to process 1.1 million triples}
 		\labelfig{all-systems-varying-windows}

### 2015-06-30 content/evaluation.tex
completed the caption for the graph figure
Fixed rendering of RIF rules

@@
 
-	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
+	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that cont ains weather observations over a period of ten days.
 	Inspired by a query in \citep{SRBench}, we formulated a rule that asserts the sensors’ information that observe a wind speed higher than a known hurricane (\reffig{SRBench-rule}).
-	\begin{figure*}
+
+	\begin{figure*}[t]
 		\centering
@@
 		\end{verbatim}
-		\caption{The RIF-Core rule inspired by SRBench}
+		\caption{}
 		\labelfig{SRBench-rule}
 	\end{figure*}
+
 	This rule involves dealing with static and dynamic data and also needs to perform background ontology reasoning to find compare against all the hurricanes that are subclasses of the specified class of hurricanes.
@@
 	We used a small schema that has 329 product types arranged in a 4-levels hierarchy.
-	The rule, shown in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
-	\begin{figure*}
+	The rule, in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
+	\begin{figure*}[t]
 		\centering
@@
 		\end{verbatim}
-		\caption{The RIF-Core rule inspired by the Berlin SPARQL benchmark}
+		\caption{}
 		\labelfig{BSBM-rule}
-	\end{figure*}
+	\end{figure*}.
 	Before we move on to the results, we noticed that for Etalis, when we write the rule with ‘and’ conjunction between the patterns, the system runs very slow (3 hours for the 1.1 million dataset with 1 second time window).
@@
 		\includegraphics[width=0.45\textwidth]{all-systems-varying-windows.png}
-		\caption{A comparison of the times taken to process 1.1 million triples}
+		\caption{}
 		\labelfig{all-systems-varying-windows}

### 2015-06-30 content/contribution/CoDeR.tex

@@
 		Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_{head}$ and the latter being the window-join $\Join$ \citep{niagaraCQ}, in order to support forward-chaining processing of fixed rules;
-		$\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}).
+		$\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}.
 		% Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
@@
 		In addition to $\text{:-}_{head}$ and $\rstreamjoin$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
-		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community to the streams of triples on which $\sigma_{triple}$ operate.
+		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}} to the streams of triples on which $\sigma_{triple}$ operates.
 		% In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
@@
 			\item[$\sigma_{triple}\,S$] \hfill \\
+				\begin{figure}[t]
+					\centering
+					\includegraphics[width=0.3\textwidth]{basic-pattern-match}
+					\caption{The triple-select operator.}
+					\labelfig{continuous datalog: basic pattern match}
+				\end{figure}
 				The triple-select operator identifies instances of triples in a stream $S$ that match some \emph{basic pattern} \citep{w3csparql}.
@@
 				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
-				\begin{figure}[t]
-					\centering
-					\includegraphics[width=0.45\textwidth]{basic-pattern-match}
-					\caption{The triple-select operator.}
-					\labelfig{continuous datalog: basic pattern match}
-				\end{figure}
 			\item[$\sigma_{external}\,S$] \hfill \\
@@
 					\centering
-					\includegraphics[width=0.45\textwidth]{datatype-function}
+					\includegraphics[width=0.3\textwidth]{datatype-function}
 					\caption{The operator encompassing all datatype functions of RIF-Core.}
@@
 					\centering
-					\includegraphics[width=0.45\textwidth]{graph-to-triple}
+					\includegraphics[width=0.3\textwidth]{graph-to-triple}
 					\caption{The graph-to-triple operator.}
@@
 					\centering
-					\includegraphics[width=0.45\textwidth]{instance-implication}
+					\includegraphics[width=0.3\textwidth]{instance-implication}
 					\caption{The instance implication operator.}

### 2015-06-30 content/contribution/continuous-datalog.tex
Fixes to defn of axioms/facts

@@
 \begin{definition}[Fact Instance]\label{def:continuous datalog: fact instance}
-A fact instance is a combination of the signature of a fact (its
-predicate $p$, subject $s$ and object $o$) and its entailment time
-$t$; a fact instance is denoted ${p(s,o):t}$, and is unique.
+A fact instance is a fact with some additional temporal annotation; it
+is the combination of a fact's signature (its predicate $p$, subject $s$
+and object $o$) and a timestamp $t$ that is used to record the time at
+which the fact was entailed. A fact instance is denoted ${p(s,o):t}$,
+and is unique.
+\end{definition}
+
+\begin{definition}[]\label{def:continuous datalog: trad}
+A fact instance may be stripped of its temporal annotation using the $trad()$
 \end{definition}
@@
 
-	Traditionally, a set of axioms $A$ entails another set of axioms $E$ in which the existence of one fact in $E$ may be justified by many distinct subsets of $A$.
-	When considering streams of facts, in which one fact may occur multiple times and be considered distinct based on its entailment time, temporal information would be lost if sets of axioms including fact instances (denoted $A_{inst}$) were to entail sets of axioms without.
-	As such, where a fact would be entailed by some set of facts $A$ that are instantiated on some stream, an instance of that fact should be entailed by the set of instances of those facts present on that stream.
-	Furthermore, each such justification in $A_{inst}$ for a fact entails a unique instance of that fact, such that each entailed instance has a single justification within the original set of axioms, and many instances of the entailed fact may occur in the entailed set of axioms $E_{inst}$.
-	This form of entailment shall be henceforth referred to as ``instance entailment'', and shall be the norm when discussing the entailments of sets of axioms that include instances of facts.
+In traditional Datalog, the set of axioms in the system consist of
+rules and facts. Taken together, a set of axioms $A$ entails another
+set of axioms $E$ in which the existence of one fact in $E$ may be
+justified by many distinct subsets of $A$.  When considering streams
+of facts, in which there may be many distinct instances of a given
+fact, temporal information would be lost if sets of axioms including
+fact instances (denoted $A_{inst}$) were to entail sets of axioms
+containing plain facts.  As such, where a fact $f$ would be entailed
+by some set of facts $A$ that are instantiated on some stream, an
+instance of $f$ should be entailed by the set of instances of those
+facts present on that stream.  Furthermore, each such justification in
+$A_{inst}$ for a fact entails a unique instance of that fact, such
+that each entailed instance has a single justification within the
+original set of axioms, and many instances of the entailed fact may
+occur in the entailed set of axioms $E_{inst}$.  This form of
+entailment shall be henceforth referred to as ``instance entailment'',
+and shall be the norm when discussing the entailments of sets of
+axioms that include instances of facts.
 

### 2015-06-30 content/contribution/R4.tex
added a new section to differentiate making Rete continuous and the specifics of reasoning in R4

@@
 	A rule document containing any number of rules is submitted to the system.
-	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards.
+	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
 	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
-	We also added a simple extension to RIF-Core to enable users to specify window constraints as part of their rules, as shown in \reffig{lst-S-RIF-Core}.
-	Further, we propose the extension of the set of PROFILEs that describe the IMPORTs of a rule set to include those definitions of RDF streams recognised by the W3C.
+	We also added a simple extension to RIF-Core in line with the extension described in \refsec{semantics}, enabling users to specify window constraints as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
+	Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
 
@@
 
-	These networks read TBox input data, and receive streaming data in the form of RDF streams and operate directly on them.
+	These networks read TBox input data, receive streaming data in the form of RDF streams and operate directly on them.
 	We chose this RDF-native approach as opposed to reusing existing technologies as in \citep{C-SPARQL,streaming-sparql} to allow full control over the low-level operators.
@@
 
-	\begin{nestedsection}{Continuous Reasoning}{implementation: continuous reasoning}
+	\begin{nestedsection}{Continuous Reasoning with Rete}{implementation: continuous rete}
 		The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, can be well fit into the stream reasoning model as it operates incrementally.
@@
 		Therefore, we place time constraints on the memories using a window-join operator instead of the traditional joins.
-
-		Data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
+	\end{nestedsection}
+	\begin{nestedsection}{Continuous Reasoning in the Semantic Web}{implementation: continuous reasoning with RDF}
+		In R4, data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
 		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.

### 2015-07-01 content/evaluation.tex
made many edits to evaluation section to clarify the nature of the experiments, as well as why the motivating case was abandoned, and added the graph for memory-consumption analysis

@@
-\begin{nestedsection}{Evaluation}{evaluation}
+\begin{nestedsection}{Evaluation, Results and Discussion}{evaluation}
 	To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning on semantic streaming data.
@@
 
-	We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that cont ains weather observations over a period of ten days.
-	Inspired by a query in \citep{SRBench}, we formulated a rule that asserts the sensors’ information that observe a wind speed higher than a known hurricane (\reffig{SRBench-rule}).
-
-	\begin{figure*}[t]
+	\begin{nestedsection}{Motivating Case}{evaluation: motivating case}
+		We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
+		Inspired by a query in \citep{SRBench}, we formulated a rule that entails that a given observation is of a weather pattern classified as a hurricane if it is of a wind speed that exceeds that of a previously classified hurricane;
+		it also entails that the wind speed observed exceeds that of a given historically recorded hurricane (\reffig{SRBench-rule}).
+		\begin{figure*}
 			\centering
 			\begin{verbatim}
-Forall ?ob ?s ?hur (
-    If And( ?ob [ssn:observedProperty -> wind_speed]
-            ?ob [ssn:observedBy -> ?s]
-            ?ob [ssn:observationResult -> ?result]
+Forall ?obs ?sensor ?result ?value ?v ?hur ?hurWind (
+    If And( ?obs [ssn:observedProperty -> wind_speed]
+    		?obs [ssn:observedBy -> ?sensor]
+            ?obs [ssn:observationResult -> ?result] 
             ?result [ssn:hasValue -> ?value]
@@
             ?hur [dbpprop:1MinWinds -> ?hurWind] 
-            External (pred:numeric-greater-than(?v ?hurWind)))
-    Then Do (Print “Wind speed at: ” ?s “ is higher than hurricane: ” ?hur))
+            External (pred:numeric-greater-than(?v ?hurWind)) )
+    Then ?obs [ssn:observedWeatherType -> yago:Hurricane111467018] )
 			\end{verbatim}
-		\caption{}
+			\caption{The RIF-Core rule inspired by SRBench.}
 			\labelfig{SRBench-rule}
 		\end{figure*}
+		This rule involves dealing with static data regarding past hurricanes, dynamic data regarding ongoing sensor readings, and also needs to perform reasoning against an ontology of background knowledge to identify the hurricanes that are instances of the specified class of hurricanes.
+		For the historical hurricane data, we prepared a small dataset taken from dbpedia, containing information about a number of hurricanes.
 
-	This rule involves dealing with static and dynamic data and also needs to perform background ontology reasoning to find compare against all the hurricanes that are subclasses of the specified class of hurricanes.
-	For the historical hurricane data, we prepared a small dataset taken from dbpedia, containing information about a number of hurricanes.
-	We push the whole stream and measure the time needed to process all the data while varying the window size. We ran this rule successfully on R4 and Jtalis getting correct results.
-	However, Sparkwave was unable to read the dataset correctly as it does not parse blank node because it builds hashtables based on the URIs of the incoming data.
+		The experiment consisted of pushing the whole stream to the compiled R4 system and measuring the time taken to process all the data, repeated for a variety the window sizes.
+		We ran this rule successfully on R4 and Jtalis getting correct results.
+		However, Sparkwave was unable to read the dataset correctly as it does not parse blank nodes;
+		this is because it uses the hash-join algorithm to improve time efficiency, as we show in \refsec{evaluation: comparative results}, but builds the hashtables based on the URIs of the incoming data.
+	\end{nestedsection}
+	\begin{nestedsection}{Comparative Results}{evaluation: comparative results}
+		In order to compare R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+		We generated a 1.1 million triples dataset containing information about 100,000 limited-time offers made available by some online marketplace.
+		We used a small schema that has 329 product types arranged in a 4-level hierarchy.
 
-	In order to compare against Sparkwave, we ran the experiment again reusing the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
-	We generated a 1.1 million triples dataset containing information about 100,000 offers.
-	We used a small schema that has 329 product types arranged in a 4-levels hierarchy.
-	The rule, in \reffig{BSBM-rule}, asks for offers for products of a given type, which means that background reasoning is needed for non-leaf products.
-	\begin{figure*}[t]
+		The rule expressed in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer, which means that background reasoning is needed for non-leaf products.
+		\begin{figure*}
 			\centering
 			\begin{verbatim}
-Forall ?offer ?product ?vendor ?price ?from ?to(
+Forall ?offer ?product ?vendor ?price ?from ?to ?delivery ?webpage(
     If And( ?offer [rdf:type -> bsbm_voc:Offer]
@@
             ?offer [dc:publisher -> ?publisher]
-            ?offer [dcc:date -> ?date])
-    Then ?product [bsbm_voc:offerPrice -> ?price])
+            ?offer [dcc:date -> ?date] )
+    Then ?product [bsbm_voc:offerPrice -> ?price] )
 			\end{verbatim}
-		\caption{}
+			\caption{The RIF-Core rule inspired by the Berlin SPARQL benchmark.}
 			\labelfig{BSBM-rule}
-	\end{figure*}.
-	Before we move on to the results, we noticed that for Etalis, when we write the rule with ‘and’ conjunction between the patterns, the system runs very slow (3 hours for the 1.1 million dataset with 1 second time window).
-	However, when we changed the `and' into `seq', the system runs 20x times faster (7.5 minutes for the same test).
-	This is because Etalis is optimized for the \emph{seq} operator.
-	In order to present results that provide a meaningful comparison between the efficiency of each system with regards to changing window sizes, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
-	We recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, but is sufficient to contrast the effect of window size on the efficiency of the three systems.
+		\end{figure*}
+		Before we move on to the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window).
+		However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
+		This is likely to be because Etalis is optimized for the \emph{seq} operator, being intended for \emph{event processing} rather than continuous querying/reasoning, which makes relevant the order of arrival of triples rather than simply their coincidence in the system.
+		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window sizes, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
+		We recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, but is sufficient to contrast the effect of window size on the time efficiency of the three systems.
 		\begin{figure}
@@
 			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows.png}
-		\caption{}
+			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various sizes over the input stream.}
 			\labelfig{all-systems-varying-windows}
@@
 		The X axis shows the time windows in milliseconds while the Y axis shows the processing time in milliseconds.
-	Generally, R4 is faster than Etalis but slower than Sparkwave.
-	The increasing difference between R4 and Sparkwave while increasing the window size is due to the fact that Sparkwave uses a hash join algorithm while the current implementation of R4 uses sequential joins, so as the beta memories (i.e time windows) grow bigger, the join needs significantly longer time to iterate over the windows in the matching process.
-	More efficient join algorithms are considered in the mean time.
 
-	Talk about Sparkwave limited exprisseviness?
+		As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small sliding windows but slower than Sparkwave, and, while the time-to-complete increases with window size in Sparkwave and R4, that of Etalis appears to remain fairly constant for the sliding-window sizes tested.
+		The growing difference between R4 and Sparkwave as the size of the sliding windows increases is likely due to the fact that Sparkwave uses a hash join algorithm while the current implementation of R4 uses loop joins;
+		as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer time to iterate over the windows in the matching process compared to the hash-based look-up of those in Sparkwave.
+		More efficient, multi-index window-joins are to be considered for R4 in future work.
+
+		Discussion about memory consumption...
+		\begin{figure}
+			\centering
+			\includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
+			\caption{A comparison of the memory consumed by the systems R4, Sparkwave and ETALIS throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
+			\labelfig{all-systems-memory-consumption}
+		\end{figure}
+	\end{nestedsection}
 

### 2015-07-01 content/personal-details-els.tex
changed title to final
Removed middle initial
set myself as the corresponding author
Fixed personal details

@@
 \begin{frontmatter}
-	\title{Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality}
+	\title{Continuous Deductive Reasoning over Streamed RDF Data}
 
@@
 	\ead{raab1g09@ecs.soton.ac.uk}
-	\author{Nicholas M. Gibbins}
+	\author{Nicholas Gibbins}
 	\ead{nmg@ecs.soton.ac.uk}
@@
 	\address{
-	          Department of Electronics and Computer Science,\\
-	          Faculty of Physical Sciences and Engineering,\\
+	          Electronics and Computer Science,\\
 	          University of Southampton,\\
-	          University Rd., Southampton, SO17 1BJ,\\
+	          Southampton, SO17 1BJ,\\
 	          United Kingdom

### 2015-07-01 content/contribution/continuous-datalog.tex
Fixed minor bugs

--- 2015-06-30 content/contribution/continuous-datalog.tex
@@
 
-	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
+%% fixme
+%	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
 
-\begin{definition}[Instant]\label{def:continuous datalog: instant}
-An instant is a discrete point in time, such that an instant ${i \in \mathbb{N}}$.
+\begin{definition}[Datalog Program]
+\label{def:continuous datalog: datalog program}
+A conventional Datalog program $\Pi = \langle R, F\rangle$ is a tuple
+which comprises a finite set of rules $R$ and a finite set of facts
+$F$. Facts are extensional atomic formulae of first-order logic, and
+rules are of the form:
 
-It should be noted that time is a continuous concept, with infinitely many divisions possible between any two discrete instants.
-However, the resolution with which such divisions can be recognised will, in practise, have a lower bound set by hardware limitations, or may be defined as part of some specific implementation.
-In these cases, anything that occurs between discrete instants is considered to have occurred at the next recognised instant.
+\[\alpha_0 :- \alpha_1 , \ldots , \alpha_n\]
+
+\noindent where $\alpha_0$ is an intensional atomic formula and
+$\alpha_1 \ldots \alpha_n$ are a mixture of intensional and
+extensional atomic formulae.
+
+For the purpose of defining Continuous Datalog in the context of the
+Semantic Web, we treat facts as 3-tuples $\langle s, p, o \rangle$,
+where $s, p, o$ are respectively the subject, predicate and object of
+an RDF triple. 
 \end{definition}
 
-\begin{definition}[Fact Instance]\label{def:continuous datalog: fact instance}
-A fact instance is a combination of the signature of a fact (its
-predicate $p$, subject $s$ and object $o$) and its entailment time
-$t$; a fact instance is denoted ${p(s,o):t}$, and is unique.
+\begin{definition}[Instant]
+\label{def:continuous datalog: instant}
+An instant is a discrete point in time, such that an instant 
+${i \in \mathbb{N}}$.
+
+It should be noted that time is a continuous concept, with infinitely
+many divisions possible between any two discrete instants.  However,
+the resolution with which such divisions can be recognised will, in
+practise, have a lower bound set by hardware limitations, or may be
+defined as part of some specific implementation.  In these cases,
+anything that occurs between discrete instants is considered to have
+occurred at the next recognised instant.
 \end{definition}
 
-\begin{definition}[Stream of Facts]\label{def:continuous datalog: stream}
-A stream of facts is a \emph{sequence} of Datalog facts ordered by
-some temporal characteristic of each instance of each fact, e.g. the
-time at which each was entailment by some process at the source of the
-stream.  The distinction between instances of facts within a
-stream is represented in Axiom~\ref{axiom:continuous datalog: window disjointness}.
+\begin{definition}[Fact Instance]
+\label{def:continuous datalog: fact instance}
+A {\em fact instance} is a fact with some additional temporal
+annotation; it is the combination of a fact's signature (its predicate
+$p$, subject $s$ and object $o$) and a timestamp $t$ (an instant) that
+is used to record the time at which the fact was asserted or
+entailed. A fact instance is denoted $\langle s, p, o \rangle \! : \!
+t$, and is unique.
+
+We represent the relationship between a fact (without temporal
+annotation) and a fact instance (with temporal annotation) with the
+function $strip$ which maps a fact instance to its underlying fact.
 \end{definition}
 
-\begin{definition}[Sliding Window]\label{def:continuous datalog: window}
-A sliding window of range ${X \in \mathbb{N}}$ over any stream $S$ at
-any given instant ${i \in \mathbb{N}}$ is the \emph{set} of instances
-of facts that occur in $S$ during the interval ${(i-X,i]}$, and is
-  denoted ${W^{S,X}_{i}}$.
+\begin{definition}[Stream of Facts]
+\label{def:continuous datalog: stream}
+A stream of facts $S$ is a \emph{sequence} of fact instances ordered
+by their timestamps; the distinction between multiple instances of a
+fact within a stream is represented in 
+Axiom~\ref{axiom:continuous datalog: window disjointness}.
+\end{definition}
+
+\begin{definition}[Sliding Window]
+\label{def:continuous datalog: window}
+A sliding window of range ${x \in \mathbb{N}}$ over any stream $S$ at
+any given instant ${i \in \mathbb{N}}$ is the \emph{set} of fact
+instances that occur in $S$ during the interval ${(i-x,i]}$, and is
+  denoted ${W^{S,x}_{i}}$.
 \end{definition}
@@
 \begin{axiom}\label{axiom:continuous datalog: window range leq 0}
-A window of range $0$ over any stream $S$ at any instant 
+A window of range $0$ over a stream $S$ at an instant 
 ${i \in \mathbb{N}}$ is always empty.
@@
 \begin{axiom}\label{axiom:continuous datalog: window composition}
-A window of any range ${X \ge 0}$ over any stream $S$ at any
-given instant ${i \in \mathbb{N}}$ is equal to the union of the
-$X$ consecutive windows of range 1 over the stream $S$ preceding and
-including that at instant $i$.
+A window of range ${x \ge 0}$ over a stream $S$ at an instant 
+${i \in \mathbb{N}}$ is equal to the union of the $x$ consecutive
+windows of range $1$ over the stream $S$ preceding and including that at
+instant $i$.
 \begin{equation*}
-W^{S,X}_{i} = \mathop{\cup}_{i=0}^{X-1} W^{S,1}_{i-i}
+W^{S,x}_{i} = \mathop{\cup}_{j=0}^{x-1} W^{S,1}_{i-j}
 \end{equation*}
@@
 \begin{axiom}\label{axiom:continuous datalog: window disjointness}
-A window of any range ${X \in \mathbb{N}}$ over any stream $S$ at any
-instant ${i \in \mathbb{N}}$ is disjoint from a second window of any
-range ${Y \in \mathbb{N}}$ over $S$ at any instant after ${i - X}$.
+A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
+instant ${i \in \mathbb{N}}$ is disjoint from a second window of 
+range ${y \in \mathbb{N}}$ over $S$ at any instant before $i - x$.
+
 \begin{equation*}
-\forall X' \geq X \, \left( W^{S,X}_{i} \cap W^{S,Y}_{i-X'} = \varnothing \right)
+W^{S,x}_{i} \cap W^{S,y}_{i-z} = \varnothing
 \end{equation*}
+
+\noindent where $z \geq x$.
 \end{axiom}
 
-	Traditionally, a set of axioms $A$ entails another set of axioms $E$ in which the existence of one fact in $E$ may be justified by many distinct subsets of $A$.
-	When considering streams of facts, in which one fact may occur multiple times and be considered distinct based on its entailment time, temporal information would be lost if sets of axioms including fact instances (denoted $A_{inst}$) were to entail sets of axioms without.
-	As such, where a fact would be entailed by some set of facts $A$ that are instantiated on some stream, an instance of that fact should be entailed by the set of instances of those facts present on that stream.
-	Furthermore, each such justification in $A_{inst}$ for a fact entails a unique instance of that fact, such that each entailed instance has a single justification within the original set of axioms, and many instances of the entailed fact may occur in the entailed set of axioms $E_{inst}$.
-	This form of entailment shall be henceforth referred to as ``instance entailment'', and shall be the norm when discussing the entailments of sets of axioms that include instances of facts.
+In Definition~\ref{def:continuous datalog: datalog program}, we took a
+Datalog program to consist of a set of rules and a set of facts.
+Taken together, the set of facts $F$ entails another set of factss $E$
+in which the existence of one fact in $E$ may be justified by many
+distinct subsets of $F$.
 
-	It should be noted that removing all temporal information from a set of axioms including fact instances $A_{inst}$ produces the set of traditional Datalog axioms $A$ instantiated in $A_{inst}$.
-	This is be represented as the function ${trad\left( A_{inst} \right) = A}$.
-	In addition, applying the ${trad()}$ function to both the original axiom instances $A_{inst}$ and the entailed axiom instances $E_{inst}$ maintains the relationship of entailment between the two sets:
-	\[ \left( A_{inst} \vDash E_{inst} \right) \rightarrow \left( A \vDash E \right) \]
+When considering streams of facts, in which there may be many distinct
+instances of a given fact, temporal information would be lost if sets
+of fact instances (denoted $F_{inst}$) were to entail sets of plain
+facts.  As such, where a fact $f$ would be entailed by some set of
+facts $F$ that are instantiated on some stream, an instance of $f$
+should be entailed by the set of instances of those facts present on
+that stream.  Furthermore, each such justification in $F_{inst}$ for a
+fact entails a unique instance of that fact, such that each entailed
+instance has a single justification within the original set of facts,
+and many instances of the entailed fact may occur in the entailed set
+of facts $E_{inst}$.  This form of entailment shall be henceforth
+referred to as {\em instance entailment}, and shall be the norm when
+discussing the entailments of sets of fact instances.
 
-	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.
+It should be noted that removing all temporal information from a set
+of fact instances $F_{inst}$ produces the set of facts $F$
+instantiated in $F_{inst}$:
 
-	\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-		If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
+\[F = \{ strip(x) | x \in F_{inst} \}\]
+
+In addition, applying the ${strip()}$ function to both the original
+fact instances $F_{inst}$ and the entailed fact instances $E_{inst}$
+maintains the relationship of entailment between the two sets:
+
+\[ \left( F_{inst} \vDash E_{inst} \right) \rightarrow \left( F \vDash E \right) \]
+
+From this description of instance entailment,
+Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be
+derived:
+
+\begin{axiom}
+\label{axiom:continuous datalog: entailment disjointness}
+
+If two sets of fact instances, $A_{inst}$ and $B_{inst}$, are
+disjoint, then a set $F_{inst1}$ of entailed fact instances (each of
+which is justified at least in part by $A_{inst}$) is necessarily
+disjoint from any set $F_{inst2}$ of entailed fact instances (each
+justified at least in part by $B_{inst}$), because entailed fact
+instances are distinct with regards to the fact instances that justify
+them.
+
 \begin{multline*}
@@
 
-	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
+Given these definitions, a Continuous Datalog program may be formally
+defined, both in terms of streams and in terms of its instantaneous
+states:
 
-	\begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
-		A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
-		This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
-		Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
-		Streams $P$ and $N$ are therefore the continuous reasoning equivalents of the streams produced by the \emph{IStream} and \emph{DStream} operators of the \emph{CQL} continuous query language \citep{CQL} respectively.
-		The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.
-		\begin{equation*}
-			CDP = \{A,S,R\} \vDash \{A',P,N\}
-		\end{equation*}
+\begin{definition}[Continuous Datalog Program]
+\label{def:continuous datalog: CDP}
+
+A Continuous Datalog Program (CDP) is a tuple $\langle R, F, S, r
+\rangle$ which comprises a finite set of rules $R$, a finite set of
+facts $F$, a stream of facts $S$ and a range $r \in \mathbb{N}$ for
+the window over $S$. This specifies that the streamed data from $S$ is
+to be considered valid from its arrival for the duration represented
+by $r$.
+
+The entailment of such a program is composed of a stream of entailed
+instances $P$ and a stream of instance negations $N$, whose contents
+are defined in 
+Axioms~\ref{axiom:continuous datalog: positive window increment}, 
+\ref{axiom:continuous datalog: negative window increment} and
+\ref{axiom:continuous datalog: entailment precedes negation}. 
+The streams $P$ and $N$ are therefore the continuous
+reasoning equivalents of the streams produced by the \emph{IStream}
+and \emph{DStream} operators of the \emph{CQL} continuous query
+language \citep{CQL} respectively.  The truth value of a CDP is
+instantaneous, with the the truth function of Continuous Datalog
+$T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in
+  \mathbb{N}}$ as arguments.
 \end{definition}
 
-	\begin{definition}[State of a ${CDP}$]\label{def:continuous datalog: CDPt}
-		The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
-		As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.
-		The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$.
+\begin{definition}[State of a CDP]
+\label{def:continuous datalog: CDPt}
+
+The state of a CDP at a given instant $i$ (denoted as ${CDP_{i}}$) is
+the union of the set of facts $F$ and the window $W^{S,r}_{i}$.  As both $F$
+and ${strip( W^{S,r}_{i} )}$ are sets of Datalog axioms, 
+${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated
+by the truth function of Datalog $T^{D}$, the result of which is the
+definition for the instantaneous truth of the ${CDP}$ at instant $i$
+by $T^{CD}$.  The instantaneous entailments of a $CDP_{i}$ are the
+union of the set $E$, the persistent entailments of $F$ (which are
+guaranteed to be entailed by every state, as Positive Datalog is
+monotonic), and the set $E_{i}$ of instances of the transient
+entailments that are justified at least in part by some subset of
+$W^{S,r}_{i}$.
+
 \begin{equation*}
@@
 \begin{equation*}
-			CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
+CDP_{i} = F \cup W^{S,r}_{i} \vDash E \cup E_{i}
 \end{equation*}
@@
 
-	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
+The following axioms define the semantics of the streams of
+entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the
+instantaneous entailments of the sequence of states of that continuous
+program.
 
-	\begin{axiom}\label{axiom:continuous datalog: positive window increment}
-		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.
+\begin{axiom}
+\label{axiom:continuous datalog: positive window increment}
+
+The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
+the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
+that are entailed by the state of the ${CDP}$ at instant $i$ that were
+not entailed by the state of ${CDP}$ at the previous instant instance
+${i-1}$.
+
 \begin{equation*}
-			W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
+W^{P,1}_{i} = \left( E \cup E_{i} \right) \setminus \left( E \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
 \end{equation*}
@@
 
-	\begin{axiom}\label{axiom:continuous datalog: negative window increment}
-		The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.
+\begin{axiom}
+\label{axiom:continuous datalog: negative window increment}
+The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
+the stream $N$ entailed by a ${CDP}$ is defined as the set of facts
+that were entailed by the state of the ${CDP}$ at the previous instant
+instance ${i-1}$ that are no longer entailed by the state of ${CDP}$
+at instant $i$.
 \begin{equation*}
@@
 
-	\begin{axiom}\label{axiom:continuous datalog: entailment precedes negation}
-		Facts may only be negated by some part of their justification leaving the sliding window $W^{S,R}_{i}$, this necessarily being after they were entailed.
-		As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$
+\begin{axiom}
+\label{axiom:continuous datalog: entailment precedes negation}
+Facts may only be negated by some part of their justification leaving
+the sliding window $W^{S,r}_{i}$, this necessarily being after they
+were entailed.  As such, a window of any range ${x \in \mathbb{N}}$
+over the stream of entailments $P$ at any given instant ${i \in
+  \mathbb{N}}$ is disjoint from a window of any range ${x \in
+  \mathbb{N}}$ over the the stream of entailment negations $N$ at any
+instant after ${i - X}$
+
+CHECK
+
 \begin{equation*}
-			\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)
+\forall X' \geq X \left( W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing \right)
 \end{equation*}
+
+\noindent where $z \geq x$.
 \end{axiom}
 
-	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes newly justified by the ${CDP}$
-	(more specifically, by some subset of the instances in $W^{S,R}_{i_{e}}$).
-	As such, the stream of entailment instances $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
-	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
-	(i.e.
-the first instant ${i_{n} \geq i_{e}}$ in which $W^{S,R}_{i_{n}}$ no longer contains the subset of fact instances first present in $W^{S,R}_{i_{e}}$ that justified the entailment instance ${p(s,o):t}$).
+From this, we can see that each instance of an entailment in some
+$E_{i}$ may take the same form as an instance of any other fact in a
+stream of facts, being $\langle s, p, o \rangle \! : \!t$, where $t$ is
+the real time associated with the instant $i_{e}$ in which
+$\langle s, p, o \rangle$ becomes newly justified by the ${CDP}$ (more
+specifically, by some subset of the instances in $W^{S,r}_{i_{e}}$).
+As such, the stream of entailment instances $P$ of a ${CDP}$ takes
+the same form as the \emph{streams of facts} already described, thus
+producing a concise stream of entailments that may be read by
+further stream processing.  In addition, the stream of negations $N$
+of a ${CDP}$ also takes this form, but differs from the streams
+discussed thus far in that it is not ordered by the entailment time
+$i_{e}$ of the entailment instances but by the time at which their
+justification \emph{ceases} to be valid (i.e. the first instant
+${i_{n} \geq i_{e}}$ in which $W^{S,r}_{i_{n}}$ no longer contains
+the subset of fact instances first present in $W^{S,r}_{i_{e}}$ that
+justified the entailment instance $\langle s, p, o \rangle \! : \!t$).
 
-	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
-	However, it can be shown from these axioms,
-	as in \ref{sec:semantics: proof for entailments},
-	that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
+Thus far the translation from instantaneous sets of entailments to
+streams of entailments has been stated axiomatically as the definition
+of those streams.  However, it can be shown from these axioms, as in
+\ref{sec:semantics: proof for entailments}, that it is also possible
+to reconstruct exactly the instantaneous set $E_{i}$ of entailment
+instances of any ${CDP}$ at any given instant from the windows of
+sufficient range over its entailed streams $P$ and $N$ at that
+instant.
+
+CHECK
+
 \[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
-	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
-	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
-	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
+
+The minimum range $x$ is such to achieve the \emph{complete} set of
+entailment instances $E_{i}$ of the ${CDP}_{i}$.  In addition, the
+minimum range $y$ is such to achieve the \emph{correct} set $E_{i}$.
+As such, a \emph{valid window} that contains only and exactly those
+instances of entailments in $E_{i}$ at every instant $i$ may be
+trivially maintained from the streams $P$ and $N$ continuously
+entailed by some ${CDP}$.
 \end{nestedsection}

### 2015-07-07 content/contribution/continuous-datalog.tex

@@
 
-	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
+	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined proof-theoretic semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
 	This comprises the extension of Datalog programs to include sources of continuously generated/reported facts, and the application of sliding window semantics over those sources, thereby providing for the interpretation of Continuous Datalog programs as sequences of instantaneous states that are each valid positive Datalog programs.
@@
 \begin{definition}[Datalog Program]
-\label{def:continuous datalog: datalog program}
+\labeldef{continuous datalog: datalog program}
 A conventional Datalog program $\Pi = \langle R, F\rangle$ is a tuple
@@
 rules are of the form:
-
-\[\alpha_0 :- \alpha_1 , \ldots , \alpha_n\]
-
-\noindent where $\alpha_0$ is an intensional atomic formula and
+\[\alpha_0 \text{ :- } \alpha_1 , \ldots , \alpha_n\]
+where $\alpha_0$ is an intensional atomic formula and
 $\alpha_1 \ldots \alpha_n$ are a mixture of intensional and
@@
 
+\begin{definition}[Entailment Justification]
+\labeldef{continuous datalog: justification}
+A justification for some entailment $f$ of some Datalog program
+${\Pi = \langle R , F \rangle \vdash f}$ is a minimal set ${J \subseteq F}$
+such that ${\Pi' = \langle R , J \rangle \vdash f}$.
+\end{definition}
+
 \begin{definition}[Instant]
-\label{def:continuous datalog: instant}
+\labeldef{continuous datalog: instant}
 An instant is a discrete point in time, such that an instant 
@@
 \begin{definition}[Fact Instance]
-\label{def:continuous datalog: fact instance}
-A {\em fact instance} is a fact with some additional temporal
-annotation; it is the combination of a fact's signature (its predicate
-$p$, subject $s$ and object $o$) and a timestamp $t$ (an instant) that
-is used to record the time at which the fact was asserted or
-entailed. A fact instance is denoted $\langle s, p, o \rangle \! : \!
-t$, and is unique.
+\labeldef{continuous datalog: fact instance}
+A {\em fact instance} is a fact with some additional temporal annotation;
+it is the combination of a fact's signature (its predicate $p$, subject
+$s$ and object $o$) and the timestamp $t$ at which the instance became true.
+A fact instance is denoted ${\langle s, p, o \rangle | t \in \mathbb{F^T}}$.
 
@@
 annotation) and a fact instance (with temporal annotation) with the
-function $strip$ which maps a fact instance to its underlying fact.
+function ${strip : \mathbb{F^T} \rightarrow \mathbb{F}}$ which maps a fact instance to its
+underlying fact ${\langle s, p, o \rangle \in \mathbb{F}}$.
 \end{definition}
@@
 \begin{definition}[Stream of Facts]
-\label{def:continuous datalog: stream}
+\labeldef{continuous datalog: stream}
 A stream of facts $S$ is a \emph{sequence} of fact instances ordered
-by their timestamps; the distinction between multiple instances of a
-fact within a stream is represented in 
+by their timestamps; in conjunction with the distinction between
+multiple instances of a fact by their timestamp, this definition justifies
 Axiom~\ref{axiom:continuous datalog: window disjointness}.
@@
 \begin{definition}[Sliding Window]
-\label{def:continuous datalog: window}
+\labeldef{continuous datalog: window}
 A sliding window of range ${x \in \mathbb{N}}$ over any stream $S$ at
@@
 range ${y \in \mathbb{N}}$ over $S$ at any instant before $i - x$.
-
 \begin{equation*}
@@
 \end{equation*}
-
-\noindent where $z \geq x$.
+where $z \geq x$.
 \end{axiom}
 
-In Definition~\ref{def:continuous datalog: datalog program}, we took a
+In \refdef{continuous datalog: datalog program}, we took a
 Datalog program to consist of a set of rules and a set of facts.
-Taken together, the set of facts $F$ entails another set of factss $E$
-in which the existence of one fact in $E$ may be justified by many
-distinct subsets of $F$.
+Taken together with the set of rules $R$, the set of facts $F$ entails
+another set of facts $E$ in which the existence of one fact in $E$ may
+be justified by many distinct subsets of $F$. In order to extend this
+definition of Datalog programs to incorporate fact insatnces,
+as in \refdef{continuous datalog: fact instance}, we define a
+time-annotated Datalog program in \refdef{continuous datalog: ta datalog program}.
 
-When considering streams of facts, in which there may be many distinct
-instances of a given fact, temporal information would be lost if sets
-of fact instances (denoted $F_{inst}$) were to entail sets of plain
-facts.  As such, where a fact $f$ would be entailed by some set of
-facts $F$ that are instantiated on some stream, an instance of $f$
-should be entailed by the set of instances of those facts present on
-that stream.  Furthermore, each such justification in $F_{inst}$ for a
-fact entails a unique instance of that fact, such that each entailed
-instance has a single justification within the original set of facts,
-and many instances of the entailed fact may occur in the entailed set
-of facts $E_{inst}$.  This form of entailment shall be henceforth
-referred to as {\em instance entailment}, and shall be the norm when
-discussing the entailments of sets of fact instances.
+\begin{definition}[Time-Annotated Datalog Program]
+\labeldef{continuous datalog: ta datalog program}
+A time-annotated Datalog program $\Pi^T = \langle R, F^T\rangle$ is a tuple
+which comprises a finite set of rules $R$ and a finite set of fact instances
+$F^T$. Fact instances are extensional atomic formulae of first-order logic
+annotated with a timestamp, as in \refdef{continuous datalog: fact instance},
+and rules are of the same form as those of conventional Datalog programs.
 
-It should be noted that removing all temporal information from a set
-of fact instances $F_{inst}$ produces the set of facts $F$
-instantiated in $F_{inst}$:
+As discussed in \refdef{continuous datalog: datalog program,continuous datalog: fact instance},
+in the context of the Semantic Web we treat fact instances as time-annotated
+3-tuples ${\langle s, p, o \rangle | t}$.
+\end{definition}
 
-\[F = \{ strip(x) | x \in F_{inst} \}\]
+When considering a time-annotated Datalog program, in which there may be
+many distinct instances of a given fact, temporal information regarding when
+instances of that fact became true would be lost if such a program were to entail
+sets of plain facts. As such, where a fact $f$ would be entailed by some Datalog
+program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle \vdash f}$,
+for some window over some stream $S$, an instance ${f | t}$ of $f$ should be
+entailed by the time-annotated Datalog program
+${\Pi^T = \langle R, W^{S,x}_i \rangle \vdash f | t}$.
 
-In addition, applying the ${strip()}$ function to both the original
-fact instances $F_{inst}$ and the entailed fact instances $E_{inst}$
-maintains the relationship of entailment between the two sets:
+It follows that applying the ${strip()}$ function to all the fact instances in
+both the set of fact instances $F^T$ in some ${\Pi^T = \langle R, F^T \rangle \vDash E^T}$
+and the set of entailed fact instances $E^T$ maintains the relationship of entailment:
+\begin{multline*}
+	\left( \langle R, F^T \rangle \vDash E^T \right) \rightarrow \\
+		\left( \langle R, \{ strip(x) : x \in F^T \} \rangle \vDash \{ strip(x) : x \in E^T \} \right)
+\end{multline*}
 
-\[ \left( F_{inst} \vDash E_{inst} \right) \rightarrow \left( F \vDash E \right) \]
+It should be noted that \refdef{continuous datalog: justification} of entailment
+justification is equally applicable to time-annotated Datalog programs,
+simply replacing entailed and justifying facts with fact instances.
+Furthermore, each justification for $f$ in some $F^T$ entails a unique
+instance ${f | t}$, such that many instances of $f$ may occur in the set
+$E^T$ of fact instances entailed by
+${\Pi^T = \langle R, W^{S,x}_i \rangle \vDash E^T}$, each with a single
+justification.  As such, entailment justification for fact instances may be
+expressed as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$,
+which maps a fact instance ${f | t}$, entailed by some
+${\Pi^T = \langle R, F^T \rangle}$, to the single minimal subset of $F^T$ that
+justifies it.
+
+This allows us to formally specify that the time $t$ at which an entailed fact
+instance ${f | t}$ becomes true is the last time at which any of its
+justifying instances become true:
+\[ t = max(\{ t' : f' | t' \in just(\Pi^T, f | t) \}) \]
+As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
+program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.
 
@@
 
-\begin{axiom}
-\label{axiom:continuous datalog: entailment disjointness}
-
-If two sets of fact instances, $A_{inst}$ and $B_{inst}$, are
-disjoint, then a set $F_{inst1}$ of entailed fact instances (each of
-which is justified at least in part by $A_{inst}$) is necessarily
-disjoint from any set $F_{inst2}$ of entailed fact instances (each
-justified at least in part by $B_{inst}$), because entailed fact
-instances are distinct with regards to the fact instances that justify
-them.
-
-\begin{multline*}
-\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
-\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
-\end{multline*}
+\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
+If two sliding windows over the same stream $S$ cover non-overlapping time periods,
+then a set ${E^T1}$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
+is necessarily disjoint from any set ${E^T2}$ of fact instances entailed by
+${\langle R, W^{S,y}_{i-z} \rangle}$, where ${z \geq x}$.
+% This is because the timestamps of the instances ${E^T1}$ are a subset of those of
+% the instances $W^{S,x}_i$, the timestamps of the instances ${E^T2}$ are a subset
+% of those of the instances $W^{S,y}_{i-z}$, fact instances are distinct by timestamp and
+% ${\{ t : f | t \in W^{S,x}_i\} \cap \{ t : f | t \in W^{S,y}_{i-z}\} = \varnothing}$.
+\[ \{ e : \langle R , W^{S,x}_i \rangle \vdash e \} \cap \{ e : \langle R , W^{S,y}_{i-z} \rangle \vdash e \} = \varnothing \]
 \end{axiom}
@@
 
-\begin{definition}[Continuous Datalog Program]
-\label{def:continuous datalog: CDP}
-
-A Continuous Datalog Program (CDP) is a tuple $\langle R, F, S, r
-\rangle$ which comprises a finite set of rules $R$, a finite set of
+\begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
+A Continuous Datalog Program $\Pi^C$ is a tuple $\langle R, F, S, r
+\rangle$ which comprises a finite set of rules $R$, a finite set of persistent
 facts $F$, a stream of facts $S$ and a range $r \in \mathbb{N}$ for
-the window over $S$. This specifies that the streamed data from $S$ is
-to be considered valid from its arrival for the duration represented
+the sliding window over $S$. This specifies that the streamed instances from $S$ are
+to be considered valid from their timestamp for the duration represented
 by $r$.
@@
 \ref{axiom:continuous datalog: entailment precedes negation}. 
-The streams $P$ and $N$ are therefore the continuous
+These specify that the streams $P$ and $N$ are the continuous
 reasoning equivalents of the streams produced by the \emph{IStream}
 and \emph{DStream} operators of the \emph{CQL} continuous query
-language \citep{CQL} respectively.  The truth value of a CDP is
-instantaneous, with the the truth function of Continuous Datalog
-$T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in
-  \mathbb{N}}$ as arguments.
+language \citep{CQL}, respectively.
+\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle P, N \rangle \]
+The truth value of a $\Pi^C$ is instantaneous, with the the truth
+function of Continuous Datalog $T^{CD}$ taking both a program ${\Pi^C}$
+and an instant ${i \in \mathbb{N}}$ as arguments.
 \end{definition}
 
-\begin{definition}[State of a CDP]
-\label{def:continuous datalog: CDPt}
+\begin{definition}[State of a $\Pi^C$]\label{def:continuous datalog: CDPt}
+The state of a $\Pi^C$ at a given instant $i$ is the time-annotated
+Datalog program ${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle}$,
+where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent facts $F$
+annotated as becoming true at the first recognised instant.
 
-The state of a CDP at a given instant $i$ (denoted as ${CDP_{i}}$) is
-the union of the set of facts $F$ and the window $W^{S,r}_{i}$.  As both $F$
-and ${strip( W^{S,r}_{i} )}$ are sets of Datalog axioms, 
-${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated
-by the truth function of Datalog $T^{D}$, the result of which is the
-definition for the instantaneous truth of the ${CDP}$ at instant $i$
-by $T^{CD}$.  The instantaneous entailments of a $CDP_{i}$ are the
-union of the set $E$, the persistent entailments of $F$ (which are
-guaranteed to be entailed by every state, as Positive Datalog is
-monotonic), and the set $E_{i}$ of instances of the transient
-entailments that are justified at least in part by some subset of
-$W^{S,r}_{i}$.
+As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of Datalog
+axioms, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
+constitutes a valid Datalog program that may be evaluated by the truth
+function of Datalog $T^{D}$, the result of which is the definition for the
+instantaneous truth of $\Pi^C$ at instant $i$ by $T^{CD}$.
+\begin{multline*}
+T^{CD} \left( \langle R, F, S, r \rangle, i \right) = \\
+	T^{D} \left( \langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \} \right)
+\end{multline*}
 
+The instantaneous entailments of a $\Pi^C_i$ are the union of the set of
+persistent entailments of ${\langle R, F^0 \rangle \vDash E^0}$ (which are
+guaranteed to be entailed by every state, as positive Datalog is monotonic),
+and the set $E^T_{i}$ of transient entailments whose justifications are at
+least partially contained in $W^{S,r}_i$.
 \begin{equation*}
-T^{CD} \left( CDP, i \right) = T^{D} \left( CDP_{i} \right)
+\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle \vDash E^0 \cup E^T_{i}
 \end{equation*}
-\begin{equation*}
-CDP_{i} = F \cup W^{S,r}_{i} \vDash E \cup E_{i}
-\end{equation*}
+$E^T_i$ is defined more formally as
+${E^T_{i} = \{ x : \left( \Pi^C_i \vdash x \right) \wedge
+\left( just(\Pi^C_i,x) \cap W^{S,r}_i \neq \varnothing \right) \}}$.
+By this definition, $E^0$ is guaranteed to be disjoint from every $E^T_i$
+for every state $\Pi^C_i$ of every Continuous Datalog Program $\Pi^C$.
 \end{definition}
@@
 The following axioms define the semantics of the streams of
-entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the
+entailments $P$ and and negations $N$ of a $\Pi^C$ in terms of the
 instantaneous entailments of the sequence of states of that continuous
@@
 \label{axiom:continuous datalog: positive window increment}
-
 The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
-the stream $P$ entailed by a ${CDP}$ is defined as the set of facts
-that are entailed by the state of the ${CDP}$ at instant $i$ that were
-not entailed by the state of ${CDP}$ at the previous instant instance
-${i-1}$.
-
+the stream $P$ entailed by a $\Pi^C$ is defined as the set of fact instances
+that are entailed by the state of the $\Pi^C$ at instant $i$ that were
+not entailed by the state of $\Pi^C$ at the previous instant ${i-1}$.
 \begin{equation*}
-W^{P,1}_{i} = \left( E \cup E_{i} \right) \setminus \left( E \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
+W^{P,1}_{i} = \left( E^0 \cup E^T_{i} \right) \setminus
+\left( E^0 \cup E^T_{i-1} \right) = E^T_{i} \setminus E^T_{i-1}
 \end{equation*}
 \end{axiom}
-
 \begin{axiom}
@@
 The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
-the stream $N$ entailed by a ${CDP}$ is defined as the set of facts
-that were entailed by the state of the ${CDP}$ at the previous instant
-instance ${i-1}$ that are no longer entailed by the state of ${CDP}$
-at instant $i$.
+the stream $N$ entailed by a $\Pi^C$ is defined as the set of fact instances
+that were entailed by the state of the $\Pi^C$ at the previous instant
+${i-1}$ that are no longer entailed by the state of $\Pi^C$ at instant $i$.
 \begin{equation*}
-W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
+W^{N,1}_{i} = \left( E^0 \cup E^T_{i-1} \right) \setminus
+\left( E^0 \cup E^T_{i} \right) = E^T_{i-1} \setminus E^T_{i}
 \end{equation*}
@@
 \label{axiom:continuous datalog: entailment precedes negation}
-Facts may only be negated by some part of their justification leaving
+Fact instances may only be negated by some part of their justification leaving
 the sliding window $W^{S,r}_{i}$, this necessarily being after they
 were entailed.  As such, a window of any range ${x \in \mathbb{N}}$
-over the stream of entailments $P$ at any given instant ${i \in
-  \mathbb{N}}$ is disjoint from a window of any range ${x \in
-  \mathbb{N}}$ over the the stream of entailment negations $N$ at any
-instant after ${i - X}$
-
-CHECK
-
+over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$
+is disjoint from a window of any range ${y \in \mathbb{N}}$ over the
+stream of entailment negations $N$ at any instant before ${i - x}$.
 \begin{equation*}
-\forall X' \geq X \left( W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing \right)
+W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing
 \end{equation*}
-
-\noindent where $z \geq x$.
+where $z \geq x$.
 \end{axiom}
 
-From this, we can see that each instance of an entailment in some
-$E_{i}$ may take the same form as an instance of any other fact in a
-stream of facts, being $\langle s, p, o \rangle \! : \!t$, where $t$ is
+NEEDS WORK\\
+--------------------------------\\
+From this, we can see that each entailed instance in some
+$E^T_{i}$ may take the same form as an instance of any other fact in a
+stream of facts, being $\langle s, p, o \rangle | t$, where $t$ is
 the real time associated with the instant $i_{e}$ in which
@@
 the subset of fact instances first present in $W^{S,r}_{i_{e}}$ that
-justified the entailment instance $\langle s, p, o \rangle \! : \!t$).
+justified the entailment instance $\langle s, p, o \rangle | t$).
+\\--------------------------------
 
@@
 \ref{sec:semantics: proof for entailments}, that it is also possible
-to reconstruct exactly the instantaneous set $E_{i}$ of entailment
-instances of any ${CDP}$ at any given instant from the windows of
-sufficient range over its entailed streams $P$ and $N$ at that
-instant.
-
-CHECK
-
-\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
-
-The minimum range $x$ is such to achieve the \emph{complete} set of
-entailment instances $E_{i}$ of the ${CDP}_{i}$.  In addition, the
-minimum range $y$ is such to achieve the \emph{correct} set $E_{i}$.
-As such, a \emph{valid window} that contains only and exactly those
-instances of entailments in $E_{i}$ at every instant $i$ may be
-trivially maintained from the streams $P$ and $N$ continuously
-entailed by some ${CDP}$.
+to reconstruct exactly the instantaneous set $E^T_{i}$ of instances
+entailed at any given instance $i$ by any $\Pi^C = \{ R, F, S, r \}$
+from the sliding windows of sufficient range over its entailed streams
+$P$ and $N$ at $i$:
+\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
+where ${y \geq x \geq r}$. The minimum range $x$ is such to achieve
+the \emph{complete} set of entailment instances $E^T_{i}$ of the
+$\Pi^C_{i}$. In addition, the minimum range $y$ is such to achieve the
+\emph{correct} set $E^T_{i}$. As such, a \emph{valid window} that
+contains only and exactly those entailed instances $E^T_{i}$ at every
+instant $i$ may be trivially maintained from the streams $P$ and $N$
+continuously entailed by some $\Pi^C$.
 \end{nestedsection}

### 2015-07-07 content/contribution/continuous-datalog.tex
changed the text more fundamentally to specify the concepts of entailment justification and instance entailment justification, time-annotation inheritance for entailed instances and the summary of the nature of the streams P and N.  Also clarified the concepts of F0 and E0

@@
 sets of plain facts. As such, where a fact $f$ would be entailed by some Datalog
-program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle \vdash f}$,
+program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle}$,
 for some window over some stream $S$, an instance ${f | t}$ of $f$ should be
@@
 $E^T$ of fact instances entailed by
-${\Pi^T = \langle R, W^{S,x}_i \rangle \vDash E^T}$, each with a single
+${\Pi^T = \langle R, W^{S,x}_i \rangle}$, each with a single
 justification.  As such, entailment justification for fact instances may be
-expressed as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$,
-which maps a fact instance ${f | t}$, entailed by some
-${\Pi^T = \langle R, F^T \rangle}$, to the single minimal subset of $F^T$ that
-justifies it.
+defined formally by \refdef{continuous datalog: instance justification}.
+\begin{definition}[Instance Entailment Justification]
+\labeldef{continuous datalog: instance justification}
+	A justification for some entailment ${f|t}$ of some time-annotated Datalog program
+	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
+	such that ${\Pi^T' = \langle R , J^T \rangle \vdash f|t}$.
+	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
+	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
+	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
+\end{definition}
 
@@
 \begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
-A Continuous Datalog Program $\Pi^C$ is a tuple $\langle R, F, S, r
-\rangle$ which comprises a finite set of rules $R$, a finite set of persistent
-facts $F$, a stream of facts $S$ and a range $r \in \mathbb{N}$ for
+A Continuous Datalog Program $\Pi^C$ is a tuple ${\langle R, F, S, r
+\rangle}$ which comprises a finite set of rules $R$, a finite set of persistent
+facts $F$, a stream of facts $S$ and a range ${r \in \mathbb{N}}$ for
 the sliding window over $S$. This specifies that the streamed instances from $S$ are
@@
 Datalog program ${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle}$,
-where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent facts $F$
-annotated as becoming true at the first recognised instant.
+where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent fact instances produced by
+annotating each fact in $f$ in $F$ as becoming true at the first recognised instant.
 
-As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of Datalog
-axioms, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
+As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
+Datalog facts, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
 constitutes a valid Datalog program that may be evaluated by the truth
@@
 \left( just(\Pi^C_i,x) \cap W^{S,r}_i \neq \varnothing \right) \}}$.
-By this definition, $E^0$ is guaranteed to be disjoint from every $E^T_i$
+By this definition and Axiom~\ref{axiom:continuous datalog: entailment disjointness},
+$E^0$ is guaranteed to be disjoint from every $E^T_i$
 for every state $\Pi^C_i$ of every Continuous Datalog Program $\Pi^C$.
@@
 
-NEEDS WORK\\
---------------------------------\\
-From this, we can see that each entailed instance in some
-$E^T_{i}$ may take the same form as an instance of any other fact in a
-stream of facts, being $\langle s, p, o \rangle | t$, where $t$ is
-the real time associated with the instant $i_{e}$ in which
-$\langle s, p, o \rangle$ becomes newly justified by the ${CDP}$ (more
-specifically, by some subset of the instances in $W^{S,r}_{i_{e}}$).
-As such, the stream of entailment instances $P$ of a ${CDP}$ takes
-the same form as the \emph{streams of facts} already described, thus
+From this, we can see that each entailed fact instance ${f|t}$ in some
+$E^T_{i}$, of the form ${f|t = \langle s, p, o \rangle | t}$, has a timestamp $t$ that is
+the real-time associated with the state $\Pi^C_{i_{e}}$ in which its particular justification for
+${f|t}$ is first complete.
+As such, the stream of entailed fact instances $P$ of a $\Pi^C$ takes precisely
+the same form as the streams of facts already described, thus
 producing a concise stream of entailments that may be read by
 further stream processing.  In addition, the stream of negations $N$
-of a ${CDP}$ also takes this form, but differs from the streams
-discussed thus far in that it is not ordered by the entailment time
-$i_{e}$ of the entailment instances but by the time at which their
-justification \emph{ceases} to be valid (i.e. the first instant
-${i_{n} \geq i_{e}}$ in which $W^{S,r}_{i_{n}}$ no longer contains
-the subset of fact instances first present in $W^{S,r}_{i_{e}}$ that
-justified the entailment instance $\langle s, p, o \rangle | t$).
-\\--------------------------------
+of a $\Pi^C$ also takes this form, but differs from the streams
+discussed thus far in that it is not ordered by the entailment/assertion time
+of the entailed fact instances, but by the real-time associated with
+the first state $\Pi^C_{i_{n}}$ in which their
+justification \emph{ceases} to be complete,
+i.e. ${just(\Pi^C_{i_{e}}, f|t) \subseteq W^{S,r}_{i_n - 1}}$ but
+${just(\Pi^C_{i_{e}}, f|t) \not\subseteq W^{S,r}_{i_n}}$.
 
@@
 \ref{sec:semantics: proof for entailments}, that it is also possible
-to reconstruct exactly the instantaneous set $E^T_{i}$ of instances
-entailed at any given instance $i$ by any $\Pi^C = \{ R, F, S, r \}$
+to reconstruct exactly the instantaneous set $E^T_{i}$, of instances
+entailed at any given instant $i$ by any $\Pi^C = \{ R, F, S, r \}$,
 from the sliding windows of sufficient range over its entailed streams

### 2015-07-08 content/contribution/continuous-datalog.tex
Added a summary of the relationship between Continuous Datalog programs, their states and conventional Datalog programs, and how the same relationships apply between their entailments.

@@
 language \citep{CQL}, respectively.
-\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle P, N \rangle \]
+\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle E, P, N \rangle \]
 The truth value of a $\Pi^C$ is instantaneous, with the the truth
@@
 From this, we can see that each entailed fact instance ${f|t}$ in some
-$E^T_{i}$, of the form ${f|t = \langle s, p, o \rangle | t}$, has a timestamp $t$ that is
-the real-time associated with the state $\Pi^C_{i_{e}}$ in which its particular justification for
-${f|t}$ is first complete.
-As such, the stream of entailed fact instances $P$ of a $\Pi^C$ takes precisely
-the same form as the streams of facts already described, thus
-producing a concise stream of entailments that may be read by
-further stream processing.  In addition, the stream of negations $N$
-of a $\Pi^C$ also takes this form, but differs from the streams
-discussed thus far in that it is not ordered by the entailment/assertion time
-of the entailed fact instances, but by the real-time associated with
-the first state $\Pi^C_{i_{n}}$ in which their
-justification \emph{ceases} to be complete,
-i.e. ${just(\Pi^C_{i_{e}}, f|t) \subseteq W^{S,r}_{i_n - 1}}$ but
-${just(\Pi^C_{i_{e}}, f|t) \not\subseteq W^{S,r}_{i_n}}$.
+$E^T_{i}$, of the form ${f|t = \langle s, p, o \rangle | t}$, has a
+timestamp $t$ that is the real-time associated with the state
+$\Pi^C_{i_{e}}$ in which its particular justification for ${f|t}$ is
+first complete. As such, the stream of entailed fact instances $P$ of
+a $\Pi^C$ takes precisely the same form as the streams of facts already
+described, thus producing a concise stream of entailments that may be
+read by further stream processing.  In addition, the stream of negations
+$N$ of a $\Pi^C$ also takes this form, but differs from the streams
+discussed thus far in that it is not ordered by the entailment/assertion
+time of the entailed fact instances, but by the real-time associated
+with the first state $\Pi^C_{i_{n}}$ in which their justification
+\emph{ceases} to be complete, i.e. ${just(\Pi^C_{i_{e}}, f|t) \subseteq W^{S,r}_{i_n - 1}}$
+but ${just(\Pi^C_{i_{e}}, f|t) \not\subseteq W^{S,r}_{i_n}}$.
 
@@
 continuously entailed by some $\Pi^C$.
+
+To sum up, a time-annotated Datalog program
+${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle}$ may be derived from
+a continuous Datalog program ${\Pi^C = \langle R, F, S, r \rangle}$ by
+applying a sliding window $W^{S,r}_i$ of range $r$ over the stream $S$ at
+instant $i$ and instantiating the facts in $F$ as $F^0$. At the same time,
+the entailments of a ${\Pi^C_i \vDash E^0 \cup E^T_i}$ may be derived from
+the entailments of the ${\Pi^C \vDash \langle E, P, N \rangle}$ as the union
+of the entailments of ${\langle R, F^0 \rangle \vDash E^0}$ and the set
+difference of the windows $W^{P,r}_i$ and $W^{N,r}_i$ of range $r$ over
+streams $P$ and $N$ at instant $i$.
+Furthermore, a conventional Datalog program ${\Pi = \langle R, F \rangle}$
+may be derived from a time-annotated Datalog program $\Pi^C_i$ by
+${strip}$ping the time annotations from the set of facts
+${F^0 \cup W^{S,r}_i}$, as the entailments of ${\Pi \vDash E}$ may be
+derived from the entailments of $\Pi^C_i$ by $strip$ping them.
+This yields the TODO RMEMBER NAME OF GRAPH in \reffig{CD-TD-D}.
+\begin{figure}
+	\centering
+	% TODO make the diagram
+	\caption{TODO}
+	\labelfig{CD-TD-D}
+\end{figure}
 \end{nestedsection}

### 2015-07-08 content/contribution/CoDeR.tex
reworded data model of CoDeR to re-focus on continuous datalog and use the right notation.

@@
 	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
-	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of axiomatic rules $A$.
+	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^C = \langle R,F,S,r \rangle \vDash \langle E,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
 	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
@@
 		The data model of CoDeR is composed of two classes of streams: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
+		In both cases, streams are multi-source multi-sink logical entities.
 
-		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any ${CDP}$ expressed by a CoDeR-based system.
-		These streams are sequences of instances of RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
-		an instance of an RDF graph is an RDF graph annotated with the time interval for which it is valid \citep{SemanticStreamingManagement,sparkwave} according to the ${CDP}$ expressed: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
-		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the streams $S$ must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the valid time interval of the graph instance in which it originated.
-		This interval, for any entailed graph/fact instance, is derived from the valid time intervals of the fact instances that justify it, as shown in \reffig{intersected-intervals}, being the graphical representation of valid times for transitive data continuously entailed by the ${CDP = \{\{c\,\text{:-}\,a,b\,.\},(\{a\}\text{:}1,\{b\}\text{:}2),3\}}$ over instants ${i = 1 \dots 4}$:
-		\begin{figure}[t]
+		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^C$ expressed by a CoDeR-based system.
+		These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
+		an instance of an RDF graph is an RDF graph annotated with a time interval: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
+		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.
+
+		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^C = \langle R, F, S, r \rangle}$ expressed.
+		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^C$.
+		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
+		\reffig{intersected-intervals} provides an example of this for some ${\Pi^C = \langle \{c\,\text{:-}\,a,b\,.\},\{\},(\{a\}\text{:}1,\{b\}\text{:}2),3\}}$ over instants ${i = 1 \dots 4}$.
+		\begin{figure}
 			\centering
@@
 		\end{align*}
-		It is in the base case of this derivation that the semantic window range $R$ of a ${CDP}$ is expressed:
-		each graph instance arriving on an input stream has a valid interval from the time of its arrival at the system for a duration equal to the range $R$.
+		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^C$.
+		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^C$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
+		As such, the annotation of streamed RDF graphs with the interval $(i_e,i_n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
 
@@
 		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
-		for a given inter-operator stream, each such RDF graph will match the same graph pattern, that being a sub-pattern of the body of one of the axiomatic rules $A$.
-		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same semantics of validity may be applied to these intermediate matches as to the true entailments of the model found on output streams, and their valid-time intervals may derived from their justifying instances in the same manner.
+		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^C$.
+		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
 		It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
-		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are the entailments of either the system or its predecessor in the stream pipeline.
+		the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.
 
-		In addition, each stream in the model is semantically distinct, with each input/output stream being from a distinct source and uniquely identifiable by the source's IRI, and each inter-operator stream containing graphs matching a specific graph pattern and uniquely identifiable by that pattern.
+		In addition, each stream in the model is semantically distinct:
+		each input/output stream is from a distinct source and uniquely identifiable by the source's IRI;
+		each inter-operator stream contains graphs matching a specific graph pattern and uniquely identifiable by that pattern.
 		As the same graph pattern may be a sub-pattern of many rule bodies, and so appear in multiple places in a processing plan, one inter-operator stream may be the input to multiple operators.
 		Furthermore, two or more operators may contribute to the same stream within the model, where the pattern of the stream in question is the union of the patterns of the streams produced by each of the operators in isolation.
-		In this case, each operator contributes results to the common stream as they are produced, thereby interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment time $i_e$ of the constituent instances.
-		As this is the intended behaviour of a stream union operator, such an operator is not needed in CoDeR-based systems.
+		In this case, each operator contributes results to the common stream as they are produced, passively interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment time $i_e$ of the constituent instances.
+		As this is the active behaviour of a stream union operator, such an operator is not needed in CoDeR.
 		However, within the algebra of CoDeR, such a stream is still denoted ${op_{application} \cup op_{application}}$.
-
-		Given that the ordering of instances in the output stream of a CoDeR-based system by the value of $i_{e}$ yields the sequence of instances in the semantic stream $P$ entailed by a ${CDP}$, and that their ordering by the value of $i_{n}$ yields the sequence of instances in the semantic stream $N$, this gives the output streams of CoDeR a dual semantics with respect to Continuous Datalog.
-		Furthermore, this dual semantics applies to all the streams of CoDeR, each expressing both the stream of entailment instances \emph{and} instance negations produced by specific pipelines of CoDeR operators applied to some set of continuous RDF data sources.
 	\end{nestedsection}

### 2015-07-08 content/evaluation.tex
added the version of the first results graph with labelled axis, and added Rehab's evaluation of the memory-consumption results.

@@
 			\centering
-			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows.png}
+			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
 			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various sizes over the input stream.}
@@
 
-		Discussion about memory consumption...
+		We also examined the memory consumption of R4, Sparkwave, and Jtalis.
+		Setting the sliding window size to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
+		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave memory needs are significantly higher.
 		\begin{figure}

### 2015-07-09 content/contribution/continuous-datalog.tex
fixed the use of primes, and added the commutative diagram (as Nick has not yet pushed a copy)

@@
 ${\Pi = \langle R , F \rangle \vdash f}$ is a minimal set ${J \subseteq F}$
-such that ${\Pi' = \langle R , J \rangle \vdash f}$.
+such that ${\Pi^\prime = \langle R , J \rangle \vdash f}$.
 \end{definition}
@@
 	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
-	such that ${\Pi^T' = \langle R , J^T \rangle \vdash f|t}$.
+	such that ${\Pi^{T\prime} = \langle R , J^T \rangle \vdash f|t}$.
 	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
@@
 justifying instances become true:
-\[ t = max(\{ t' : f' | t' \in just(\Pi^T, f | t) \}) \]
+\[ t = max(\{ t^\prime : f^\prime | t^\prime \in just(\Pi^T, f | t) \}) \]
 As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
@@
 
-As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
-Datalog facts, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
+As both $F$ and ${W_i = \{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
+Datalog facts, the tuple ${\langle R, F \cup W_i}$
 constitutes a valid Datalog program that may be evaluated by the truth
@@
 derived from the entailments of $\Pi^C_i$ by $strip$ping them.
-This yields the TODO RMEMBER NAME OF GRAPH in \reffig{CD-TD-D}.
+This yields the commutative diagram in \reffig{CD-TD-D}.
 \begin{figure}
 	\centering
-	% TODO make the diagram
-	\caption{TODO}
+	$
+	\begin{CD}
+		{\Pi^C = \langle R, F, S, r \rangle} @>\vDash>> {\langle E, P, N \rangle} \\
+		@V{window}VV @VV{window}V \\
+		{\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle} @>\vDash>> {E^0 \cup E^T_i} \\
+		@V{strip}VV @VV{strip}V \\
+		{\Pi = \langle R, F \cup W_i \rangle} @>\vDash>> {E \cup E_i}
+	\end{CD}
+	$
+	\caption{Commutative diagram from Continuous Datalog programs to Datalog entailments.}
 	\labelfig{CD-TD-D}

### 2015-07-09 content/contribution/R4.tex
first pass of editing to bring the description of R4 more in line with those of continuous datalog and CoDeR

@@
 	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
-	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams.
+	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams, constructed according to the Rete pattern matching algorithm \citep{forgy79}.
 	The system architecture is shown in \reffig{R4-architecture}.
@@
 	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
-	We also added a simple extension to RIF-Core in line with the extension described in \refsec{semantics}, enabling users to specify window constraints as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
+	We also added a simple extension to RIF-Core in line with the extension described in \refsec{semantics}, enabling users to specify sliding window constraints as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
 	Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
@@
 
-	These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete networks.
+	These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete dataflow networks.
 	The optimizer employs simple known heuristics to generate a good plan.
-	These heuristics include sharing nodes between rules where possible, avoiding Cartesian products as much as possible by joining patterns that have common variables, pushing more selective patterns (the ones with less variables) earlier in the network to minimize intermediate results.
+	These heuristics include: sharing nodes between rules where the sub-graph patterns their results match are variants of each other; avoiding Cartesian products as much as possible by joining patterns that have common variables earlier in the beta network; and pushing more selective patterns (those with fewer variables) earlier in the alpha network to minimize intermediate results.
 
 	The optimizer instantiates a Rete network for background reasoning if required and another network for generic rules.
-	The first network feeds into the second one and is also re-entrant to enable iterative inference of results, e.g. the calculation of transitive closure.
+	The first network feeds into the second one and the two together are ``re-entrant'', meaning that the first network takes as input the stream of entailments produced, to enable iterative inference of results such as the calculation of transitive closure.
 
-	These networks read TBox input data, receive streaming data in the form of RDF streams and operate directly on them.
+	These networks receive streaming data in the form of sequences of RDF graphs, and operate directly on each of these graphs incrementally.
 	We chose this RDF-native approach as opposed to reusing existing technologies as in \citep{C-SPARQL,streaming-sparql} to allow full control over the low-level operators.
@@
 		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
-		Source nodes also time stamps the incoming tuples with system timestamp if they are not already stamped from their origins.
+		Source nodes also timestamp the incoming graphs with the system time at their arrival, if they were not already stamped at their origin.
 		They propagate the data to the successor nodes, which are the alpha nodes.
 
-		Alpha nodes are single-entry nodes that form a discrimination network.
-		In R4, alpha nodes can receive input from any number of nodes through its single entry point.
-		Each input triple is matched against some conditions, and is either dropped if there is no match, or propagated downstream to the successor nodes in the case of a match.
+		Alpha nodes are single-entry nodes that form a discrimination network, as shown in \reffig{example-rete-network}.
+		In R4, alpha nodes can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
+		Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to the successor node(s) in the case of a match.
 		As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
-		The first alpha node in the network propagates its output to a special node called the left input adaptor, while the other alpha nodes propagate their output to join nodes.
+
+		The first alpha node in the network propagates matched quads to a special node called the left input adaptor, while the other alpha nodes propagate their output directly to join nodes.
 		The left input adaptor node is responsible of preparing the incoming triples to enter the beta network from the left input of the first join node in a given join-tree.
-		It creates a new token i.e. a partial match, defined as a list of triples, for each incoming triple and adds this triple as the first item in the list.
-		Tokens are then sent to the first join node in the beta network.
+		It creates a new token that represents a partial match, defined as a list of triples, for each rule body for which the incoming quad is a sub-graph match, then adds this quad as the first item in each list.
+		Tokens are then sent to the first join nodes in the beta network.
 
-		Beta nodes are two-entry nodes that are responsible for joining the branches of the alpha network.
+		Beta nodes are two-input nodes that are responsible for joining the branches of the alpha network, as demonstrated in \reffig{example-rete-network}.
 		In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
 		Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing triples received from alpha nodes.
-		Join nodes are beta nodes that match tuples from both sides according to some conditions, e.g. a shared variable binding.
+		Join nodes are beta nodes that match graphs from both sides according to some conditions, e.g. a shared variable binding.
 		As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
@@
 		The new token is stamped with the earliest timestamp of its triples, and then propagated to the next join node or to the terminal node if it is the last join.
-		At this point, the join node performs a garbage collection over its left window, ensuring that items with timestamps that are smaller that current time minus the semantic window size are deleted.
+		At this point, the join node performs a garbage collection over its left window, ensuring that items with timestamps that are smaller that current time minus the semantic window range are deleted.
 		Conversely, when the join node is right activated by receiving a triple from an alpha node, it adds the new triple to the right window, matches it against the elements of the left window, and performs a garbage collection over the right window.
 
-		This method of annotating tokens with the earliest timestamp of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
-		However, the ordering of tokens in their streams and their propagation through the synchronous system expresses those annotation semantics implicitly.
-		This leaves the timestamp of a token, in combination with the size of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances in CoDeR.
+		This method of annotating tokens with the earliest timestamp of those of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
+		However, the ordering of tokens in their streams and the system time at which they are propagated through the system expresses those annotation semantics implicitly.
+		This leaves the timestamp of a token, in combination with the range $r$ of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances as in CoDeR.
 		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
 
- 		Finally, terminal nodes receive tokens from the last node in the network and are responsible of producing the new entailed triples, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
-
- 		Explain an example Rete network.
+ 		Finally, terminal nodes receive tokens from the last node in a join-tree and are responsible for producing entailed triples, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
+ 		The output of terminal nodes is both the output of the network itself, and re-entered to the first nodes of the alpha network to support iterative inference.
 

### 2015-07-10 content/contribution/continuous-datalog.tex
Incremental changes to datalog description

--- 2015-07-07 content/contribution/continuous-datalog.tex
@@
 The first requirement of stream reasoning is a well-defined semantics.
-This must take into account the continuous nature of a subset of the
+that takes into account both the continuous nature of a subset of the
 base axioms of a stream reasoning problem, as well as its
@@
 
-	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined proof-theoretic semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
-	This comprises the extension of Datalog programs to include sources of continuously generated/reported facts, and the application of sliding window semantics over those sources, thereby providing for the interpretation of Continuous Datalog programs as sequences of instantaneous states that are each valid positive Datalog programs.
-	As such, these states each inherit the semantics of positive Datalog.
-	Furthermore, it involves the continuous derivation of entailments from this sequence of instantaneous states as a set of streams of facts;
-	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
+Continuous Datalog is an extension of positive Datalog with
+stream-processing semantics, with a view to accommodating the aims of
+the latter within the well-defined model-theoretic semantics of the
+former, which are shared with RIF-Core \cite{w3crifbld} and therefore
+support reasoning at the expressivity of OWL 2 RL.  This comprises the
+extension of Datalog programs to include sources of continuously
+generated/reported facts, and the application of sliding window
+semantics over those sources, thereby providing for the interpretation
+of Continuous Datalog programs as sequences of instantaneous states
+that are each valid positive Datalog programs.  As such, these states
+each inherit the semantics of positive Datalog.  Furthermore, it
+involves the continuous derivation of entailments from this sequence
+of instantaneous states as a set of streams of facts; these streams
+share a data model with the streams of continuous queries, as defined
+by the stream processing community, thereby allowing the integration
+of Continuous Datalog programs into stream processing pipelines.
 
@@
 
-\begin{definition}[Datalog Program]
+\begin{definition}[Datalog]
 \labeldef{continuous datalog: datalog program}
+
 A conventional Datalog program $\Pi = \langle R, F\rangle$ is a tuple
-which comprises a finite set of rules $R$ and a finite set of facts
-$F$. Facts are extensional atomic formulae of first-order logic, and
-rules are of the form:
-\[\alpha_0 \text{ :- } \alpha_1 , \ldots , \alpha_n\]
+which comprises a finite set of rules $R$ and a finite set of ground
+facts $F$. Facts are extensional atomic formulae of first-order logic,
+and rules are of the form:
+
+\[\alpha_0 :- \alpha_1 , \ldots , \alpha_n\]
 where $\alpha_0$ is an intensional atomic formula and
@@
 sets of plain facts. As such, where a fact $f$ would be entailed by some Datalog
-program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle}$,
+program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle \vdash f}$,
 for some window over some stream $S$, an instance ${f | t}$ of $f$ should be
@@
 $E^T$ of fact instances entailed by
-${\Pi^T = \langle R, W^{S,x}_i \rangle}$, each with a single
+${\Pi^T = \langle R, W^{S,x}_i \rangle \vDash E^T}$, each with a single
 justification.  As such, entailment justification for fact instances may be
-defined formally by \refdef{continuous datalog: instance justification}.
-\begin{definition}[Instance Entailment Justification]
-\labeldef{continuous datalog: instance justification}
-	A justification for some entailment ${f|t}$ of some time-annotated Datalog program
-	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
-	such that ${\Pi^T' = \langle R , J^T \rangle \vdash f|t}$.
-	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
-	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
-	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
-\end{definition}
+expressed as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$,
+which maps a fact instance ${f | t}$, entailed by some
+${\Pi^T = \langle R, F^T \rangle}$, to the single minimal subset of $F^T$ that
+justifies it.
 
@@
 \begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
-A Continuous Datalog Program $\Pi^C$ is a tuple ${\langle R, F, S, r
-\rangle}$ which comprises a finite set of rules $R$, a finite set of persistent
-facts $F$, a stream of facts $S$ and a range ${r \in \mathbb{N}}$ for
+A Continuous Datalog Program $\Pi^C$ is a tuple $\langle R, F, S, r
+\rangle$ which comprises a finite set of rules $R$, a finite set of persistent
+facts $F$, a stream of facts $S$ and a range $r \in \mathbb{N}$ for
 the sliding window over $S$. This specifies that the streamed instances from $S$ are
@@
 Datalog program ${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle}$,
-where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent fact instances produced by
-annotating each fact in $f$ in $F$ as becoming true at the first recognised instant.
+where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent facts $F$
+annotated as becoming true at the first recognised instant.
 
-As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
-Datalog facts, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
+As both $F$ and ${\{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of Datalog
+axioms, the tuple ${\langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \}}$
 constitutes a valid Datalog program that may be evaluated by the truth
@@
 \left( just(\Pi^C_i,x) \cap W^{S,r}_i \neq \varnothing \right) \}}$.
-By this definition and Axiom~\ref{axiom:continuous datalog: entailment disjointness},
-$E^0$ is guaranteed to be disjoint from every $E^T_i$
+By this definition, $E^0$ is guaranteed to be disjoint from every $E^T_i$
 for every state $\Pi^C_i$ of every Continuous Datalog Program $\Pi^C$.
@@
 
-From this, we can see that each entailed fact instance ${f|t}$ in some
-$E^T_{i}$, of the form ${f|t = \langle s, p, o \rangle | t}$, has a timestamp $t$ that is
-the real-time associated with the state $\Pi^C_{i_{e}}$ in which its particular justification for
-${f|t}$ is first complete.
-As such, the stream of entailed fact instances $P$ of a $\Pi^C$ takes precisely
-the same form as the streams of facts already described, thus
+NEEDS WORK\\
+--------------------------------\\
+From this, we can see that each entailed instance in some
+$E^T_{i}$ may take the same form as an instance of any other fact in a
+stream of facts, being $\langle s, p, o \rangle | t$, where $t$ is
+the real time associated with the instant $i_{e}$ in which
+$\langle s, p, o \rangle$ becomes newly justified by the ${CDP}$ (more
+specifically, by some subset of the instances in $W^{S,r}_{i_{e}}$).
+As such, the stream of entailment instances $P$ of a ${CDP}$ takes
+the same form as the \emph{streams of facts} already described, thus
 producing a concise stream of entailments that may be read by
 further stream processing.  In addition, the stream of negations $N$
-of a $\Pi^C$ also takes this form, but differs from the streams
-discussed thus far in that it is not ordered by the entailment/assertion time
-of the entailed fact instances, but by the real-time associated with
-the first state $\Pi^C_{i_{n}}$ in which their
-justification \emph{ceases} to be complete,
-i.e. ${just(\Pi^C_{i_{e}}, f|t) \subseteq W^{S,r}_{i_n - 1}}$ but
-${just(\Pi^C_{i_{e}}, f|t) \not\subseteq W^{S,r}_{i_n}}$.
+of a ${CDP}$ also takes this form, but differs from the streams
+discussed thus far in that it is not ordered by the entailment time
+$i_{e}$ of the entailment instances but by the time at which their
+justification \emph{ceases} to be valid (i.e. the first instant
+${i_{n} \geq i_{e}}$ in which $W^{S,r}_{i_{n}}$ no longer contains
+the subset of fact instances first present in $W^{S,r}_{i_{e}}$ that
+justified the entailment instance $\langle s, p, o \rangle | t$).
+\\--------------------------------
 
@@
 \ref{sec:semantics: proof for entailments}, that it is also possible
-to reconstruct exactly the instantaneous set $E^T_{i}$, of instances
-entailed at any given instant $i$ by any $\Pi^C = \{ R, F, S, r \}$,
+to reconstruct exactly the instantaneous set $E^T_{i}$ of instances
+entailed at any given instance $i$ by any $\Pi^C = \{ R, F, S, r \}$
 from the sliding windows of sufficient range over its entailed streams

### 2015-07-10 content/contribution/CoDeR.tex

@@
 	\begin{nestedsection}{Data Model}{model: data}
-		The data model of CoDeR is composed of two classes of streams: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
-		In both cases, streams are multi-source multi-sink logical entities.
+		The data model of CoDeR is composed of multiple-source multiple-sink streams.
+		These streams are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
 
@@
 		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
-		\reffig{intersected-intervals} provides an example of this for some ${\Pi^C = \langle \{c\,\text{:-}\,a,b\,.\},\{\},(\{a\}\text{:}1,\{b\}\text{:}2),3\}}$ over instants ${i = 1 \dots 4}$.
+		\reffig{intersected-intervals} provides an example of this for some ${\Pi^C = \langle \{c\,\text{:-}\,a,b\,.\},\{\},(\{a\}|1,\{b\}|2),3\}}$ over instants ${i = 1 \dots 4}$.
 		\begin{figure}
@@
 		\begin{align*}
-			CDP_{1} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \} \\
-			CDP_{2} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \, b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \, b\text{:}2. \, c\text{:}2. \} \\
-			CDP_{3} & = \{c\,\text{:-}\,a,b.\} \cup \{ a\text{:}1. \, b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, a\text{:}1. \, b\text{:}2. \, c\text{:}2. \} \\
-			CDP_{4} & = \{c\,\text{:-}\,a,b.\} \cup \{ b\text{:}2. \} & \vDash & \{ c\,\text{:-}\,a,b. \, b\text{:}2. \}
+			\Pi^C_{1} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \} \rangle & \vDash & \{ \} \\
+			\Pi^C_{2} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
+			\Pi^C_{3} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
+			\Pi^C_{4} & = \langle \{c\,\text{:-}\,a,b.\}, \{ b|2. \} & \vDash & \{ \}
 		\end{align*}
@@
 		% Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
-		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by $\rstreamjoin$.
+		Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, which we represent algebraically by $\rstreamjoin$.
 		% In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
 		In addition to $\text{:-}_{head}$ and $\rstreamjoin$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
-		Finally, as basic patterns match individual triples, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate streams $S$ from the streams of graphs recommended by the RSP community\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}} to the streams of triples on which $\sigma_{triple}$ operates.
+		Finally, as basic patterns match individual triples and CoDeR's streams are composed of graphs, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate input streams to the streams of triples on which $\sigma_{triple}$ operates.
 		% In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
@@
 		As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
-		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${CDP = \{ \{b\,\text{:-}\,a.\}, S, R \}}$, where $a$ and $b$ are RDF triples.
+		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^C = \{ \{b\,\text{:-}\,a.\}, \{\}, S, r \}}$, where $a$ and $b$ are RDF triples.
 		\begin{equation}\labeleqn{basic-example-plan}
@@
 		\begin{description}
-			\item[$\sigma_{triple}\,S$] \hfill \\
+			\item[$\sigma_{triple}\,X$] \hfill \\
 				\begin{figure}[t]
@@
 				\end{figure}
-				The triple-select operator identifies instances of triples in a stream $S$ that match some \emph{basic pattern} \citep{w3csparql}.
+				The triple-select operator identifies instances of triples in a stream $X$ that match some \emph{basic pattern} \citep{w3csparql}.
 				This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to semantic data.
 				Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the basic pattern to be matched.
-				It takes an inter-operator stream $S$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				It takes an inter-operator stream $X$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
 				The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
-			\item[$\sigma_{external}\,S$] \hfill \\
+			\item[$\sigma_{external}\,X$] \hfill \\
 				\begin{figure}[b]
@@
 				This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
-				It filters elements of its input stream $S$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
+				It filters elements of its input stream $X$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
 				thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern}, as specified in SPARQL.
-				As such, it takes an inter-operator stream $S$ of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+				As such, it takes an inter-operator stream $X$ of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
 				The contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
-			\item[${{}_g{\pi_t}\,S}$] \hfill \\
+			\item[${{}_g{\pi_t}\,X}$] \hfill \\
 				\begin{figure}[t]
@@
 				The graph-to-triples operator ${{}_g{\pi_t}}$ is a unary operator that deconstructs graph instances into sequences of triple instances, the union of the triples instantiated therein being equal to the contents of the original graph.
-				It takes a heterogeneous (\emph{input/output}) stream $S$ of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of RDF triple instances.
+				It takes a heterogeneous (\emph{input/output}) stream $X$ of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of RDF triple instances.
 				Each triple instance produced is assigned the same valid-time interval as the graph instance in whose graph the instantiated triple was contained.
@@
 				\end{multline*}
-			\item[$\text{:-}_{head}\,S$] \hfill \\
+			\item[$\text{:-}_{head}\,X$] \hfill \\
 				\begin{figure}[b]
@@
 				\end{figure}
-				Instance implication is a one-pass operator that takes as input an inter-operator stream $S$ whose graph pattern is the body of one of the axiomatic rules $A$ of the expressed ${CDP}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
+				Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^C$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
 				Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
@@
 			% \item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
-			\item[${S_1\,\rstreamjoin\,S_2}$] \hfill \\
+			\item[${X\,\rstreamjoin\,Y}$] \hfill \\
 				\begin{figure}[t]
@@
 				The implementations of SteMs in the literature show a great variety of behaviour regarding window maintenance \citep{SteMs}, in the most extreme case ceasing to be a wholly stream-to-stream operator and supporting the incorporation of ``static'' data into stream processing systems.
-				In the minimal algebra of CoDeR, however, SteMs simply support the building of a \emph{valid window} (as defined in \refsec{semantics}) from its first inter-operator stream $S_1$, and the probing of that window with data from its second inter-operator stream $S_2$ to produce an inter-operator stream of the homogeneously joined graph instances.
-				Furthermore, the nature of probing in CoDeR is strictly limited to that of the natural join between the variables in the patterns for streams $S_1$ and $S_2$.
+				In the minimal algebra of CoDeR, however, SteMs simply support the building of a \emph{valid window} (as defined in \refsec{semantics}) from its first inter-operator stream $X$, and the probing of that window with data from its second inter-operator stream $Y$ to produce an inter-operator stream of the homogeneously joined graph instances.
+				Furthermore, the nature of probing in CoDeR is strictly limited to that of the natural join between the variables in the patterns for streams $X$ and $Y$.
 				Finally, in contrast to the direct inheritance of valid-times in unary operators of CoDeR, the valid-time interval assigned to each graph produced by a SteM is the intersection of the valid-times of the two joined graphs.
@@
 		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
-		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $R$ over the system input streams $S$ defined in a given ${CDP = \{A,S,R\}}$.
+		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^C = \{R,F,S,r\}}$.
 		It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
-		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $A$ of a ${CDP}$, all facts that are logically entailed by that ${CDP}$ at a given instant will be persisted within that system at that instant,
+		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^C$ at a given instant will be persisted within that system at that instant,
 		i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
 	\end{nestedsection}
-
-%	\subimport{conclusions/}{future-work.tex}
 \end{nestedsection}

### 2015-07-10 content/contribution/continuous-datalog.tex
Revised definitions for time

@@
 
-A conventional Datalog program $\Pi = \langle R, F\rangle$ is a tuple
-which comprises a finite set of rules $R$ and a finite set of ground
-facts $F$. Facts are extensional atomic formulae of first-order logic,
-and rules are of the form:
+A conventional Datalog program $\Pi = \langle R, F\rangle$ comprises a
+set of clauses, divided into a finite set of rules $R$ and a finite
+set of ground facts $F$. Facts are formulae of first-order logic, and
+rules are of the form:
 
-\[\alpha_0 :- \alpha_1 , \ldots , \alpha_n\]
-where $\alpha_0$ is an intensional atomic formula and
-$\alpha_1 \ldots \alpha_n$ are a mixture of intensional and
-extensional atomic formulae.
+\[\alpha_0 :\!-\ \alpha_1 , \ldots , \alpha_n\]
+where $\alpha_i$ are atomic formulae.
 
-For the purpose of defining Continuous Datalog in the context of the
-Semantic Web, we treat facts as 3-tuples $\langle s, p, o \rangle$,
-where $s, p, o$ are respectively the subject, predicate and object of
-an RDF triple. 
+The Herbrand base $\hat{H}$ of $\Pi$ is the set of all facts that can
+be expressed using the predicates in $\Pi$; for a program $\Pi =
+\langle R, F \rangle$, $F \subseteq \hat{H}$. Note that we do not
+distinguish between the predicates in $F$ and those in the heads of
+the rules in $R$ (as extensional and intensional predicates), as is
+the case in the model theory for Datalog presented in
+\cite{datalog-basics}; in this respect, our treatment follows that of
+the model theory for Prolog presented in \cite{prolog-semantics}. A
+Herbrand interpretation $\mathcal{I}$ of $\Pi$ is any subset of
+$\hat{H}$, and a Herbrand interpretation $\mathcal{I}$ that satisfies
+the set of clauses in program $\Pi$ is a Herbrand model of $\Pi$,
+written $\models_{\mathcal{I}} \Pi$.
+
+We define the concept of entailment in Datalog as follows. A fact $f$
+is entailed by a program $\Pi$ iff each interpretation satisfying
+$\Pi$ (satisfying every clause in $\Pi$) also satisfies $f$, written
+$\Pi \models f$. The entailment of $\Pi$ is therefore the set of all
+facts that logically follow from $\Pi$:
+
+\[cons(\Pi) = \{ f \in \hat{H} | \Pi \models f \}\]
 \end{definition}
 
-\begin{definition}[Entailment Justification]
+\begin{definition}[Justification]
 \labeldef{continuous datalog: justification}
-A justification for some entailment $f$ of some Datalog program
-${\Pi = \langle R , F \rangle \vdash f}$ is a minimal set ${J \subseteq F}$
-such that ${\Pi^\prime = \langle R , J \rangle \vdash f}$.
+A {\em justification} for the entailment $E$ of some Datalog program
+${\langle R , F \rangle \models E}$ is a minimal set
+${J \subseteq F}$ such that ${\langle R , J \rangle \models E}$.
 \end{definition}
@@
 \labeldef{continuous datalog: instant}
-An instant is a discrete point in time, such that an instant 
-${i \in \mathbb{N}}$.
-
-It should be noted that time is a continuous concept, with infinitely
-many divisions possible between any two discrete instants.  However,
-the resolution with which such divisions can be recognised will, in
-practise, have a lower bound set by hardware limitations, or may be
-defined as part of some specific implementation.  In these cases,
-anything that occurs between discrete instants is considered to have
-occurred at the next recognised instant.
+We assume a discrete, ordered time domain $\mathbb{T}$; an instant $i
+\in \mathbb{T}$ is a point in time. It should be noted that time is a
+continuous concept, with infinitely many divisions possible between
+any two discrete instants.  However, the resolution with which such
+divisions can be recognised will, in practise, have a lower bound set
+by hardware limitations, or may be defined as part of some specific
+implementation.  In these cases, anything that occurs between discrete
+instants is considered to have occurred at the next recognised
+instant.
 \end{definition}
@@
 \labeldef{continuous datalog: fact instance}
-A {\em fact instance} is a fact with some additional temporal annotation;
-it is the combination of a fact's signature (its predicate $p$, subject
-$s$ and object $o$) and the timestamp $t$ at which the instance became true.
-A fact instance is denoted ${\langle s, p, o \rangle | t \in \mathbb{F^T}}$.
+A {\em fact instance} is a fact with some additional temporal
+annotation; it is a tuple containing a fact $f$ and the timestamp
+$t \in \mathbb{T}$ at which $f$ became true, written
+$\langle f, t \rangle$. Two fact instances containing the same fact
+and timestamp are considered identical.
 
@@
 annotation) and a fact instance (with temporal annotation) with the
-function ${strip : \mathbb{F^T} \rightarrow \mathbb{F}}$ which maps a fact instance to its
-underlying fact ${\langle s, p, o \rangle \in \mathbb{F}}$.
+function ${strip : \hat{H} \times \mathbb{T} \rightarrow \hat{H}}$
+which maps a fact instance to its underlying fact.
 \end{definition}
@@
 \labeldef{continuous datalog: window}
+
 A sliding window of range ${x \in \mathbb{N}}$ over any stream $S$ at
 any given instant ${i \in \mathbb{N}}$ is the \emph{set} of fact
-instances that occur in $S$ during the interval ${(i-x,i]}$, and is
-denoted ${W^{S,x}_{i}}$.
+instances that occur in $S$ during the interval ${(i-x,i]}$, written
+${W^{S,x}_{i}}$.
+
+\[ W^{s,x}_i = \{ \langle f, t \rangle | \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]
+
 \end{definition}
@@
 another set of facts $E$ in which the existence of one fact in $E$ may
-be justified by many distinct subsets of $F$. In order to extend this
+be justified by many distinct subsets of $F$.
+
+In order to extend this
 definition of Datalog programs to incorporate fact insatnces,

### 2015-07-10 content/evaluation.tex
minor edits and corrected "window size" to "window range" throughout

@@
 
-		The experiment consisted of pushing the whole stream to the compiled R4 system and measuring the time taken to process all the data, repeated for a variety the window sizes.
+		The experiment consisted of pushing the whole stream to the compiled R4 system and measuring the time taken to process all the data, repeated for a variety the window ranges.
 		We ran this rule successfully on R4 and Jtalis getting correct results.
@@
 	\begin{nestedsection}{Comparative Results}{evaluation: comparative results}
-		In order to compare R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+		In order to compare the efficiency of R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
 		We generated a 1.1 million triples dataset containing information about 100,000 limited-time offers made available by some online marketplace.
@@
 
-		The rule expressed in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer, which means that background reasoning is needed for non-leaf products.
+		The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
+		As offers may be associated with product super-types, background reasoning is needed to determine the offer price for specific product sub-types.
 		\begin{figure*}
@@
 		\end{figure*}
-		Before we move on to the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window).
+
+		Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
 		However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
-		This is likely to be because Etalis is optimized for the \emph{seq} operator, being intended for \emph{event processing} rather than continuous querying/reasoning, which makes relevant the order of arrival of triples rather than simply their coincidence in the system.
-		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window sizes, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
-		We recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, but is sufficient to contrast the effect of window size on the time efficiency of the three systems.
+		Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
+		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
+		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time- and memory-efficiency of the three systems.
 		\begin{figure}
@@
 			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
-			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various sizes over the input stream.}
+			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
 			\labelfig{all-systems-varying-windows}
 		\end{figure}
-		The X axis shows the time windows in milliseconds while the Y axis shows the processing time in milliseconds.
 
-		As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small sliding windows but slower than Sparkwave, and, while the time-to-complete increases with window size in Sparkwave and R4, that of Etalis appears to remain fairly constant for the sliding-window sizes tested.
-		The growing difference between R4 and Sparkwave as the size of the sliding windows increases is likely due to the fact that Sparkwave uses a hash join algorithm while the current implementation of R4 uses loop joins;
-		as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer time to iterate over the windows in the matching process compared to the hash-based look-up of those in Sparkwave.
-		More efficient, multi-index window-joins are to be considered for R4 in future work.
+		As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small sliding windows but slower than Sparkwave, and, while the time-to-complete increases with window range in both Sparkwave and R4, that of Etalis appears to remain fairly constant for the sliding-window ranges tested.
+		The growing difference between R4 and Sparkwave as the range of the sliding windows increases is likely due to the fact that Sparkwave uses a hash join algorithm with garbage-collection window-maintenance while the current implementation of R4 uses loop joins over lists maintained by indexing over timestamps;
+		as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer to iteratively check for matches compared to the hash-based look-up of those in Sparkwave.
+		More time-efficient, multi-index window-joins are to be considered for R4 in future work.
 
-		We also examined the memory consumption of R4, Sparkwave, and Jtalis.
-		Setting the sliding window size to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
+		We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
+		Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
 		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave memory needs are significantly higher.
+		Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to the manual garbage collection performed.
 		\begin{figure}

### 2015-07-10 content/abstract.tex
Added abstract

@@
-Continuous Datalog semantics
-
-CoDeR Logical Model (minimal algebra and stream data model) for Continuous Deductive Reasoning
-
-Physical Model and Windowed Rete Algorithm
-
-R4 (\& Squall)
-
-Evaluation (against Sparkwave \cite{sparkwave} and CQELS)
+In this paper, we introduce Continuous Datalog, an extension of
+conventional Datalog that adds the concept of streams of
+temporally-annotated facts to conventional Datalog and defines the
+notion of entailment over a windowed stream. We go on to define CoDeR,
+an abstract model for continuous deductive reasoning that provides
+both a data stream model based on Continuous Datalog and a minimal
+algebra for expressing queries and programs.  Finally, we describe the
+R4 stream reasoner, which implements the CoDeR model as a Rete-based
+dataflow network and provide a brief comparative evaluation of R4
+against ETALIS and Sparkwave.

### 2015-07-12 content/contribution/continuous-datalog.tex
Fixed definitions for Datalog, time. Reordered to move window definitions after TA-Datalog

@@
 
-%% fixme
-%	In regards to Continuous Datalog, an ``instant'', a ``stream of facts'' and a ``sliding window'' over a stream are defined as follows:
-
 \begin{definition}[Datalog]
@@
 \[cons(\Pi) = \{ f \in \hat{H} | \Pi \models f \}\]
+
+%For convenience, we will write $\Pi \models E$ to indicate that $E =
+%cons(\Pi)$ is the set of all facts entailed by $\Pi$.
+
 \end{definition}
@@
 \labeldef{continuous datalog: justification}
-A {\em justification} for the entailment $E$ of some Datalog program
-${\langle R , F \rangle \models E}$ is a minimal set
-${J \subseteq F}$ such that ${\langle R , J \rangle \models E}$.
+
+We define a {\em justification} of an entailment of a Datalog program
+as a minimal subset of the program that is sufficient for the
+entailment to hold. More precisely, for a Datalog program
+$\Pi = \langle R, F \rangle$ and an entailment $E$ where
+$\langle R, F \rangle \models E$, the program $\langle J_R, J_F \rangle$
+with $J_R \subseteq R$ and $J_F \subseteq F$ is a justification for $E$
+iff $\langle J_R, J_F \rangle \models E$ and, for all $J'_R \subset J_R$ or $J'_F \subset J_F$, $\langle J'_R, J'_F \rangle \not\models E$.
+
+We can simplify this notion of justification by making the assumption
+that the justification can be defined in terms of a minimal subset of
+the ground facts of the program; the rules of the program are
+therefore implicitly considered to be part of every justification. We
+refer to this simplified justification as a {\em fact justification},
+and define it as follows: given a Datalog program $\Pi = \langle R, F
+\rangle$ and an entailment $E$ where $\langle R, F \rangle \models E$,
+the set of facts $J \subseteq F$ is a fact justification for $E$ iff
+$\langle R, J \rangle \models E$ and, for all $J' \subset J$, $\langle R, J' \rangle
+\not\models E$.
 \end{definition}
 
-\begin{definition}[Instant]
+\begin{definition}[Time Domain]
 \labeldef{continuous datalog: instant}
-We assume a discrete, ordered time domain $\mathbb{T}$; an instant $i
-\in \mathbb{T}$ is a point in time. It should be noted that time is a
-continuous concept, with infinitely many divisions possible between
-any two discrete instants.  However, the resolution with which such
-divisions can be recognised will, in practise, have a lower bound set
-by hardware limitations, or may be defined as part of some specific
-implementation.  In these cases, anything that occurs between discrete
-instants is considered to have occurred at the next recognised
-instant.
+
+Let $\mathbb{T}$ be a discrete time domain with a total order under
+$\leqslant$; an element of $\mathbb{T}$ is an {\em instant}.
+$\mathbb{T}$ has a lower bound $0$, such that for all $i \in
+\mathbb{T}$, $0 \leqslant i$. For convenience, we write $i+n$ where $i
+\in \mathbb{T}$ to indicate the $n$th successor of instant $i$,
+similarly $i-n$.
+
+%It should be noted that time is a continuous concept, with infinitely
+%many divisions possible between any two discrete instants.  However,
+%the resolution with which such divisions can be recognised will, in
+%practise, have a lower bound set by hardware limitations, or may be
+%defined as part of some specific implementation.  In these cases,
+%anything that occurs between discrete instants is considered to have
+%occurred at the next recognised instant.
 \end{definition}
@@
  \labeldef{continuous datalog: fact instance}
+
 A {\em fact instance} is a fact with some additional temporal
-annotation; it is a tuple containing a fact $f$ and the timestamp
-$t \in \mathbb{T}$ at which $f$ became true, written
-$\langle f, t \rangle$. Two fact instances containing the same fact
-and timestamp are considered identical.
+annotation; it is a tuple containing a fact $f$ and an instant $t \in
+\mathbb{T}$, written $\langle f, t \rangle$. Two fact instances
+containing the same fact and timestamp are considered identical; two
+fact instances containing the same fact but different timestamps are
+considered to be distinct.
 
@@
 function ${strip : \hat{H} \times \mathbb{T} \rightarrow \hat{H}}$
-which maps a fact instance to its underlying fact.
+which maps a fact instance to its underlying fact:
+
+\[ strip(\langle f, t\rangle) = f \]
+
+For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
+of fact instances and $F = \{ strip(f) | f \in F^T \}$.
+
 \end{definition}
+
+\begin{definition}[Time-Annotated Datalog]
+\labeldef{continuous datalog: ta datalog program}
+
+A time-annotated Datalog program $\Pi^T = \langle R, F^T\rangle$ is a
+tuple which comprises a finite set of rules $R$ and a finite set of
+fact instances $F^T \subseteq \hat{H} \times \mathbb{T}$. Fact
+instances are as defined in \refdef{continuous datalog: fact
+  instance}, and rules are of the same form as in conventional
+Datalog programs.
+
+When considering a time-annotated Datalog program, in which there may
+be many distinct instances of a given fact, temporal information
+regarding when instances of that fact became true would be lost if
+such a program were to entail sets of plain facts without temporal
+annotations. We therefore extend the model theory for Datalog from
+\refdef{continuous datalog: datalog program} by considering the
+Herbrand base of all fact instances $\hat{H}^T$ (all facts at all
+instants). As before, a fact $f$ is entailed by a program $\Pi^T$ iff
+each interpretation satisfying every clause in $\Pi^T$ also satisfies
+$f$. The entailment of $\Pi^T$ is therefore:
+
+\[ cons^T(\Pi^T) = \{ f \in \hat{H}^T | \Pi^T \models f \} \]
+
+It follows that applying the ${strip()}$ function to all the fact
+instances in both the set of fact instances $F^T$ in some ${\Pi^T =
+  \langle R, F^T \rangle}$ and the set of entailed fact instances
+$cons^T(\Pi^T)$ maintains the relationship of entailment; if a fact
+instance $f$ is entailed by $\langle R, F^T\rangle$, then $strip(f)$
+will be entailed by $\langle R, strip(F^T) \rangle$.
+\end{definition}
+
+\begin{definition}[Time-Annotated Justification]
+\labeldef{continuous datalog: instance justification}
+
+The justification of entailments (as in \refdef{continuous datalog:
+  justification}) is equally applicable to time-annotated Datalog
+programs, with facts in the entailment and the justification for that
+entailment being replaced with fact instances.
+
+Furthermore, each justification for $f$ in some $F^T$
+entails a unique instance ${f | t}$, such that many instances of $f$
+may occur in the set $E^T$ of fact instances entailed by ${\Pi^T =
+  \langle R, W^{S,x}_i \rangle}$, each with a single justification.
+
+As such, entailment justification for fact instances may be defined
+formally by \refdef{continuous datalog: instance justification}.
+
+A justification for some entailment ${f|t}$ of some time-annotated Datalog program
+	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
+	such that ${\Pi^{T\prime} = \langle R , J^T \rangle \vdash f|t}$.
+	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
+	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
+	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
+\end{definition}
+
+This allows us to formally specify that the time $t$ at which an entailed fact
+instance ${f | t}$ becomes true is the last time at which any of its
+justifying instances become true:
+\[ t = max(\{ t^\prime : f^\prime | t^\prime \in just(\Pi^T, f | t) \}) \]
+As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
+program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.
 
@@
 
-In \refdef{continuous datalog: datalog program}, we took a
-Datalog program to consist of a set of rules and a set of facts.
-Taken together with the set of rules $R$, the set of facts $F$ entails
-another set of facts $E$ in which the existence of one fact in $E$ may
-be justified by many distinct subsets of $F$.
-
-In order to extend this
-definition of Datalog programs to incorporate fact insatnces,
-as in \refdef{continuous datalog: fact instance}, we define a
-time-annotated Datalog program in \refdef{continuous datalog: ta datalog program}.
-
-\begin{definition}[Time-Annotated Datalog Program]
-\labeldef{continuous datalog: ta datalog program}
-A time-annotated Datalog program $\Pi^T = \langle R, F^T\rangle$ is a tuple
-which comprises a finite set of rules $R$ and a finite set of fact instances
-$F^T$. Fact instances are extensional atomic formulae of first-order logic
-annotated with a timestamp, as in \refdef{continuous datalog: fact instance},
-and rules are of the same form as those of conventional Datalog programs.
-
-As discussed in \refdef{continuous datalog: datalog program,continuous datalog: fact instance},
-in the context of the Semantic Web we treat fact instances as time-annotated
-3-tuples ${\langle s, p, o \rangle | t}$.
-\end{definition}
-
-When considering a time-annotated Datalog program, in which there may be
-many distinct instances of a given fact, temporal information regarding when
-instances of that fact became true would be lost if such a program were to entail
-sets of plain facts. As such, where a fact $f$ would be entailed by some Datalog
-program ${\Pi = \langle R, \{ strip(x) : x \in W^{S,x}_i \} \rangle}$,
-for some window over some stream $S$, an instance ${f | t}$ of $f$ should be
-entailed by the time-annotated Datalog program
-${\Pi^T = \langle R, W^{S,x}_i \rangle \vdash f | t}$.
-
-It follows that applying the ${strip()}$ function to all the fact instances in
-both the set of fact instances $F^T$ in some ${\Pi^T = \langle R, F^T \rangle \vDash E^T}$
-and the set of entailed fact instances $E^T$ maintains the relationship of entailment:
-\begin{multline*}
-	\left( \langle R, F^T \rangle \vDash E^T \right) \rightarrow \\
-		\left( \langle R, \{ strip(x) : x \in F^T \} \rangle \vDash \{ strip(x) : x \in E^T \} \right)
-\end{multline*}
-
-It should be noted that \refdef{continuous datalog: justification} of entailment
-justification is equally applicable to time-annotated Datalog programs,
-simply replacing entailed and justifying facts with fact instances.
-Furthermore, each justification for $f$ in some $F^T$ entails a unique
-instance ${f | t}$, such that many instances of $f$ may occur in the set
-$E^T$ of fact instances entailed by
-${\Pi^T = \langle R, W^{S,x}_i \rangle}$, each with a single
-justification.  As such, entailment justification for fact instances may be
-defined formally by \refdef{continuous datalog: instance justification}.
-\begin{definition}[Instance Entailment Justification]
-\labeldef{continuous datalog: instance justification}
-	A justification for some entailment ${f|t}$ of some time-annotated Datalog program
-	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
-	such that ${\Pi^{T\prime} = \langle R , J^T \rangle \vdash f|t}$.
-	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
-	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
-	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
-\end{definition}
-
-This allows us to formally specify that the time $t$ at which an entailed fact
-instance ${f | t}$ becomes true is the last time at which any of its
-justifying instances become true:
-\[ t = max(\{ t^\prime : f^\prime | t^\prime \in just(\Pi^T, f | t) \}) \]
-As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
-program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.
+In \refdef{continuous datalog: datalog program}, we took a Datalog
+program to consist of a set of rules and a set of facts.  In order to
+extend this definition of Datalog programs to incorporate fact
+instances, as in \refdef{continuous datalog: fact instance}, we define
+a time-annotated Datalog program as follows:
 

### 2015-07-12 content/contribution/continuous-datalog.tex
Fixed up to the end of defn of TA-Datalog

@@
 
-\[\alpha_0 :\!-\ \alpha_1 , \ldots , \alpha_n\]
-where $\alpha_i$ are atomic formulae.
+\[ \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n \]
+
+\noindent where $\alpha_i$ are atomic formulae. Note that we have excluded any
+universally quantified variables for clarity.
 
@@
 
-We define the concept of entailment in Datalog as follows. A fact $f$
+We define the concept of entailment in Datalog as follows. A fact $e$
 is entailed by a program $\Pi$ iff each interpretation satisfying
-$\Pi$ (satisfying every clause in $\Pi$) also satisfies $f$, written
-$\Pi \models f$. The entailment of $\Pi$ is therefore the set of all
-facts that logically follow from $\Pi$:
-
-\[cons(\Pi) = \{ f \in \hat{H} | \Pi \models f \}\]
-
-%For convenience, we will write $\Pi \models E$ to indicate that $E =
-%cons(\Pi)$ is the set of all facts entailed by $\Pi$.
-
+$\Pi$ (satisfying every clause in $\Pi$) also satisfies $e$, written
+$\Pi \models e$. For a a program $\Pi$ and a set of entailed facts
+$E$, $\Pi \models E$ iff $\Pi \models e$ for every $e \in E$. For
+convenience, we refer to the set of all facts that are entailed by a
+program (the intersection of all the Herbrand models of the program)
+as the {\em entailment of a program}, written $cons(\Pi)$ and defined
+as:
+\begin{align*}
+  cons(\Pi) &= \{ e \in \hat{H} \mid \Pi \models e \} \\
+  &= \bigcap_\mathcal{I} \models_\mathcal{I} \Pi
+\end{align*}
 \end{definition}
@@
 
-We define a {\em justification} of an entailment of a Datalog program
-as a minimal subset of the program that is sufficient for the
-entailment to hold. More precisely, for a Datalog program
-$\Pi = \langle R, F \rangle$ and an entailment $E$ where
-$\langle R, F \rangle \models E$, the program $\langle J_R, J_F \rangle$
-with $J_R \subseteq R$ and $J_F \subseteq F$ is a justification for $E$
-iff $\langle J_R, J_F \rangle \models E$ and, for all $J'_R \subset J_R$ or $J'_F \subset J_F$, $\langle J'_R, J'_F \rangle \not\models E$.
+We define a {\em justification} of a set of facts that are entailed by
+a Datalog program as a minimal subset of the program that is
+sufficient for the entailment to hold. More precisely, for a Datalog
+program $\Pi = \langle R, F \rangle$ and a set of entailed facts $E$
+where $\langle R, F \rangle \models E$, the program $\langle J_R, J_F
+\rangle$ with $J_R \subseteq R$ and $J_F \subseteq F$ is a
+justification for $E$ iff $\langle J_R, J_F \rangle \models E$ and,
+for all $J'_R \subset J_R$ or $J'_F \subset J_F$, $\langle J'_R, J'_F
+\rangle \not\models E$.
 
-We can simplify this notion of justification by making the assumption
-that the justification can be defined in terms of a minimal subset of
-the ground facts of the program; the rules of the program are
-therefore implicitly considered to be part of every justification. We
-refer to this simplified justification as a {\em fact justification},
-and define it as follows: given a Datalog program $\Pi = \langle R, F
-\rangle$ and an entailment $E$ where $\langle R, F \rangle \models E$,
-the set of facts $J \subseteq F$ is a fact justification for $E$ iff
-$\langle R, J \rangle \models E$ and, for all $J' \subset J$, $\langle R, J' \rangle
-\not\models E$.
+We simplify this notion of justification by making the assumption that
+a justification can be defined in terms of a minimal subset of the
+ground facts of the program; the rules of the program are therefore
+implicitly considered to be part of every justification. Any
+subsequent references to {\em justification} are to this simplified
+sense, which is defined as follows: given a Datalog program $\Pi =
+\langle R, F \rangle$ and a set of entailed facts $E$ where $\langle
+R, F \rangle \models E$, the set of facts $J \subseteq F$ is a
+justification for $E$ iff $\langle R, J \rangle \models E$ and, for
+all $J' \subset J$, $\langle R, J' \rangle \not\models E$. Note that a
+given set of entailed facts may not have a unique justification.
+
 \end{definition}
@@
 \in \mathbb{T}$ to indicate the $n$th successor of instant $i$,
-similarly $i-n$.
+and $i-n$ to indicate the $n$th predecessor of instant $i$.
 
-%It should be noted that time is a continuous concept, with infinitely
-%many divisions possible between any two discrete instants.  However,
-%the resolution with which such divisions can be recognised will, in
-%practise, have a lower bound set by hardware limitations, or may be
-%defined as part of some specific implementation.  In these cases,
-%anything that occurs between discrete instants is considered to have
-%occurred at the next recognised instant.
 \end{definition}
@@
 annotation) and a fact instance (with temporal annotation) with the
-function ${strip : \hat{H} \times \mathbb{T} \rightarrow \hat{H}}$
-which maps a fact instance to its underlying fact:
-
+function $strip()$ which maps a fact instance to its underlying fact:
 \[ strip(\langle f, t\rangle) = f \]
@@
 For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
-of fact instances and $F = \{ strip(f) | f \in F^T \}$.
+of fact instances and $F = \{ strip(f) \mid f \in F^T \}$.
 
 \end{definition}
+
+In \refdef{continuous datalog: datalog program}, we took a Datalog
+program to consist of a set of rules and a set of facts.  In order to
+extend this definition of Datalog programs to incorporate fact
+instances, we define a Time-Annotated Datalog program as follows:
 
@@
 
-A time-annotated Datalog program $\Pi^T = \langle R, F^T\rangle$ is a
-tuple which comprises a finite set of rules $R$ and a finite set of
-fact instances $F^T \subseteq \hat{H} \times \mathbb{T}$. Fact
-instances are as defined in \refdef{continuous datalog: fact
-  instance}, and rules are of the same form as in conventional
-Datalog programs.
+A Time-Annotated Datalog program (henceforth abbreviated TA-Datalog)
+$\Pi^T = \langle R, F^T\rangle$ is a tuple which comprises a finite
+set of rules $R$ and a finite set of fact instances $F^T \subseteq
+\hat{H} \times \mathbb{T}$. Fact instances are as defined in
+\refdef{continuous datalog: fact instance}, and rules are of the same
+form as in conventional Datalog programs.
 
-When considering a time-annotated Datalog program, in which there may
+When considering a TA-Datalog program, in which there may
 be many distinct instances of a given fact, temporal information
@@
 Herbrand base of all fact instances $\hat{H}^T$ (all facts at all
-instants). As before, a fact $f$ is entailed by a program $\Pi^T$ iff
-each interpretation satisfying every clause in $\Pi^T$ also satisfies
-$f$. The entailment of $\Pi^T$ is therefore:
+instants). As before, a fact instance $e$ is entailed by a program
+$\Pi^T$ iff each interpretation satisfying every clause in $\Pi^T$
+also satisfies $e$. The entailment of the program $\Pi^T$ is therefore:
+\[ cons^T(\Pi^T) = \{ e \in \hat{H}^T \mid \Pi^T \models e \} \]
 
-\[ cons^T(\Pi^T) = \{ f \in \hat{H}^T | \Pi^T \models f \} \]
+Informally, the set of instants appearing in some set $E^T$ of fact
+instances entailed by some program ${\langle R, F^T \rangle}$ will be
+a subset of the instances appearing the fact instances in $F^T$.  In
+addition, the instant $t$ in a fact instance $\langle e, t \rangle$
+that is entailed by a TA-Datalog program $\Pi^T = \langle R^T, F^T
+\rangle$ will be bounded from above by the maximum instant in any fact
+instance in $F^T$. A TA-Datalog program may entail fact instances that
+are only distinguished by their instants (and are therefore considered
+distinct from each other, as per \refdef{continuous datalog: fact
+  instance}).
+
+Consider the following example TA-Datalog program:
+\begin{align*}
+  \Pi^T = \langle {} & \{ a \Leftarrow b \land c \}, \\
+  & \{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle  \} \rangle
+\end{align*}
+
+The entailment of the program $\Pi^T$ will be:
+
+\[cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}\]
 
@@
 instances in both the set of fact instances $F^T$ in some ${\Pi^T =
-  \langle R, F^T \rangle}$ and the set of entailed fact instances
-$cons^T(\Pi^T)$ maintains the relationship of entailment; if a fact
-instance $f$ is entailed by $\langle R, F^T\rangle$, then $strip(f)$
-will be entailed by $\langle R, strip(F^T) \rangle$.
+  \langle R, F^T \rangle}$ and the entailment of the program $\Pi^T$
+(the set of fact instances $cons^T(\Pi^T)$) maintains the relationship
+of entailment; if $\langle R, F^T\rangle \models \langle e, t
+\rangle$, then $\langle R, strip(F^T) \rangle \models strip(\langle e,
+t \rangle)$.
+
+This relationship is shown in \reffig{TD-D}.
+
+\begin{figure}[htb]
+  \[
+  \begin{CD}
+\Pi^T = \langle R, F^T \rangle @>\models>> cons^T(\Pi^T) \\
+@V{strip}VV @V{strip}VV \\
+\Pi = \langle R, F \rangle @>\models>> cons(\Pi)
+  \end{CD}
+  \]
+\caption{Relationship between Datalog and TA-Datalog entailments}
+\labelfig{TD-D}
+\end{figure}
+
+Reconsidering the example TA-Datalog program above, we may convert it
+into the following equivalent conventional Datalog program:
+\[
+  \Pi = \langle  \{ a \Leftarrow b \land c \}, \{ a, b \} \rangle
+\]
+Trivially, the entailment of this program $\Pi$ will therefore be:
+\[cons(\Pi) = \{ c \}\]
 \end{definition}
@@
 
-The justification of entailments (as in \refdef{continuous datalog:
-  justification}) is equally applicable to time-annotated Datalog
+The justification of entailments is equally applicable to TA-Datalog
 programs, with facts in the entailment and the justification for that
-entailment being replaced with fact instances.
+entailment being replaced with fact instances.  Following the
+simplified notion of justification in \refdef{continuous datalog:
+  justification}, we define the justification of a set of fact
+instances that are entailed by a TA-Datalog as a minimal subset of
+the ground fact instances of the program that are sufficient for
+the entailment to hold.
 
-Furthermore, each justification for $f$ in some $F^T$
-entails a unique instance ${f | t}$, such that many instances of $f$
-may occur in the set $E^T$ of fact instances entailed by ${\Pi^T =
-  \langle R, W^{S,x}_i \rangle}$, each with a single justification.
+More formally, for a TA-Datalog program $\Pi^T = \langle R, F^T \rangle$
+and a set of entailed fact instances $E^T$ where $\langle R, F^T \rangle \models E^T$, the set of fact instances $J^T \subseteq F^T$ is a justification for
+$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime} \not\models E^T$.
 
-As such, entailment justification for fact instances may be defined
-formally by \refdef{continuous datalog: instance justification}.
+For an entailed fact instance $\langle e, t\rangle$ and a
+justification of that fact instance $J^T$, the instant $t$ at which
+$\langle e, t \rangle$ becomes true is take to be the latest instant
+at which any of the fact instances in the justification became true:
+\[ t = max(\{ t' \mid \langle j', t' \rangle \in J^T \}) \]
 
-A justification for some entailment ${f|t}$ of some time-annotated Datalog program
-	${\Pi^T = \langle R , F^T \rangle \vdash f|t}$ is the single minimal set ${J^T \subseteq F^T}$
-	such that ${\Pi^{T\prime} = \langle R , J^T \rangle \vdash f|t}$.
-	The mapping of the fact instance ${f | t}$ to the minimal subset $J^T$
-	within ${\Pi^T = \langle R, F^T \rangle}$ that justifies it is expressed
-	as the function ${just : \mathbb{\Pi^T} , \mathbb{F^T} \rightarrow \{ \mathbb{F^T} \}}$.
+Note that, although an entailed fact instance $\langle e, t\rangle$
+may have multiple justifications, all the justifications for $\langle
+e, t \rangle$ will share the same maximum instant; a justification
+with a different maximum instant $t' \neq t$ would entail a different
+fact instance $\langle e, t' \rangle$. Considering again the example
+TA-Datalog program in \refdef{continuous datalog: ta datalog program},
+the entailed fact instance $\langle c, 2 \rangle$ would have the
+following set of justifications:
+\[
+  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
+\]
+
 \end{definition}
 
-This allows us to formally specify that the time $t$ at which an entailed fact
-instance ${f | t}$ becomes true is the last time at which any of its
-justifying instances become true:
-\[ t = max(\{ t^\prime : f^\prime | t^\prime \in just(\Pi^T, f | t) \}) \]
-As such, the set of timestamps $T$ for some set $E^T$ of instances entailed by some
-program ${\langle R, F^T \rangle}$ is a subset of the timestamps for the instances in $F^T$.
+Having defined the notion of time annotation in Datalog, we now extend
+this concept to that of Continuous Datalog (henceforth C-Datalog).
 
-\begin{definition}[Stream of Facts]
+\begin{definition}[Streams]
 \labeldef{continuous datalog: stream}
-A stream of facts $S$ is a \emph{sequence} of fact instances ordered
-by their timestamps; in conjunction with the distinction between
+A {\em stream} is a sequence of fact instances ordered
+by their instants; in conjunction with the distinction between
 multiple instances of a fact by their timestamp, this definition justifies
@@
 
-A sliding window of range ${x \in \mathbb{N}}$ over any stream $S$ at
-any given instant ${i \in \mathbb{N}}$ is the \emph{set} of fact
+A sliding window of range ${x \in \mathbb{N}}$ over a stream $S$ at
+an instant ${i \in \mathbb{T}}$ is the \emph{set} of fact
 instances that occur in $S$ during the interval ${(i-x,i]}$, written
-${W^{S,x}_{i}}$.
+${W^{S,x}_{i}}$, and defined as follows:
 
-\[ W^{s,x}_i = \{ \langle f, t \rangle | \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]
+\[ W^{S,x}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]
 
@@
 A window of range $0$ over a stream $S$ at an instant 
-${i \in \mathbb{N}}$ is always empty.
+${i \in \mathbb{T}}$ is always empty.
 \begin{equation*}
@@
 \begin{axiom}\label{axiom:continuous datalog: window composition}
-A window of range ${x \ge 0}$ over a stream $S$ at an instant 
-${i \in \mathbb{N}}$ is equal to the union of the $x$ consecutive
+A window of range ${x > 0}$ over a stream $S$ at an instant 
+${i \in \mathbb{T}}$ is equal to the union of the $x$ consecutive
 windows of range $1$ over the stream $S$ preceding and including that at
@@
 A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
-instant ${i \in \mathbb{N}}$ is disjoint from a second window of 
+instant ${i \in \mathbb{T}}$ is disjoint from a second window of 
 range ${y \in \mathbb{N}}$ over $S$ at any instant before $i - x$.
@@
 \end{axiom}
-
-In \refdef{continuous datalog: datalog program}, we took a Datalog
-program to consist of a set of rules and a set of facts.  In order to
-extend this definition of Datalog programs to incorporate fact
-instances, as in \refdef{continuous datalog: fact instance}, we define
-a time-annotated Datalog program as follows:
 

### 2015-07-13 content/contribution/continuous-datalog.tex
Various changes

@@
 
-In addition, it can be shown that these streams may be trivially
-manipulated to retrieve the instantaneous sets of entailments produced
-by the more semantically well-defined offerings from the reasoning
-community.
+In addition, we show that these streams may be manipulated to retrieve
+the instantaneous sets of entailments produced by the more
+semantically well-defined offerings from the reasoning community.
+
+%%% what does this mean? ^
 
@@
 $\Pi$ (satisfying every clause in $\Pi$) also satisfies $e$, written
-$\Pi \models e$. For a a program $\Pi$ and a set of entailed facts
+$\Pi \models e$. For a program $\Pi$ and a set of entailed facts
 $E$, $\Pi \models E$ iff $\Pi \models e$ for every $e \in E$. For
@@
 
-When considering a TA-Datalog program, in which there may
-be many distinct instances of a given fact, temporal information
-regarding when instances of that fact became true would be lost if
-such a program were to entail sets of plain facts without temporal
+When considering a TA-Datalog program, in which there may be many
+distinct instances of a given fact, temporal information regarding
+when instances of that fact became true would be lost if such a
+program were to entail sets of plain facts without temporal
 annotations. We therefore extend the model theory for Datalog from
@@
 $\Pi^T$ iff each interpretation satisfying every clause in $\Pi^T$
-also satisfies $e$. The entailment of the program $\Pi^T$ is therefore:
+also satisfies $e$, written $\Pi^T \models e$. The {\em entailment of the
+program} $\Pi^T$ is therefore:
 \[ cons^T(\Pi^T) = \{ e \in \hat{H}^T \mid \Pi^T \models e \} \]
@@
 Consider the following example TA-Datalog program:
-\begin{align*}
-  \Pi^T = \langle {} & \{ a \Leftarrow b \land c \}, \\
-  & \{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle  \} \rangle
-\end{align*}
+\[
+\Pi^T = \langle \{ c \Leftarrow a \land b \}, \{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle  \} \rangle
+\]
 
 The entailment of the program $\Pi^T$ will be:
-
 \[cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}\]
@@
 \rangle$, then $\langle R, strip(F^T) \rangle \models strip(\langle e,
-t \rangle)$.
-
-This relationship is shown in \reffig{TD-D}.
+t \rangle)$. This relationship is shown in \reffig{TD-D}.
 
@@
 Reconsidering the example TA-Datalog program above, we may convert it
-into the following equivalent conventional Datalog program:
-\[
-  \Pi = \langle  \{ a \Leftarrow b \land c \}, \{ a, b \} \rangle
-\]
+into the following equivalent conventional Datalog program $\Pi$:
+\begin{align*}
+\Pi &= \langle \{ c \Leftarrow a \land b \}, strip(\{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle \}) \rangle \\
+    &= \langle \{ c \Leftarrow a \land b \}, \{ a, b \} \rangle
+\end{align*}
 Trivially, the entailment of this program $\Pi$ will therefore be:
-\[cons(\Pi) = \{ c \}\]
+\begin{align*}
+cons(\Pi) &= strip(\{\langle c, 1 \rangle, \langle c, 2 \rangle\}) \\
+ &= \{ c \}  
+\end{align*}
 \end{definition}
@@
 
-Having defined the notion of time annotation in Datalog, we now extend
-this concept to that of Continuous Datalog (henceforth C-Datalog).
+Having defined the notion of time annotation in Datalog, we now define
+the notions of streams and windows before extending to Continuous Datalog.
 
@@
 A {\em stream} is a sequence of fact instances ordered
-by their instants; in conjunction with the distinction between
-multiple instances of a fact by their timestamp, this definition justifies
-Axiom~\ref{axiom:continuous datalog: window disjointness}.
+by their instants.
+%FIXME: does this mean that a stream never contains duplicates?
+%; in conjunction with the distinction between
+%multiple instances of a fact by their timestamp, this definition justifies
+%Axiom~\ref{axiom:continuous datalog: window disjointness}.
 \end{definition}
 
-\begin{definition}[Sliding Window]
+\begin{definition}[Window]
 \labeldef{continuous datalog: window}
 
-A sliding window of range ${x \in \mathbb{N}}$ over a stream $S$ at
-an instant ${i \in \mathbb{T}}$ is the \emph{set} of fact
-instances that occur in $S$ during the interval ${(i-x,i]}$, written
-${W^{S,x}_{i}}$, and defined as follows:
-
+A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an instant
+${i \in \mathbb{T}}$ is the \emph{set} of fact instances that occur in
+$S$ during the interval ${(i-x,i]}$, written ${W^{S,x}_{i}}$, and
+defined as follows:
 \[ W^{S,x}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]  
-
 \end{definition}
@@
 derived:
-
 \begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
   If two sliding windows over the same stream $S$ cover non-overlapping time periods,
-then a set ${E^T1}$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
-is necessarily disjoint from any set ${E^T2}$ of fact instances entailed by
-${\langle R, W^{S,y}_{i-z} \rangle}$, where ${z \geq x}$.
+then a set $E^T_1$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
+is necessarily disjoint from any set $E^T_2$ of fact instances entailed by
+$\langle R, W^{S,y}_{i-z} \rangle$, where ${z \geq x}$.
 % This is because the timestamps of the instances ${E^T1}$ are a subset of those of
@@
 
-\begin{definition}[Continuous Datalog Program]\label{def:continuous datalog: CDP}
-A Continuous Datalog Program $\Pi^C$ is a tuple ${\langle R, F, S, r
-\rangle}$ which comprises a finite set of rules $R$, a finite set of persistent
-facts $F$, a stream of facts $S$ and a range ${r \in \mathbb{N}}$ for
-the sliding window over $S$. This specifies that the streamed instances from $S$ are
-to be considered valid from their timestamp for the duration represented
-by $r$.
+\begin{definition}[Continuous Datalog Program]
+\labeldef{continuous datalog: CDP}
 
-The entailment of such a program is composed of a stream of entailed
-instances $P$ and a stream of instance negations $N$, whose contents
+A Continuous Datalog program (henceforth abbreviated C-Datalog)
+$\Pi^C = \langle R, F^0, S, r \rangle$ is a tuple which comprises a
+finite set of rules $R$, a finite set of static fact instances $F^0$
+all of which have the instant $0$, a stream $S$ and a range ${r \in
+\mathbb{N}}$ for the window over $S$.
+
+%This specifies that the streamed instances from $S$
+%are to be considered valid from their timestamp for the duration
+%represented by $r$.
+
+We consider the entailment of the C-Datalog program $\Pi^C$ not just
+as a set of fact instances (as for TA-Datalog), but in three parts: a
+set of fact instances $E^0$ entailed by the fact instances in $F^0$, a
+stream $P$ of fact instances entailed by the fact instances in $F^0$
+and S, and a stream $N$ of fact instance negations. $P$ and $N$ can be
+considered as the continuous reasoning equivalents of the streams
+produced by the \emph{IStream} and \emph{DStream} operators of the
+\emph{CQL} continuous query language \citep{CQL}, respectively.
+
+whose contents
 are defined in 
@@
 \ref{axiom:continuous datalog: entailment precedes negation}. 
-These specify that the streams $P$ and $N$ are the continuous
-reasoning equivalents of the streams produced by the \emph{IStream}
-and \emph{DStream} operators of the \emph{CQL} continuous query
-language \citep{CQL}, respectively.
+These specify that the streams 
 \[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle E, P, N \rangle \]

### 2015-07-13 content/contribution/continuous-datalog.tex
Fixed minor formatting isues

@@
 
-Continuous Datalog is an extension of positive Datalog with
-stream-processing semantics, with a view to accommodating the aims of
-the latter within the well-defined model-theoretic semantics of the
-former, which are shared with RIF-Core \cite{w3crifbld} and therefore
-support reasoning at the expressivity of OWL 2 RL.  This comprises the
-extension of Datalog programs to include sources of continuously
-generated/reported facts, and the application of sliding window
-semantics over those sources, thereby providing for the interpretation
-of Continuous Datalog programs as sequences of instantaneous states
-that are each valid positive Datalog programs.  As such, these states
-each inherit the semantics of positive Datalog.  Furthermore, it
-involves the continuous derivation of entailments from this sequence
-of instantaneous states as a set of streams of facts; these streams
-share a data model with the streams of continuous queries, as defined
-by the stream processing community, thereby allowing the integration
-of Continuous Datalog programs into stream processing pipelines.
+Continuous Datalog (henceforth C-Datalog) is an extension of positive
+Datalog with stream-processing semantics, with a view to accommodating
+the aims of the latter within the well-defined model-theoretic
+semantics of the former, which are shared with RIF-Core
+\cite{w3crifbld} and therefore support reasoning at the expressivity
+of OWL 2 RL. As part of its application to Semantic Web reasoning, we
+assume that the atomic formulae of the Datalog programs described
+below correspond to RDF triples: $\langle subj, pred, obj\rangle$.
+
+C-Datalog differs from conventional Datalog through the addition of
+sources of continuously generated or reported facts (i.e streams), and
+the application of sliding window semantics over those sources,
+thereby providing for the interpretation of Continuous Datalog
+programs as sequences of instantaneous states that are each valid
+positive Datalog programs.  As such, these states each inherit the
+semantics of positive Datalog.  Furthermore, it involves the
+continuous derivation of entailments from this sequence of
+instantaneous states as a set of streams of facts; these streams share
+a data model with the streams of continuous queries, as defined by the
+stream processing community, thereby allowing the integration of
+Continuous Datalog programs into stream processing pipelines.
 
@@
 \labeldef{continuous datalog: stream}
-A {\em stream} is a sequence of fact instances ordered
-by their instants.
-%FIXME: does this mean that a stream never contains duplicates?
-%; in conjunction with the distinction between
-%multiple instances of a fact by their timestamp, this definition justifies
-%Axiom~\ref{axiom:continuous datalog: window disjointness}.
+
+A {\em stream} is an unbounded sequence of fact instances ordered by
+their instants; the instants are non-decreasing in the stream, such
+that for any two fact instances $\langle f_i, t_i \rangle$ and
+$\langle f_j, t_j\rangle$ where the former appears on the stream
+  before the latter, $t_i \leqslant t_j$.
 \end{definition}
@@
 derived:
+
 \begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
@@
 % ${\{ t : f | t \in W^{S,x}_i\} \cap \{ t : f | t \in W^{S,y}_{i-z}\} = \varnothing}$.
-\[ \{ e : \langle R , W^{S,x}_i \rangle \vdash e \} \cap \{ e : \langle R , W^{S,y}_{i-z} \rangle \vdash e \} = \varnothing \]
+\[
+\{ e : \langle R , W^{S,x}_i \rangle \vdash e \} \cap \{ e : \langle R , W^{S,y}_{i-z} \rangle \vdash e \} = \varnothing
+\]
 \end{axiom}
@@
 finite set of rules $R$, a finite set of static fact instances $F^0$
-all of which have the instant $0$, a stream $S$ and a range ${r \in
-\mathbb{N}}$ for the window over $S$.
+(all of which have the instant $0$), a stream $S$ and a range
+${r \in \mathbb{N}}$ for a window over $S$.
 
@@
 
-We consider the entailment of the C-Datalog program $\Pi^C$ not just
+We consider the entailment of the program $\Pi^C$ not just
 as a set of fact instances (as for TA-Datalog), but in three parts: a
@@
 stream $P$ of fact instances entailed by the fact instances in $F^0$
-and S, and a stream $N$ of fact instance negations. $P$ and $N$ can be
-considered as the continuous reasoning equivalents of the streams
-produced by the \emph{IStream} and \emph{DStream} operators of the
-\emph{CQL} continuous query language \citep{CQL}, respectively.
+and S, and a stream $N$ of fact instance negations: $\langle R, F^0,
+S, r \rangle \models \langle E^0, P, N \rangle$. $P$ and $N$ can be
+considered to be the continuous reasoning equivalents of the streams
+produced by the \emph{IStream} and \emph{DStream} operators
+respectively of the \emph{CQL} continuous query language \citep{CQL},
+which create streams based on the additions to and deletions from a
+relation.
 
-whose contents
-are defined in 
-Axioms~\ref{axiom:continuous datalog: positive window increment}, 
-\ref{axiom:continuous datalog: negative window increment} and
-\ref{axiom:continuous datalog: entailment precedes negation}. 
-These specify that the streams 
-\[ \Pi^C = \langle R, F, S, r \rangle \vDash \langle E, P, N \rangle \]
+In order to
+
+\end{definition}
+
+\begin{definition}[Instanteous State of a C-Datalog program]
+\label{def:continuous datalog: CDPt}
+
+Given a C-Datalog program $\Pi^C$, we can
+
 The truth value of a $\Pi^C$ is instantaneous, with the the truth
@@
 and an instant ${i \in \mathbb{N}}$ as arguments.
-\end{definition}
 
-\begin{definition}[State of a $\Pi^C$]\label{def:continuous datalog: CDPt}
 The state of a $\Pi^C$ at a given instant $i$ is the time-annotated

### 2015-07-13 content/contribution/continuous-datalog.tex
minor fixes to typos and descriptions up to definition of streams

--- 2015-07-13 content/contribution/continuous-datalog.tex
@@
 \begin{nestedsection}{Continuous Datalog}{semantics}
-The first requirement of stream reasoning is a well-defined semantics.
+The first requirement of stream reasoning is a well-defined semantics
 that takes into account both the continuous nature of a subset of the
@@
 The Herbrand base $\hat{H}$ of $\Pi$ is the set of all facts that can
-be expressed using the predicates in $\Pi$; for a program $\Pi =
+be expressed using the predicates and constants in $\Pi$; for a program $\Pi =
 \langle R, F \rangle$, $F \subseteq \hat{H}$. Note that we do not
@@
 A {\em fact instance} is a fact with some additional temporal
-annotation; it is a tuple containing a fact $f$ and an instant $t \in
+annotation regarding the instant from which it is recognised to be true;
+it is a tuple containing a fact $f$ and an instant $t \in
 \mathbb{T}$, written $\langle f, t \rangle$. Two fact instances
@@
 instances entailed by some program ${\langle R, F^T \rangle}$ will be
-a subset of the instances appearing the fact instances in $F^T$.  In
+a subset of the instants appearing in the fact instances in $F^T$.  In
 addition, the instant $t$ in a fact instance $\langle e, t \rangle$
@@
 justification of that fact instance $J^T$, the instant $t$ at which
-$\langle e, t \rangle$ becomes true is take to be the latest instant
+$\langle e, t \rangle$ becomes true is taken to be the latest instant
 at which any of the fact instances in the justification became true:

### 2015-07-13 content/conclusions.tex
started conclusions, not sure if it is any good...

@@
 \begin{nestedsection}{Conclusions}{conclusions}
+	We have presented C-Datalog, a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog.
+	C-Datalog differs from other temporal interpretations of Datalog \citep{TODO} in that it forgoes the concepts of conventional temporal reasoning in favour of the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	C-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program, the derivation of concise streams of facts as the entailment of a continuous program over time.
+	Additionally, it supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program.
+
 	We are more semantically sound than other RDF Stream Processing offerings and more ``continuous'' than other continuous reasoning offerings, without sacrificing processing efficiency.

### 2015-07-14 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
migrated the proof to new notation

@@
 	\begin{thrm}\labelhyp{continuous datalog: entailment} \hfill \\
-		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
+		For any ${\Pi^{C,r} = \langle R, F^0, S, r\} \models \langle P, N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_{i}}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \geq r}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E^T_{i}$ transitively entailed by the ${\Pi^{C,r}}$ at that instant $i$:
 		\begin{equation*}
-			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i}
+			W^{P,x}_{i} \setminus W^{N,y}_{i} = E_{i}
 		\end{equation*}
@@
 	\begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
-		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
+		For any ${\Pi^{C,r} \models \langle P,N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_i}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the set $E^T_i$ of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
 		\begin{equation*}
-			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d}
+			W^{P,x}_{i} \setminus W^{N,y}_{i} = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d}
 		\end{equation*}
@@
 	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
-		Assume that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$.
-		Remember that ${W^{S,R}_{i} \cap W^{S,R}_{i-R} = \varnothing}$, ${W^{S,R}_{i} \vDash E_{i}}$ and ${W^{S,R}_{i-R} \vDash E_{i-R}}$.
-		Therefore:
+		Assume that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${x \geq r}$ for some ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ and ${y \geq x}$ at every instant ${i \in \mathbb{T}}$.
+		It follows that:
 		\begin{align*}
-			E_{i} & = W^{P,X}_{i} \setminus W^{N,Y}_{i} \\
-			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d} \\
-			& = \mathop{\cup}^{X}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			\text{Axiom~\ref{axiom:continuous datalog: entailment disjointness}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\
-			& = E_{i} && \qedhere
+			E^T_{i} & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
+			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d} \\
+			& = \mathop{\cup}^{x}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
+			& = \left( E^T_{i} \setminus E^T_{i-x} \right) \cup \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
+			\text{Axiom~\ref{axiom:continuous datalog: entailment disjointness}} & = E^T_{i} \cup \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
+			& = E^T_{i} \cup \left( E^T_{i} \setminus \mathop{\cap}^{x-1}_{d=1} E^T_{i-d} \right) \\
+			& = E^T_{i} && \qedhere
 		\end{align*}
@@
 
-	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${X \in \mathbb{N}}$, regardless of the Range of of the ${CDP}$ entailing $P$ and $N$.
+	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${x \in \mathbb{N}}$, regardless of the range of the ${\Pi^{C,r}}$ entailing $P$ and $N$.
 
 	\begin{proof}[Proof of Lemma~\ref{hyp:continuous datalog: entailment lemma} by Induction]
-		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${X = k}$ and ${Y \geq X}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${X = k + 1}$, where ${k \in \mathbb{N}}$.
+		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${x = k}$ and ${y \geq x}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${x = k + 1}$, where ${k \in \mathbb{N}}$.
 
@@
 		\begin{align*}
-			W^{P,X}_{i} \setminus W^{P,Y}_{i} & = \\
-			\text{Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{X-1}_{d=0}W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \\
-			& = \mathop{\cup}^{X-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \right) \\
-			& = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{Y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-			\text{Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } Y \geq X & = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right)
+			W^{P,x}_{i} \setminus W^{P,y}_{i} & = \\
+			\text{Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{x-1}_{d=0}W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0}W^{N,1}_{i-j} \\
+			& = \mathop{\cup}^{x-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0}W^{N,1}_{i-j} \right) \\
+			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
+			\text{Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } y \geq x & = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right)
 		\end{align*}
-		The general case also shows that any negative entailments within the window of range $Y$ over $N$ not in the window of range $X$ over $N$ do not influence the composition of the set of values ${W^{P,X}_{i} \setminus W^{P,Y}_{i}}$.
+		The general case also shows that any negative entailments within the window of range $y$ over $N$ not in the window of range $x$ over $N$ do not influence the composition of the set of values ${W^{P,x}_{i} \setminus W^{N,y}_{i}}$.
 
-		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${X = 1}$ (therefore ${Y \geq 1}$).
+		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${x = 1}$ (therefore ${y \geq 1}$).
 		\begin{align*}
-			E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-1} & = W^{P,1}_{i} \setminus W^{N,Y}_{i} \\
+			E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-1} & = W^{P,1}_{i} \setminus W^{N,y}_{i} \\
 			\text{\textbf{General Case}} & = W^{P,1}_{i} \setminus W^{N,1}_{i} \\
-			\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E_{i} \setminus E_{i-1} \right) \setminus \left( E_{i-1} \setminus E_{i} \right) \\
-			& = E_{i} \setminus \left( E_{i-1} \setminus \left( E_{i-1} \setminus E_{i} \right) \right) \\
-			& = E_{i} \setminus \left( E_{i-1} \cap E_{i} \right) \\
-			& = \left( E_{i} \setminus E_{i-1} \right) \cup \left( E_{i} \setminus E_{i} \right) \\
-			& = E_{i} \setminus E_{i-1} \\
-			& = E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-d} && \qedhere
+			\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E^T_{i} \setminus E^T_{i-1} \right) \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) \\
+			& = E^T_{i} \setminus \left( E^T_{i-1} \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) \right) \\
+			& = E^T_{i} \setminus \left( E^T_{i-1} \cap E^T_{i} \right) \\
+			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \cup \left( E^T_{i} \setminus E^T_{i} \right) \\
+			& = E^T_{i} \setminus E^T_{i-1} \\
+			& = E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} && \qedhere
 		\end{align*}
 
-		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
+		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${x = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${x = k}$.
 		\begin{figure*}[p]
@@
 			\begin{align*}
-				E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
-				\text{\textbf{General Case}} & = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
-				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
+				E^t_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
+				\text{\textbf{General Case}} & = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
+				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-j} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
+				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=-1} W^{N,1}_{i-j} \right) \\
+				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
+				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
 				& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
-				\text{\textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}:}& = \left( \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus \left( E_{i} \setminus E_{i+1} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus \left( E_{i} \cap E_{i+1} \right) \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = \left( E_{i+1} \cap E^C_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right)\\
-				& = E_{i+1} \cap \left( E^C_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = E_{i+1} \setminus \left( E_{i} \setminus \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = E_{i+1} \setminus \left( E_{i} \cap \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = E_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E_{i-d} \\
-				& = E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} && \qedhere
+				\text{\textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}:}& = \left( \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \\
+				& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \\
+				& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+				& = \left( E^T_{i+1} \setminus \left( E^T_{i} \cap E^T_{i+1} \right) \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+				& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+				& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+				& = \left( E^T_{i+1} \cap \left( E^T_{i} \right^C \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right)\\
+				& = E^T_{i+1} \cap \left( \left( E^T_{i} \right)^C \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+				& = E^T_{i+1} \setminus \left( E^T_{i} \setminus \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+				& = E^T_{i+1} \setminus \left( E^T_{i} \cap \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+				& = E^T_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E^T_{i-d} \\
+				& = E^T_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} && \qedhere
 			\end{align*}

### 2015-07-14 content/contribution/continuous-datalog.tex
moved the proof of the theorem from appendix to section on c-datalog, then commented out to be re-introduced if desired.  Reformatted appendix to a set of figures, being general case, base case and inductive case
Revised defn of C-Datalog, added initial defn of WC-Datalog

--- 2015-07-13 content/contribution/continuous-datalog.tex
@@
 the application of sliding window semantics over those sources,
-thereby providing for the interpretation of Continuous Datalog
-programs as sequences of instantaneous states that are each valid
-positive Datalog programs.  As such, these states each inherit the
-semantics of positive Datalog.  Furthermore, it involves the
-continuous derivation of entailments from this sequence of
-instantaneous states as a set of streams of facts; these streams share
-a data model with the streams of continuous queries, as defined by the
-stream processing community, thereby allowing the integration of
-Continuous Datalog programs into stream processing pipelines.
-
-In addition, we show that these streams may be manipulated to retrieve
-the instantaneous sets of entailments produced by the more
-semantically well-defined offerings from the reasoning community.
-
-%%% what does this mean? ^
+thereby providing for the interpretation of C-Datalog programs as
+sequences of instantaneous states that are each valid positive Datalog
+programs.  As such, these states each inherit the semantics of
+positive Datalog.  Furthermore, it involves the continuous derivation
+of entailments from this sequence of instantaneous states as a stream
+of facts; these streams share a data model with the streams of
+continuous queries, as defined by the stream processing community,
+thereby allowing the integration of C-Datalog programs into stream
+processing pipelines. Finally, we show that these streams may be
+manipulated to retrieve the instantaneous sets of entailments produced
+by the equivalent Datalog programs.
 
@@
 
-A Time-Annotated Datalog program (henceforth abbreviated TA-Datalog)
-$\Pi^T = \langle R, F^T\rangle$ is a tuple which comprises a finite
-set of rules $R$ and a finite set of fact instances $F^T \subseteq
-\hat{H} \times \mathbb{T}$. Fact instances are as defined in
-\refdef{continuous datalog: fact instance}, and rules are of the same
-form as in conventional Datalog programs.
+A {\em Time-Annotated Datalog program} (henceforth abbreviated
+TA-Datalog program) $\Pi^T = \langle R, F^T\rangle$ is a tuple which
+comprises a finite set of rules $R$ and a finite set of fact instances
+$F^T \subseteq \hat{H} \times \mathbb{T}$. Fact instances are as
+defined in \refdef{continuous datalog: fact instance}, and rules are
+of the same form as in conventional Datalog programs.
 
@@
 \refdef{continuous datalog: datalog program} by considering the
-Herbrand base of all fact instances $\hat{H}^T$ (all facts at all
+Herbrand base of all fact instances $\hat{H}^T$ (all facts with all
 instants). As before, a fact instance $e$ is entailed by a program
@@
 also satisfies $e$, written $\Pi^T \models e$. The {\em entailment of the
-program} $\Pi^T$ is therefore:
+TA-Datalog program} $\Pi^T$ is therefore:
 \[ cons^T(\Pi^T) = \{ e \in \hat{H}^T \mid \Pi^T \models e \} \]
@@
 
-Reconsidering the example TA-Datalog program above, we may convert it
-into the following equivalent conventional Datalog program $\Pi$:
+Reconsidering the example TA-Datalog program above, we may reduce it
+to the following equivalent conventional Datalog program $\Pi$:
 \begin{align*}
@@
 
-A {\em stream} is an unbounded sequence of fact instances ordered by
-their instants; the instants are non-decreasing in the stream, such
-that for any two fact instances $\langle f_i, t_i \rangle$ and
-$\langle f_j, t_j\rangle$ where the former appears on the stream
-  before the latter, $t_i \leqslant t_j$.
+A {\em stream} $S$ is an unbounded, ordered sequence of fact
+instances. We indicate that a fact instance $\langle f, t\rangle$
+appears in $S$ by writing $\langle f, r \rangle \in S$. We indicate
+that (in a given stream) fact instance $f_i$ appears before fact
+instance $f_j$ by writing $f_i \prec f_j$.
+
+Streams may be (but are not necessarily) ordered by the instants in
+their fact instances, with the instants non-decreasing in the stream,
+such that for any two fact instances $\langle f_i, t_i \rangle$ and
+$\langle f_j, t_j\rangle$ where $\langle f_i, t_i \rangle \prec
+\langle f_j, t_j \rangle$, $t_i \leqslant t_j$; we refer to such
+streams as {\em in-order streams}, and to streams whose fact instances
+are not ordered by their instants as {\em out-of-order streams}.
 \end{definition}
@@
 ${i \in \mathbb{T}}$ is the \emph{set} of fact instances that occur in
-$S$ during the interval ${(i-x,i]}$, written ${W^{S,x}_{i}}$, and
-defined as follows:
+$S$ with instances in the interval ${(i-x,i]}$, written
+  ${W^{S,x}_{i}}$, and defined as follows:
 \[ W^{S,x}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]  
@@
 
-A Continuous Datalog program (henceforth abbreviated C-Datalog)
-$\Pi^C = \langle R, F^0, S, r \rangle$ is a tuple which comprises a
-finite set of rules $R$, a finite set of static fact instances $F^0$
-(all of which have the instant $0$), a stream $S$ and a range
-${r \in \mathbb{N}}$ for a window over $S$.
+We initially take a {\em Continuous Datalog program} (henceforth a
+C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
+which comprises a finite set of rules $R$, a finite set of static fact
+instances $F^0$ (all of which have the instant $0$) and an in-order
+stream of fact instances $S$.
 
-%This specifies that the streamed instances from $S$
-%are to be considered valid from their timestamp for the duration
-%represented by $r$.
+The model theory for a such a C-Datalog program broadly follows that
+for a TA-Datalog program; we again take the Herbrand base of all fact
+instances $\hat{H}$. Disregarding the order of fact instances in $S$,
+a fact instance $e$ is entailed by a program $\Pi^C$ iff each
+interpretation satisfying every clause in $\Pi^C$. However, we take
+the entailment of the program $\Pi^C$ to be an in-order stream $E^S$,
+where the order in which entailed fact instances appear in $E^S$
+depends on the order in which the fact instances in their
+justifications appear in $S$. Given that that instant of an entailed
+fact instance is the maximum of the instants in the fact instances
+that form the justification for that entailed fact instance (as in
+\refdef{continuous datalog: instance justification}), $E^S$ will also
+be an in-order stream.
 
-We consider the entailment of the program $\Pi^C$ not just
-as a set of fact instances (as for TA-Datalog), but in three parts: a
-set of fact instances $E^0$ entailed by the fact instances in $F^0$, a
-stream $P$ of fact instances entailed by the fact instances in $F^0$
-and S, and a stream $N$ of fact instance negations: $\langle R, F^0,
-S, r \rangle \models \langle E^0, P, N \rangle$. $P$ and $N$ can be
-considered to be the continuous reasoning equivalents of the streams
-produced by the \emph{IStream} and \emph{DStream} operators
-respectively of the \emph{CQL} continuous query language \citep{CQL},
-which create streams based on the additions to and deletions from a
-relation.
+However, this definition of a C-Datalog program does not take into
+account any notion of windowing over the stream $S$. We define a {\em
+  Windowed Continuous Datalog program} $\Pi^{WC} = \langle R, F^0, S,
+r \rangle$ to be a tuple with components as defined above for a
+C-Datalog program, but with the addition of $r \in \mathbb{N}$ for the
+range of the window over $S$, which effectively specifies that the
+fact instances from $S$ are to contribute to the entailments of $\Pi^{WC}$
+from their instant for a duration of $r$.
+
+We consider the entailment of the WC-Datalog program $\Pi^{WC}$ not as
+a single stream of fact instances (as for C-Datalog), but in three
+parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle$
+that is the entailment of the TA-Datalog program $\langle R,
+F^0\rangle$, a stream $P$ of fact instances entailed by the fact
+instances in $F^0$ and $S$, and a stream $N$ of fact instance negations:
+$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
+and $N$ can be considered to be the continuous reasoning equivalents
+of the streams produced by the \emph{IStream} and \emph{DStream}
+operators respectively of the \emph{CQL} continuous query language
+\citep{CQL}, which create streams based on the additions to and
+deletions from a relation.
 
@@
 
-Given a C-Datalog program $\Pi^C$, we can
-
-The truth value of a $\Pi^C$ is instantaneous, with the the truth
-function of Continuous Datalog $T^{CD}$ taking both a program ${\Pi^C}$
-and an instant ${i \in \mathbb{N}}$ as arguments.
-
-The state of a $\Pi^C$ at a given instant $i$ is the time-annotated
-Datalog program ${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle}$,
-where ${F^0 = \{ f | 0 : f \in F \}}$ is the set of persistent fact instances produced by
-annotating each fact in $f$ in $F$ as becoming true at the first recognised instant.
-
-As both $F$ and ${W_i = \{ strip(x) : x \in W^{S,r}_{i} \}}$ are sets of conventional
-Datalog facts, the tuple ${\langle R, F \cup W_i}$
-constitutes a valid Datalog program that may be evaluated by the truth
-function of Datalog $T^{D}$, the result of which is the definition for the
-instantaneous truth of $\Pi^C$ at instant $i$ by $T^{CD}$.
-\begin{multline*}
-T^{CD} \left( \langle R, F, S, r \rangle, i \right) = \\
-	T^{D} \left( \langle R, F \cup \{ strip(x) : x \in W^{S,r}_{i} \} \right)
-\end{multline*}
+We can reduce a C-Datalog program $\Pi^C = \langle R, F^0, S, r
+\rangle$ to its TA-Datalog equivalent by taking the window $W^{S,r}_i$
+over stream $S$ of range $r \in \mathbb{N}$ at instant $i \in
+\mathbb{T}$; we refer to this as the {\em instantaneous state} of the
+C-Datalog program, written as $\Pi^C_i = \langle R, F^0 \cup W^{s,r}_i
+\rangle$. As previously shown in \refdef{continuous datalog: ta
+  datalog program}, this may then be further reduced to the Datalog
+program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.
 

### 2015-07-14 content/contribution/continuous-datalog.tex

@@
 
+% \begin{thrm}\labelhyp{continuous datalog: entailment} \hfill \\
+
+% For any ${\Pi^{C,r} \models \langle P, N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_{i}}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \geq r}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the transitive portion $E^T_{i}$ of the entailment of the state of the ${\Pi^{C,r}}$ at that instant $i$:
+
+% \begin{equation*}
+%   W^{P,x}_{i} \setminus W^{N,y}_{i} = E^T_{i}
+% \end{equation*}
+
+% \end{thrm}
+
+% \begin{proof}
+
+% In order to prove \refhyp{continuous datalog: entailment} for the specific case of ${x \geq r}$, for some range $r$ of the window over $S$ in some ${\Pi^{C,r} = \langle R,F^0,S,r \rangle}$, a general case must first be proven for all ${x \geq 1}$.
+% This general case constitutes the Lemma \ref{hyp:continuous datalog: entailment lemma}, which provides an equivalence between the entailments of discrete states of $\Pi^{C,r}$ and windows over the streamed entailment ${\langle P, N \rangle}$ for all window ranges ${x \geq 1}$.
+
+% \begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
+
+% For any ${\Pi^{C,r} \models \langle P,N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_i}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \geq 1}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the transitive portion $E^T_i$ of the entailment of the state of $\Pi^{C,r}$ at that instant $i$ less the intersection of the transitive portions of the entailments of all previous states of $\Pi^{C,r}$, from instant ${i-1}$ up to and including ${i-x}$:
+
+% \begin{equation*}
+%   W^{P,x}_{i} \setminus W^{N,y}_{i} = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d}
+% \end{equation*}
+
+% \end{lem}
+
+% The inductive proof of Lemma \ref{hyp:continuous datalog: entailment lemma} is available in \ref{app:semantics: proof for entailment}:
+% a \emph{general case} (\reffig{semantics: induction general case}) eliminates the range ${y \geq x}$ over $N$ as a factor;
+% a \emph{base case} (\reffig{semantics: induction base case}) is shown to hold when ${x = 1}$;
+% and the \emph{inductive case} (\reffig{semantics: induction inductive case}) holds when ${x = k + 1}$, assuming the lemma holds for ${x = k}$.
+% Given this lemma, the derivation of \refhyp{continuous datalog: entailment} presented in \reffig{continuous datalog: proof for entailment} follows.
+
+% \begin{figure*}
+
+% \centering
+
+% \begin{align*}
+%   E^T_{i} & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
+%   & = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d}
+%   && \text{, Lemma~\ref{hyp:continuous datalog: entailment lemma}} \\
+%   & = \mathop{\cup}^{x}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
+%   & = \left( E^T_{i} \setminus E^T_{i-x} \right) \cup
+%   \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
+%   & = E^T_{i} \cup \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right)
+%   && \text{, Axiom~\ref{axiom:continuous datalog: entailment disjointness}} \\
+%   & = E^T_{i} \cup \left( E^T_{i} \setminus \mathop{\cap}^{x-1}_{d=1} E^T_{i-d} \right) \\
+%   & = E^T_{i} && \qedhere
+% \end{align*}
+
+% \caption{Derivation for proof of \refhyp{continuous datalog: entailment}}
+% \labelfig{continuous datalog: proof for entailment}
+
+% \end{figure*}
+
+% \end{proof}
+
 To sum up, a time-annotated Datalog program

### 2015-07-14 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex

--- 2015-06-25 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
@@
-\begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
-	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
-		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
-		\begin{equation*}
-			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i}
-		\end{equation*}
-	\end{hyp}
-	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
-	\begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
-		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
-		\begin{equation*}
-			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d}
-		\end{equation*}
-	\end{lem}
+\begin{nestedappendix}{Entailment Semantics: Proof of Inductive Lemma}{semantics: proof for entailment}
+	\begin{figure*}
+		\centering
+		\caption{General Case: expresses the set difference $W^{P,x}_{i} \setminus W^{N,y}_{i}$ in terms of windows of range 1, eliminating $y$ as a variable.}
+		\labelfig{semantics: induction general case}
+		\begin{align*}
+			\mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
+			& = \mathop{\cup}^{x-1}_{d=0} W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} && \text{, Axiom~\ref{axiom:continuous datalog: window composition}} \\
+			& = \mathop{\cup}^{x-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} \right) \\
+			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
+			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{, Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } y \geq x
+		\end{align*}
+	\end{figure*}
 
-	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
-		Assume that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$.
-		Remember that ${W^{S,R}_{i} \cap W^{S,R}_{i-R} = \varnothing}$, ${W^{S,R}_{i} \vDash E_{i}}$ and ${W^{S,R}_{i-R} \vDash E_{i-R}}$.
-		Therefore:
+	\begin{figure*}
+		\centering
+		\caption{Base Case: proves Lemma~\ref{hyp:continuous datalog: entailment lemma} for the case ${x = 1}$.}
+		\labelfig{semantics: induction base case}
 		\begin{align*}
-			E_{i} & = W^{P,X}_{i} \setminus W^{N,Y}_{i} \\
-			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d} \\
-			& = \mathop{\cup}^{X}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			\text{Axiom~\ref{axiom:continuous datalog: entailment disjointness}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
-			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\
-			& = E_{i} && \qedhere
+			E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} & = W^{P,1}_{i} \setminus W^{N,y}_{i} \\
+			& = W^{P,1}_{i} \setminus W^{N,1}_{i} && \text{, \textbf{General Case}} \\
+			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) && \text{, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
+			& = E^T_{i} \setminus \left( E^T_{i-1} \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) \right) \\
+			& = E^T_{i} \setminus \left( E^T_{i-1} \cap E^T_{i} \right) \\
+			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \cup \left( E^T_{i} \setminus E^T_{i} \right) \\
+			& = E^T_{i} \setminus E^T_{i-1} \\
+			& = E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} && \qedhere
 		\end{align*}
-	\end{proof}
+	\end{figure*}
 
-	Lemma~\ref{hyp:continuous datalog: entailment lemma} may be shown to hold for all ${X \in \mathbb{N}}$, regardless of the Range of of the ${CDP}$ entailing $P$ and $N$.
-
-	\begin{proof}[Proof of Lemma~\ref{hyp:continuous datalog: entailment lemma} by Induction]
-		\textbf{Inductive Hypothesis:} Given that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for windows of range ${X = k}$ and ${Y \geq X}$ over the entailed streams $P$ and $N$, Lemma~\ref{hyp:continuous datalog: entailment lemma} will hold for windows of range ${X = k + 1}$, where ${k \in \mathbb{N}}$.
-
-		Firstly, the \textbf{general case} can be defined in terms of sub-windows of range $1$ by Axiom~\ref{axiom:continuous datalog: window composition} over the entailed streams $P$ and $N$, for which Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment} provide an equivalence to sets of entailments.
-		\begin{align*}
-			W^{P,X}_{i} \setminus W^{P,Y}_{i} & = \\
-			\text{Axiom~\ref{axiom:continuous datalog: window composition}} & = \mathop{\cup}^{X-1}_{d=0}W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \\
-			& = \mathop{\cup}^{X-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{Y-1}_{j=0}W^{N,1}_{i-j} \right) \\
-			& = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{Y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-			\text{Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } Y \geq X & = \mathop{\cup}^{X-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right)
-		\end{align*}
-		The general case also shows that any negative entailments within the window of range $Y$ over $N$ not in the window of range $X$ over $N$ do not influence the composition of the set of values ${W^{P,X}_{i} \setminus W^{P,Y}_{i}}$.
-
-		Next, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{base case}, where ${X = 1}$ (therefore ${Y \geq 1}$).
-		\begin{align*}
-			E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-1} & = W^{P,1}_{i} \setminus W^{N,Y}_{i} \\
-			\text{\textbf{General Case}} & = W^{P,1}_{i} \setminus W^{N,1}_{i} \\
-			\text{Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} & = \left( E_{i} \setminus E_{i-1} \right) \setminus \left( E_{i-1} \setminus E_{i} \right) \\
-			& = E_{i} \setminus \left( E_{i-1} \setminus \left( E_{i-1} \setminus E_{i} \right) \right) \\
-			& = E_{i} \setminus \left( E_{i-1} \cap E_{i} \right) \\
-			& = \left( E_{i} \setminus E_{i-1} \right) \cup \left( E_{i} \setminus E_{i} \right) \\
-			& = E_{i} \setminus E_{i-1} \\
-			& = E_{i} \setminus \mathop{\cap}^{1}_{d=1} E_{i-d} && \qedhere
-		\end{align*}
-
-		Finally, Lemma~\ref{hyp:continuous datalog: entailment lemma} may be proven for the \textbf{inductive case} as in \reffig{continuous datalog: entailment lemma inductive case}, where ${X = k + 1}$ and Lemma~\ref{hyp:continuous datalog: entailment lemma} holds when ${X = k}$.
-		\begin{figure*}[p]
+	\begin{figure*}
 		\centering
-			\caption[Proof of Entailment Lemma for Continuous Datalog]{Proof of Inductive Case of the Entailment Lemma for Continuous Datalog.}
+		\caption{Inductive Case: proves Lemma~\ref{hyp:continuous datalog: entailment lemma} for the case ${x = k + 1}$, assuming that it holds for ${x = k}$.}
 		\labelfig{continuous datalog: entailment lemma inductive case}
 		\begin{align*}
-				E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
-				\text{\textbf{General Case}} & = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-				& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=-1} W^{N,1}_{i-j} \right) \\
-				& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{i}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
-				& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
+			E^t_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
+			& = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{, \textbf{General Case}} \\
+			& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-j} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
+			& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=-1} W^{N,1}_{i-j} \right) \\
+			& = \left( W^{P,1}_{i+1} \setminus W^{N,1}_{i+1} \right) \cup \mathop{\cup}^{k-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{d}_{j=0} W^{N,1}_{i-j} \setminus W^{N,1}_{i+1} \right) \\
+			& = \left( W^{P,1}_{i+1} \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \right) \setminus W^{N,1}_{i+1} \\
 			& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
-				\text{\textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}:}& = \left( \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \setminus \left( E_{i} \setminus E_{i+1} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \right) \cup \left( E_{i} \setminus \left( E_{i} \setminus E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus \left( E_{i} \cap E_{i+1} \right) \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( \left( E_{i} \cap E_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = \left( E_{i+1} \setminus E_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = \left( E_{i+1} \cap E^C_{i} \right) \cup \left( E_{i+1} \cap \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right)\\
-				& = E_{i+1} \cap \left( E^C_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = E_{i+1} \setminus \left( E_{i} \setminus \left( E_{i} \setminus \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \right) \\
-				& = E_{i+1} \setminus \left( E_{i} \cap \mathop{\cap}^{k}_{d=1} E_{i-d} \right) \\
-				& = E_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E_{i-d} \\
-				& = E_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E_{i+1-d} && \qedhere
+			& = \left( \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) && \text{, \textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
+			& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \\
+			& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+			& = \left( E^T_{i+1} \setminus \left( E^T_{i} \cap E^T_{i+1} \right) \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+			& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( \left( E^T_{i} \cap E^T_{i+1} \right) \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+			& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+			& = \left( E^T_{i+1} \cap \left( E^T_{i} \right^C \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right)\\
+			& = E^T_{i+1} \cap \left( \left( E^T_{i} \right)^C \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+			& = E^T_{i+1} \setminus \left( E^T_{i} \setminus \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
+			& = E^T_{i+1} \setminus \left( E^T_{i} \cap \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \\
+			& = E^T_{i+1} \setminus \mathop{\cap}^{k}_{d=0} E^T_{i-d} \\
+			& = E^T_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} && \qedhere
 		\end{align*}
 	\end{figure*}
-	\end{proof}
-\end{nestedsection}
+\end{nestedappendix}

### 2015-07-14 content/contribution/continuous-datalog.tex
fixed axioms 5-7

@@
 instances. We indicate that a fact instance $\langle f, t\rangle$
-appears in $S$ by writing $\langle f, r \rangle \in S$. We indicate
+appears in $S$ by writing $\langle f, t \rangle \in S$. We indicate
 that (in a given stream) fact instance $f_i$ appears before fact
@@
 
-From this description of instance entailment,
-Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be
-derived:
-
-\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-  If two sliding windows over the same stream $S$ cover non-overlapping time periods,
-then a set $E^T_1$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
-is necessarily disjoint from any set $E^T_2$ of fact instances entailed by
-$\langle R, W^{S,y}_{i-z} \rangle$, where ${z \geq x}$.
-% This is because the timestamps of the instances ${E^T1}$ are a subset of those of
-% the instances $W^{S,x}_i$, the timestamps of the instances ${E^T2}$ are a subset
-% of those of the instances $W^{S,y}_{i-z}$, fact instances are distinct by timestamp and
-% ${\{ t : f | t \in W^{S,x}_i\} \cap \{ t : f | t \in W^{S,y}_{i-z}\} = \varnothing}$.
-\[
-\{ e : \langle R , W^{S,x}_i \rangle \vdash e \} \cap \{ e : \langle R , W^{S,y}_{i-z} \rangle \vdash e \} = \varnothing
-\]
-\end{axiom}
-
 Given these definitions, a Continuous Datalog program may be formally
@@
 account any notion of windowing over the stream $S$. We define a {\em
-  Windowed Continuous Datalog program} $\Pi^{WC} = \langle R, F^0, S,
+  Windowed Continuous Datalog program} $\Pi^{C, r} = \langle R, F^0, S,
 r \rangle$ to be a tuple with components as defined above for a
@@
 
-We consider the entailment of the WC-Datalog program $\Pi^{WC}$ not as
-a single stream of fact instances (as for C-Datalog), but in three
+We consider the entailment of the WC-Datalog program $\Pi^{C,r}$ not
+as a single stream of fact instances (as for C-Datalog), but in three
 parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle$
@@
 F^0\rangle$, a stream $P$ of fact instances entailed by the fact
-instances in $F^0$ and $S$, and a stream $N$ of fact instance negations:
-$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
-and $N$ can be considered to be the continuous reasoning equivalents
-of the streams produced by the \emph{IStream} and \emph{DStream}
-operators respectively of the \emph{CQL} continuous query language
-\citep{CQL}, which create streams based on the additions to and
-deletions from a relation.
+instances in $F^0$ and $S$, and a stream $N$ of fact instance
+negations: $\langle R, F^0, S, r \rangle \models \langle E^0, P, N
+\rangle$. $P$ and $N$ can be considered to be the continuous reasoning
+equivalents of the streams produced by the \emph{IStream} and
+\emph{DStream} operators respectively of the \emph{CQL} continuous
+query language \citep{CQL}, which create streams based on the
+additions to and deletions from a relation. Note that $P$ is an
+in-order stream whose members are a subset of those in $E^S$ for an
+equivalent C-Datalog programe, while $N$ may be an out-of-order
+stream.
 
+%%%FIXME
 In order to
@@
 
-\begin{definition}[Instanteous State of a C-Datalog program]
+%%%FIXME
+From this description of instance entailment,
+Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be
+derived:
+
+\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
+  If two sliding windows over the same stream $S$ cover non-overlapping time periods,
+then a set $E^T_1$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
+is necessarily disjoint from any set $E^T_2$ of fact instances entailed by
+$\langle R, W^{S,y}_{i-z} \rangle$, where ${z \geq x}$.
+% This is because the timestamps of the instances ${E^T1}$ are a subset of those of
+% the instances $W^{S,x}_i$, the timestamps of the instances ${E^T2}$ are a subset
+% of those of the instances $W^{S,y}_{i-z}$, fact instances are distinct by timestamp and
+% ${\{ t : f | t \in W^{S,x}_i\} \cap \{ t : f | t \in W^{S,y}_{i-z}\} = \varnothing}$.
+\[
+\{ e \mid \langle R , W^{S,x}_i \rangle \models e \} \cap \{ e \mid \langle R , W^{S,y}_{i-z} \rangle \models e \} = \varnothing
+\]
+\end{axiom}
+
+\begin{definition}[Instanteous State of a WC-Datalog program]
 \label{def:continuous datalog: CDPt}
 
-We can reduce a C-Datalog program $\Pi^C = \langle R, F^0, S, r
+We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
 \rangle$ to its TA-Datalog equivalent by taking the window $W^{S,r}_i$
@@
 \mathbb{T}$; we refer to this as the {\em instantaneous state} of the
-C-Datalog program, written as $\Pi^C_i = \langle R, F^0 \cup W^{s,r}_i
+WC-Datalog program, written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i
 \rangle$. As previously shown in \refdef{continuous datalog: ta
@@
 
-The instantaneous entailments of a $\Pi^C_i$ are the union of the set of
-persistent entailments of ${\langle R, F^0 \rangle \vDash E^0}$ (which are
-guaranteed to be entailed by every state, as positive Datalog is monotonic),
-and the set $E^T_{i}$ of transient entailments whose justifications are at
-least partially contained in $W^{S,r}_i$.
-\begin{equation*}
-\Pi^C_i = \langle R, F^0 \cup W^{S,r}_{i} \rangle \vDash E^0 \cup E^T_{i}
-\end{equation*}
-$E^T_i$ is defined more formally as
-${E^T_{i} = \{ x : \left( \Pi^C_i \vdash x \right) \wedge
-\left( just(\Pi^C_i,x) \cap W^{S,r}_i \neq \varnothing \right) \}}$.
+The instantaneous entailments of a $\Pi^{C,r}_i$ are the union of the set
+of persistent entailments of ${\langle R, F^0 \rangle \vDash E^0}$
+(which are guaranteed to be entailed by every state, as positive
+Datalog is monotonic), and the set $E^T_{i}$ of transient entailments
+whose justifications are at least partially contained in
+$W^{S,r}_i$. We define $E^T_i$ as:
+\[
+E^T_i = \{ x \mid \Pi^{WC}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
+\]
+\noindent where $J^T$ is a justification for the entailment of x by $\Pi^{WC}_i$.
+
 By this definition and Axiom~\ref{axiom:continuous datalog: entailment disjointness},
 $E^0$ is guaranteed to be disjoint from every $E^T_i$
-for every state $\Pi^C_i$ of every Continuous Datalog Program $\Pi^C$.
+for every state $\Pi^{C,r}_i$ of every WC-Datalog program $\Pi^{C,r}$.
 \end{definition}
@@
 The following axioms define the semantics of the streams of
-entailments $P$ and and negations $N$ of a $\Pi^C$ in terms of the
+entailments $P$ and and negations $N$ of a $\Pi^{C,r}$ in terms of the
 instantaneous entailments of the sequence of states of that continuous
@@
 \label{axiom:continuous datalog: positive window increment}
-The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
-the stream $P$ entailed by a $\Pi^C$ is defined as the set of fact instances
-that are entailed by the state of the $\Pi^C$ at instant $i$ that were
-not entailed by the state of $\Pi^C$ at the previous instant ${i-1}$.
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $P$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that are entailed by the state
+$\Pi^{C,r}_i$ that were not entailed by the state $\Pi^{C,r}_{i-1}$.
 \begin{equation*}
@@
 \end{axiom}
+
 \begin{axiom}
 \label{axiom:continuous datalog: negative window increment}
-The window of range 1 at any given instant ${i \in \mathbb{N}}$ over
-the stream $N$ entailed by a $\Pi^C$ is defined as the set of fact instances
-that were entailed by the state of the $\Pi^C$ at the previous instant
-${i-1}$ that are no longer entailed by the state of $\Pi^C$ at instant $i$.
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $N$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that were entailed by the state
+$\Pi^{C,r}_{i-1}$ that are no longer entailed by the state
+$\Pi^{C,r}_i$.
 \begin{equation*}
@@
 were entailed.  As such, a window of any range ${x \in \mathbb{N}}$
-over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$
+over the stream of entailments $P$ at any given instant ${i \in \mathbb{T}}$
 is disjoint from a window of any range ${y \in \mathbb{N}}$ over the
@@
 
-From this, we can see that each entailed fact instance ${f|t}$ in some
-$E^T_{i}$, of the form ${f|t = \langle s, p, o \rangle | t}$, has a
-timestamp $t$ that is the real-time associated with the state
-$\Pi^C_{i_{e}}$ in which its particular justification for ${f|t}$ is
-first complete. As such, the stream of entailed fact instances $P$ of
-a $\Pi^C$ takes precisely the same form as the streams of facts already
+From this, we can see that each entailed fact instance $\langle f, e
+\rangle$ in some $E^T_{i}$ has an instant $e$ that is associated with
+the state $\Pi^{C,r}_e$ in which a justification for $\langle f, e
+\rangle$ is first completed.
+
+As such, the stream of entailed fact instances $P$ of
+a $\Pi^{C,r}$ takes precisely the same form as the streams of facts already
 described, thus producing a concise stream of entailments that may be

### 2015-07-14 content/conclusions.tex
tidied up the conclusions on C-Datalog, done most of conclusions for CoDeR

--- 2015-06-10 content/conclusions.tex
@@
 \begin{nestedsection}{Conclusions}{conclusions}
+	We have presented C-Datalog, a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
+	In this way it differs from other temporal interpretations of Datalog \citep{TODO} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	C-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of CQL \citep{CQL}.
+	In addition to this, however, it also supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program via the application of sliding windows, leading to the well-founded definition of a \emph{valid window}.
+
+	We have also presented CoDeR, a processing model for Continuous Deductive Reasoning over RDF that expresses the semantics of C-Datalog.
+	This processing model is composed of a stream-based data model with a dual semantics and a stream-to-stream extension of a semantic interpretation of the relational algebra.
+	The data model is composed of streams of RDF graphs annotated both with the time at which they became true and the time at which they will cease to be true, with the ordering of graphs by the former not necessarily being the same as the ordering by the latter, hence providing a dual semantics for the stream.
+	As such, the sliding window inherent in a continuous program may be represented in the annotation of each graph in the streamed input, rather than sliding window operators in the algebra.
+	Instead, the algebra makes use of \emph{valid window} operators to maintain the set of graphs that are both statically applicable at a given point in the processing \emph{and} considered to be true at a given moment.
+	It also de-constructs the conventional window-join operator of stream-to-stream algebras for continuous querying into the union of the outputs of complementary State Modules \citep{SteMs}, thereby supporting greater variety in the plans that express a given continuous program.
+
 	We are more semantically sound than other RDF Stream Processing offerings and more ``continuous'' than other continuous reasoning offerings, without sacrificing processing efficiency.

### 2015-07-14 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
fixed application notation and bracketing error in inductive proof

@@
 			\mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{i}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
-			& = \mathop{\cup}^{x-1}_{d=0} W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} && \text{, Axiom~\ref{axiom:continuous datalog: window composition}} \\
+			& = \mathop{\cup}^{x-1}_{d=0} W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} && \text{by Axiom~\ref{axiom:continuous datalog: window composition}} \\
 			& = \mathop{\cup}^{x-1}_{d=0} \left( W^{P,1}_{i-d} \setminus \mathop{\cup}^{y-1}_{j=0} W^{N,1}_{i-j} \right) \\
 			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{y-1}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
-			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{, Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } y \geq x
+			& = \mathop{\cup}^{x-1}_{d=0}\mathop{\cap}^{d}_{j=0} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{by Axiom~\ref{axiom:continuous datalog: entailment precedes negation}, } y \geq x
 		\end{align*}
@@
 			E^T_{i} \setminus \mathop{\cap}^{1}_{d=1} E^T_{i-d} & = W^{P,1}_{i} \setminus W^{N,y}_{i} \\
-			& = W^{P,1}_{i} \setminus W^{N,1}_{i} && \text{, \textbf{General Case}} \\
-			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) && \text{, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
+			& = W^{P,1}_{i} \setminus W^{N,1}_{i} && \text{by \textbf{General Case}} \\
+			& = \left( E^T_{i} \setminus E^T_{i-1} \right) \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) && \text{by Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
 			& = E^T_{i} \setminus \left( E^T_{i-1} \setminus \left( E^T_{i-1} \setminus E^T_{i} \right) \right) \\
@@
 		\begin{align*}
-			E^t_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
-			& = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{, \textbf{General Case}} \\
+			E^T_{i+1} \setminus \mathop{\cap}^{k+1}_{d=1} E^T_{i+1-d} & = W^{P,k+1}_{i+1} \setminus W^{P,k+1}_{i+1} \\
+			& = \mathop{\cup}^{k-1}_{d=-1}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) && \text{by \textbf{General Case}} \\
 			& = \mathop{\cap}^{-1}_{j=-1} \left( W^{P,1}_{i-j} \setminus W^{N,1}_{i-j} \right) \cup \mathop{\cup}^{k-1}_{d=0}\mathop{\cap}^{d}_{j=-1} \left( W^{P,1}_{i-d} \setminus W^{N,1}_{i-j} \right) \\
@@
 			& = \left( W^{P,1}_{i+1} \cup \left( W^{P,k}_{i} \setminus W^{N,k}_{i} \right) \right) \setminus W^{N,1}_{i+1} \\
-			& = \left( \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) && \text{, \textbf{Inductive Step}, Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}} \\
+			& = \left( \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) &&
+				\begin{aligned}
+					&\text{by \textbf{Inductive Step}} \\
+					&\text{and Axioms~\ref{axiom:continuous datalog: positive window increment}~and~\ref{axiom:continuous datalog: negative window increment}}
+				\end{aligned} \\
 			& = \left( E^T_{i+1} \setminus E^T_{i} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \setminus \left( E^T_{i} \setminus E^T_{i+1} \right) \right) \\
@@
 			& = \left( E^T_{i+1} \setminus E^T_{i} \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\
-			& = \left( E^T_{i+1} \cap \left( E^T_{i} \right^C \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right)\\
+			& = \left( E^T_{i+1} \cap \left( E^T_{i} \right)^C \right) \cup \left( E^T_{i+1} \cap \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right)\\
 			& = E^T_{i+1} \cap \left( \left( E^T_{i} \right)^C \cup \left( E^T_{i} \setminus \mathop{\cap}^{k}_{d=1} E^T_{i-d} \right) \right) \\

### 2015-07-14 content/contribution/continuous-datalog.tex
Fixed defn of streams

@@
 
-A {\em stream} $S$ is an unbounded, ordered sequence of fact
-instances. We indicate that a fact instance $\langle f, t\rangle$
-appears in $S$ by writing $\langle f, t \rangle \in S$. We indicate
-that (in a given stream) fact instance $f_i$ appears before fact
-instance $f_j$ by writing $f_i \prec f_j$.
+A {\em stream} $S$ is an unbounded sequence of fact instances. We
+define a sequence as a mapping from fact instances to instants, where
+the instant records the time at which the fact instance appears on the
+stream. This has the consequence that a particular fact instance
+(i.e. a fact accompanied by a particular instant that is its time
+annotation) may appear only once in a given stream. For convenience,
+we indicate that a fact instance $\langle f, t\rangle$ appears in $S$
+by writing $\langle f, t \rangle \in S$. We indicate that a fact
+instance $f_i$ appears strictly before a fact instance $f_j$ on stream
+$S$ (i.e. $S(f_i) < S(f_j)$) by writing $f_i \prec f_j$.
 
@@
 $\langle f_j, t_j\rangle$ where $\langle f_i, t_i \rangle \prec
-\langle f_j, t_j \rangle$, $t_i \leqslant t_j$; we refer to such
+\langle f_j, t_j \rangle$, $t_i < t_j$; we refer to such
 streams as {\em in-order streams}, and to streams whose fact instances
@@
 
-A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an instant
-${i \in \mathbb{T}}$ is the \emph{set} of fact instances that occur in
-$S$ with instances in the interval ${(i-x,i]}$, written
-  ${W^{S,x}_{i}}$, and defined as follows:
-\[ W^{S,x}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i-x) < t \leqslant i \}\]  
+A window of range ${r \in \mathbb{N}}$ over a stream $S$ at an instant
+${i \in \mathbb{T}}$ is the \emph{set} of fact instances that appear in
+$S$ in the interval ${(i-r,i]}$, written ${W^{S,r}_{i}}$, and defined as follows:
+\[
+W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
+\]  
 \end{definition}
@@
 range of the window over $S$, which effectively specifies that the
-fact instances from $S$ are to contribute to the entailments of $\Pi^{WC}$
+fact instances from $S$ are to contribute to the entailments of $\Pi^{C,r}$
 from their instant for a duration of $r$.
@@
 as a single stream of fact instances (as for C-Datalog), but in three
-parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle$
+parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
 that is the entailment of the TA-Datalog program $\langle R,
@@
 stream.
-
-%%%FIXME
-In order to
-
 \end{definition}
 
-%%%FIXME
-From this description of instance entailment,
-Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be
-derived:
+From the definition of entailment for TA-Datalog in \refdef{continuous
+  datalog: ta datalog program} and the definition of windows in
+\refdef{continuous datalog: window}, we can derive the following:
 
 \begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-  If two sliding windows over the same stream $S$ cover non-overlapping time periods,
-then a set $E^T_1$ of fact instances entailed by ${\langle R, W^{S,x}_i \rangle}$
-is necessarily disjoint from any set $E^T_2$ of fact instances entailed by
-$\langle R, W^{S,y}_{i-z} \rangle$, where ${z \geq x}$.
-% This is because the timestamps of the instances ${E^T1}$ are a subset of those of
-% the instances $W^{S,x}_i$, the timestamps of the instances ${E^T2}$ are a subset
-% of those of the instances $W^{S,y}_{i-z}$, fact instances are distinct by timestamp and
-% ${\{ t : f | t \in W^{S,x}_i\} \cap \{ t : f | t \in W^{S,y}_{i-z}\} = \varnothing}$.
+
+If two sliding windows over the same in-order stream $S$ cover
+non-overlapping time periods, then the set of fact instances entailed
+by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
+set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
+where ${z \geq x}$.
 \[
@@
 \[
-E^T_i = \{ x \mid \Pi^{WC}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
+E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
 \]
-\noindent where $J^T$ is a justification for the entailment of x by $\Pi^{WC}_i$.
 
-By this definition and Axiom~\ref{axiom:continuous datalog: entailment disjointness},
-$E^0$ is guaranteed to be disjoint from every $E^T_i$
-for every state $\Pi^{C,r}_i$ of every WC-Datalog program $\Pi^{C,r}$.
+\noindent where $J^T$ is a justification for the entailment of x by
+$\Pi^{C,r}_i$. By this definition and Axiom~\ref{axiom:continuous
+  datalog: entailment disjointness}, $E^0$ is guaranteed to be
+disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
+WC-Datalog program $\Pi^{C,r}$.
 \end{definition}
@@
 \label{axiom:continuous datalog: entailment precedes negation}
-Fact instances may only be negated by some part of their justification leaving
-the sliding window $W^{S,r}_{i}$, this necessarily being after they
-were entailed.  As such, a window of any range ${x \in \mathbb{N}}$
-over the stream of entailments $P$ at any given instant ${i \in \mathbb{T}}$
-is disjoint from a window of any range ${y \in \mathbb{N}}$ over the
-stream of entailment negations $N$ at any instant before ${i - x}$.
+Fact instances may only be negated by some part of their justification
+leaving the sliding window $W^{S,r}_{i}$, this necessarily being after
+they were entailed.  As such, a window of any range ${x \in
+  \mathbb{N}}$ over the stream of entailments $P$ at any given instant
+${i \in \mathbb{T}}$ is disjoint from a window of any range ${y \in
+  \mathbb{N}}$ over the stream of entailment negations $N$ at any
+instant before ${i - x}$.
 \begin{equation*}
@@
 
-From this, we can see that each entailed fact instance $\langle f, e
-\rangle$ in some $E^T_{i}$ has an instant $e$ that is associated with
-the state $\Pi^{C,r}_e$ in which a justification for $\langle f, e
-\rangle$ is first completed.
+As noted above in \refdef{continuous datalog: CDP}, the stream of
+negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
+out-of-order stream in which fact instances appear on the stream
 
-As such, the stream of entailed fact instances $P$ of
-a $\Pi^{C,r}$ takes precisely the same form as the streams of facts already
-described, thus producing a concise stream of entailments that may be
-read by further stream processing.  In addition, the stream of negations
-$N$ of a $\Pi^C$ also takes this form, but differs from the streams
+also takes this form, but differs from the streams
 discussed thus far in that it is not ordered by the entailment/assertion

### 2015-07-14 content/contribution/continuous-datalog.tex
A few changes

--- 2015-07-14 content/contribution/continuous-datalog.tex
@@
 a fact instance $e$ is entailed by a program $\Pi^C$ iff each
-interpretation satisfying every clause in $\Pi^C$. However, we take
-the entailment of the program $\Pi^C$ to be an in-order stream $E^S$,
-where the order in which entailed fact instances appear in $E^S$
-depends on the order in which the fact instances in their
-justifications appear in $S$. Given that that instant of an entailed
+interpretation satisfying every clause in $\Pi^C$ also satisfies $e$.
+However, we take the entailment of the program $\Pi^C$ to be an in-order
+stream $E^S$, where the order in which entailed fact instances appear in
+$E^S$ depends on the order in which the fact instances in their
+justifications appear in $S$. Given that that the instant of an entailed
 fact instance is the maximum of the instants in the fact instances
@@
 as a single stream of fact instances (as for C-Datalog), but in three
-parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle$
+parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
 that is the entailment of the TA-Datalog program $\langle R,
-F^0\rangle$, a stream $P$ of fact instances entailed by the fact
-instances in $F^0$ and $S$, and a stream $N$ of fact instance
-negations: $\langle R, F^0, S, r \rangle \models \langle E^0, P, N
-\rangle$. $P$ and $N$ can be considered to be the continuous reasoning
-equivalents of the streams produced by the \emph{IStream} and
-\emph{DStream} operators respectively of the \emph{CQL} continuous
-query language \citep{CQL}, which create streams based on the
-additions to and deletions from a relation. Note that $P$ is an
-in-order stream whose members are a subset of those in $E^S$ for an
-equivalent C-Datalog programe, while $N$ may be an out-of-order
-stream.
+F^0\rangle$; a stream $P$ of fact instances entailed by the rules $R$ and
+the fact instances in $F^0$ and $S$, ordered by the instant at which they
+became justified (i.e. their instants); and a stream $N$ of those same fact
+instances, ordered by the instant at which they ceased to be justified:
+$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
+and $N$ can be considered to be the continuous reasoning equivalents of
+the streams produced by the \emph{IStream} and\emph{DStream} operators
+respectively of the \emph{CQL} continuous query language \citep{CQL},
+which create streams based on the additions to and deletions from a
+relation. Note that $P$ is an in-order stream whose members are a subset
+of those in $E^S$ for an equivalent C-Datalog program, while $N$ may be
+an out-of-order stream.
 
@@
 We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
-\rangle$ to its TA-Datalog equivalent by taking the window $W^{S,r}_i$
-over stream $S$ of range $r \in \mathbb{N}$ at instant $i \in
-\mathbb{T}$; we refer to this as the {\em instantaneous state} of the
-WC-Datalog program, written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i
-\rangle$. As previously shown in \refdef{continuous datalog: ta
-  datalog program}, this may then be further reduced to the Datalog
-program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.
+\rangle$ at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
+taking the window $W^{S,r}_i$ over stream $S$ of range $r \in \mathbb{N}$;
+we refer to this as the {\em instantaneous state} of the WC-Datalog program,
+written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle$. As
+previously shown in \refdef{continuous datalog: ta datalog program}, this
+may then be further reduced to the Datalog program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.
 
-The instantaneous entailments of a $\Pi^{C,r}_i$ are the union of the set
-of persistent entailments of ${\langle R, F^0 \rangle \vDash E^0}$
-(which are guaranteed to be entailed by every state, as positive
-Datalog is monotonic), and the set $E^T_{i}$ of transient entailments
-whose justifications are at least partially contained in
-$W^{S,r}_i$. We define $E^T_i$ as:
+The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
+semantically distinct portions: the \emph{persistent} portion is the
+entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
+program, which is guaranteed to be part of the entailment of every state
+due to the monotonicity of positive Datalog; the \emph{transitive} portion
+is the set $E^T_{i}$ of entailed fact instances whose justifications are at
+least partially contained in $W^{S,r}_i$. We define $E^T_i$ as:
 \[
@@
 \]
-\noindent where $J^T$ is a justification for the entailment of x by $\Pi^{WC}_i$.
+where $J^T$ is a justification for the entailment of $x$ by $\Pi^{WC}_i$.
 

### 2015-07-14 content/contribution/continuous-datalog.tex

--- 2015-07-14 content/contribution/continuous-datalog.tex
@@
 that is the entailment of the TA-Datalog program $\langle R,
-F^0\rangle$, a stream $P$ of fact instances entailed by the fact
-instances in $F^0$ and $S$, and a stream $N$ of fact instance
-negations: $\langle R, F^0, S, r \rangle \models \langle E^0, P, N
-\rangle$. $P$ and $N$ can be considered to be the continuous reasoning
-equivalents of the streams produced by the \emph{IStream} and
-\emph{DStream} operators respectively of the \emph{CQL} continuous
-query language \citep{CQL}, which create streams based on the
-additions to and deletions from a relation. Note that $P$ is an
-in-order stream whose members are a subset of those in $E^S$ for an
-equivalent C-Datalog programe, while $N$ may be an out-of-order
-stream.
+F^0\rangle$; a stream $P$ of fact instances entailed by the rules $R$ and
+the fact instances in $F^0$ and $S$, ordered by the instant at which they
+became justified (i.e. their instants); and a stream $N$ of those same fact
+instances, ordered by the instant at which they ceased to be justified:
+$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
+and $N$ can be considered to be the continuous reasoning equivalents of
+the streams produced by the \emph{IStream} and\emph{DStream} operators
+respectively of the \emph{CQL} continuous query language \citep{CQL},
+which create streams based on the additions to and deletions from a
+relation. Note that $P$ is an in-order stream whose members are a subset
+of those in $E^S$ for an equivalent C-Datalog program, while $N$ may be
+an out-of-order stream.
+
+%%%FIXME
+In order to
+
 \end{definition}

### 2015-07-14 content/contribution/continuous-datalog.tex
Hopefully final changes to this section.

@@
 and a set of entailed fact instances $E^T$ where $\langle R, F^T \rangle \models E^T$, the set of fact instances $J^T \subseteq F^T$ is a justification for
-$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime} \not\models E^T$.
+$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime}\rangle \not\models E^T$.
 
@@
 interpretation satisfying every clause in $\Pi^C$ also satisfies $e$.
-However, we take the entailment of the program $\Pi^C$ to be an in-order
-stream $E^S$, where the order in which entailed fact instances appear in
-$E^S$ depends on the order in which the fact instances in their
-justifications appear in $S$. Given that that the instant of an entailed
-fact instance is the maximum of the instants in the fact instances
-that form the justification for that entailed fact instance (as in
-\refdef{continuous datalog: instance justification}), $E^S$ will also
-be an in-order stream.
+However, we take the {\em entailment of the C-Datalog program} $\Pi^C$
+to be an in-order stream $E^S$, where the order in which entailed fact
+instances appear in $E^S$ depends on the order in which the fact
+instances in their justifications appear in $S$. Given that that the
+instant of an entailed fact instance is the maximum of the instants in
+the fact instances that form the justification for that entailed fact
+instance (as in \refdef{continuous datalog: instance justification}),
+$E^S$ will also be an in-order stream.
 
@@
 negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
-out-of-order stream in which fact instances appear on the stream
-
-also takes this form, but differs from the streams
-discussed thus far in that it is not ordered by the entailment/assertion
-time of the entailed fact instances, but by the real-time associated
-with the first state $\Pi^C_{i_{n}}$ in which their justification
-\emph{ceases} to be complete, i.e. ${just(\Pi^C_{i_{e}}, f|t) \subseteq W^{S,r}_{i_n - 1}}$
-but ${just(\Pi^C_{i_{e}}, f|t) \not\subseteq W^{S,r}_{i_n}}$.
+out-of-order stream in which fact instances appear on the stream in
+the order in which they cease to have any complete justifications.
 
@@
 streams of entailments has been stated axiomatically as the definition
-of those streams.  However, it can be shown from these axioms, as in
-\ref{sec:semantics: proof for entailments}, that it is also possible
-to reconstruct exactly the instantaneous set $E^T_{i}$, of instances
-entailed at any given instant $i$ by any $\Pi^C = \{ R, F, S, r \}$,
-from the sliding windows of sufficient range over its entailed streams
-$P$ and $N$ at $i$:
+of those streams.
+
+However, it follows from these axioms that the instantaneous set
+$E^T_{i}$ of instances entailed at any given instant $i$ by any
+$\Pi^{C,r} = \langle R, F^0, S, r \rangle$ may be reconstructed from
+windows of sufficient range over its entailed streams $P$ and $N$ at
+$i$:
 \[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
@@
 the \emph{complete} set of entailment instances $E^T_{i}$ of the
-$\Pi^C_{i}$. In addition, the minimum range $y$ is such to achieve the
-\emph{correct} set $E^T_{i}$. As such, a \emph{valid window} that
+$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
+\emph{sound} set $E^T_{i}$. As such, a \emph{valid window} that
 contains only and exactly those entailed instances $E^T_{i}$ at every
 instant $i$ may be trivially maintained from the streams $P$ and $N$
-continuously entailed by some $\Pi^C$.
+continuously entailed by some $\Pi^{C,r}$.
 
@@
 
-To sum up, a time-annotated Datalog program
-${\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle}$ may be derived from
-a continuous Datalog program ${\Pi^C = \langle R, F, S, r \rangle}$ by
-applying a sliding window $W^{S,r}_i$ of range $r$ over the stream $S$ at
-instant $i$ and instantiating the facts in $F$ as $F^0$. At the same time,
-the entailments of a ${\Pi^C_i \vDash E^0 \cup E^T_i}$ may be derived from
-the entailments of the ${\Pi^C \vDash \langle E, P, N \rangle}$ as the union
-of the entailments of ${\langle R, F^0 \rangle \vDash E^0}$ and the set
-difference of the windows $W^{P,r}_i$ and $W^{N,r}_i$ of range $r$ over
-streams $P$ and $N$ at instant $i$.
-Furthermore, a conventional Datalog program ${\Pi = \langle R, F \rangle}$
-may be derived from a time-annotated Datalog program $\Pi^C_i$ by
-${strip}$ping the time annotations from the set of facts
-${F^0 \cup W^{S,r}_i}$, as the entailments of ${\Pi \vDash E}$ may be
-derived from the entailments of $\Pi^C_i$ by $strip$ping them.
-This yields the commutative diagram in \reffig{CD-TD-D}.
-\begin{figure}
+In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
+F^0, S \rangle$, we can apply to it a window of range $r$ to yield a
+Windowed Continuous Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
+\rangle$. We can evaluate this window at an instant $i \in \mathbb{T}$
+to yield a Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup
+W^{S,r}_i \rangle$, from we which can then strip the time annotations
+to yield a conventional Datalog program $\Pi = \langle R, F
+\rangle$. The relationship between these *-Datalog programs and their
+entailments is expressed in the commutative diagram in
+\reffig{CD-TD-D}.
+
+\begin{figure*}
 \centering
@@
 \begin{CD}
-	{\Pi^C = \langle R, F, S, r \rangle} @>\vDash>> {\langle E, P, N \rangle} \\
-	@V{window}VV @VV{window}V \\
-	{\Pi^C_i = \langle R, F^0 \cup W^{S,r}_i \rangle} @>\vDash>> {E^0 \cup E^T_i} \\
+	{\Pi^C = \langle R, F, S \rangle} @>\models>> {E^S} \\
+	@V{add\ window}VV \\
+        {\Pi^{C,r} = \langle R, F^0, S, r \rangle} @>\models>> {\langle cons^T(\langle R, F^0\rangle), P, N \rangle} \\
+        @V{evaluate\ window}VV @VV{evaluate\ window}V \\
+        {\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle} @>\models>> {cons^T(\langle R, F^0 \rangle) \cup (W^{P,r}_i \setminus W^{N,r}_i)} \\
+	@VVV @VVV \\
+        {\Pi^T = \langle R, F^T \rangle } @>\models>> {cons^T(\Pi^T)} \\
 	@V{strip}VV @VV{strip}V \\
-	{\Pi = \langle R, F \cup W_i \rangle} @>\vDash>> {E \cup E_i}
+        {\Pi = \langle R, F \rangle } @>\models>> {cons(\Pi)}
 \end{CD}
@@
 \labelfig{CD-TD-D}
-\end{figure}
+\end{figure*}
 \end{nestedsection}

### 2015-07-14 content/contribution/CoDeR.tex
Cleaning up §3, minor changes elsewhere.

@@
 	CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
-	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^C = \langle R,F,S,r \rangle \vDash \langle E,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
+	It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^{C,r} = \langle R,F^0,S,r \rangle \models \langle E^0,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
 	It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
@@
 
-		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^C$ expressed by a CoDeR-based system.
+		In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^{C,r}$ expressed by a CoDeR-based system.
 		These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
@@
 
-		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^C = \langle R, F, S, r \rangle}$ expressed.
-		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^C$.
+		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
+		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$.
 		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
-		\reffig{intersected-intervals} provides an example of this for some ${\Pi^C = \langle \{c\,\text{:-}\,a,b\,.\},\{\},(\{a\}|1,\{b\}|2),3\}}$ over instants ${i = 1 \dots 4}$.
+		\reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
 		\begin{figure}
 			\centering
-			\includegraphics[width=0.45\textwidth]{intersected-intervals}
+			\includegraphics[width=0.25\textwidth]{intersected-intervals}
 			\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
@@
 		\begin{align*}
-			\Pi^C_{1} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \} \rangle & \vDash & \{ \} \\
-			\Pi^C_{2} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
-			\Pi^C_{3} & = \langle \{c\,\text{:-}\,a,b.\}, \{ a|1. \, b|2. \} \rangle & \vDash & \{ c|2. \} \\
-			\Pi^C_{4} & = \langle \{c\,\text{:-}\,a,b.\}, \{ b|2. \} & \vDash & \{ \}
+			\Pi^{C,r}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
+			\Pi^{C,r}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+			\Pi^{C,r}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+			\Pi^{C,r}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
 		\end{align*}
-		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^C$.
-		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^C$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
+		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
+		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
 		As such, the annotation of streamed RDF graphs with the interval $(i_e,i_n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
@@
 		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
-		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^C$.
+		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
 		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
@@
 		As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
-		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^C = \{ \{b\,\text{:-}\,a.\}, \{\}, S, r \}}$, where $a$ and $b$ are RDF triples.
+		For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^{C,r} = \langle \{b \Leftarrow a\}, \{\}, S, r \rangle}$, where $a$ and $b$ are RDF triples.
 		\begin{equation}\labeleqn{basic-example-plan}
@@
 				\end{figure}
-				Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^C$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
+				Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^{C,r}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
 				Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
@@
 		Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
-		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^C = \{R,F,S,r\}}$.
+		as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^{C,r} = \langle R,F^0,S,r\rangle}$.
 		It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
-		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^C$ at a given instant will be persisted within that system at that instant,
+		It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
 		i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.

### 2015-07-14 content/evaluation.tex
Cleaning up §3, minor changes elsewhere.

@@
 		Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
-		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating ETALIS to use \emph{seq} instead of \emph{And}.
+		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating Etalis to use \emph{seq} instead of \emph{And}.
 		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time- and memory-efficiency of the three systems.
@@
 			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
-			\caption{A comparison of the times taken by the systems R4, Sparkwave and ETALIS to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
+			\caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
 			\labelfig{all-systems-varying-windows}
@@
 			\includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
-			\caption{A comparison of the memory consumed by the systems R4, Sparkwave and ETALIS throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
+			\caption{A comparison of the memory consumed by the systems R4, Sparkwave and Etalis throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
 			\labelfig{all-systems-memory-consumption}

### 2015-07-14 content/personal-details-ieee.tex
Added the CORRECT article class, and set up the template correctly apart from the problem listed below.  Also fixed the details for Rehab in all the personal-details-*.tex
PROBLEMS
- inductive step still overflows both column and page
- abstract or the rest of the frontmatter does not format properly in twocolumn mode
Removed middle initial
Cleaning up §3, minor changes elsewhere.

@@
 %\keywords   {Deductive Reasoning, Streamed Data, Semantic Data, Continuous Processing, Plan Optimisation}
-\author    {\IEEEauthorblockN{David L. Monks, Rehab ???, Dr. Nicholas M. Gibbins}
+\author    {\IEEEauthorblockN{David L. Monks, Rehab Albeladi, Nicholas Gibbins}
             \IEEEauthorblockA{
-                              Web and Internet Science\\
                               Electronics and Computer Science\\
-                              Physical Sciences and Engineering\\
                               University of Southampton\\       
-                              \{dm11g08,r???,nmg\}@ecs.soton.ac.uk
+                              Southampton, SO17 1BJ\\
+                              United Kingdom \\
+                              \{dm11g08,raab1g09,nmg\}@ecs.soton.ac.uk
                               }
@@
 \date{\today}
+
+\maketitle
+
+\begin{abstract}
+	\input{abstract.tex}
+\end{abstract}

### 2015-07-14 content/personal-details-ios.tex
Cleaning up §3, minor changes elsewhere.

@@
 	\address{
-	          Department of Electronics and Computer Science,\\
-	          Faculty of Physical Sciences and Engineering,\\
+	          Electronics and Computer Science,\\
 	          University of Southampton,\\
-	          University Rd., Southampton, SO17 1BJ,\\
+	          Southampton, SO17 1BJ,\\
 	          United Kingdom\\

### 2015-07-14 content/conclusions.tex

--- 2015-06-10 content/conclusions.tex
@@
 \begin{nestedsection}{Conclusions}{conclusions}
-	We are more semantically sound than other RDF Stream Processing offerings and more ``continuous'' than other continuous reasoning offerings, without sacrificing processing efficiency.
+	We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
+	In this way it differs from other temporal interpretations of Datalog \citep{TODO} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.
+	In addition to this, however, it also supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program via the application of windows to the concise streams, leading to the well-founded definition of a \emph{valid window} over the combined streams $P$ and $N$, being the outputs of the reasoning equivalents of the \emph{IStream} and \emph{DStream}.
+	Thus WC-Datalog represents a well-founded semantics not only for continuous reasoning, but for coherent and concise streamed entailments of continuous programs, suitable for consumption in further stream processing.
 
-%	\subimport{conclusions/}{future-work.tex}
+	In order to express this semantics we have developed CoDeR, a processing model for Continuous Deductive Reasoning over RDF.
+	This processing model is composed of a stream-based data model with a dual semantics and a stream-to-stream extension of a semantic interpretation of the relational algebra.
+	The data model is composed of streams of RDF graphs annotated both with the time at which they became true and the time at which they will cease to be true, with the ordering of graphs by the former being the sequence of facts in $P$ and the ordering by the latter being the sequence of facts in $N$, hence providing a dual semantics for the stream.
+	As such, the window inherent in a continuous program may be represented in the annotation of each graph in the streamed input, rather than ranged window operators in the algebra.
+	Instead, the algebra makes use of \emph{valid window} operators to maintain the set of graphs that are both considered to be true at a given moment \emph{and} applicable at a given point in the processing.
+	It also de-constructs the conventional window-join operator of stream-to-stream algebras for continuous querying into the union of the outputs of complementary State Modules (\emph{SteMs}) \citep{SteMs}, to which it is logically and functionally equivalent, thereby supporting greater variety in the possible plans that express a given continuous program.
+
+	Finally, we have implemented the CoDeR model in the synchronous RDF stream-reasoning system R4.
+	R4 utilises a variant of the Rete pattern matching algorithm \citep{forgy79} to produce processing plans for sets of rules expressed in RIF-Core \citep{w3crifcore}, the W3C recommendation for expressing sets of rules with the semantics of positive Datalog and the rule language used to express the set of entailment rules of OWL 2 RL \citep{w3cowl2profiles}.
+	R4 is sound and complete with respect to the semantics of WC-Datalog by virtue of implementing CoDeR.
+	Given that many state-of-the-art stream-querying and continuous-reasoning systems have varying semantics, and so the expected results of each system will vary for the same query definition, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and ETALIS \citep{EP-SPARQL}.
+	We have demonstrated that R4 is comparable in both respects to these RDF stream-querying systems when applying low-range windows to high-throughput streams, despite Sparkwave and ETALIS supporting more limited/less explicitly well-founded reasoning than that of R4.
+
+	Moving forward, we plan to improve the time-efficiency of R4 by investigating more efficient join algorithms within its implementations of SteMs, as well as improving the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
+reducing the latency from receipt of some new graph to the production of entailments justified in part by that new graph.
 \end{nestedsection}

### 2015-07-15 content/conclusions.tex
Minor fixes to bibliography, fixed missing citation in conclusion. Minor stylistic tweaks to template.

@@
 	We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
-	In this way it differs from other temporal interpretations of Datalog \citep{TODO} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	In this way it differs from other temporal interpretations of Datalog \citep{OrgunWadge92,Tuzhilin93} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing and casting a continuous program as a series of discrete states.
 	WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.

### 2015-07-15 content/conclusions.tex
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

--- 2015-07-14 content/conclusions.tex
@@
 	We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
-	In this way it differs from other temporal interpretations of Datalog \citep{TODO} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	In this way it differs from other temporal interpretations of Datalog \citep{orgun88chronolog} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
 	WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.
@@
 	R4 is sound and complete with respect to the semantics of WC-Datalog by virtue of implementing CoDeR.
-	Given that many state-of-the-art stream-querying and continuous-reasoning systems have varying semantics, and so the expected results of each system will vary for the same query definition, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and ETALIS \citep{EP-SPARQL}.
-	We have demonstrated that R4 is comparable in both respects to these RDF stream-querying systems when applying low-range windows to high-throughput streams, despite Sparkwave and ETALIS supporting more limited/less explicitly well-founded reasoning than that of R4.
+	Given that many state-of-the-art stream-querying and continuous-reasoning systems have varying semantics, and so the expected results of each system will vary for the same query definition, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and Etalis \citep{EP-SPARQL}.
+	We have demonstrated that R4 is comparable in both respects to these RDF stream-querying systems when applying low-range windows to high-throughput streams, despite Sparkwave and Etalis supporting more limited/less explicitly well-founded reasoning than that of R4.
 
-	Moving forward, we plan to improve the time-efficiency of R4 by investigating more efficient join algorithms within its implementations of SteMs, as well as improving the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
-reducing the latency from receipt of some new graph to the production of entailments justified in part by that new graph.
+	Moving forward, we plan to improve the time-efficiency of R4 by investigating more efficient join algorithms within its implementations of SteMs.
+	At a higher level, we also plan to investigate alternate planing algorithms to improve the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
+reduce the latency from receipt of some new graph to the production of entailments justified in part by that new graph.
+	These primarily focus on the use of SteMs outside of the composition of window-joins, either taking an adaptive approach \citep{eddies,CACQ,CQELS} or utilising the wider array of static plans such an operator affords.
 \end{nestedsection}

### 2015-07-15 content/contribution/CoDeR.tex
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

@@
 		These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
-		an instance of an RDF graph is an RDF graph annotated with a time interval: ${\langle \{(s,p,o),\dots,(s',p',o')\},i_{e},i_{n} \rangle}$.
+		an instance of an RDF graph is an RDF graph annotated with a time interval ${(e,n]}$: ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
 		As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.
 
-		The interval ${(i_e,i_n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
-		For graphs on the input stream, $i_e$ is the time at which a given graph was asserted, and ${i_n = i_e + r}$ is the time at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$.
-		For graphs on the output stream, $i_e$ is the latest time at which any part of the justification of a given graph was asserted, and $i_n$ is the earliest time at which any part of its justification will leave the sliding window.
+		The interval ${(e,n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
+		For graphs on the input stream, $e$ is the instant at which a given graph was asserted, and ${n = e + r}$ is the instant at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$ (or become ``negated'').
+		For graphs on the output stream, $e$ is the latest instant at which any part of the justification of a given graph was asserted, and $n$ is the earliest instant at which any part of its justification will leave the sliding window.
 		\reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
@@
 			\centering
-			\includegraphics[width=0.25\textwidth]{intersected-intervals}
-			\caption{Derivation of valid-time for entailed fact instance from those of its justifying fact instances.}
+			\includegraphics[width=0.3\textwidth]{intersected-intervals}
+			\caption{Derivation of valid-time for entailed fact instance from those of its justification.}
 			\labelfig{intersected-intervals}
@@
 		It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
-		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $i_{n}$ at which they \emph{ceased} to be justified by $W^{S,r}_{i_n}$, the output stream of a CoDeR-based system also expresses the stream $N$.
-		As such, the annotation of streamed RDF graphs with the interval $(i_e,i_n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
+		However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $n$ at which $W^{S,r}_{n}$ \emph{ceased} to justify them, the output stream of a CoDeR-based system also expresses the stream $N$.
+		As such, the annotation of streamed RDF graphs with the interval $(e,n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
 
@@
 		These streams are sequences of the intermediate results of processing between steps in the production of entailments.
-		As the production of entailments in CoDeR is cast as a pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
+		As the production of entailments in CoDeR is cast as an iterative pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
 		an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
-		As these instances of RDF graphs may be considered to be ``entailed'' by the presence of valid instances of their constituent RDF triples within the system, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
+		These instances of RDF graphs could be considered to be ``entailed'' by the sub-pattern of the stream on which they appear and the presence of valid instances of their constituent RDF triples within the system;
+		as such, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
 		It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
@@
 		each input/output stream is from a distinct source and uniquely identifiable by the source's IRI;
-		each inter-operator stream contains graphs matching a specific graph pattern and uniquely identifiable by that pattern.
+		each inter-operator stream contains graphs matching a specific graph pattern and are uniquely identifiable by that pattern.
 		As the same graph pattern may be a sub-pattern of many rule bodies, and so appear in multiple places in a processing plan, one inter-operator stream may be the input to multiple operators.
 		Furthermore, two or more operators may contribute to the same stream within the model, where the pattern of the stream in question is the union of the patterns of the streams produced by each of the operators in isolation.
-		In this case, each operator contributes results to the common stream as they are produced, passively interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment time $i_e$ of the constituent instances.
+		In this case, each operator contributes results to the common stream as they are produced, passively interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment instant $e$ of the constituent instances.
 		As this is the active behaviour of a stream union operator, such an operator is not needed in CoDeR.

### 2015-07-15 content/contribution/R4.tex
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

@@
-\begin{nestedsection}{R4: Rule-based Reasoning over RDF streams using Rete}{implementation}
-	As a proof of concept, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
+\begin{nestedsection}{R4 System: Rule-based Reasoning over RDF streams using Rete}{implementation}
+	As a proof of concept for the CoDeR model, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
 	Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams, constructed according to the Rete pattern matching algorithm \citep{forgy79}.
@@
 	\end{figure}
-
-	A rule document containing any number of rules is submitted to the system.
-	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
-	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
-	We also added a simple extension to RIF-Core in line with the extension described in \refsec{semantics}, enabling users to specify sliding window constraints as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
-	Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
 
@@
 
-	These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete dataflow networks.
+	A rule document containing any number of rules is submitted to the system.
+	We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
+	The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
+	We also added a simple extension to RIF-Core in line with the extension of Datalog described in \refsec{semantics}, enabling users to specify a window range as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
+	Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
+
+	These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete-like dataflow networks.
 	The optimizer employs simple known heuristics to generate a good plan.
@@
 	The optimizer instantiates a Rete network for background reasoning if required and another network for generic rules.
-	The first network feeds into the second one and the two together are ``re-entrant'', meaning that the first network takes as input the stream of entailments produced, to enable iterative inference of results such as the calculation of transitive closure.
+	The first network feeds into the second one and the two together are ``re-entrant'', meaning that the first network takes as input the stream of entailments produced by both the first and the second, to enable iterative inference of results such as the calculation of transitive closure.
 
@@
 	We chose this RDF-native approach as opposed to reusing existing technologies as in \citep{C-SPARQL,streaming-sparql} to allow full control over the low-level operators.
-	This can offer maximum optimization opportunities such as adaptively optimizing the network topologies, which is our future work.
+	This can offer maximum optimization opportunities such as adaptively optimizing the network topologies, which is part of our future work.
 
@@
 		The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, can be well fit into the stream reasoning model as it operates incrementally.
-		The RETE algorithm can process large data sets efficiently because it avoids iterating over both data elements (facts or working memory) and over the production rules.
+		The RETE algorithm can process large data sets efficiently because it avoids iterating over both data elements (facts, or ``working memory'') and over the production rules.
 		To avoid iteration over data elements, the RETE algorithm stores with each condition (or pattern), a list of the data elements that it matches.
@@
 		In a streaming context, this is infeasible as streams can grow without bounds.
-		Therefore, we place time constraints on the memories using a window-join operator instead of the traditional joins.
+		Therefore, we place time constraints on the memories using a window-join operator instead of the traditional incremental joins.
 	\end{nestedsection}
 	\begin{nestedsection}{Continuous Reasoning in the Semantic Web}{implementation: continuous reasoning with RDF}
-		In R4, data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
-		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
-		Source nodes also timestamp the incoming graphs with the system time at their arrival, if they were not already stamped at their origin.
-		They propagate the data to the successor nodes, which are the alpha nodes.
-
-		Alpha nodes are single-entry nodes that form a discrimination network, as shown in \reffig{example-rete-network}.
-		In R4, alpha nodes can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
-		Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to the successor node(s) in the case of a match.
-		As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
-
-		The first alpha node in the network propagates matched quads to a special node called the left input adaptor, while the other alpha nodes propagate their output directly to join nodes.
-		The left input adaptor node is responsible of preparing the incoming triples to enter the beta network from the left input of the first join node in a given join-tree.
-		It creates a new token that represents a partial match, defined as a list of triples, for each rule body for which the incoming quad is a sub-graph match, then adds this quad as the first item in each list.
-		Tokens are then sent to the first join nodes in the beta network.
-
-		Beta nodes are two-input nodes that are responsible for joining the branches of the alpha network, as demonstrated in \reffig{example-rete-network}.
-		In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
-		Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing triples received from alpha nodes.
-		Join nodes are beta nodes that match graphs from both sides according to some conditions, e.g. a shared variable binding.
-		As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
-		In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
-		These windows are implemented as priorities queues and their items are ordered according to their timestamps.
-
-		When a join node is left activated, i.e.
-it receives a token through its left input, it first adds the new token to its left window, then it iterates over its right window elements trying to find a match.
-		When a match is found, a new token is created by adding the right triple to the triples list of the left token.
-		The new token is stamped with the earliest timestamp of its triples, and then propagated to the next join node or to the terminal node if it is the last join.
-		At this point, the join node performs a garbage collection over its left window, ensuring that items with timestamps that are smaller that current time minus the semantic window range are deleted.
-		Conversely, when the join node is right activated by receiving a triple from an alpha node, it adds the new triple to the right window, matches it against the elements of the left window, and performs a garbage collection over the right window.
-
-		This method of annotating tokens with the earliest timestamp of those of its components initially appears to violate the annotation semantics laid out in \refsec{semantics}.
-		However, the ordering of tokens in their streams and the system time at which they are propagated through the system expresses those annotation semantics implicitly.
-		This leaves the timestamp of a token, in combination with the range $r$ of the semantic window associated with the rule set, able to express the negation time $i_n$ of instances as in CoDeR.
-		As such, the interpretation of beta nodes in R4 is an implementation of the window-join operator characterised as a pair of SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
-
- 		Finally, terminal nodes receive tokens from the last node in a join-tree and are responsible for producing entailed triples, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
- 		The output of terminal nodes is both the output of the network itself, and re-entered to the first nodes of the alpha network to support iterative inference.
-
 		\begin{figure*}[t]
@@
  		\end{figure*}
+		In R4, data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
+		Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
+		Source nodes also annotate the incoming quads with the system time at their arrival if their parent graph was not already annotated at its origin, otherwise passing the annotation of the graph to the resulting quads.
+		This annotation $a$ relates to their instant of assertion $e$ within the CoDeR data model, and the instant $n$ at which they will leave the window whose range $r$ is defined in the system's rule set may be derived by ${n = a + r}$.
+		Source nodes propagate the annotated quads to their successor nodes, which are the alpha nodes.
+
+		Alpha nodes are single-entry nodes that form a discrimination network, as shown in \reffig{example-rete-network}.
+		In R4, alpha nodes can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
+		Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to its successor node(s) in the case of a match.
+		As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
+
+		The first alpha node in the network propagates matched quads to a special node called the left input adaptor, while the other alpha nodes propagate their output directly to beta nodes.
+		The left input adaptor node is responsible of preparing the incoming quads to enter the beta network from the left input of the first beta node in a given join-tree.
+		It creates a new \emph{token} that represents a partial match, defined as a list of quads, for each rule body for which the incoming quad is a sub-graph match, then adds this quad as the first item in each list.
+		It also annotates each token with the same time as was the quad for which it was generated.
+		Tokens are then sent to the first nodes in the beta network.
+
+		Beta nodes are two-input nodes that are responsible for joining the branches of the alpha network, as demonstrated in \reffig{example-rete-network}.
+		In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
+		Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing quads received from alpha nodes.
+		Join nodes are beta nodes that match graphs from both sides according to some conditions, e.g. a shared variable binding.
+		As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
+		In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
+		These windows are implemented as priorities queues in which quads are ordered according to their time annotations, and pruned by ``popping'' from the queue and deleting any quads whose annotation $a \leqslant i - r$, where $i$ is the current system time and $r$ is the range of the semantic window.
+
+		When a join node is left activated, i.e.
+it receives a token through its left input, it first adds the new token to its left window, then it prunes its right window before iterating over the remaining quads to find any matches.
+		When a match is found, a new token is created by duplicating the left token and adding the right quad to the new token's quad list.
+		The new token is annotated with the \emph{earliest} time with which its component quads were annotated, and then propagated to the next beta node, or to the terminal node if it is in the root node of a join tree.
+		Conversely, when the join node is right activated by receiving a quad from an alpha node, it adds the new quad to the right window, prunes the left window then matches the quad against the remaining tokens.
+
+		This method of annotating tokens with the earliest assertion time of those of its components initially appears to violate the annotation semantics for entailed fact instances laid out in \refsec{semantics,model}.
+		However, the ordering of tokens in their streams and the system time $i$ at which they are propagated through the system expresses implicitly CoDeR's entailment time $e$ of entailed instances.
+		This leaves the annotation $a$ of a token in R4, in combination with the range $r$ of the semantic window associated with the rule set, able to express CoDeR's negation time ${n}$ of entailed instances by the same calculation as those of asserted instances: ${n = a + r}$.
+		As such, the interpretation of beta nodes in R4 is a valid implementation of the window-join operator characterised as a pair of CoDeR SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
+
+		Finally, terminal nodes receive tokens from the root nodes of join-trees and are responsible for producing entailed graphs, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
+		As with the $\text{:-}_{head}$ operator, the annotation of these entailed graphs is directly inherited from the completed token from which it is produced.
+ 		The union of the output of all terminal nodes is both the output of the system itself, and re-entered as an input stream to the source nodes to support iterative inference.
 	\end{nestedsection}

### 2015-07-15 content/contribution/continuous-datalog.tex
removed rogue FIXME (actual final change to section 2)
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

@@
-\begin{nestedsection}{Continuous Datalog}{semantics}
+\begin{nestedsection}{Semantics: Windowed Continuous Datalog}{semantics}
 The first requirement of stream reasoning is a well-defined semantics
@@
 
-Continuous Datalog (henceforth C-Datalog) is an extension of positive
+Windowed Continuous Datalog (henceforth WC-Datalog) is an extension of positive
 Datalog with stream-processing semantics, with a view to accommodating
@@
 
-C-Datalog differs from conventional Datalog through the addition of
+WC-Datalog differs from conventional Datalog through the addition of
 sources of continuously generated or reported facts (i.e streams), and
-the application of sliding window semantics over those sources,
-thereby providing for the interpretation of C-Datalog programs as
-sequences of instantaneous states that are each valid positive Datalog
+the application of windowing semantics over those sources,
+thereby providing for the interpretation of WC-Datalog programs as
+sequences of instantaneous states that may each be reduced to valid positive Datalog
 programs.  As such, these states each inherit the semantics of
@@
 continuous queries, as defined by the stream processing community,
-thereby allowing the integration of C-Datalog programs into stream
+thereby allowing the integration of WC-Datalog programs into stream
 processing pipelines. Finally, we show that these streams may be
@@
 
-\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
 A window of range $0$ over a stream $S$ at an instant 
@@
 \end{equation*}
-\end{axiom}
+\end{axiom}}
 
-\begin{axiom}\label{axiom:continuous datalog: window composition}
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window start leq 0}
+A window of any range $x \in \mathbb{N}$ over a stream $S$ at (or before)
+the first instant of time is always empty.
+\begin{equation*}
+W^{S,x}_{i} = \varnothing
+\end{equation*}
+where $i \leqslant 0$.
+\end{axiom}}
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window composition}
 A window of range ${x > 0}$ over a stream $S$ at an instant 
@@
 \end{equation*}
-\end{axiom}
+\end{axiom}}
 
-\begin{axiom}\label{axiom:continuous datalog: window disjointness}
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window disjointness}
 A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
@@
 \end{equation*}
-where $z \geq x$.
-\end{axiom}
+where $z \geqslant x$.
+\end{axiom}}
 
@@
 an out-of-order stream.
-
-%%%FIXME
-In order to
-
 \end{definition}
@@
 
-\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
 If two sliding windows over the same in-order stream $S$ cover
@@
 set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
-where ${z \geq x}$.
+where ${z \geqslant x}$.
 \[
@@
 \]
-\end{axiom}
+\end{axiom}}
 
@@
 is the set $E^T_{i}$ of entailed fact instances whose justifications are at
-least partially contained in $W^{S,r}_i$. We define $E^T_i$ as:
+least partially contained in $W^{S,r}_i$. {\nobreak We define $E^T_i$ as:
 \[
@@
 \]
-
-\noindent where $J^T$ is a justification for the entailment of x by
-$\Pi^{C,r}_i$. By this definition and Axiom~\ref{axiom:continuous
+where $J^T$ is a justification for the entailment of x by
+$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
   datalog: entailment disjointness}, $E^0$ is guaranteed to be
@@
 \end{definition}
-
-The following axioms define the semantics of the streams of
-entailments $P$ and and negations $N$ of a $\Pi^{C,r}$ in terms of the
-instantaneous entailments of the sequence of states of that continuous
-program.
-
-\begin{axiom}
-\label{axiom:continuous datalog: positive window increment}
-The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
-the stream $P$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
-as the set of fact instances that are entailed by the state
-$\Pi^{C,r}_i$ that were not entailed by the state $\Pi^{C,r}_{i-1}$.
-\begin{equation*}
-W^{P,1}_{i} = \left( E^0 \cup E^T_{i} \right) \setminus
-\left( E^0 \cup E^T_{i-1} \right) = E^T_{i} \setminus E^T_{i-1}
-\end{equation*}
-\end{axiom}
-
-\begin{axiom}
-\label{axiom:continuous datalog: negative window increment}
-The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
-the stream $N$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
-as the set of fact instances that were entailed by the state
-$\Pi^{C,r}_{i-1}$ that are no longer entailed by the state
-$\Pi^{C,r}_i$.
-\begin{equation*}
-W^{N,1}_{i} = \left( E^0 \cup E^T_{i-1} \right) \setminus
-\left( E^0 \cup E^T_{i} \right) = E^T_{i-1} \setminus E^T_{i}
-\end{equation*}
-\end{axiom}
-
-\begin{axiom}
-\label{axiom:continuous datalog: entailment precedes negation}
-Fact instances may only be negated by some part of their justification
-leaving the sliding window $W^{S,r}_{i}$, this necessarily being after
-they were entailed.  As such, a window of any range ${x \in
-  \mathbb{N}}$ over the stream of entailments $P$ at any given instant
-${i \in \mathbb{T}}$ is disjoint from a window of any range ${y \in
-  \mathbb{N}}$ over the stream of entailment negations $N$ at any
-instant before ${i - x}$.
-\begin{equation*}
-  W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing
-\end{equation*}
-where $z \geq x$.
-\end{axiom}
-
-As noted above in \refdef{continuous datalog: CDP}, the stream of
-negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
-out-of-order stream in which fact instances appear on the stream in
-the order in which they cease to have any complete justifications.
-
-Thus far the translation from instantaneous sets of entailments to
-streams of entailments has been stated axiomatically as the definition
-of those streams.
-
-However, it follows from these axioms that the instantaneous set
-$E^T_{i}$ of instances entailed at any given instant $i$ by any
-$\Pi^{C,r} = \langle R, F^0, S, r \rangle$ may be reconstructed from
-windows of sufficient range over its entailed streams $P$ and $N$ at
-$i$:
-\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
-where ${y \geq x \geq r}$. The minimum range $x$ is such to achieve
-the \emph{complete} set of entailment instances $E^T_{i}$ of the
-$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
-\emph{sound} set $E^T_{i}$. As such, a \emph{valid window} that
-contains only and exactly those entailed instances $E^T_{i}$ at every
-instant $i$ may be trivially maintained from the streams $P$ and $N$
-continuously entailed by some $\Pi^{C,r}$.
-
-% \begin{thrm}\labelhyp{continuous datalog: entailment} \hfill \\
-
-% For any ${\Pi^{C,r} \models \langle P, N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_{i}}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \geq r}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the transitive portion $E^T_{i}$ of the entailment of the state of the ${\Pi^{C,r}}$ at that instant $i$:
-
-% \begin{equation*}
-%   W^{P,x}_{i} \setminus W^{N,y}_{i} = E^T_{i}
-% \end{equation*}
-
-% \end{thrm}
-
-% \begin{proof}
-
-% In order to prove \refhyp{continuous datalog: entailment} for the specific case of ${x \geq r}$, for some range $r$ of the window over $S$ in some ${\Pi^{C,r} = \langle R,F^0,S,r \rangle}$, a general case must first be proven for all ${x \geq 1}$.
-% This general case constitutes the Lemma \ref{hyp:continuous datalog: entailment lemma}, which provides an equivalence between the entailments of discrete states of $\Pi^{C,r}$ and windows over the streamed entailment ${\langle P, N \rangle}$ for all window ranges ${x \geq 1}$.
-
-% \begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
-
-% For any ${\Pi^{C,r} \models \langle P,N \rangle}$, where ${\Pi^{C,r}_{i} \models E^0 \cup E^T_i}$ for any ${i \in \mathbb{T}}$, the set of facts in a window of any range ${x \geq 1}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${y \geq x}$ over the stream $N$ at the same instant $i$ is equal to the transitive portion $E^T_i$ of the entailment of the state of $\Pi^{C,r}$ at that instant $i$ less the intersection of the transitive portions of the entailments of all previous states of $\Pi^{C,r}$, from instant ${i-1}$ up to and including ${i-x}$:
-
-% \begin{equation*}
-%   W^{P,x}_{i} \setminus W^{N,y}_{i} = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d}
-% \end{equation*}
-
-% \end{lem}
-
-% The inductive proof of Lemma \ref{hyp:continuous datalog: entailment lemma} is available in \ref{app:semantics: proof for entailment}:
-% a \emph{general case} (\reffig{semantics: induction general case}) eliminates the range ${y \geq x}$ over $N$ as a factor;
-% a \emph{base case} (\reffig{semantics: induction base case}) is shown to hold when ${x = 1}$;
-% and the \emph{inductive case} (\reffig{semantics: induction inductive case}) holds when ${x = k + 1}$, assuming the lemma holds for ${x = k}$.
-% Given this lemma, the derivation of \refhyp{continuous datalog: entailment} presented in \reffig{continuous datalog: proof for entailment} follows.
-
-% \begin{figure*}
-
-% \centering
-
-% \begin{align*}
-%   E^T_{i} & = W^{P,x}_{i} \setminus W^{N,y}_{i} \\
-%   & = E^T_{i} \setminus \mathop{\cap}^{x}_{d=1} E^T_{i-d}
-%   && \text{, Lemma~\ref{hyp:continuous datalog: entailment lemma}} \\
-%   & = \mathop{\cup}^{x}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
-%   & = \left( E^T_{i} \setminus E^T_{i-x} \right) \cup
-%   \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right) \\
-%   & = E^T_{i} \cup \mathop{\cup}^{x-1}_{d=1} \left( E^T_{i} \setminus E^T_{i-d} \right)
-%   && \text{, Axiom~\ref{axiom:continuous datalog: entailment disjointness}} \\
-%   & = E^T_{i} \cup \left( E^T_{i} \setminus \mathop{\cap}^{x-1}_{d=1} E^T_{i-d} \right) \\
-%   & = E^T_{i} && \qedhere
-% \end{align*}
-
-% \caption{Derivation for proof of \refhyp{continuous datalog: entailment}}
-% \labelfig{continuous datalog: proof for entailment}
-
-% \end{figure*}
-
-% \end{proof}
-
-In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
-F^0, S \rangle$, we can apply to it a window of range $r$ to yield a
-Windowed Continuous Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
-\rangle$. We can evaluate this window at an instant $i \in \mathbb{T}$
-to yield a Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup
-W^{S,r}_i \rangle$, from we which can then strip the time annotations
-to yield a conventional Datalog program $\Pi = \langle R, F
-\rangle$. The relationship between these *-Datalog programs and their
-entailments is expressed in the commutative diagram in
-\reffig{CD-TD-D}.
 
@@
 \end{figure*}
+
+The following axioms define the semantics of the streams of
+entailments $P$ and and negations $N$ of a $\Pi^{C,r}$ in terms of the
+instantaneous entailments of the sequence of states of that continuous
+program.
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: positive window increment}
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $P$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that are entailed by the state
+$\Pi^{C,r}_i$ that were not entailed by the state $\Pi^{C,r}_{i-1}$.
+\begin{equation*}
+W^{P,1}_{i} = \left( E^0 \cup E^T_{i} \right) \setminus
+\left( E^0 \cup E^T_{i-1} \right) = E^T_{i} \setminus E^T_{i-1}
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: negative window increment}
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $N$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that were entailed by the state
+$\Pi^{C,r}_{i-1}$ that are no longer entailed by the state
+$\Pi^{C,r}_i$.
+\begin{equation*}
+W^{N,1}_{i} = \left( E^0 \cup E^T_{i-1} \right) \setminus
+\left( E^0 \cup E^T_{i} \right) = E^T_{i-1} \setminus E^T_{i}
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: entailment precedes negation}
+Fact instances may only be negated by some part of their justification
+leaving the sliding window $W^{S,r}_{i}$, this necessarily being after
+they were entailed.  As such, a window of any range ${x \in
+  \mathbb{N}}$ over the stream of entailments $P$ at any given instant
+${i \in \mathbb{T}}$ is disjoint from a window of any range ${y \in
+  \mathbb{N}}$ over the stream of entailment negations $N$ at any
+instant before ${i - x}$.
+\begin{equation*}
+  W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing
+\end{equation*}
+where $z \geqslant x$.
+\end{axiom}}
+
+As noted above in \refdef{continuous datalog: CDP}, the stream of
+negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
+out-of-order stream in which fact instances appear on the stream in
+the order in which they cease to have any complete justifications.
+
+Thus far the translation from instantaneous sets of entailments to
+streams of entailments has been stated axiomatically as the definition
+of those streams.
+
+However, it follows from these axioms that the instantaneous set
+$E^T_{i}$ of instances entailed at any given instant $i$ by any
+$\Pi^{C,r} = \langle R, F^0, S, r \rangle$ may be reconstructed from
+windows of sufficient range over its entailed streams $P$ and $N$ at
+$i$:
+\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
+where ${y \geqslant x \geqslant r}$. The minimum range $x$ is such to achieve
+the \emph{complete} set of entailment instances $E^T_{i}$ of the
+$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
+\emph{sound} set $E^T_{i}$. As such, a \emph{valid window} that
+contains only and exactly those entailed instances $E^T_{i}$ at every
+instant $i$ may be trivially maintained from the streams $P$ and $N$
+continuously entailed by some $\Pi^{C,r}$.
+
+In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
+F^0, S \rangle$, we can apply to it a window of range $r$ to yield a
+Windowed Continuous Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
+\rangle$. We can evaluate this window at an instant $i \in \mathbb{T}$
+to yield a Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup
+W^{S,r}_i \rangle$, from we which can then strip the time annotations
+to yield a conventional Datalog program $\Pi = \langle R, F
+\rangle$. The relationship between these *-Datalog programs and their
+entailments is expressed in the commutative diagram in
+\reffig{CD-TD-D}.
+
 \end{nestedsection}

### 2015-07-15 content/evaluation.tex
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

@@
 \begin{nestedsection}{Evaluation, Results and Discussion}{evaluation}
-	To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning on semantic streaming data.
+	To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning to some extent on semantic streaming data.
 	These include Etalis \citep{EP-SPARQL}, Sparkwave \citep{sparkwave}, Streaming knowledge bases \citep{walavalkar08streamingkb}, and the incremental reasoner presented in \citep{inc-reasoning-background-knowledge}.
@@
 	\begin{nestedsection}{Motivating Case}{evaluation: motivating case}
-		We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
-		Inspired by a query in \citep{SRBench}, we formulated a rule that entails that a given observation is of a weather pattern classified as a hurricane if it is of a wind speed that exceeds that of a previously classified hurricane;
-		it also entails that the wind speed observed exceeds that of a given historically recorded hurricane (\reffig{SRBench-rule}).
-		\begin{figure*}
+		\begin{figure*}[t]
 			\centering
@@
 		\end{figure*}
+
+		We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
+		Inspired by a query in \citep{SRBench}, we formulated a rule that entails that a given observation is of a weather pattern classified as a hurricane if it is of a wind speed that exceeds that of a previously classified hurricane;
+		it also entails that the wind speed observed exceeds that of a given historically recorded hurricane (\reffig{SRBench-rule}).
 		This rule involves dealing with static data regarding past hurricanes, dynamic data regarding ongoing sensor readings, and also needs to perform reasoning against an ontology of background knowledge to identify the hurricanes that are instances of the specified class of hurricanes.
@@
 		However, Sparkwave was unable to read the dataset correctly as it does not parse blank nodes;
-		this is because it uses the hash-join algorithm to improve time efficiency, as we show in \refsec{evaluation: comparative results}, but builds the hashtables based on the URIs of the incoming data.
+		this is because it uses the hash-join algorithm to improve time efficiency, as we show in \refsec{evaluation: comparative results}, but builds the hashtables based on the URIs of the incoming data as which blank nodes are not valid.
 	\end{nestedsection}
 	\begin{nestedsection}{Comparative Results}{evaluation: comparative results}
-		In order to compare the efficiency of R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
-		We generated a 1.1 million triples dataset containing information about 100,000 limited-time offers made available by some online marketplace.
-		We used a small schema that has 329 product types arranged in a 4-level hierarchy.
-
-		The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
-		As offers may be associated with product super-types, background reasoning is needed to determine the offer price for specific product sub-types.
-		\begin{figure*}
+		\begin{figure*}[t]
 			\centering
@@
 
+		In order to compare the efficiency of R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+		We used another 1.1 million triples dataset, this time using the provided data generator, containing information about 100,000 limited-time offers made available by some fictional online marketplace.
+		We used a small schema that has 329 product types arranged in a 4-level hierarchy.
+
+		The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
+		As offers may be associated with product super-types, some small degree of background reasoning is needed to determine the offer price for specific product sub-types.
+
 		Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
@@
 		In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating Etalis to use \emph{seq} instead of \emph{And}.
-		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time- and memory-efficiency of the three systems.
+		It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time-efficiency and memory-consumption of the three systems.
 		\begin{figure}
@@
 			\includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
-			\caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given sliding windows of various ranges over the input stream.}
+			\caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given windows of various ranges over the input stream.}
 			\labelfig{all-systems-varying-windows}
@@
 
-		As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small sliding windows but slower than Sparkwave, and, while the time-to-complete increases with window range in both Sparkwave and R4, that of Etalis appears to remain fairly constant for the sliding-window ranges tested.
-		The growing difference between R4 and Sparkwave as the range of the sliding windows increases is likely due to the fact that Sparkwave uses a hash join algorithm with garbage-collection window-maintenance while the current implementation of R4 uses loop joins over lists maintained by indexing over timestamps;
+		As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small windows but slower than Sparkwave, and, while the time-to-complete increases with window range in both Sparkwave and R4, that of Etalis appears to remain fairly constant for the window ranges tested.
+		The growing difference between R4 and Sparkwave as the range of the windows increases is likely due to the fact that Sparkwave uses a hash-join algorithm with garbage-collection window-maintenance while the current implementation of R4 uses loop-joins over lists maintained by indexing over time annotations;
 		as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer to iteratively check for matches compared to the hash-based look-up of those in Sparkwave.
@@
 		Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
-		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave memory needs are significantly higher.
-		Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to the manual garbage collection performed.
+		It shows that, while R4 and Jtalis follow similar patterns, Sparkwave's memory needs are significantly higher and continue to increase over time, a behaviour that should not occur when applying windows over streams.
+		Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to an interaction between the manual garbage collection performed by Sparkwave and the underlying JVM garbage collector.
 		\begin{figure}

### 2015-07-15 content/introduction.tex
Some final alterations to bring the entire paper more fluidly in line, terminologically, and to balance the placement of figures more evenly through the document.

@@
 	We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
-	Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems EP-SPARQL and Sparkwave.
+	Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems Etalis and Sparkwave.
 	Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Continuous Datalog.

### 2015-07-15 content/conclusions.tex

@@
 	We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
-	In this way it differs from other temporal interpretations of Datalog \citep{orgun88chronolog} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing, casting a continuous program as a series of discrete states.
+	In this way it differs from other temporal interpretations of Datalog \citep{OrgunWadge92,Tuzhilin93} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing and casting a continuous program as a series of discrete states.
 	WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.

### 2015-07-15 article.tex
TESTED and WORKING: re-organised folder to work without latexbuild.sh, instead running off "$ make", and possible to run easily using "$ pdflatex article.tex" and "$ bibtex article".
Also got rid of unused images and tex files

\documentclass[twocolumn,preprint,3p,number]{template/elsarticle}
% \documentclass[preprint,5p,number]{elsarticle}

\journal{Web Semantics}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amscd}

\usepackage{hyperref}

\usepackage{subfig}
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.pdf} % ,.jpeg,.jpg,.png}

\theoremstyle{plain}
\newtheorem{thrm}{Theorem}
\newtheorem{lem}{Lemma}[thrm]
\newtheorem{axiom}{Axiom}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

% Document relocation and referencing
\usepackage{import}
\usepackage{template/labelledrefs}

\include{template/ebnf}
\include{template/outerjoins}

\makeatletter
\def\@author#1{\g@addto@macro\elsauthors{\normalsize%
    \def\baselinestretch{1}%
    \upshape\authorsep#1\unskip\textsuperscript{%
      \ifx\@fnmark\@empty\else\unskip\sep\@fnmark\let\sep=,\fi
      \ifx\@corref\@empty\else\unskip\sep\@corref\let\sep=,\fi
      }%
    \def\authorsep{\unskip,\space}%
    \global\let\@fnmark\@empty
    \global\let\@corref\@empty  %% Added
    \global\let\sep\@empty}%
    \@eadauthor={#1}
}
\makeatother

\begin{document}

	\subimport{content/}{personal-details-els.tex}

	\subimport{content/}{introduction.tex}
	% \subimport{../content/}{preliminaries.tex}
	\subimport{content/}{contribution.tex}
	\subimport{content/}{evaluation.tex}
	\subimport{content/}{conclusions.tex}
%	\subimport{../content/}{future-work.tex}

%	\subimport{../content/}{appendices.tex}
  \section*{References}
	\bibliographystyle{template/elsarticle-harv}
	\bibliography{bibliography}
\end{document}

### 2015-07-15 latexbuild.sh

@@ -1,85 +1,0 @@
-#!/bin/bash
-cd template
-docs=""
-count=0
-while [ $# != 0 ]; do
-	tflag=false
-	dflag=false
-	t=""
-	if [[ $1 ==  "-t"* ]]; then
-		tflag=true
-		t=`echo "$1" | sed -e 's/-t//'`
-		if [ -z $t ]; then
-			shift
-			t="$1"
-		fi
-		t=${t}"-wrapper"
-		shift
-		cat > "$t.tex" << EOF
-\\input{preamble-thesis.tex}
-\\begin{document}
-	\\begin{nestedsection}{$t}{main}
-EOF
-		count=$((count+1))
-	elif [[ $1 == "-d" ]]; then
-		dflag=true
-		t=`echo "$2" | sed -e 's/.*\///'`
-		t=${t}"-wrapper"
-		cat > "$t.tex" << EOF
-\\input{preamble-thesis.tex}
-\\begin{document}
-EOF
-		count=$((count+1))
-	fi
-
-	while [[ $1 ==  "-d" ]]; do
-		shift
-		cat >> "$t.tex" << EOF
-		\\subimport{../content/}{$1.tex}
-EOF
-		shift
-	done;
-	
-	if $tflag ; then
-		cat >> "$t.tex" << EOF
-	\\end{nestedsection}
-	\\bibliographystyle{apalike}
-	\\bibliography{../bibliography}
-\\end{document}
-EOF
-	elif $dflag ; then
-		cat >> "$t.tex" << EOF
-	\\input{bibliography}
-\\end{document}
-EOF
-	fi
-	
-	if [[ -z $t ]]; then
-		t="$1"
-		count=$((count+1))
-		shift
-	fi
-
-	docs="$docs $t"
-done
-
-echo \*\*\* Processing $count documents: $docs \*\*\*
-
-for t in $docs ; do
-	echo \*\*\* "$t" about to be processed \*\*\*
-	make -f ../makefile DOC="$t"
-	echo \*\*\* "$t" processed \*\*\*
-
-	tFinal=`echo "$t" | sed -e 's/-wrapper$//'`
-	mv "$t.pdf" "../$tFinal.pdf"
-	open "../$tFinal.pdf"
-	start "../$tFinal.pdf"
-done
-wrappers=`ls -1 *-wrapper* 2>/dev/null`
-rm -f $wrappers
-if [[ -n $wrappers ]] ; then
[…]

### 2015-07-15 article.tex
migrated everything into a single document as much as possible, with elsarticle.cls and the various ".bst"s and all images in the root directory.  got rid of all the usepackages apart from the AMS packages, hyperref and the etoolbox.  migrated the one use of subfigs to minipages

@@
-\documentclass[twocolumn,preprint,3p,number]{template/elsarticle}
+\documentclass[twocolumn,preprint,3p,number]{elsarticle}
 % \documentclass[preprint,5p,number]{elsarticle}
@@
 
-\usepackage{subfig}
-\graphicspath{{images/}}
+\usepackage{etoolbox}
+
 \DeclareGraphicsExtensions{.pdf} % ,.jpeg,.jpg,.png}
 
+%% THEOREMS
 \theoremstyle{plain}
@@
 
-% Document relocation and referencing
-\usepackage{import}
-\usepackage{template/labelledrefs}
+%% Referencing and sections
+\makeatletter
+\newcounter{nestingdepth}
+\ifcsundef{chapter}{ %
+  \setcounter{nestingdepth}{1} %
+}{ %
+  \setcounter{nestingdepth}{0} %
+}
 
-\include{template/ebnf}
-\include{template/outerjoins}
+\newenvironment{nestedsection}[2]{
+  \ifnum\value{nestingdepth}=0
+    \chapter{#1}
+  \else
+    \ifnum\value{nestingdepth}=1
+      \section{#1}
+    \else
+      \ifnum\value{nestingdepth}=2
+        \subsection{#1}
+      \else
+        \ifnum\value{nestingdepth}=3
+          \subsubsection{#1}
+        \else
+          \ifnum\value{nestingdepth}=4
+            \paragraph{#1}
+          \else
+            \PackageError{nestedsections}{Maximum nesting level exceeded!}{uh oh!}
+          \fi
+        \fi
+      \fi
+    \fi
+  \fi
+  \addtocounter{nestingdepth}{1}
+  \label{sec:#2}
+}{\addtocounter{nestingdepth}{-1}}
 
+\newenvironment{untrackednestedsection}[1]{
+  \ifnum\value{nestingdepth}=0
+    \chapter*{#1}\def\sectiontype{chapter}
+  \else
+    \ifnum\value{nestingdepth}=1
+      \section*{#1}\def\sectiontype{section}
+    \else
+      \ifnum\value{nestingdepth}=2
+        \subsection*{#1}\def\sectiontype{subsection}
+      \else
+        \ifnum\value{nestingdepth}=3
+          \subsubsection*{#1}\def\sectiontype{subsubsection}
+        \else
+          \ifnum\value{nestingdepth}=4
+            \paragraph*{#1}\def\sectiontype{paragraph}
+          \else
+            \PackageError{nestedsections}{Maximum nesting level exceeded!}{uh oh!}
+          \fi
+        \fi
+      \fi
+    \fi
+  \fi
+  \addtocounter{nestingdepth}{1}
+}{\addtocounter{nestingdepth}{-1}}
+
+\newenvironment{nestedappendix}[2]{
+  \ifnum\value{nestingdepth}=0
+    \chapter{#1}
+  \else
+    \ifnum\value{nestingdepth}=1
+      \section{#1}
+    \else
+      \ifnum\value{nestingdepth}=2
+        \subsection{#1}
+      \else
+        \ifnum\value{nestingdepth}=3
+          \subsubsection{#1}
+        \else
+          \ifnum\value{nestingdepth}=4
+            \paragraph{#1}
+          \else
+            \PackageError{nestedsections}{Maximum nesting level exceeded!}{uh oh!}
+          \fi
+        \fi
+      \fi
+    \fi
+  \fi
+  \addtocounter{nestingdepth}{1}
+  \label{app:#2}
+}{\addtocounter{nestingdepth}{-1}}
+
+\newcommand\@addnestedcontentsline[3]{
+  \ifnum\value{nestingdepth}=1
+    \label{sec:#3}\addcontentsline{#1}{chapter}{#2}
+  \else
+    \ifnum\value{nestingdepth}=2
+      \label{sec:#3}\addcontentsline{#1}{section}{#2}
+    \else
+      \ifnum\value{nestingdepth}=3
+        \label{sec:#3}\addcontentsline{#1}{subsection}{#2}
+      \else
+        \ifnum\value{nestingdepth}=4
+          \label{sec:#3}\addcontentsline{#1}{subsubsection}{#2}
+        \else
+          \ifnum\value{nestingdepth}=5
+            \label{sec:#3}\addcontentsline{#1}{paragraph}{#2}
+          \else
+            \PackageError{nestedsections}{Maximum nesting level exceeded!}{uh oh!}
+          \fi
+        \fi
+      \fi
+    \fi
+  \fi
+}
+
+\newenvironment{nestedsection*}[2]{
+  \begin{untrackednestedsection}{#1}{#2}
+    \@addnestedcontentsline{toc}{#1}{#2}
+}{\end{untrackednestedsection}}
+
+\def\@multiref[#1]#2{%
+  \@listfirstref[#1]#2,\relax%
+}
+\def\@listfirstref[#1]#2,#3\relax{%
+      \ifx\relax#3\relax  %
+         \def\next{~\@prefref[#1]{#2}}%
+      \else
+         \def\next{s~\@prefref[#1]{#2}\@listrefs[#1]#3\relax}%
+      \fi%
+      \next%
+}
+\def\@listrefs[#1]#2,#3\relax{%
+      \ifx\relax#3\relax  %
+         \def\next{\@endmultiref[#1]{#2}}%
+      \else
+         \def\next{\@listref[#1]{#2}\@listrefs[#1]#3\relax}%
+      \fi%
+      \next%
+}
+\def\@listref[#1]#2{, \@prefref[#1]{#2}}
+\def\@endmultiref[#1]#2{ and~\@prefref[#1]{#2}}
+\def\@prefref[#1]#2{%
+  \ifx\relax#1\relax%
+      \def\next{\ref{#2}}%
+   \else%
+      \def\next{\ref{#1:#2}}%
+   \fi%
+   \next%
+}
+
+\def\labelfig#1{\label{fig:#1}}
+\def\reffig#1{Figure\@multiref[fig]{#1}}
+\def\labeltab#1{\label{tab:#1}}
+\def\reftab#1{Table\@multiref[tab]{#1}}
+\def\refsec#1{Section\@multiref[sec]{#1}}
+\def\refchap#1{Chapter\@multiref[sec]{#1}}
+\def\refapp#1{Appendix~\@prefref[app]{#1}}
+\def\labeleqn#1{\label{eqn:#1}}
+\def\refeqn#1{Equation\@multiref[eqn]{#1}}
+\def\labelalg#1{\label{alg:#1}}
+\def\refalg#1{Algorithm\@multiref[alg]{#1}}
+\def\labelhyp#1{\label{hyp:#1}}
+\def\refhyp#1{Hypothesis~\@prefref[hyp]{#1}}
+\def\labelqstn#1{\label{qstn:#1}}
+\def\refqstn#1{Question\@multiref[qstn]{#1}}
+\def\labeldef#1{\label{def:#1}}
+\def\refdef#1{Definition\@multiref[def]{#1}}
+\def\labelcha#1{\label{cha:#1}}
+\def\refcha#1{challenge of \@prefref[cha]{#1}}
+\makeatother
+
+%% EBNF
+\newenvironment{ebnf}
+  {\newcommand{\NTerm}[1]{##1}
+   \newcommand{\Rule}[2]{\item[##1]
+        ::=\; ##2}
+   \newcommand{\OrRule}[1]{\par
+        \ \ \ \textbar\ ##1 }
+   \newcommand{\Token}[1]{\mbox{\small{`\textsf{\textbf{##1}}'}}}
+   \newcommand{\ZeroOrMore}[1]{{##1}*}
+   \newcommand{\OneOrMore}[1]{{##1}+}
+   \newcommand{\OrList}[1]{(\ ##1\ )}
+   \newcommand{\OrSep}{\ \textbar\ }
+   \newcommand{\Optional}[1]{{##1}?}
+   \begin{flushleft}
+   \begin{list}{}{
+     \unboldmath
+     \setlength{\parsep}{0cm}
+     \setlength{\itemsep}{0cm}
+     \setlength{\parskip}{0cm}
+     \renewcommand{\makelabel}[1]{##1\hfil}
+     \setlength{\labelwidth}{2cm}
+     \setlength{\leftmargin}{3cm}
+%     \settowidth{\itemindent}{::= \;\;\;}
+%     \setlength{\itemindent}{0cm}
+%     \settowidth{\listparindent}{\textbar \;\;\;\;}
+%     \setlength{\listparindent}{0pt - \listparindent}
+%     \setlength{\leftmargin}{\labelwidth + \labelsep - \itemindent}
+}
+   \begin{sloppypar}}
+   {\end{sloppypar}\end{list}\end{flushleft}}
+
+%% Outer Joins
+\def\ojoin{\setbox0=\hbox{$\Join$}%
+  \rule{.25em}{.4pt}\llap{\rule[.46em]{.25em}{.4pt}}}
+\def\leftouterjoin{\mathbin{\ojoin\mkern-8mu\Join}}
+\def\rightouterjoin{\mathbin{\Join\mkern-8mu\ojoin}}
+\def\fullouterjoin{\mathbin{\ojoin\mkern-8mu\Join\mkern-8mu\ojoin}}
+
+\def\hjoin{\setbox0=\hbox{$\rhd$}%
+  \rule{.25em}{.4pt}\llap{\rule[.46em]{.25em}{.4pt}}}
+\def\lstreamjoin{\mathbin{%
+  \setbox0=\hbox{$\lhd$}%
+    \rule{.25em}{.4pt}\llap{\rule[.46em]{.25em}{.4pt}}%
+  \mkern-8mu\lhd}}
+\def\rstreamjoin{\mathbin{\rhd\mkern-2mu%
+  \setbox0=\hbox{$\rhd$}%
+    \rule[0.21em]{.5em}{.4pt}%
+}}
+
+%% FIX CORREF
 \makeatletter
@@
 
-	\subimport{content/}{personal-details-els.tex}
+% Personal Details
+\begin{frontmatter}
+  \title{Continuous Deductive Reasoning over Streamed RDF Data}
 
-	\subimport{content/}{introduction.tex}
-	% \subimport{../content/}{preliminaries.tex}
-	\subimport{content/}{contribution.tex}
-	\subimport{content/}{evaluation.tex}
-	\subimport{content/}{conclusions.tex}
-%	\subimport{../content/}{future-work.tex}
+  \author{David L. Monks\corref{cor1}}
+  \ead{dm11g08@ecs.soton.ac.uk}
+  \cortext[cor1]{Corresponding author.}
+  \author{Rehab Albeladi}
+  \ead{raab1g09@ecs.soton.ac.uk}
+  \author{Nicholas Gibbins}
+  \ead{nmg@ecs.soton.ac.uk}
+  \address{
+            Electronics and Computer Science,\\
+            University of Southampton,\\
+            Southampton, SO17 1BJ,\\
+            United Kingdom
+  }
 
-%	\subimport{../content/}{appendices.tex}
+  \begin{abstract}
+    In this paper, we introduce Continuous Datalog, an extension of
+    conventional Datalog that adds the concept of streams of
+    temporally-annotated facts to conventional Datalog and defines the
+    notion of entailment over a windowed stream. We go on to define CoDeR,
+    an abstract model for continuous deductive reasoning that provides
+    both a data stream model based on Continuous Datalog and a minimal
+    algebra for expressing queries and programs.  Finally, we describe the
+    R4 stream reasoner, which implements the CoDeR model as a Rete-based
+    dataflow network and provide a brief comparative evaluation of R4
+    against ETALIS and Sparkwave.
+
+  \end{abstract}
+
+  \begin{keyword}
+        Deductive Reasoning\sep Streamed Data\sep Semantic Data\sep Continuous Processing\sep Plan Optimisation
+  \end{keyword}
+
+  \date{\today}
+
+\end{frontmatter}
+
+\begin{nestedsection}{Introduction}{intro}
+  % general framing of the problem (why reason over streams)
+  As the Semantic Web becomes more extensive, with more people providing and describing their data in the semantic formats recommended by the W3C, the variety in the nature of the data expressed continues to approach that of previous, more established data formats.
+  This includes the growing number of use cases for the continuous processing of streams of semantic data, in which large volumes of time-sensitive data must be processed in near-real time, with the aim of minimising the volume of data transported, duplicated and stored across the Web whilst maximising the variety of problems that can be solved.
+  With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
+  This may be characterised more precisely as continuous deductive reasoning over streamed RDF data.
+
+  The value of such processing is the same as that of reasoning over comparatively static data, that being that the division of knowledge into particular knowledge and general knowledge allows for the expression of a greater volume of information in a lesser volume of data.
+  In addition, this division allows the application of a single domain of general knowledge to a variety of distinct sets of particular knowledge, and the application of many distinct domains of general knowledge to a single set of particular knowledge.
+  This is a particularly useful characteristic in the context of semantic stream processing, in which all current use cases consider the application of some static domain of knowledge to some stream of particular knowledge;
+  thus, with stream reasoning the volume of streamed data may be dramatically reduced by extracting the static domain of general knowledge that underpins the data and providing that domain as a static ontology, then streaming only the minimal particular knowledge that would be necessary to deduce the full data set.
+
+  % motivating scenario
+  Take, for example, the case of extensive sensor networks with many and varied sensors, such as those served by the SemsorGrid4Env project\footnote{\url{http://www.semsorgrid4env.eu/}}.
+  Any stand-alone semantic representation of this data may contain classification information for each sensor, according to what they measure, as well as how frequently and how accurately, as well as other factors.
+  It may also contain explicit data concerning events that are reported by no one sensor, and are instead extrapolated from the base sensor data.
+  In both of these cases, the classification criteria and extrapolation process for macro-events may be expressed as general knowledge that is applicable to all sensor readings.
+  As such, a consumer of the streams produced using SemsorGrid4Env could download the appropriate ontologies once, then use them to reason over a reduced set of streamed data containing only the base readings from the sensor network, thus obtaining the classification and macro-event information desired without having to stream it explicitly.
+
+  % *brief* summary of state of the art (to be expanded in related work)
+  Recent research has approached the task of stream reasoning from two different directions.
+  %   RDF stream processing and reasoning
+  The RDF stream processing community has proposed extensions to their semantic query languages to support the iterative application of queries to the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}.
+  %   issues with existing art: no explicit semantics for reasoning
+  While this approach is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, the query-by-query windowing semantics of the base systems do not translate to a coherent semantics for multi-rule stream reasoning.
+  On the other hand, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
+  While these algorithms do support a coherent semantics for continuous reasoning, this semantics casts the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams, and so preclude the further processing of entailments in a stream-processing manner.
+  As such, stream reasoning in its current form either lacks a coherent semantics for continuous truth and entailment, or is necessarily a terminal rather than intermediate process in any streamed data pipeline.
+
+  % Rete as non-streaming implementationexample (application to dataflow networks)
+  Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
+  These systems send data from operator to operator, with `memories' compiled for two-pass operators in which to retain the specific data received, in much the same way as sliding windows are used in stream processing.
+  This makes them functionally similar to many stream-to-stream processing systems, and have even been a foundation for some such as Sparkwave \citep{sparkwave}.
+  However, as \citep{forgy79} does not consider the semantics of stream processing, and those semantics are not clearly defined elsewhere for reasoning, such a semantics must still be defined before any stream reasoning system based on Rete may claim to adhere to them.
+
+  % overview
+  In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Continuous Datalog.
+  As positive Datalog is also the semantic equivalent of RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, this continuous semantics may be applied to the continuous application of OWL 2 ontologies to streamed RDF data.
+  Next, we present the abstract processing model CoDeR that provides for the expression of processing plans that satisfy the semantics of Continuous Datalog.
+  We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
+  Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems Etalis and Sparkwave.
+  Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Continuous Datalog.
+\end{nestedsection}
+
+\begin{nestedsection}{Windowed Continuous Datalog}{semantics}
+The first requirement of stream reasoning is a well-defined semantics
+that takes into account both the continuous nature of a subset of the
+base axioms of a stream reasoning problem, as well as its
+instantaneous truth and entailments, in line with the work of the
+reasoning community.  However, it must also provide a continuous
+interpretation of those instantaneous entailments in order to
+integrate with other stream-processing technologies.  More
+specifically, an appropriate semantics for stream reasoning on the
+Semantic Web must: allow the expression of OWL 2 ontologies to some
+degree of expressivity; not sacrifice the precise semantics thereof
+whilst incorporating the semantics of stream processing; and allow for
+the continuous derivation of entailments in a concise manner.
+
+Windowed Continuous Datalog (henceforth WC-Datalog) is an extension of positive
+Datalog with stream-processing semantics, with a view to accommodating
+the aims of the latter within the well-defined model-theoretic
+semantics of the former, which are shared with RIF-Core
+\cite{w3crifbld} and therefore support reasoning at the expressivity
+of OWL 2 RL. As part of its application to Semantic Web reasoning, we
+assume that the atomic formulae of the Datalog programs described
+below correspond to RDF triples: $\langle subj, pred, obj\rangle$.
+
+WC-Datalog differs from conventional Datalog through the addition of
+sources of continuously generated or reported facts (i.e streams), and
+the application of windowing semantics over those sources,
+thereby providing for the interpretation of WC-Datalog programs as
+sequences of instantaneous states that may each be reduced to valid positive Datalog
+programs.  As such, these states each inherit the semantics of
+positive Datalog.  Furthermore, it involves the continuous derivation
+of entailments from this sequence of instantaneous states as a stream
+of facts; these streams share a data model with the streams of
+continuous queries, as defined by the stream processing community,
+thereby allowing the integration of WC-Datalog programs into stream
+processing pipelines. Finally, we show that these streams may be
+manipulated to retrieve the instantaneous sets of entailments produced
+by the equivalent Datalog programs.
+
+\begin{definition}[Datalog]
+\labeldef{continuous datalog: datalog program}
+
+A conventional Datalog program $\Pi = \langle R, F\rangle$ comprises a
+set of clauses, divided into a finite set of rules $R$ and a finite
+set of ground facts $F$. Facts are formulae of first-order logic, and
+rules are of the form:
+
+\[ \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n \]
+
+\noindent where $\alpha_i$ are atomic formulae. Note that we have excluded any
+universally quantified variables for clarity.
+
+The Herbrand base $\hat{H}$ of $\Pi$ is the set of all facts that can
+be expressed using the predicates and constants in $\Pi$; for a program $\Pi =
+\langle R, F \rangle$, $F \subseteq \hat{H}$. Note that we do not
+distinguish between the predicates in $F$ and those in the heads of
+the rules in $R$ (as extensional and intensional predicates), as is
+the case in the model theory for Datalog presented in
+\cite{datalog-basics}; in this respect, our treatment follows that of
+the model theory for Prolog presented in \cite{prolog-semantics}. A
+Herbrand interpretation $\mathcal{I}$ of $\Pi$ is any subset of
+$\hat{H}$, and a Herbrand interpretation $\mathcal{I}$ that satisfies
+the set of clauses in program $\Pi$ is a Herbrand model of $\Pi$,
+written $\models_{\mathcal{I}} \Pi$.
+
+We define the concept of entailment in Datalog as follows. A fact $e$
+is entailed by a program $\Pi$ iff each interpretation satisfying
+$\Pi$ (satisfying every clause in $\Pi$) also satisfies $e$, written
+$\Pi \models e$. For a program $\Pi$ and a set of entailed facts
+$E$, $\Pi \models E$ iff $\Pi \models e$ for every $e \in E$. For
+convenience, we refer to the set of all facts that are entailed by a
+program (the intersection of all the Herbrand models of the program)
+as the {\em entailment of a program}, written $cons(\Pi)$ and defined
+as:
+\begin{align*}
+  cons(\Pi) &= \{ e \in \hat{H} \mid \Pi \models e \} \\
+  &= \bigcap_\mathcal{I} \models_\mathcal{I} \Pi
+\end{align*}
+\end{definition}
+
+\begin{definition}[Justification]
+\labeldef{continuous datalog: justification}
+
+We define a {\em justification} of a set of facts that are entailed by
+a Datalog program as a minimal subset of the program that is
+sufficient for the entailment to hold. More precisely, for a Datalog
+program $\Pi = \langle R, F \rangle$ and a set of entailed facts $E$
+where $\langle R, F \rangle \models E$, the program $\langle J_R, J_F
+\rangle$ with $J_R \subseteq R$ and $J_F \subseteq F$ is a
+justification for $E$ iff $\langle J_R, J_F \rangle \models E$ and,
+for all $J'_R \subset J_R$ or $J'_F \subset J_F$, $\langle J'_R, J'_F
+\rangle \not\models E$.
+
+We simplify this notion of justification by making the assumption that
+a justification can be defined in terms of a minimal subset of the
+ground facts of the program; the rules of the program are therefore
+implicitly considered to be part of every justification. Any
+subsequent references to {\em justification} are to this simplified
+sense, which is defined as follows: given a Datalog program $\Pi =
+\langle R, F \rangle$ and a set of entailed facts $E$ where $\langle
+R, F \rangle \models E$, the set of facts $J \subseteq F$ is a
+justification for $E$ iff $\langle R, J \rangle \models E$ and, for
+all $J' \subset J$, $\langle R, J' \rangle \not\models E$. Note that a
+given set of entailed facts may not have a unique justification.
+
+\end{definition}
+
+\begin{definition}[Time Domain]
+\labeldef{continuous datalog: instant}
+
+Let $\mathbb{T}$ be a discrete time domain with a total order under
+$\leqslant$; an element of $\mathbb{T}$ is an {\em instant}.
+$\mathbb{T}$ has a lower bound $0$, such that for all $i \in
+\mathbb{T}$, $0 \leqslant i$. For convenience, we write $i+n$ where $i
+\in \mathbb{T}$ to indicate the $n$th successor of instant $i$,
+and $i-n$ to indicate the $n$th predecessor of instant $i$.
+
+\end{definition}
+
+\begin{definition}[Fact Instance]
+ \labeldef{continuous datalog: fact instance}
+
+A {\em fact instance} is a fact with some additional temporal
+annotation regarding the instant from which it is recognised to be true;
+it is a tuple containing a fact $f$ and an instant $t \in
+\mathbb{T}$, written $\langle f, t \rangle$. Two fact instances
+containing the same fact and timestamp are considered identical; two
+fact instances containing the same fact but different timestamps are
+considered to be distinct.
+
+We represent the relationship between a fact (without temporal
+annotation) and a fact instance (with temporal annotation) with the
+function $strip()$ which maps a fact instance to its underlying fact:
+\[ strip(\langle f, t\rangle) = f \]
+
+For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
+of fact instances and $F = \{ strip(f) \mid f \in F^T \}$.
+
+\end{definition}
+
+In \refdef{continuous datalog: datalog program}, we took a Datalog
+program to consist of a set of rules and a set of facts.  In order to
+extend this definition of Datalog programs to incorporate fact
+instances, we define a Time-Annotated Datalog program as follows:
+
+\begin{definition}[Time-Annotated Datalog]
+\labeldef{continuous datalog: ta datalog program}
+
+A {\em Time-Annotated Datalog program} (henceforth abbreviated
+TA-Datalog program) $\Pi^T = \langle R, F^T\rangle$ is a tuple which
+comprises a finite set of rules $R$ and a finite set of fact instances
+$F^T \subseteq \hat{H} \times \mathbb{T}$. Fact instances are as
+defined in \refdef{continuous datalog: fact instance}, and rules are
+of the same form as in conventional Datalog programs.
+
+When considering a TA-Datalog program, in which there may be many
+distinct instances of a given fact, temporal information regarding
+when instances of that fact became true would be lost if such a
+program were to entail sets of plain facts without temporal
+annotations. We therefore extend the model theory for Datalog from
+\refdef{continuous datalog: datalog program} by considering the
+Herbrand base of all fact instances $\hat{H}^T$ (all facts with all
+instants). As before, a fact instance $e$ is entailed by a program
+$\Pi^T$ iff each interpretation satisfying every clause in $\Pi^T$
+also satisfies $e$, written $\Pi^T \models e$. The {\em entailment of the
+TA-Datalog program} $\Pi^T$ is therefore:
+\[ cons^T(\Pi^T) = \{ e \in \hat{H}^T \mid \Pi^T \models e \} \]
+
+Informally, the set of instants appearing in some set $E^T$ of fact
+instances entailed by some program ${\langle R, F^T \rangle}$ will be
+a subset of the instants appearing in the fact instances in $F^T$.  In
+addition, the instant $t$ in a fact instance $\langle e, t \rangle$
+that is entailed by a TA-Datalog program $\Pi^T = \langle R^T, F^T
+\rangle$ will be bounded from above by the maximum instant in any fact
+instance in $F^T$. A TA-Datalog program may entail fact instances that
+are only distinguished by their instants (and are therefore considered
+distinct from each other, as per \refdef{continuous datalog: fact
+  instance}).
+
+Consider the following example TA-Datalog program:
+\[
+\Pi^T = \langle \{ c \Leftarrow a \land b \}, \{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle  \} \rangle
+\]
+
+The entailment of the program $\Pi^T$ will be:
+\[cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}\]
+
+It follows that applying the ${strip()}$ function to all the fact
+instances in both the set of fact instances $F^T$ in some ${\Pi^T =
+  \langle R, F^T \rangle}$ and the entailment of the program $\Pi^T$
+(the set of fact instances $cons^T(\Pi^T)$) maintains the relationship
+of entailment; if $\langle R, F^T\rangle \models \langle e, t
+\rangle$, then $\langle R, strip(F^T) \rangle \models strip(\langle e,
+t \rangle)$. This relationship is shown in \reffig{TD-D}.
+
+\begin{figure}[htb]
+  \[
+  \begin{CD}
+\Pi^T = \langle R, F^T \rangle @>\models>> cons^T(\Pi^T) \\
+@V{strip}VV @V{strip}VV \\
+\Pi = \langle R, F \rangle @>\models>> cons(\Pi)
+  \end{CD}
+  \]
+\caption{Relationship between Datalog and TA-Datalog entailments}
+\labelfig{TD-D}
+\end{figure}
+
+Reconsidering the example TA-Datalog program above, we may reduce it
+to the following equivalent conventional Datalog program $\Pi$:
+\begin{align*}
+\Pi &= \langle \{ c \Leftarrow a \land b \}, strip(\{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle \}) \rangle \\
+    &= \langle \{ c \Leftarrow a \land b \}, \{ a, b \} \rangle
+\end{align*}
+Trivially, the entailment of this program $\Pi$ will therefore be:
+\begin{align*}
+cons(\Pi) &= strip(\{\langle c, 1 \rangle, \langle c, 2 \rangle\}) \\
+ &= \{ c \}  
+\end{align*}
+\end{definition}
+
+\begin{definition}[Time-Annotated Justification]
+\labeldef{continuous datalog: instance justification}
+
+The justification of entailments is equally applicable to TA-Datalog
+programs, with facts in the entailment and the justification for that
+entailment being replaced with fact instances.  Following the
+simplified notion of justification in \refdef{continuous datalog:
+  justification}, we define the justification of a set of fact
+instances that are entailed by a TA-Datalog as a minimal subset of
+the ground fact instances of the program that are sufficient for
+the entailment to hold.
+
+More formally, for a TA-Datalog program $\Pi^T = \langle R, F^T \rangle$
+and a set of entailed fact instances $E^T$ where $\langle R, F^T \rangle \models E^T$, the set of fact instances $J^T \subseteq F^T$ is a justification for
+$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime}\rangle \not\models E^T$.
+
+For an entailed fact instance $\langle e, t\rangle$ and a
+justification of that fact instance $J^T$, the instant $t$ at which
+$\langle e, t \rangle$ becomes true is taken to be the latest instant
+at which any of the fact instances in the justification became true:
+\[ t = max(\{ t' \mid \langle j', t' \rangle \in J^T \}) \]
+
+Note that, although an entailed fact instance $\langle e, t\rangle$
+may have multiple justifications, all the justifications for $\langle
+e, t \rangle$ will share the same maximum instant; a justification
+with a different maximum instant $t' \neq t$ would entail a different
+fact instance $\langle e, t' \rangle$. Considering again the example
+TA-Datalog program in \refdef{continuous datalog: ta datalog program},
+the entailed fact instance $\langle c, 2 \rangle$ would have the
+following set of justifications:
+\[
+  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
+\]
+
+\end{definition}
+
+Having defined the notion of time annotation in Datalog, we now define
+the notions of streams and windows before extending to Continuous Datalog.
+
+\begin{definition}[Streams]
+\labeldef{continuous datalog: stream}
+
+A {\em stream} $S$ is an unbounded sequence of fact instances. We
+define a sequence as a mapping from fact instances to instants, where
+the instant records the time at which the fact instance appears on the
+stream. This has the consequence that a particular fact instance
+(i.e. a fact accompanied by a particular instant that is its time
+annotation) may appear only once in a given stream. For convenience,
+we indicate that a fact instance $\langle f, t\rangle$ appears in $S$
+by writing $\langle f, t \rangle \in S$. We indicate that a fact
+instance $f_i$ appears strictly before a fact instance $f_j$ on stream
+$S$ (i.e. $S(f_i) < S(f_j)$) by writing $f_i \prec f_j$.
+
+Streams may be (but are not necessarily) ordered by the instants in
+their fact instances, with the instants non-decreasing in the stream,
+such that for any two fact instances $\langle f_i, t_i \rangle$ and
+$\langle f_j, t_j\rangle$ where $\langle f_i, t_i \rangle \prec
+\langle f_j, t_j \rangle$, $t_i < t_j$; we refer to such
+streams as {\em in-order streams}, and to streams whose fact instances
+are not ordered by their instants as {\em out-of-order streams}.
+\end{definition}
+
+\begin{definition}[Window]
+\labeldef{continuous datalog: window}
+
+A window of range ${r \in \mathbb{N}}$ over a stream $S$ at an instant
+${i \in \mathbb{T}}$ is the \emph{set} of fact instances that appear in
+$S$ in the interval ${(i-r,i]}$, written ${W^{S,r}_{i}}$, and defined as follows:
+\[
+W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
+\]  
+\end{definition}
+
+The following axioms define the semantics of streams and windows over
+those streams in Continuous Datalog:
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
+A window of range $0$ over a stream $S$ at an instant 
+${i \in \mathbb{T}}$ is always empty.
+\begin{equation*}
+W^{S,0}_{i} = \varnothing
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window start leq 0}
+A window of any range $x \in \mathbb{N}$ over a stream $S$ at (or before)
+the first instant of time is always empty.
+\begin{equation*}
+W^{S,x}_{i} = \varnothing
+\end{equation*}
+where $i \leqslant 0$.
+\end{axiom}}
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window composition}
+A window of range ${x > 0}$ over a stream $S$ at an instant 
+${i \in \mathbb{T}}$ is equal to the union of the $x$ consecutive
+windows of range $1$ over the stream $S$ preceding and including that at
+instant $i$.
+\begin{equation*}
+W^{S,x}_{i} = \mathop{\cup}_{j=0}^{x-1} W^{S,1}_{i-j}
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: window disjointness}
+A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
+instant ${i \in \mathbb{T}}$ is disjoint from a second window of 
+range ${y \in \mathbb{N}}$ over $S$ at any instant before $i - x$.
+\begin{equation*}
+W^{S,x}_{i} \cap W^{S,y}_{i-z} = \varnothing
+\end{equation*}
+where $z \geqslant x$.
+\end{axiom}}
+
+Given these definitions, a Continuous Datalog program may be formally
+defined, both in terms of streams and in terms of its instantaneous
+states:
+
+\begin{definition}[Continuous Datalog Program]
+\labeldef{continuous datalog: CDP}
+
+We initially take a {\em Continuous Datalog program} (henceforth a
+C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
+which comprises a finite set of rules $R$, a finite set of static fact
+instances $F^0$ (all of which have the instant $0$) and an in-order
+stream of fact instances $S$.
+
+The model theory for a such a C-Datalog program broadly follows that
+for a TA-Datalog program; we again take the Herbrand base of all fact
+instances $\hat{H}$. Disregarding the order of fact instances in $S$,
+a fact instance $e$ is entailed by a program $\Pi^C$ iff each
+interpretation satisfying every clause in $\Pi^C$ also satisfies $e$.
+However, we take the {\em entailment of the C-Datalog program} $\Pi^C$
+to be an in-order stream $E^S$, where the order in which entailed fact
+instances appear in $E^S$ depends on the order in which the fact
+instances in their justifications appear in $S$. Given that that the
+instant of an entailed fact instance is the maximum of the instants in
+the fact instances that form the justification for that entailed fact
+instance (as in \refdef{continuous datalog: instance justification}),
+$E^S$ will also be an in-order stream.
+
+However, this definition of a C-Datalog program does not take into
+account any notion of windowing over the stream $S$. We define a {\em
+  Windowed Continuous Datalog program} $\Pi^{C, r} = \langle R, F^0, S,
+r \rangle$ to be a tuple with components as defined above for a
+C-Datalog program, but with the addition of $r \in \mathbb{N}$ for the
+range of the window over $S$, which effectively specifies that the
+fact instances from $S$ are to contribute to the entailments of $\Pi^{C,r}$
+from their instant for a duration of $r$.
+
+We consider the entailment of the WC-Datalog program $\Pi^{C,r}$ not
+as a single stream of fact instances (as for C-Datalog), but in three
+parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
+that is the entailment of the TA-Datalog program $\langle R,
+F^0\rangle$; a stream $P$ of fact instances entailed by the rules $R$ and
+the fact instances in $F^0$ and $S$, ordered by the instant at which they
+became justified (i.e. their instants); and a stream $N$ of those same fact
+instances, ordered by the instant at which they ceased to be justified:
+$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
+and $N$ can be considered to be the continuous reasoning equivalents of
+the streams produced by the \emph{IStream} and\emph{DStream} operators
+respectively of the \emph{CQL} continuous query language \citep{CQL},
+which create streams based on the additions to and deletions from a
+relation. Note that $P$ is an in-order stream whose members are a subset
+of those in $E^S$ for an equivalent C-Datalog program, while $N$ may be
+an out-of-order stream.
+\end{definition}
+
+From the definition of entailment for TA-Datalog in \refdef{continuous
+  datalog: ta datalog program} and the definition of windows in
+\refdef{continuous datalog: window}, we can derive the following:
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
+If two sliding windows over the same in-order stream $S$ cover
+non-overlapping time periods, then the set of fact instances entailed
+by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
+set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
+where ${z \geqslant x}$.
+\[
+\{ e \mid \langle R , W^{S,x}_i \rangle \models e \} \cap \{ e \mid \langle R , W^{S,y}_{i-z} \rangle \models e \} = \varnothing
+\]
+\end{axiom}}
+
+\begin{definition}[Instanteous State of a WC-Datalog program]
+\label{def:continuous datalog: CDPt}
+
+We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
+\rangle$ at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
+taking the window $W^{S,r}_i$ over stream $S$ of range $r \in \mathbb{N}$;
+we refer to this as the {\em instantaneous state} of the WC-Datalog program,
+written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle$. As
+previously shown in \refdef{continuous datalog: ta datalog program}, this
+may then be further reduced to the Datalog program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.
+
+The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
+semantically distinct portions: the \emph{persistent} portion is the
+entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
+program, which is guaranteed to be part of the entailment of every state
+due to the monotonicity of positive Datalog; the \emph{transitive} portion
+is the set $E^T_{i}$ of entailed fact instances whose justifications are at
+least partially contained in $W^{S,r}_i$. {\nobreak We define $E^T_i$ as:
+\[
+E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
+\]
+where $J^T$ is a justification for the entailment of x by
+$\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
+  datalog: entailment disjointness}, $E^0$ is guaranteed to be
+disjoint from every $E^T_i$ for every state $\Pi^{C,r}_i$ of every
+WC-Datalog program $\Pi^{C,r}$.
+\end{definition}
+
+\begin{figure*}
+\centering
+$
+\begin{CD}
+  {\Pi^C = \langle R, F, S \rangle} @>\models>> {E^S} \\
+  @V{add\ window}VV \\
+        {\Pi^{C,r} = \langle R, F^0, S, r \rangle} @>\models>> {\langle cons^T(\langle R, F^0\rangle), P, N \rangle} \\
+        @V{evaluate\ window}VV @VV{evaluate\ window}V \\
+        {\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle} @>\models>> {cons^T(\langle R, F^0 \rangle) \cup (W^{P,r}_i \setminus W^{N,r}_i)} \\
+  @VVV @VVV \\
+        {\Pi^T = \langle R, F^T \rangle } @>\models>> {cons^T(\Pi^T)} \\
+  @V{strip}VV @VV{strip}V \\
+        {\Pi = \langle R, F \rangle } @>\models>> {cons(\Pi)}
+\end{CD}
+$
+\caption{Commutative diagram from Continuous Datalog programs to Datalog entailments.}
+\labelfig{CD-TD-D}
+\end{figure*}
+
+The following axioms define the semantics of the streams of
+entailments $P$ and and negations $N$ of a $\Pi^{C,r}$ in terms of the
+instantaneous entailments of the sequence of states of that continuous
+program.
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: positive window increment}
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $P$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that are entailed by the state
+$\Pi^{C,r}_i$ that were not entailed by the state $\Pi^{C,r}_{i-1}$.
+\begin{equation*}
+W^{P,1}_{i} = \left( E^0 \cup E^T_{i} \right) \setminus
+\left( E^0 \cup E^T_{i-1} \right) = E^T_{i} \setminus E^T_{i-1}
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: negative window increment}
+The window of range 1 at any given instant ${i \in \mathbb{T}}$ over
+the stream $N$ entailed by a WC-Datalog program $\Pi^{C,r}$ is defined
+as the set of fact instances that were entailed by the state
+$\Pi^{C,r}_{i-1}$ that are no longer entailed by the state
+$\Pi^{C,r}_i$.
+\begin{equation*}
+W^{N,1}_{i} = \left( E^0 \cup E^T_{i-1} \right) \setminus
+\left( E^0 \cup E^T_{i} \right) = E^T_{i-1} \setminus E^T_{i}
+\end{equation*}
+\end{axiom}}
+
+{\nobreak\begin{axiom}
+\label{axiom:continuous datalog: entailment precedes negation}
+Fact instances may only be negated by some part of their justification
+leaving the sliding window $W^{S,r}_{i}$, this necessarily being after
+they were entailed.  As such, a window of any range ${x \in
+  \mathbb{N}}$ over the stream of entailments $P$ at any given instant
+${i \in \mathbb{T}}$ is disjoint from a window of any range ${y \in
+  \mathbb{N}}$ over the stream of entailment negations $N$ at any
+instant before ${i - x}$.
+\begin{equation*}
+  W^{P,x}_{i} \cap W^{N,y}_{i-z} = \varnothing
+\end{equation*}
+where $z \geqslant x$.
+\end{axiom}}
+
+As noted above in \refdef{continuous datalog: CDP}, the stream of
+negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
+out-of-order stream in which fact instances appear on the stream in
+the order in which they cease to have any complete justifications.
+
+Thus far the translation from instantaneous sets of entailments to
+streams of entailments has been stated axiomatically as the definition
+of those streams.
+
+However, it follows from these axioms that the instantaneous set
+$E^T_{i}$ of instances entailed at any given instant $i$ by any
+$\Pi^{C,r} = \langle R, F^0, S, r \rangle$ may be reconstructed from
+windows of sufficient range over its entailed streams $P$ and $N$ at
+$i$:
+\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
+where ${y \geqslant x \geqslant r}$. The minimum range $x$ is such to achieve
+the \emph{complete} set of entailment instances $E^T_{i}$ of the
+$\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
+\emph{sound} set $E^T_{i}$. As such, a \emph{valid window} that
+contains only and exactly those entailed instances $E^T_{i}$ at every
+instant $i$ may be trivially maintained from the streams $P$ and $N$
+continuously entailed by some $\Pi^{C,r}$.
+
+In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
+F^0, S \rangle$, we can apply to it a window of range $r$ to yield a
+Windowed Continuous Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
+\rangle$. We can evaluate this window at an instant $i \in \mathbb{T}$
+to yield a Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup
+W^{S,r}_i \rangle$, from we which can then strip the time annotations
+to yield a conventional Datalog program $\Pi = \langle R, F
+\rangle$. The relationship between these *-Datalog programs and their
+entailments is expressed in the commutative diagram in
+\reffig{CD-TD-D}.
+
+\end{nestedsection}
+
+\begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
+  CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
+  It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^{C,r} = \langle R,F^0,S,r \rangle \models \langle E^0,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
+  It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
+
+  % Stream-based Data Model
+  \begin{nestedsection}{Data Model}{model: data}
+    The data model of CoDeR is composed of multiple-source multiple-sink streams.
+    These streams are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
+
+    In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^{C,r}$ expressed by a CoDeR-based system.
+    These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
+    an instance of an RDF graph is an RDF graph annotated with a time interval ${(e,n]}$: ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
+    As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.
+
+    The interval ${(e,n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
+    For graphs on the input stream, $e$ is the instant at which a given graph was asserted, and ${n = e + r}$ is the instant at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$ (or become ``negated'').
+    For graphs on the output stream, $e$ is the latest instant at which any part of the justification of a given graph was asserted, and $n$ is the earliest instant at which any part of its justification will leave the sliding window.
+    \reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.3\textwidth]{intersected-intervals}
+      \caption{Derivation of valid-time for entailed fact instance from those of its justification.}
+      \labelfig{intersected-intervals}
+    \end{figure}
+    \begin{align*}
+      \Pi^{C,r}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
+      \Pi^{C,r}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+      \Pi^{C,r}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+      \Pi^{C,r}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
+    \end{align*}
+    It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
+    However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $n$ at which $W^{S,r}_{n}$ \emph{ceased} to justify them, the output stream of a CoDeR-based system also expresses the stream $N$.
+    As such, the annotation of streamed RDF graphs with the interval $(e,n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
+
+    In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
+    These streams are sequences of the intermediate results of processing between steps in the production of entailments.
+    As the production of entailments in CoDeR is cast as an iterative pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
+    an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
+    These instances of RDF graphs could be considered to be ``entailed'' by the sub-pattern of the stream on which they appear and the presence of valid instances of their constituent RDF triples within the system;
+    as such, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
+    It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
+    the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.
+
+    In addition, each stream in the model is semantically distinct:
+    each input/output stream is from a distinct source and uniquely identifiable by the source's IRI;
+    each inter-operator stream contains graphs matching a specific graph pattern and are uniquely identifiable by that pattern.
+    As the same graph pattern may be a sub-pattern of many rule bodies, and so appear in multiple places in a processing plan, one inter-operator stream may be the input to multiple operators.
+    Furthermore, two or more operators may contribute to the same stream within the model, where the pattern of the stream in question is the union of the patterns of the streams produced by each of the operators in isolation.
+    In this case, each operator contributes results to the common stream as they are produced, passively interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment instant $e$ of the constituent instances.
+    As this is the active behaviour of a stream union operator, such an operator is not needed in CoDeR.
+    However, within the algebra of CoDeR, such a stream is still denoted ${op_{application} \cup op_{application}}$.
+  \end{nestedsection}
+
+  % Minimal Algebra
+  \begin{nestedsection}{Minimal Algebra}{model: algebra}
+    The minimal algebra of CoDeR is based on that of Positive Datalog, that being ${\{\bot, \text{:-}, \wedge\}}$.
+    Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_{head}$ and the latter being the window-join $\Join$ \citep{niagaraCQ}, in order to support forward-chaining processing of fixed rules;
+    $\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}.
+    % Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
+    Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, which we represent algebraically by $\rstreamjoin$.
+    % In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
+    In addition to $\text{:-}_{head}$ and $\rstreamjoin$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
+    Finally, as basic patterns match individual triples and CoDeR's streams are composed of graphs, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate input streams to the streams of triples on which $\sigma_{triple}$ operates.
+    % In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
+    In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, \rstreamjoin, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
+
+    As all operators are stream-to-stream in nature, they each take one or two streams as input and produce a single stream (though they may contribute to any number of union streams).
+    As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
+    For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^{C,r} = \langle \{b \Leftarrow a\}, \{\}, S, r \rangle}$, where $a$ and $b$ are RDF triples.
+    \begin{equation}\labeleqn{basic-example-plan}
+      \text{:-}_b \left( \sigma_a \left( {}_g{\pi_t}\,S \right) \right)
+    \end{equation}
+
+    \begin{description}
+      \item[$\sigma_{triple}\,X$] \hfill \\
+        \begin{figure}[t]
+          \centering
+          \includegraphics[width=0.3\textwidth]{basic-pattern-match}
+          \caption{The triple-select operator.}
+          \labelfig{continuous datalog: basic pattern match}
+        \end{figure}
+        The triple-select operator identifies instances of triples in a stream $X$ that match some \emph{basic pattern} \citep{w3csparql}.
+        This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to semantic data.
+        Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the basic pattern to be matched.
+        It takes an inter-operator stream $X$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+        The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
+      \item[$\sigma_{external}\,X$] \hfill \\
+        \begin{figure}[b]
+          \centering
+          \includegraphics[width=0.3\textwidth]{datatype-function}
+          \caption{The operator encompassing all datatype functions of RIF-Core.}
+          \labelfig{continuous datalog: datatype function}
+        \end{figure}
+        This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
+        It filters elements of its input stream $X$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
+        thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern}, as specified in SPARQL.
+        As such, it takes an inter-operator stream $X$ of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
+        The contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
+      \item[${{}_g{\pi_t}\,X}$] \hfill \\
+        \begin{figure}[t]
+          \centering
+          \includegraphics[width=0.3\textwidth]{graph-to-triple}
+          \caption{The graph-to-triple operator.}
+          \labelfig{continuous datalog: graph-to-triple}
+        \end{figure}
+        The graph-to-triples operator ${{}_g{\pi_t}}$ is a unary operator that deconstructs graph instances into sequences of triple instances, the union of the triples instantiated therein being equal to the contents of the original graph.
+        It takes a heterogeneous (\emph{input/output}) stream $X$ of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of RDF triple instances.
+        Each triple instance produced is assigned the same valid-time interval as the graph instance in whose graph the instantiated triple was contained.
+        Consider, for example, the input/output stream containing an instance of a graph containing the RDF triples $a$ and $b$ and valid from instant $t_e$ to instant $t_n$:
+        \begin{multline*}
+          {}_g{\pi_t} (\dots,\langle \{a,b\},t_e,t_n \rangle,\dots) = \\
+            (\dots,\langle a,t_e,t_n \rangle,\langle b,t_e,t_n \rangle,\dots)
+        \end{multline*}
+      \item[$\text{:-}_{head}\,X$] \hfill \\
+        \begin{figure}[b]
+          \centering
+          \includegraphics[width=0.3\textwidth]{instance-implication}
+          \caption{The instance implication operator.}
+          \labelfig{continuous datalog: instance implication}
+        \end{figure}
+        Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^{C,r}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
+        Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
+        This is similar to the CONSTRUCT clause of SPARQL, or the final step of \emph{conflict set resolution} within the Rete algorithm \citep{forgy79}.
+        In addition, each produced instance inherits its valid-time interval from the graph instance that implied it.
+      % \item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
+      \item[${X\,\rstreamjoin\,Y}$] \hfill \\
+        \begin{figure}[t]
+          \centering
+          \begin{minipage}[b]{0.45\textwidth}\includegraphics[width=\textwidth]{SteM}\end{minipage}\\\hfill\\
+          \begin{minipage}[b]{0.45\textwidth}\includegraphics[width=\textwidth]{window-join}\end{minipage}
+          \caption{Exploded views of two-pass joining operators.\hspace{\textwidth}Top: the SteM operator. Bottom: a typical window-join.}
+          \labelfig{continuous datalog: SteM}
+        \end{figure}
+        State Modules (\emph{SteM}s) are asymmetrical binary operators that each perform the dual function of maintaining a window of graph instances and ``joining'' each graph instance on a stream against those within the maintained window (known as ``probing'' the window).
+        These are the same two functions performed for each window in a symmetric window-join, extended from Rete \emph{beta nodes} in \citep{ReteDBMS}, with results of each probe being produced to the same result stream as they occur.
+        Window-joins are, therefore, functionally identical to the union of two complementary SteMs:
+        % ${A \Join B = \left( A\,{}_b{\Join_p}\,B \right) \cup \left( B\,{}_b{\Join_p}\,A \right)}$, where $A$ and $B$ are inter-operator streams.
+        ${A \Join B = \left( A\,\rstreamjoin\,B \right) \cup \left( B\,\rstreamjoin\,A \right)}$, where $A$ and $B$ are inter-operator streams.
+        In addition, it should be noted that SteMs express the semantics of the sequential intersection operator ${seq}$ proposed in EP-SPARQL \citep{EP-SPARQL}, in that they join new data from one stream with the prior (valid) contents of another, thereby supporting rudimentary event processing.
+
+        The implementations of SteMs in the literature show a great variety of behaviour regarding window maintenance \citep{SteMs}, in the most extreme case ceasing to be a wholly stream-to-stream operator and supporting the incorporation of ``static'' data into stream processing systems.
+        In the minimal algebra of CoDeR, however, SteMs simply support the building of a \emph{valid window} (as defined in \refsec{semantics}) from its first inter-operator stream $X$, and the probing of that window with data from its second inter-operator stream $Y$ to produce an inter-operator stream of the homogeneously joined graph instances.
+        Furthermore, the nature of probing in CoDeR is strictly limited to that of the natural join between the variables in the patterns for streams $X$ and $Y$.
+        Finally, in contrast to the direct inheritance of valid-times in unary operators of CoDeR, the valid-time interval assigned to each graph produced by a SteM is the intersection of the valid-times of the two joined graphs.
+    \end{description}
+
+    % As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+    As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+    Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
+    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^{C,r} = \langle R,F^0,S,r\rangle}$.
+    It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
+    It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
+    i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
+  \end{nestedsection}
+\end{nestedsection}
+
+\begin{nestedsection}{R4 System: Rule-based Reasoning over RDF streams using Rete}{implementation}
+  As a proof of concept for the CoDeR model, we have implemented R4: a Rule-based Reasoner for RDF-streams using Rete.
+  Reasoning in R4 is performed continuously and incrementally as dataflow networks that work directly on RDF streams, constructed according to the Rete pattern matching algorithm \citep{forgy79}.
+  The system architecture is shown in \reffig{R4-architecture}.
+  \begin{figure}[b]
+    \centering
+    \includegraphics[width=0.45\textwidth]{R4-architecture.png}
+    \caption{The system architecture of R4.}
+    \labelfig{R4-architecture}
+  \end{figure}
+
+  \begin{figure*}[t]
+    \centering
+    \begin{ebnf}
+      \Rule{Document}{
+              \Optional{\NTerm{IRIMETA}} 
+              \Token{Document} 
+              \Token{(} 
+              \Optional{\NTerm{Base}} 
+              \ZeroOrMore{\NTerm{Prefix}} 
+              \ZeroOrMore{\NTerm{Import}}
+              \Optional{\NTerm{Window}}
+              \Optional{\NTerm{Group}} 
+              \Token{)}
+      }
+      \Rule{Base}{
+              \Token{Base} 
+              \Token{(} 
+              \NTerm{ANGLEBRACKIRI} 
+              \Token{)}
+      }
+      \Rule{Prefix}{
+              \Token{Prefix} 
+              \Token{(} 
+              \NTerm{Name} 
+              \NTerm{ANGLEBRACKIRI} \Token{)}
+      }
+      \Rule{Import}{
+              \Optional{\NTerm{IRIMETA}} 
+              \Token{Import} 
+              \Token{(} 
+              \NTerm{LOCATOR} 
+              \Optional{\NTerm{PROFILE}} 
+              \Token{)}
+      }
+      \Rule{Group}{
+              \Optional{\NTerm{IRIMETA}} 
+              \Token{Group} 
+              \Token{(} 
+              \ZeroOrMore{\OrList{\NTerm{RULE} \OrSep \NTerm{Group}}} 
+              \Token{)}
+      }
+      \Rule{RULE}{
+              \Optional{\NTerm{IRIMETA}}
+              \Token{Forall}
+              \OneOrMore{\NTerm{Variable}}
+              \Token{(}
+              \NTerm{CLAUSE}
+              \Token{)}
+      }
+      \OrRule{
+              \NTerm{CLAUSE}
+      }
+      \Rule{CLAUSE}{
+              \OrList{\NTerm{Implies} \OrSep \NTerm{ATOMIC}}
+      }
+      \Rule{Implies}{
+              \Optional{\NTerm{IRIMETA}}
+              \OrList{\NTerm{ATOMIC} \OrSep \Token{And} \Token{(} \ZeroOrMore{ATOMIC} \Token{)}}
+              \Token{:-}
+              \NTerm{FORMULA}
+      }
+      \Rule{LOCATOR}{
+              \NTerm{ANGLEBRACKURI}
+      }
+      \Rule{PROFILE}{
+              \NTerm{ANGLEBRACKURI} 
+      }
+      \Rule{Window}{
+              \NTerm{Number}
+              \NTerm{TimeUnit}
+      }
+      \Rule{TimeUnit}{
+              \OrList{\Token{ms} \OrSep \Token{s} \OrSep \Token{m} \OrSep \Token{h}}
+      }
+    \end{ebnf}
+    \caption{Streaming RIF-Core Grammar}
+    \labelfig{lst-S-RIF-Core}
+  \end{figure*}
+
+  A rule document containing any number of rules is submitted to the system.
+  We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
+  The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
+  We also added a simple extension to RIF-Core in line with the extension of Datalog described in \refsec{semantics}, enabling users to specify a window range as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
+  Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
+
+  These rules are translated by the rule parser into a set of objects that are then used by the network optimizer to generate the Rete-like dataflow networks.
+  The optimizer employs simple known heuristics to generate a good plan.
+  These heuristics include: sharing nodes between rules where the sub-graph patterns their results match are variants of each other; avoiding Cartesian products as much as possible by joining patterns that have common variables earlier in the beta network; and pushing more selective patterns (those with fewer variables) earlier in the alpha network to minimize intermediate results.
+
+  The optimizer instantiates a Rete network for background reasoning if required and another network for generic rules.
+  The first network feeds into the second one and the two together are ``re-entrant'', meaning that the first network takes as input the stream of entailments produced by both the first and the second, to enable iterative inference of results such as the calculation of transitive closure.
+
+  These networks receive streaming data in the form of sequences of RDF graphs, and operate directly on each of these graphs incrementally.
+  We chose this RDF-native approach as opposed to reusing existing technologies as in \citep{C-SPARQL,streaming-sparql} to allow full control over the low-level operators.
+  This can offer maximum optimization opportunities such as adaptively optimizing the network topologies, which is part of our future work.
+
+  \begin{nestedsection}{Continuous Reasoning with Rete}{implementation: continuous rete}
+    The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, can be well fit into the stream reasoning model as it operates incrementally.
+    The RETE algorithm can process large data sets efficiently because it avoids iterating over both data elements (facts, or ``working memory'') and over the production rules.
+    To avoid iteration over data elements, the RETE algorithm stores with each condition (or pattern), a list of the data elements that it matches.
+    These lists are updated when the working memory changes, in a forward chaining process.
+    To avoid iteration over the production rules, the RETE algorithm creates a dataflow tree-structured network to represent the rules.
+
+    However, the Rete algorithm is memory-intensive as it stores all partial matches, trading memory for speed.
+    In a streaming context, this is infeasible as streams can grow without bounds.
+    Therefore, we place time constraints on the memories using a window-join operator instead of the traditional incremental joins.
+  \end{nestedsection}
+  \begin{nestedsection}{Continuous Reasoning in the Semantic Web}{implementation: continuous reasoning with RDF}
+    \begin{figure*}[t]
+      \centering
+      \includegraphics[width=0.9\textwidth]{example-rete-network.png}
+      \caption{An example Rete network with corresponding (RDFS) rules.}
+      \labelfig{example-rete-network}
+    \end{figure*}
+    In R4, data enters the network through the source nodes, which convert streams of graphs into streams of quads, by extracting the graph ID and attaching it to every triple in this graph.
+    Source nodes are, therefore, implementations of the graph-to-stream operator ${{}_g{\pi_t}}$ of CoDeR.
+    Source nodes also annotate the incoming quads with the system time at their arrival if their parent graph was not already annotated at its origin, otherwise passing the annotation of the graph to the resulting quads.
+    This annotation $a$ relates to their instant of assertion $e$ within the CoDeR data model, and the instant $n$ at which they will leave the window whose range $r$ is defined in the system's rule set may be derived by ${n = a + r}$.
+    Source nodes propagate the annotated quads to their successor nodes, which are the alpha nodes.
+
+    Alpha nodes are single-entry nodes that form a discrimination network, as shown in \reffig{example-rete-network}.
+    In R4, alpha nodes can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
+    Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to its successor node(s) in the case of a match.
+    As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
+
+    The first alpha node in the network propagates matched quads to a special node called the left input adaptor, while the other alpha nodes propagate their output directly to beta nodes.
+    The left input adaptor node is responsible of preparing the incoming quads to enter the beta network from the left input of the first beta node in a given join-tree.
+    It creates a new \emph{token} that represents a partial match, defined as a list of quads, for each rule body for which the incoming quad is a sub-graph match, then adds this quad as the first item in each list.
+    It also annotates each token with the same time as was the quad for which it was generated.
+    Tokens are then sent to the first nodes in the beta network.
+
+    Beta nodes are two-input nodes that are responsible for joining the branches of the alpha network, as demonstrated in \reffig{example-rete-network}.
+    In our implementation, as in the original Rete algorithm, beta nodes form a left-deep tree.
+    Each beta node maintains a left memory, which is a beta memory storing tokens received from other beta nodes (or from the left input adaptor node in case of the first beta node) and a right memory, which is an alpha memory storing quads received from alpha nodes.
+    Join nodes are beta nodes that match graphs from both sides according to some conditions, e.g. a shared variable binding.
+    As explained earlier, we use stream-to-stream window-join operators to avoid storing and operating over all partial results.
+    In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
+    These windows are implemented as priorities queues in which quads are ordered according to their time annotations, and pruned by ``popping'' from the queue and deleting any quads whose annotation $a \leqslant i - r$, where $i$ is the current system time and $r$ is the range of the semantic window.
+
+    When a join node is left activated, i.e.
+it receives a token through its left input, it first adds the new token to its left window, then it prunes its right window before iterating over the remaining quads to find any matches.
+    When a match is found, a new token is created by duplicating the left token and adding the right quad to the new token's quad list.
+    The new token is annotated with the \emph{earliest} time with which its component quads were annotated, and then propagated to the next beta node, or to the terminal node if it is in the root node of a join tree.
+    Conversely, when the join node is right activated by receiving a quad from an alpha node, it adds the new quad to the right window, prunes the left window then matches the quad against the remaining tokens.
+
+    This method of annotating tokens with the earliest assertion time of those of its components initially appears to violate the annotation semantics for entailed fact instances laid out in \refsec{semantics,model}.
+    However, the ordering of tokens in their streams and the system time $i$ at which they are propagated through the system expresses implicitly CoDeR's entailment time $e$ of entailed instances.
+    This leaves the annotation $a$ of a token in R4, in combination with the range $r$ of the semantic window associated with the rule set, able to express CoDeR's negation time ${n}$ of entailed instances by the same calculation as those of asserted instances: ${n = a + r}$.
+    As such, the interpretation of beta nodes in R4 is a valid implementation of the window-join operator characterised as a pair of CoDeR SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
+
+    Finally, terminal nodes receive tokens from the root nodes of join-trees and are responsible for producing entailed graphs, directly implementing the instance entailment operator $\text{:-}_{head}$ of CoDeR.
+    As with the $\text{:-}_{head}$ operator, the annotation of these entailed graphs is directly inherited from the completed token from which it is produced.
+    The union of the output of all terminal nodes is both the output of the system itself, and re-entered as an input stream to the source nodes to support iterative inference.
+  \end{nestedsection}
+\end{nestedsection}
+
+\begin{nestedsection}{Evaluation, Results and Discussion}{evaluation}
+  To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning to some extent on semantic streaming data.
+  These include Etalis \citep{EP-SPARQL}, Sparkwave \citep{sparkwave}, Streaming knowledge bases \citep{walavalkar08streamingkb}, and the incremental reasoner presented in \citep{inc-reasoning-background-knowledge}.
+  To the best of our knowledge, the latter two implementations were never made public, so we compare R4 against Etalis and Sparkwave only.
+  The experiments are performed on an Intel Core i5 computer with 3.2 GHz and 8 GB memory.
+  We set up Jtalis (the Java wrapper of Etalis) over SWI-Prolog v7.2.1 and installed Sparkwave v0.5.1.
+
+  \begin{nestedsection}{Motivating Case}{evaluation: motivating case}
+    \begin{figure*}[t]
+      \centering
+      \begin{verbatim}
+Forall ?obs ?sensor ?result ?value ?v ?hur ?hurWind (
+    If And( ?obs [ssn:observedProperty -> wind_speed]
+        ?obs [ssn:observedBy -> ?sensor]
+            ?obs [ssn:observationResult -> ?result] 
+            ?result [ssn:hasValue -> ?value]
+            ?value [ssnExt:hasQuantityValue -> ?v]
+            ?hur [rdf:type -> yago:Hurricane111467018]
+            ?hur [dbpprop:1MinWinds -> ?hurWind] 
+            External (pred:numeric-greater-than(?v ?hurWind)) )
+    Then ?obs [ssn:observedWeatherType -> yago:Hurricane111467018] )
+      \end{verbatim}
+      \caption{The RIF-Core rule inspired by SRBench.}
+      \labelfig{SRBench-rule}
+    \end{figure*}
+
+    We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
+    Inspired by a query in \citep{SRBench}, we formulated a rule that entails that a given observation is of a weather pattern classified as a hurricane if it is of a wind speed that exceeds that of a previously classified hurricane;
+    it also entails that the wind speed observed exceeds that of a given historically recorded hurricane (\reffig{SRBench-rule}).
+    This rule involves dealing with static data regarding past hurricanes, dynamic data regarding ongoing sensor readings, and also needs to perform reasoning against an ontology of background knowledge to identify the hurricanes that are instances of the specified class of hurricanes.
+    For the historical hurricane data, we prepared a small dataset taken from dbpedia, containing information about a number of hurricanes.
+
+    The experiment consisted of pushing the whole stream to the compiled R4 system and measuring the time taken to process all the data, repeated for a variety the window ranges.
+    We ran this rule successfully on R4 and Jtalis getting correct results.
+    However, Sparkwave was unable to read the dataset correctly as it does not parse blank nodes;
+    this is because it uses the hash-join algorithm to improve time efficiency, as we show in \refsec{evaluation: comparative results}, but builds the hashtables based on the URIs of the incoming data as which blank nodes are not valid.
+  \end{nestedsection}
+  \begin{nestedsection}{Comparative Results}{evaluation: comparative results}
+    \begin{figure*}[t]
+      \centering
+      \begin{verbatim}
+Forall ?offer ?product ?vendor ?price ?from ?to ?delivery ?webpage(
+    If And( ?offer [rdf:type -> bsbm_voc:Offer]
+            ?offer [bsbm_voc:product -> ?product]
+            ?offer [bsbm_voc:vendor -> ?vendor]
+            ?offer [bsbm_voc:price -> ?price]
+            ?offer [bsbm_voc:validFrom -> ?from]
+            ?offer [bsbm_voc:validTo -> ?to]
+            ?offer [bsbm_voc:deliveryDays -> ?delivery]
+            ?offer [bsbm_voc:offerWebpage -> ?webpage]
+            ?offer [dc:publisher -> ?publisher]
+            ?offer [dcc:date -> ?date] )
+    Then ?product [bsbm_voc:offerPrice -> ?price] )
+      \end{verbatim}
+      \caption{The RIF-Core rule inspired by the Berlin SPARQL benchmark.}
+      \labelfig{BSBM-rule}
+    \end{figure*}
+
+    In order to compare the efficiency of R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+    We used another 1.1 million triples dataset, this time using the provided data generator, containing information about 100,000 limited-time offers made available by some fictional online marketplace.
+    We used a small schema that has 329 product types arranged in a 4-level hierarchy.
+
+    The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
+    As offers may be associated with product super-types, some small degree of background reasoning is needed to determine the offer price for specific product sub-types.
+
+    Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
+    However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
+    Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
+    In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating Etalis to use \emph{seq} instead of \emph{And}.
+    It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time-efficiency and memory-consumption of the three systems.
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
+      \caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given windows of various ranges over the input stream.}
+      \labelfig{all-systems-varying-windows}
+    \end{figure}
+
+    As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small windows but slower than Sparkwave, and, while the time-to-complete increases with window range in both Sparkwave and R4, that of Etalis appears to remain fairly constant for the window ranges tested.
+    The growing difference between R4 and Sparkwave as the range of the windows increases is likely due to the fact that Sparkwave uses a hash-join algorithm with garbage-collection window-maintenance while the current implementation of R4 uses loop-joins over lists maintained by indexing over time annotations;
+    as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer to iteratively check for matches compared to the hash-based look-up of those in Sparkwave.
+    More time-efficient, multi-index window-joins are to be considered for R4 in future work.
+
+    We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
+    Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
+    It shows that, while R4 and Jtalis follow similar patterns, Sparkwave's memory needs are significantly higher and continue to increase over time, a behaviour that should not occur when applying windows over streams.
+    Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to an interaction between the manual garbage collection performed by Sparkwave and the underlying JVM garbage collector.
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
+      \caption{A comparison of the memory consumed by the systems R4, Sparkwave and Etalis throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
+      \labelfig{all-systems-memory-consumption}
+    \end{figure}
+  \end{nestedsection}
+\end{nestedsection}
+
+\begin{nestedsection}{Conclusions}{conclusions}
+  We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
+  In this way it differs from other temporal interpretations of Datalog \citep{OrgunWadge92,Tuzhilin93} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing and casting a continuous program as a series of discrete states.
+  WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.
+  In addition to this, however, it also supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program via the application of windows to the concise streams, leading to the well-founded definition of a \emph{valid window} over the combined streams $P$ and $N$, being the outputs of the reasoning equivalents of the \emph{IStream} and \emph{DStream}.
+  Thus WC-Datalog represents a well-founded semantics not only for continuous reasoning, but for coherent and concise streamed entailments of continuous programs, suitable for consumption in further stream processing.
+
+  In order to express this semantics we have developed CoDeR, a processing model for Continuous Deductive Reasoning over RDF.
+  This processing model is composed of a stream-based data model with a dual semantics and a stream-to-stream extension of a semantic interpretation of the relational algebra.
+  The data model is composed of streams of RDF graphs annotated both with the time at which they became true and the time at which they will cease to be true, with the ordering of graphs by the former being the sequence of facts in $P$ and the ordering by the latter being the sequence of facts in $N$, hence providing a dual semantics for the stream.
+  As such, the window inherent in a continuous program may be represented in the annotation of each graph in the streamed input, rather than ranged window operators in the algebra.
+  Instead, the algebra makes use of \emph{valid window} operators to maintain the set of graphs that are both considered to be true at a given moment \emph{and} applicable at a given point in the processing.
+  It also de-constructs the conventional window-join operator of stream-to-stream algebras for continuous querying into the union of the outputs of complementary State Modules (\emph{SteMs}) \citep{SteMs}, to which it is logically and functionally equivalent, thereby supporting greater variety in the possible plans that express a given continuous program.
+
+  Finally, we have implemented the CoDeR model in the synchronous RDF stream-reasoning system R4.
+  R4 utilises a variant of the Rete pattern matching algorithm \citep{forgy79} to produce processing plans for sets of rules expressed in RIF-Core \citep{w3crifcore}, the W3C recommendation for expressing sets of rules with the semantics of positive Datalog and the rule language used to express the set of entailment rules of OWL 2 RL \citep{w3cowl2profiles}.
+  R4 is sound and complete with respect to the semantics of WC-Datalog by virtue of implementing CoDeR.
+  Given that many state-of-the-art stream-querying and continuous-reasoning systems have varying semantics, and so the expected results of each system will vary for the same query definition, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and Etalis \citep{EP-SPARQL}.
+  We have demonstrated that R4 is comparable in both respects to these RDF stream-querying systems when applying low-range windows to high-throughput streams, despite Sparkwave and Etalis supporting more limited/less explicitly well-founded reasoning than that of R4.
+
+  Moving forward, we plan to improve the time-efficiency of R4 by investigating more efficient join algorithms within its implementations of SteMs.
+  At a higher level, we also plan to investigate alternate planing algorithms to improve the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
+reduce the latency from receipt of some new graph to the production of entailments justified in part by that new graph.
+  These primarily focus on the use of SteMs outside of the composition of window-joins, either taking an adaptive approach \citep{eddies,CACQ,CQELS} or utilising the wider array of static plans such an operator affords.
+\end{nestedsection}
+
   \section*{References}
-	\bibliographystyle{template/elsarticle-harv}
+	\bibliographystyle{elsarticle-harv}
 	\bibliography{bibliography}

### 2015-11-03 article.tex
clarified the distinction between general/particular knowledge and static/streamed data, and that between general/particular knowledge and e.g. TBox/ABox

@@
   In addition, this division allows the application of a single domain of general knowledge to a variety of distinct sets of particular knowledge, and the application of many distinct domains of general knowledge to a single set of particular knowledge.
-  This is a particularly useful characteristic in the context of semantic stream processing, in which all current use cases consider the application of some static domain of knowledge to some stream of particular knowledge;
+  It should be noted that this division of general and particular knowledge in the expression of a reasoning problem depends on the generalisation of that problem:
+  in Description Logics, it may be considered that the TBox of a knowledge base describes general knowledge and the ABox describes particular knowledge;
+  however, the fundamental rules by which entailments are derived from knowledge bases by Description Logics constitute a more general knowledge than that of the TBox of some knowledge base, by which the entirety of a knowledge base may be considered particular knowledge.
+  In the latter case, a distinction may still be made between a TBox and ABox in the context of streaming, with the former being expressed in static data and the latter being streamed or vice versa, though this is dependent on the rate at which the TBox and ABox are expected to change.
+  In the existing stream processing literature, the vast majority of current use cases consider the application of some \emph{specific} static domain of general knowledge (e.g.
+a specific TBox) to some stream of particular knowledge;
   thus, with stream reasoning the volume of streamed data may be dramatically reduced by extracting the static domain of general knowledge that underpins the data and providing that domain as a static ontology, then streaming only the minimal particular knowledge that would be necessary to deduce the full data set.

### 2015-11-03 article.tex
split mention of state of the art from brief description/problems, for clarity

@@
   %   RDF stream processing and reasoning
-  The RDF stream processing community has proposed extensions to their semantic query languages to support the iterative application of queries to the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}.
+  The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the iterative application of queries to the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}.
+  Meanwhile, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
+
   %   issues with existing art: no explicit semantics for reasoning
-  While this approach is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, the query-by-query windowing semantics of the base systems do not translate to a coherent semantics for multi-rule stream reasoning.
-  On the other hand, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
-  While these algorithms do support a coherent semantics for continuous reasoning, this semantics casts the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams, and so preclude the further processing of entailments in a stream-processing manner.
-  As such, stream reasoning in its current form either lacks a coherent semantics for continuous truth and entailment, or is necessarily a terminal rather than intermediate process in any streamed data pipeline.
+  While the iterative application of queries to the results of others is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, the query-by-query windowing semantics of the base stream-querying systems do not translate to a coherent semantics for multi-rule stream reasoning.
+  This is because the windowing semantics proposed for stream querying involve applying the entirety of a ``problem'' (be it a single query or a unified body of general knowledge) to a bounded subset of the data in a stream (a window), henceforth referred to as \emph{global windowing};
+  the iterative application of individually windowed queries both introduces and \emph{fails to properly consider} the semantics of Complex Event Processing, leading to inconsistencies between the results of identical rule sets expressed in the cited systems \citep{LARS}.
+
+  Truth-maintenance algorithms, on the other hand, do support a coherent semantics for continuous reasoning.
+  However, this semantics casts the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams.
+  Without research to extend these semantics to entail concise streams that maintain in their own right the coherence of the existing semantics, these algorithms preclude the further processing of entailments in a stream-processing manner.
 

### 2015-11-03 article.tex
introduced LARS and explained what it is meant for and why it is not a suitable foundation for new approaches to stream reasoning

@@
   Without research to extend these semantics to entail concise streams that maintain in their own right the coherence of the existing semantics, these algorithms preclude the further processing of entailments in a stream-processing manner.
+
+  As such, until recently, stream reasoning either lacked a coherent semantics for continuous truth and entailment, or was necessarily a terminus rather than intermediate process in any streamed data pipeline.
+  However, the work by \citet{LARS} has produced the LARS framework, by which the semantics of stream processing systems may be expressed and formally analysed.
+  LARS extends the semantics of Answer Set Programming (itself an extension of Datalog semantics with features such as Negation as Failure) with the notion of streamed data with a point-in-time semantics.
+  Paired with this is the notion of a logical windowing operator that supports the expression of windows as part of the general knowledge of a reasoning problem, as is necessary to express reasoning as performed by the iterative application of continuous queries in the manner already discussed.
+  Though this form of windowing also allows the expression of temporally-aware rules in the manner of Complex Event Processing, it does \emph{not} align with or support the original intent of \emph{global windowing} in stream querying.
+  Furthermore, the expression of granular windowing such as that expressible in LARS has not been defined in existing Semantic Web reasoning languages such as RIF \citep{w3crif} and OWL \citep{w3cowl2}.
+  Finally, the complexity of entailment over programs as defined in LARS makes it unsuitable as a semantics on which to \emph{base} the expression of reasoning in real-time:
+  the complexity of entailment in LARS is PSPACE, or co-NP given restrictions on window-nesting, where real-time processing of a task requires that task to be at most polynomial in its time complexity.
 

### 2015-11-03 article.tex
refinement of the description of the Rete algorithm in reference to stream reasoning and semantics

@@
   % Rete as non-streaming implementationexample (application to dataflow networks)
-  Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
+  Progressing from reasoning semantics to data processing, Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
   These systems send data from operator to operator, with `memories' compiled for two-pass operators in which to retain the specific data received, in much the same way as sliding windows are used in stream processing.
-  This makes them functionally similar to many stream-to-stream processing systems, and have even been a foundation for some such as Sparkwave \citep{sparkwave}.
-  However, as \citep{forgy79} does not consider the semantics of stream processing, and those semantics are not clearly defined elsewhere for reasoning, such a semantics must still be defined before any stream reasoning system based on Rete may claim to adhere to them.
+  This makes them functionally similar to many stream-to-stream processing systems, and they have even been a foundation for some such as Sparkwave \citep{sparkwave}.
+  However, \citet{forgy79} did not incorporate the semantics of stream processing into the model underlying the Rete algorithm.
+  In addition, more recent data-flow-based models either don't consider ``reasoning'' more expressive than RDFS \citep{sparkwave}, or are based on rule/query-level windowing \citep{C-SPARQL,walavalkar08streamingkb}, with problems as already discussed.
 

### 2015-11-03 article.tex
extended the summation of contributions to more clearly lay out the scope/aim of the semantics and express its advantages re: integration with existing expressions of Semantic Web reasoning

@@
   % overview
-  In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Continuous Datalog.
-  As positive Datalog is also the semantic equivalent of RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, this continuous semantics may be applied to the continuous application of OWL 2 ontologies to streamed RDF data.
-  Next, we present the abstract processing model CoDeR that provides for the expression of processing plans that satisfy the semantics of Continuous Datalog.
+  In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Windowed Continuous Datalog.
+  The scope of this semantics is to be an expression of inference, as currently described by Semantic Web reasoning languages, that incorporates streamed semantic data and is of low enough computational time-complexity to be feasibly computed continuously in real-time.
+  As entailment in positive Datalog has polynomial-time complexity and is the equivalent of entailment in RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, the continuous semantics of Windowed Continuous Datalog may be used to apply OWL 2 ontologies to streams of RDF data.
+  Next, we present the abstract processing model CoDeR that provides for the expression of processing plans that satisfy the semantics of Windowed Continuous Datalog.
   We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
   Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems Etalis and Sparkwave.
-  Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Continuous Datalog.
+  Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Windowed Continuous Datalog.
 \end{nestedsection}

### 2015-11-03 article.tex
MINOR: added the word "windowing" before "semantics" in start of WC-Datalog

@@
 degree of expressivity; not sacrifice the precise semantics thereof
-whilst incorporating the semantics of stream processing; and allow for
+whilst incorporating the windowing semantics of stream processing; and allow for
 the continuous derivation of entailments in a concise manner.

### 2015-11-03 article.tex
MINOR: removed "noindent" and improved wording of what WC-Datalog provides for

@@
 programs.  As such, these states each inherit the semantics of
-positive Datalog.  Furthermore, it involves the continuous derivation
+positive Datalog.  Furthermore, WC-Datalog provides for the continuous derivation
 of entailments from this sequence of instantaneous states as a stream
@@
 rules are of the form:
-
 \[ \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n \]
-
-\noindent where $\alpha_i$ are atomic formulae. Note that we have excluded any
+where $\alpha_i$ are atomic formulae. Note that we have excluded any
 universally quantified variables for clarity.

### 2015-11-03 article.tex
NEEDS WORK: referenced RSP community definition of streams, needs reason for using a subsumed semantics

@@
 of entailments from this sequence of instantaneous states as a stream
-of facts; these streams share a data model with the streams of
-continuous queries, as defined by the stream processing community,
-thereby allowing the integration of WC-Datalog programs into stream
-processing pipelines. Finally, we show that these streams may be
+of facts; the data model of these streams is subsumed by that currently proposed
+by the RDF stream processing community\footnote{As presented at
+\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}
+on \mbox{03/11/2015}.}, thereby allowing the integration of WC-Datalog programs
+into stream processing pipelines. Finally, we show that these streams may be
 manipulated to retrieve the instantaneous sets of entailments produced

### 2015-11-03 article.tex
Major reshuffle of TA-* section, bringing justification ahead of examples, and defining satisfiability of datalog rules by fact instances.  Also neatened up markup (mostly using \prime instead of ')

@@
 set of clauses, divided into a finite set of rules $R$ and a finite
-set of ground facts $F$. Facts are formulae of first-order logic, and
+set of ground facts $F$. Facts are atomic formulae of first-order logic, and
 rules are of the form:
@@
 where $\langle R, F \rangle \models E$, the program $\langle J_R, J_F
-\rangle$ with $J_R \subseteq R$ and $J_F \subseteq F$ is a
+\rangle$ with ${J_R \subseteq R}$ and ${J_F \subseteq F}$ is a
 justification for $E$ iff $\langle J_R, J_F \rangle \models E$ and,
-for all $J'_R \subset J_R$ or $J'_F \subset J_F$, $\langle J'_R, J'_F
-\rangle \not\models E$.
+for all ${J^\prime_R \subset J_R}$ or ${J^\prime_F \subset J_F}$, ${\langle J^\prime_R, J^\prime_F
+\rangle \not\models E}$.
 
@@
 justification for $E$ iff $\langle R, J \rangle \models E$ and, for
-all $J' \subset J$, $\langle R, J' \rangle \not\models E$. Note that a
+all $J^\prime \subset J$, $\langle R, J^\prime \rangle \not\models E$. Note that a
 given set of entailed facts may not have a unique justification.
@@
 Herbrand base of all fact instances $\hat{H}^T$ (all facts with all
-instants). As before, a fact instance $e$ is entailed by a program
-$\Pi^T$ iff each interpretation satisfying every clause in $\Pi^T$
-also satisfies $e$, written $\Pi^T \models e$. The {\em entailment of the
-TA-Datalog program} $\Pi^T$ is therefore:
-\[ cons^T(\Pi^T) = \{ e \in \hat{H}^T \mid \Pi^T \models e \} \]
-
-Informally, the set of instants appearing in some set $E^T$ of fact
-instances entailed by some program ${\langle R, F^T \rangle}$ will be
-a subset of the instants appearing in the fact instances in $F^T$.  In
-addition, the instant $t$ in a fact instance $\langle e, t \rangle$
-that is entailed by a TA-Datalog program $\Pi^T = \langle R^T, F^T
-\rangle$ will be bounded from above by the maximum instant in any fact
-instance in $F^T$. A TA-Datalog program may entail fact instances that
-are only distinguished by their instants (and are therefore considered
-distinct from each other, as per \refdef{continuous datalog: fact
-  instance}).
+instants). As before, a fact instance $\langle e, t_e \rangle$ is 
+entailed by a program $\Pi^T$ iff each interpretation
+$\mathcal{I}^T \subseteq \hat{H}^T$ satisfying every
+clause (rule in $R$ and fact instance in $F^T$) in $\Pi^T$ also satisfies
+$\langle e, t_e \rangle$, written $\Pi^T \models \langle e, t_e \rangle$.
+As with facts in positive Datalog, we say that an interpretation
+$\mathcal{I}^T$ satisfies a fact instance $\langle f, t \rangle \in F^T$
+iff $\langle f, t \rangle \in \mathcal{I}^T$.  We say the interpretation
+$\mathcal{I}^T$ satisfies a rule
+$(r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n) \in R$ iff,
+given $\{\langle a_1, t_1 \rangle, \dots, \langle a_n, t_n \rangle\} \subseteq \mathcal{I}^T$
+such that $\{a_1, \dots, a_n\}$ satisfies the body of $r$,
+$\langle a_0, t_0 \rangle \in \mathcal{I}^T$ (where $t_0 = max(t_1,\dots,t_n)$,
+see \refdef{continuous datalog: instance justification}) such that
+$\{a_0, \dots, a_n\}$ satisfies $r$.  The {\em entailment of the TA-Datalog
+program} $\Pi^T$ is therefore:
+\[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
+\Pi^T \models \langle e, t_e \rangle \} \]
+
+\end{definition}
+
+\begin{definition}[Time-Annotated Justification]
+\labeldef{continuous datalog: instance justification}
+
+The justification of entailments is equally applicable to TA-Datalog
+programs, with facts in the entailment and the justification for that
+entailment being replaced with fact instances.  Following the
+simplified notion of justification in \refdef{continuous datalog:
+  justification}, we define the justification of a set of fact
+instances that are entailed by a TA-Datalog as a minimal subset of
+the ground fact instances of the program that are sufficient for
+the entailment to hold.
+
+More formally, for a TA-Datalog program $\Pi^T = \langle R, F^T \rangle$
+and a set of entailed fact instances $E^T$ where $\langle R, F^T \rangle \models E^T$, the set of fact instances $J^T \subseteq F^T$ is a justification for
+$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime}\rangle \not\models E^T$.
+
+For an entailed fact instance $\langle e, t_e\rangle$ and a
+justification of that fact instance $J^T$, the instant $t_e$ at which
+$\langle e, t_e \rangle$ becomes true is, intuitively, taken to be the
+latest instant at which any of the fact instances in the justification
+became true:
+\[ t_e = max(\{ t \mid \langle j, t \rangle \in J^T \}) \]
+
+Note that, although an entailed fact instance $\langle e, t_e\rangle$
+may have multiple justifications, all the justifications for $\langle
+e, t_e \rangle$ will share the same maximum instant; a justification
+with a different maximum instant $t^\prime \neq t_e$ would entail a different
+fact instance $\langle e, t^\prime \rangle$.
+
+\end{definition}
+
+From this definition of entailment in TA-Datalog, the set of instants
+appearing in some set $E^T$ of fact instances entailed by some program
+${\langle R, F^T \rangle}$ will be a subset of the instants appearing
+in the fact instances in $F^T$.
 
 Consider the following example TA-Datalog program:
-\[
+\begin{equation}
   \Pi^T = \langle \{ c \Leftarrow a \land b \}, \{ \langle a, 1 \rangle, \langle a, 2 \rangle, \langle b, 1 \rangle, \langle b, 2 \rangle  \} \rangle
-\]
+\end{equation}
 
 The entailment of the program $\Pi^T$ will be:
-\[cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}\]
+\begin{equation}
+  cons^T(\Pi^T) = \{\langle c, 1 \rangle, \langle c, 2 \rangle\}
+\end{equation}
+with the following justifications:
+\begin{equation}
+  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
+\end{equation}
 
@@
 (the set of fact instances $cons^T(\Pi^T)$) maintains the relationship
-of entailment; if $\langle R, F^T\rangle \models \langle e, t
+of entailment; if $\langle R, F^T\rangle \models \langle e, t_e
 \rangle$, then $\langle R, strip(F^T) \rangle \models strip(\langle e,
-t \rangle)$. This relationship is shown in \reffig{TD-D}.
+t_e \rangle)$. This relationship is shown in \reffig{TD-D}.
 
@@
 \end{align*}
-\end{definition}
-
-\begin{definition}[Time-Annotated Justification]
-\labeldef{continuous datalog: instance justification}
-
-The justification of entailments is equally applicable to TA-Datalog
-programs, with facts in the entailment and the justification for that
-entailment being replaced with fact instances.  Following the
-simplified notion of justification in \refdef{continuous datalog:
-  justification}, we define the justification of a set of fact
-instances that are entailed by a TA-Datalog as a minimal subset of
-the ground fact instances of the program that are sufficient for
-the entailment to hold.
-
-More formally, for a TA-Datalog program $\Pi^T = \langle R, F^T \rangle$
-and a set of entailed fact instances $E^T$ where $\langle R, F^T \rangle \models E^T$, the set of fact instances $J^T \subseteq F^T$ is a justification for
-$E^T$ iff $\langle R, J^T \rangle \models E^T$ and, for all $J^{T\prime} \subset J^T$, $\langle R, J^{T\prime}\rangle \not\models E^T$.
-
-For an entailed fact instance $\langle e, t\rangle$ and a
-justification of that fact instance $J^T$, the instant $t$ at which
-$\langle e, t \rangle$ becomes true is taken to be the latest instant
-at which any of the fact instances in the justification became true:
-\[ t = max(\{ t' \mid \langle j', t' \rangle \in J^T \}) \]
-
-Note that, although an entailed fact instance $\langle e, t\rangle$
-may have multiple justifications, all the justifications for $\langle
-e, t \rangle$ will share the same maximum instant; a justification
-with a different maximum instant $t' \neq t$ would entail a different
-fact instance $\langle e, t' \rangle$. Considering again the example
-TA-Datalog program in \refdef{continuous datalog: ta datalog program},
-the entailed fact instance $\langle c, 2 \rangle$ would have the
-following set of justifications:
-\[
-  \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
-\]
-
-\end{definition}
 

### 2015-11-10 article.tex
defined instant +/- number as instant
fixed notation for instances in a stream to be a succession of instances, not ambiguous variables
fixed minor wording and definition issues in Time and Fact Instance to define the behavior of 0-1 when talking about instants and keep wording consistent

@@
 \in \mathbb{T}$ to indicate the $n$th successor of instant $i$,
-and $i-n$ to indicate the $n$th predecessor of instant $i$.
+and $i-n$ to indicate the $n$th predecessor of instant $i$, where
+${+ : \mathbb{T} \times \mathbb{N} \rightarrow \mathbb{T}}$ and
+${- : \mathbb{T} \times \mathbb{N} \rightarrow \mathbb{T}}$.  Furthermore,
+in the case where $i < n$, ${i - n = 0}$.
 
@@
 \mathbb{T}$, written $\langle f, t \rangle$. Two fact instances
-containing the same fact and timestamp are considered identical; two
-fact instances containing the same fact but different timestamps are
+containing the same fact and instant are considered identical; two
+fact instances containing the same fact but different instants are
 considered to be distinct.
@@
 by writing $\langle f, t \rangle \in S$. We indicate that a fact
-instance $f_i$ appears strictly before a fact instance $f_j$ on stream
-$S$ (i.e. $S(f_i) < S(f_j)$) by writing $f_i \prec f_j$.
+instance $\langle f_i, t_i \rangle$ appears strictly before a fact
+instance $\langle f_j, t_j \rangle$ on stream $S$ (i.e.
+$S\left(\langle f_i, t_i \rangle\right) < S\left(\langle f_j, t_j \rangle\right)$)
+by writing $\langle f_i, t_i \rangle \prec \langle f_j, t_j \rangle$.
 

### 2015-11-10 article.tex
clarified definition of satisfaction in TA Datalog and made wording consistent regarding the open-ended semantics of truth in fact-instances.

@@
 entailed by a program $\Pi^T$ iff each interpretation
-$\mathcal{I}^T \subseteq \hat{H}^T$ satisfying every
-clause (rule in $R$ and fact instance in $F^T$) in $\Pi^T$ also satisfies
-$\langle e, t_e \rangle$, written $\Pi^T \models \langle e, t_e \rangle$.
-As with facts in positive Datalog, we say that an interpretation
-$\mathcal{I}^T$ satisfies a fact instance $\langle f, t \rangle \in F^T$
-iff $\langle f, t \rangle \in \mathcal{I}^T$.  We say the interpretation
-$\mathcal{I}^T$ satisfies a rule
-$(r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n) \in R$ iff,
-given $\{\langle a_1, t_1 \rangle, \dots, \langle a_n, t_n \rangle\} \subseteq \mathcal{I}^T$
-such that $\{a_1, \dots, a_n\}$ satisfies the body of $r$,
-$\langle a_0, t_0 \rangle \in \mathcal{I}^T$ (where $t_0 = max(t_1,\dots,t_n)$,
-see \refdef{continuous datalog: instance justification}) such that
-$\{a_0, \dots, a_n\}$ satisfies $r$.  The {\em entailment of the TA-Datalog
-program} $\Pi^T$ is therefore:
+$\mathcal{I}^T \subseteq \hat{H}^T$ satisfying every clause in $\Pi^T$
+also satisfies $\langle e, t_e \rangle$, written
+$\Pi^T \models \langle e, t_e \rangle$.  However, we must extend the
+definition of satisfaction from that for positive Datalog to incorporate
+the notion of time.  The extension must support the satisfaction of the
+rules $R$ of a TA-Datalog program by sets of fact instances, despite the
+former containing no notion of fact annotation, and must additionally
+uphold the semantics of temporal annotation described in
+\refdef{continuous datalog: fact instance}.
+
+To this end, we say that an interpretation $\mathcal{I}^T$ satisfies a
+fact instance $\langle f, t \rangle$ iff
+$\langle f, t \rangle \in \mathcal{I}^T$, as in positive Datalog.
+We also say that the interpretation $\mathcal{I}^T$ satisfies a rule
+$r = \alpha_0 \Leftarrow \alpha_1 \land \ldots \land \alpha_n$ iff, for
+every \emph{extended variable substitution} $\bar{s}$ by which
+$\{\langle \bar{s}\left( \alpha_1 \right), t_1 \rangle, \dots, \langle \bar{s}\left( \alpha_n \right), t_n \rangle\} \subseteq \mathcal{I}^T$, there exists a fact instance
+$\langle \bar{s}\left( \alpha_0 \right), max\left(t_1, \dots, t_n \right) \rangle \in \mathcal{I}^T$.
+The meaning of the restriction on the instant associated with the instance of
+$\bar{s}\left( \alpha_0 \right)$ in $\mathcal{I}^T$ is better described in terms
+of an entailment and its time-annotated justification, as per
+\refdef{continuous datalog: instance justification}.
+
+The {\em entailment of the TA-Datalog program} $\Pi^T$ is therefore:
 \[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
@@
 For an entailed fact instance $\langle e, t_e\rangle$ and a
-justification of that fact instance $J^T$, the instant $t_e$ at which
-$\langle e, t_e \rangle$ becomes true is, intuitively, taken to be the
-latest instant at which any of the fact instances in the justification
+justification of that fact instance $J^T$, the instant $t_e$ from which
+$\langle e, t_e \rangle$ is recognised to be true is, intuitively, taken to be the
+latest instant from which any of the fact instances in the justification
 became true:

### 2015-11-10 article.tex
gathering of all justifications in example of TA Datalog
expanded discussion of streams and windows in reference to those of the RSP community, and provided definitions of sliding windows and window states

@@
 \end{equation}
-with the following justifications:
-\begin{equation}
+with the following justifications for $\langle c, 1 \rangle$ and $\langle c, 2 \rangle$ respectively:
+\begin{gather}
+  \{ \ \{ \langle a, 1 \rangle, \langle b, 1 \rangle \} \ \} \\
   \{ \ \{ \langle a, 1 \rangle, \langle b, 2 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 1 \rangle \},  \{ \langle a, 2 \rangle, \langle b, 2 \rangle \} \  \}
-\end{equation}
+\end{gather}
 
@@
 Having defined the notion of time annotation in Datalog, we now define
-the notions of streams and windows before extending to Continuous Datalog.
+a notion of streams that is closely related to that currently being
+refined by the RSP community.  It differs in two key ways.  Firstly,
+elements in streams are not timestamped \emph{RDF graphs} but instances of
+individual \emph{RDF triples} (facts). Secondly, fact instances are
+annotated for the purposes of distinguishing them from instances of the
+same fact asserted/entailed at different times, rather than for the
+purpose of ordering them as is the case with timestamped graphs.  This
+ordering is instead expressed in the streams themselves, as detailed
+formally in \refdef{continuous datalog: stream}, where a fact instance
+may occur on many streams and be mapped by each stream to a different
+instant.
+Each stream in WC-Datalog may be assigned a distinct semantics, in a
+similar manner to the distinct predication of timestamps proposed by
+the RSP community; they are, therefore, comparably expressive in that
+regard, but also represent the manner in which data occurs \emph{in a
+stream at a given time} in real-world scenarios.
 
@@
 are not ordered by their instants as {\em out-of-order streams}.
+
 \end{definition}
 
-\begin{definition}[Window]
+Based on this definition of streams, we go on to define the notion of
+sliding windows over streams, as well as states of those windows at
+specific instants.  We do not consider count-based window functions, as
+these serve a semantic purpose at high specificity, e.g. \emph{retain
+the last ten buses to stop at a specific stop}.  In contrast, as will be
+formally described in \refdef{continuous datalog: CDP}, we consider only
+the application of the \emph{entirety of a rule set} within the
+consecutive states of one (or more) window(s); in this context, an
+example application of a count-based window might be to \emph{determine
+the implications of the last hundred facts reported by a sensor network,
+given some background knowledge regarding that network}.
+
+\begin{definition}[Sliding Window]
+\labeldef{continuous datalog: sliding window}
+
+A \emph{sliding-window} definition consists of a stream $S$ that the window is
+said to ``slide over'', and a range ${r \in \mathbb{N}}$ that is the
+number of consecutive instants that the window covers.  This is written
+$W^{S,r}$.
+\end{definition}
+
+\begin{definition}[Window State]
 \labeldef{continuous datalog: window}
 
-A window of range ${r \in \mathbb{N}}$ over a stream $S$ at an instant
-${i \in \mathbb{T}}$ is the \emph{set} of fact instances that appear in
-$S$ in the interval ${(i-r,i]}$, written ${W^{S,r}_{i}}$, and defined as follows:
+The \emph{state} of a sliding-window $W^{S,r}$ at an instant
+${i \in \mathbb{T}}$ is the \emph{set} of fact instances that are
+associated in $S$ with any instant in the interval ${(i-r,i]}$, and is
+denoted ${W^{S,r}_{i}}$.
+The state of a sliding-window is defined as follows:
 \[

### 2015-11-10 article.tex
expanded discussion of axioms regarding boundary cases of windows of range 0 and at instant 0

@@
 {\nobreak\begin{axiom}\label{axiom:continuous datalog: window range leq 0}
-A window of range $0$ over a stream $S$ at an instant 
+The state of a window of range $0$ over a stream $S$ at an instant 
 ${i \in \mathbb{T}}$ is always empty.
@@
 
+Axiom \ref{axiom:continuous datalog: window range leq 0} is trivially
+true by reference to the interval in \refdef{continuous datalog: window},
+and the definition of numeric intervals.  The state of a window of range
+$0$ over stream $S$ at interval $i$ is composed of fact instances that
+are each mapped by $S$ to any instant in the interval ${(i,i]}$.  This
+interval is defined as the set of numbers ${\{ x \mid i < x \leqslant i \}}$,
+which is the empty set for any $i$.
+
 {\nobreak\begin{axiom}\label{axiom:continuous datalog: window start leq 0}
-A window of any range $x \in \mathbb{N}$ over a stream $S$ at (or before)
-the first instant of time is always empty.
+A window of any range $x \in \mathbb{N}$ over a stream $S$ at the
+first instant of time is always empty.
 \begin{equation*}
-W^{S,x}_{i} = \varnothing
+W^{S,x}_{0} = \varnothing
 \end{equation*}
-where $i \leqslant 0$.
 \end{axiom}}
+
+Axiom \ref{axiom:continuous datalog: window start leq 0} follows from
+Axiom \ref{axiom:continuous datalog: window range leq 0} and the definition
+of the $-$ operator with regards to instants in
+\refdef{continuous datalog: instant}.  The state of a window of range
+$x$ over stream $S$ at the first interval is composed of fact instances that
+are each mapped by $S$ to any instant in the interval ${(0-x,0]}$. As the instant
+${0 - x = 0}$, the interval of instants in $W^{S,x}_{0}$ evaluates to ${(0,0]}$,
+which makes it empty by Axiom \ref{axiom:continuous datalog: window range leq 0}.
 
@@
 {\nobreak\begin{axiom}\label{axiom:continuous datalog: window disjointness}
-A window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
-instant ${i \in \mathbb{T}}$ is disjoint from a second window of 
+A the state of a window of range ${x \in \mathbb{N}}$ over a stream $S$ at an
+instant ${i \in \mathbb{T}}$ is disjoint from a second state of any window of 
 range ${y \in \mathbb{N}}$ over $S$ at any instant before $i - x$.

### 2015-11-10 article.tex
significant changes to the discussion of (Windowed) Continuous Datalog to cover multiple streams, multiple windows, and pipelining programs.

@@
 
-\begin{definition}[Continuous Datalog Program]
+\begin{definition}[(Windowed) Continuous Datalog Program]
 \labeldef{continuous datalog: CDP}
@@
 C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
-which comprises a finite set of rules $R$, a finite set of static fact
-instances $F^0$ (all of which have the instant $0$) and an in-order
-stream of fact instances $S$.
+which comprises a finite set of rules $R$, a finite set of persistent/static
+fact instances $F^0$ (all of which have the instant $0$, meaning that they
+cannot appear in any state of any window, by
+\refdef{continuous datalog: window} and Axioms~\ref{axiom:continuous datalog: window range leq 0} and
+\ref{axiom:continuous datalog: window start leq 0}) and an in-order stream
+of fact instances $S$.  It should be noted that, in practice, data may be
+drawn from more than one source, and so $S$ may be a union of
+the mappings represented by $n$ in-order streams $S_0,\dots,S_n$ without altering
+the semantics presented here.
 
@@
 instances $\hat{H}$. Disregarding the order of fact instances in $S$,
-a fact instance $e$ is entailed by a program $\Pi^C$ iff each
-interpretation satisfying every clause in $\Pi^C$ also satisfies $e$.
-However, we take the {\em entailment of the C-Datalog program} $\Pi^C$
-to be an in-order stream $E^S$, where the order in which entailed fact
-instances appear in $E^S$ depends on the order in which the fact
-instances in their justifications appear in $S$. Given that that the
-instant of an entailed fact instance is the maximum of the instants in
-the fact instances that form the justification for that entailed fact
-instance (as in \refdef{continuous datalog: instance justification}),
-$E^S$ will also be an in-order stream.
+a fact instance $\langle e, t \rangle$ is entailed by a program $\Pi^C$
+iff each interpretation satisfying every clause in $\Pi^C$ also
+satisfies $\langle e, t \rangle$. However, we take the {\em entailment
+of the C-Datalog program} $\Pi^C$ to be an in-order stream $E^S$ rather
+than a set of fact instances. Given that the instant associated with an
+entailed fact instance is the maximum of those associated with
+any part of its justification (as in
+\refdef{continuous datalog: instance justification}), and $E^S$ is an
+in-order stream, $E^S$ expresses the times at which each entailed fact
+instance was first justified by $\Pi^C$.
 
 However, this definition of a C-Datalog program does not take into
-account any notion of windowing over the stream $S$. We define a {\em
-  Windowed Continuous Datalog program} $\Pi^{C, r} = \langle R, F^0, S,
-r \rangle$ to be a tuple with components as defined above for a
-C-Datalog program, but with the addition of $r \in \mathbb{N}$ for the
-range of the window over $S$, which effectively specifies that the
-fact instances from $S$ are to contribute to the entailments of $\Pi^{C,r}$
-from their instant for a duration of $r$.
+account any notion of windowing over the stream $S$, and so any fact of
+which an instance occurs in $S$ will be considered true from the instant
+of its occurrence in $S$.  While this is technically acceptable in the
+case of sets of rules without constraints, if a constraint exists and its
+body becomes satisfied by some $\Pi^C$ at some instant $i$ then the
+program will be unsatisfiable from that instant on, making $E^S$ empty
+from then on.  As such, we define a {\em Windowed Continuous Datalog program}
+$\Pi^{C, r} = \langle R, F^0, \{\langle S_0, r_0 \rangle,\dots,\langle S_n, r_n \rangle\} \rangle$
+(where $r = max(r_0,\dots,r_n)$) to be a tuple with components $R$ and $F^0$ as defined above for a
+C-Datalog program, but with each part of $S = S_0 \cup \dots \cup S_n$
+associated with the range $r_x \in \mathbb{N}$ of a sliding window over
+$S_x$, for $x \in [0,n]$. This specifies that the fact instances
+from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
+they occur in $S_x$ for a duration of $r_x$ consecutive instants.
 
 We consider the entailment of the WC-Datalog program $\Pi^{C,r}$ not
-as a single stream of fact instances (as for C-Datalog), but in three
-parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
-that is the entailment of the TA-Datalog program $\langle R,
-F^0\rangle$; a stream $P$ of fact instances entailed by the rules $R$ and
-the fact instances in $F^0$ and $S$, ordered by the instant at which they
-became justified (i.e. their instants); and a stream $N$ of those same fact
-instances, ordered by the instant at which they ceased to be justified:
-$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
-and $N$ can be considered to be the continuous reasoning equivalents of
-the streams produced by the \emph{IStream} and\emph{DStream} operators
-respectively of the \emph{CQL} continuous query language \citep{CQL},
-which create streams based on the additions to and deletions from a
-relation. Note that $P$ is an in-order stream whose members are a subset
-of those in $E^S$ for an equivalent C-Datalog program, while $N$ may be
-an out-of-order stream.
+as a single stream of fact instances (as for C-Datalog), but a triple: a
+set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$ that is the
+entailment of the TA-Datalog program $\langle R, F^0\rangle$; a stream
+$P$ of fact instances entailed by the rules $R$ and the fact instances in
+$F^0$ and $S$, ordered by the instant at which they became justified (i.e.
+their instants); and a stream $N$ of those same fact instances, ordered by
+the instant at which they \emph{ceased} to be justified (i.e. some part of
+their justification left a sliding window $W^{S_x,r_x}$). We express this
+entailment $\Pi^{C,r} \models \langle E^0, P, N \rangle$.
+It should be noted that $P$ and $N$ can be considered to be the continuous
+reasoning equivalents of the streams produced by the \emph{IStream} and
+\emph{DStream} operators, respectively, of the \emph{CQL} continuous query
+language \citep{CQL}, which create streams based on the additions to and
+deletions from a relation.
+
+Note that $P$ is an in-order stream whose members are a subset
+of those in $E^S$ for an equivalent C-Datalog program, up until the instant $i$
+at which a contradiction occurs in $S$, while $N$ may be an out-of-order
+stream. Furthermore, the members of $P$ from the instant $i+1$ are a subset of
+those in $E^{S\prime}$, where $\langle R,F^0,S^\prime \rangle \models E^{S\prime}$
+and $S^\prime$ is the in-order stream composed of the mappings in $S$ to instants
+greater than $i$. In this way, a WC-Datalog program is able to \emph{recover}
+from contradictions in $S$.
+
 \end{definition}
@@
 {\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-If two sliding windows over the same in-order stream $S$ cover
+If two states of windows over the same in-order stream $S$ cover
 non-overlapping time periods, then the set of fact instances entailed
@@
 
-We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
-\rangle$ at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
-taking the window $W^{S,r}_i$ over stream $S$ of range $r \in \mathbb{N}$;
-we refer to this as the {\em instantaneous state} of the WC-Datalog program,
-written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle$. As
-previously shown in \refdef{continuous datalog: ta datalog program}, this
-may then be further reduced to the Datalog program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.
+We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
+\{ \langle S_0, r_0 \rangle,\dots,\langle S_n, r_n \rangle \} \rangle$
+at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
+applying the rules $R$ to the union of the persistent facts $F^0$ and all
+finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
+$r_x \in \mathbb{N}$, for all $x \in [0,n]$; we refer to this as the {\em instantaneous state} of
+the WC-Datalog program, written
+$\Pi^{C,r}_i = \langle R, F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i \rangle$.
+As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
+reduced to the Datalog program
+$\Pi = \langle R, strip(F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i) \rangle$.
 
@@
 entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
-program, which is guaranteed to be part of the entailment of every state
-due to the monotonicity of positive Datalog; the \emph{transitive} portion
+program, which is guaranteed to be part of the entailment of every consistent
+state due to the monotonicity of positive Datalog; the \emph{transient} portion
 is the set $E^T_{i}$ of entailed fact instances whose justifications are at
-least partially contained in $W^{S,r}_i$. {\nobreak We define $E^T_i$ as:
+least partially contained in $W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i$.
+{\nobreak We define $E^T_i$ as:
 \[
-E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
+E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cap \left( W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i \right) \neq \varnothing \}
 \]
-where $J^T$ is a justification for the entailment of x by
+where $J^T$ is a justification for the entailment of $x$ by
 $\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous
@@
 WC-Datalog program $\Pi^{C,r}$.
+
+An \emph{inconsistent state} of a C-Datalog program is any instantaneous
+state $\Pi^{C,r}_i$ that is unsatisfiable, i.e. that contains at least
+one contradiction.  Such a state has no entailment.
+
 \end{definition}
@@
 
+% It should be noted that a the existence of a contradiction in some
+% C-Datalog program $\langle R, F^0, S \rangle$ does not imply that any
+% \emph{state} of the WC-Datalog $\Pi^{C,r} = \langle R, F^0, S, r \rangle$ will be
+% inconsistent. Where a state $\Pi^{C,r}_i$ of $\Pi^{C,r}$ \emph{is} inconsistent,
+% the transient portion $E^T_i$ of its entailment is empty, as an unsatisfiable
+% TA-Datalog program entails nothing.  In this case, therefore, \emph{all}
+% fact instances transiently entailed by $\Pi^{C,r}_{i-1}$ occur in $N$ at the $i$,
+% and no fact instances will occur in $P$ from $i$ until the first instant $j > i$ at
+% which $\Pi^{C,r}_j$ is satisfiable.  In reference to the stream $E^S$ entailed by
+% the C-Datalog program $\langle R, F^0, S \rangle$, therefore, $P$ is an in-order
+% stream whose members \emph{that are mapped to instants between $i$ } are a subset of those in $E^S$
+% to instants greater than $i$
+
 {\nobreak\begin{axiom}
@@
 leaving the sliding window $W^{S,r}_{i}$, this necessarily being after
-they were entailed.  As such, a window of any range ${x \in
+they were justified.  As such, a window of any range ${x \in
   \mathbb{N}}$ over the stream of entailments $P$ at any given instant
@@
 
-As noted above in \refdef{continuous datalog: CDP}, the stream of
-negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
-out-of-order stream in which fact instances appear on the stream in
-the order in which they cease to have any complete justifications.
+% As noted above in \refdef{continuous datalog: CDP}, the stream of
+% negations $N$ of a WC-Datalog program $\Pi^{C,r}$ may be an
+% out-of-order stream in which fact instances appear on the stream in
+% the order in which they cease to have any complete justifications.
 
@@
 of those streams.
-
 However, it follows from these axioms that the instantaneous set
-$E^T_{i}$ of instances entailed at any given instant $i$ by any
-$\Pi^{C,r} = \langle R, F^0, S, r \rangle$ may be reconstructed from
-windows of sufficient range over its entailed streams $P$ and $N$ at
+$E^T_{i}$ of instances entailed at any given instant $i$ by any single-window program
+$\Pi^{C,r} = \langle R, F^0, \{ \langle S, r \rangle \} \rangle$ may be reconstructed from
+windows of sufficient range {\nobreak over its entailed streams $P$ and $N$ at
 $i$:
 \[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
-where ${y \geqslant x \geqslant r}$. The minimum range $x$ is such to achieve
+where ${y \geqslant x \geqslant r}$}. The minimum range $x$ is such to achieve
 the \emph{complete} set of entailment instances $E^T_{i}$ of the
 $\Pi^{C,r}_{i}$. In addition, the minimum range $y$ is such to achieve the
-\emph{sound} set $E^T_{i}$. As such, a \emph{valid window} that
-contains only and exactly those entailed instances $E^T_{i}$ at every
-instant $i$ may be trivially maintained from the streams $P$ and $N$
-continuously entailed by some $\Pi^{C,r}$.
+\emph{sound} set $E^T_{i}$.  It should be noted that the proof holds in the
+case of $n$ windows of varying range, with $x \geqslant max(r_0,\dots,r_n)$.
+We capture this derivation in the form of a \emph{valid window}, as per
+\refdef{continuous datalog: valid window}.
+
+\begin{definition}[Valid Window] \labeldef{continuous datalog: valid window}
+
+A \emph{valid window} contains only and exactly those instances $E^T_{i}$ entailed by
+the state $\Pi^{C,r}_i$ at every instant $i \in \mathbb{T}$.  Such a window may be
+trivially maintained from the streams $P$ and $N$ continuously entailed
+by some $\Pi^{C,r}$: adding all fact instances occurring in $P$ at $i$ to the state
+of the window at $i-1$, then removing all fact instances occurring on $N$ at $i$ from
+the resultant set of fact instances, produces the state of the window at $i$.  Such
+a window obviously has no definite range but is taken over two streams (e.g. $P$ and
+$N$), and so is denoted $W^{P \setminus N}$.
+
+\end{definition}
+
+The decomposition of streams into more sparse streams without loss of
+information allows us to define the state of a valid window
+$W^{P \setminus N}$ as a union of the states of sliding windows over
+each of the singleton streams $S_{\langle f, t \rangle} =
+(\langle f, t \rangle \mapsto t^\prime)$, where
+$S(\langle f, t \rangle) = t^\prime$.  The range of the sliding
+window over each $S_{\langle f, t \rangle}$ is then the difference between
+the instant at which $\langle f, t \rangle$ occurs in $P$ and that
+at which it occurs in $N$:
+\[
+  W^{P \setminus N}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in W^{S_{\langle f, t \rangle},N(\langle f, t \rangle) - P(\langle f, t \rangle)}_i\}
+\]
+Treated this way, the entailment of one WC-Datalog program, e.g. that in
+\refeqn{continuous datalog: example WC-Datalog program},
+may then be taken as a part or all of the data in a successive WC-Datalog
+program, without modification of the existing semantics or loss of information
+regarding the validity of the entailments.  However, in the interest
+of being concise, such a successive WC-Datalog program may express valid
+windows alongside any sliding windows. For example, 
+a program composed of the set of rules $R^\prime \neq R$, the set of base facts
+$F^{0\prime} \neq F^0$, a sliding window of range $x \geqslant r$ over the
+in-order stream $S^\prime \neq S$, and the entailment of
+$\Pi^{C,r} \models \langle E^0, P, N \rangle$, would be expressed as in \refeqn{continuous datalog: successive WC-Datalog program}.
+\begin{gather}
+  \labeleqn{continuous datalog: example WC-Datalog program}
+    \Pi^{C,r} \langle R, F^0, \{W^{S,r}\} \rangle \models \langle E^0, P, N \rangle \\
+  \labeleqn{continuous datalog: successive WC-Datalog program}
+    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{P \setminus N}, W^{S^\prime,x} \} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
+\end{gather}
 
 In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
-F^0, S \rangle$, we can apply to it a window of range $r$ to yield a
-Windowed Continuous Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
-\rangle$. We can evaluate this window at an instant $i \in \mathbb{T}$
-to yield a Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup
-W^{S,r}_i \rangle$, from we which can then strip the time annotations
-to yield a conventional Datalog program $\Pi = \langle R, F
-\rangle$. The relationship between these *-Datalog programs and their
+F^0, S_0 \cup \dots \cup S_n \rangle$, we can apply to it a set of
+windows of range $r_0,\dots,r_n$ to yield a Windowed Continuous Datalog
+program $\Pi^{C,r} = \langle R, F^0, \{ W^{S_0, r_0}, \dots,
+W^{S_n, r_n} \} \rangle$. We can evaluate these windows at
+an instant $i \in \mathbb{T}$ to yield a Time-Annotated Datalog program
+$\Pi^T = \langle R, F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i
+\rangle$, from we which can then strip the time annotations to yield a
+conventional Datalog program $\Pi = \langle R, F \rangle$. The
+relationship between these \ast-Datalog programs and their
 entailments is expressed in the commutative diagram in

### 2015-11-17 article.tex
INCOMPLETE Updated introduction to fight our corner a little more, added in some pertinent references

@@
   With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
-  This may be characterised more precisely as continuous deductive reasoning over streamed RDF data.
+  In the most recent literature, this has been characterised as a layer of higher complexity processing on top of Complex Event Processing (CEP) \citep{mileo15webSR}, deeming it suitable only for application to low-throughput macro-event streams.
+  However, Semantic Web reasoning languages provide many varying degrees of expressivity corresponding to to varying levels of complexity.
+  Furthermore, in many cases static ontologies are applicable to streamed data (such as the description and classification of readings from sensor networks, for example), and therefore do not require the features unique to CEP (and temporal reasoning \citep{OrgunWadge92,Tuzhilin93}), rather requiring a concept of data becoming less relevant over time.
+  A prime expression of low-complexity reasoning that is capable of such tasks in a static context is the rule language profile of OWL 2 (OWL 2 RL) \citep{w3cowl2profiles}.
+  The continuous application of reasoning of this nature to streamed RDF data may obviously be characterised as `stream reasoning', but occupies the same general complexity bracket as stream querying in the hierarchy presented by \citep{mileo15webSR}, making it applicable to high-throughput streams.
+  It is this low-complexity continuous reasoning that is the focus of this paper.
 
@@
   in Description Logics, it may be considered that the TBox of a knowledge base describes general knowledge and the ABox describes particular knowledge;
-  however, the fundamental rules by which entailments are derived from knowledge bases by Description Logics constitute a more general knowledge than that of the TBox of some knowledge base, by which the entirety of a knowledge base may be considered particular knowledge.
+  however, the fundamental rules by which entailments are derived from knowledge bases by Description Logics constitute a more general knowledge than that of the TBox of some specific knowledge base, by which the entirety of a knowledge base may be considered particular knowledge.
   In the latter case, a distinction may still be made between a TBox and ABox in the context of streaming, with the former being expressed in static data and the latter being streamed or vice versa, though this is dependent on the rate at which the TBox and ABox are expected to change.
@@
   As such, a consumer of the streams produced using SemsorGrid4Env could download the appropriate ontologies once, then use them to reason over a reduced set of streamed data containing only the base readings from the sensor network, thus obtaining the classification and macro-event information desired without having to stream it explicitly.
+  Furthermore, sensor readings are often superseded by later readings of the same type from the same sensor, and so older readings may be discarded as time progresses.
 
@@
   %   RDF stream processing and reasoning
-  The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the iterative application of queries to the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}.
+  The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the \emph{iterative/recursive execution} of queries over the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}, in which each query expresses a rule in a logic program.
+  This reflects precisely the approach to reasoning using Datalog in deductive databases, which has been characterised as the ``QUOTE'' \citep{DATALOG CHARACTERISATION}.
   Meanwhile, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
@@
   %   issues with existing art: no explicit semantics for reasoning
-  While the iterative application of queries to the results of others is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, the query-by-query windowing semantics of the base stream-querying systems do not translate to a coherent semantics for multi-rule stream reasoning.
-  This is because the windowing semantics proposed for stream querying involve applying the entirety of a ``problem'' (be it a single query or a unified body of general knowledge) to a bounded subset of the data in a stream (a window), henceforth referred to as \emph{global windowing};
-  the iterative application of individually windowed queries both introduces and \emph{fails to properly consider} the semantics of Complex Event Processing, leading to inconsistencies between the results of identical rule sets expressed in the cited systems \citep{LARS}.
-
-  Truth-maintenance algorithms, on the other hand, do support a coherent semantics for continuous reasoning.
-  However, this semantics casts the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams.
+  While the iterative execution of queries as rules is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, as practised in deductive databases, the query-by-query windowing semantics of most stream-querying systems do not translate to a coherent semantics for multi-rule stream reasoning of the kind investigated in this paper.
+  This is because the windowing semantics proposed for stream querying involve applying the entirety of a ``problem'' (be it a single query or a unified body of general knowledge, as we propose) to a bounded subset of the data in a stream (a window).
+  This semantics is henceforth referred to as \emph{global windowing}.
+  The iterative execution of individually windowed queries, however, both introduces and \emph{fails to properly consider} the semantics of Complex Event Processing, leading to inconsistencies between the results of identical rule sets expressed in the cited systems \citep{LARS}.
+
+  Truth-maintenance algorithms for continuously changing knowledge bases and logic programs, on the other hand, do support coherent semantics for continuous reasoning.
+  However, these semantics cast the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams.
   Without research to extend these semantics to entail concise streams that maintain in their own right the coherence of the existing semantics, these algorithms preclude the further processing of entailments in a stream-processing manner.
 
-  As such, until recently, stream reasoning either lacked a coherent semantics for continuous truth and entailment, or was necessarily a terminus rather than intermediate process in any streamed data pipeline.
-  However, the work by \citet{LARS} has produced the LARS framework, by which the semantics of stream processing systems may be expressed and formally analysed.
-  LARS extends the semantics of Answer Set Programming (itself an extension of Datalog semantics with features such as Negation as Failure) with the notion of streamed data with a point-in-time semantics.
-  Paired with this is the notion of a logical windowing operator that supports the expression of windows as part of the general knowledge of a reasoning problem, as is necessary to express reasoning as performed by the iterative application of continuous queries in the manner already discussed.
-  Though this form of windowing also allows the expression of temporally-aware rules in the manner of Complex Event Processing, it does \emph{not} align with or support the original intent of \emph{global windowing} in stream querying.
+  As such, until recently, stream reasoning either lacked a coherent semantics for continuous truth and entailment, or was necessarily a terminus rather than intermediate process in any streamed data pipeline \citep{valle09streamReasoningSummary}.
+  However, the work by \citet{LARS} has produced the LARS framework, by which the reasoning semantics of stream processing systems may be expressed and formally analysed.
+  LARS extends the semantics of Answer Set Programming (itself an extension of Datalog semantics with features such as Negation as Failure and aggregation) with the notion of streamed data with a point-in-time semantics.
+  Paired with this is the notion of a logical windowing operator that supports the expression of windows as part of the general knowledge of a reasoning problem, as is necessary to express reasoning as performed by the iterative execution of continuous queries in the manner already discussed.
+  Though this form of windowing also allows the expression of temporally-aware rules in the manner of Complex Event Processing, it does \emph{not} align with or support the expression of \emph{global windowing} derived from stream \emph{querying}.
   Furthermore, the expression of granular windowing such as that expressible in LARS has not been defined in existing Semantic Web reasoning languages such as RIF \citep{w3crif} and OWL \citep{w3cowl2}.
   Finally, the complexity of entailment over programs as defined in LARS makes it unsuitable as a semantics on which to \emph{base} the expression of reasoning in real-time:
-  the complexity of entailment in LARS is PSPACE, or co-NP given restrictions on window-nesting, where real-time processing of a task requires that task to be at most polynomial in its time complexity.
+  the complexity of entailment in LARS is PSPACE, or co-NP given restrictions on window-nesting, which is generally considered too high a complexity for near-real-time entailment over high-throughput streams \citep{mileo15webSR}.
 
@@
   In addition, more recent data-flow-based models either don't consider ``reasoning'' more expressive than RDFS \citep{sparkwave}, or are based on rule/query-level windowing \citep{C-SPARQL,walavalkar08streamingkb}, with problems as already discussed.
+  Despite this, data-flow-based models and the vast body of associated literature present a firm foundation for the processing of reasoning tasks in a flexible and scalable manner.
 
@@
   In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Windowed Continuous Datalog.
-  The scope of this semantics is to be an expression of inference, as currently described by Semantic Web reasoning languages, that incorporates streamed semantic data and is of low enough computational time-complexity to be feasibly computed continuously in real-time.
+  The scope of this semantics is to be an expression of inference that extends the static form currently described by Semantic Web reasoning languages with global reasoning semantics, thereby incorporating potentially high-throughput streams of semantic data but resulting in a low enough computational time-complexity that entailments be feasibly computed continuously in real-time.
   As entailment in positive Datalog has polynomial-time complexity and is the equivalent of entailment in RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, the continuous semantics of Windowed Continuous Datalog may be used to apply OWL 2 ontologies to streams of RDF data.
-  Next, we present the abstract processing model CoDeR that provides for the expression of processing plans that satisfy the semantics of Windowed Continuous Datalog.
+  Next, we present the abstract processing model CoDeR that provides for the expression of data-flow processing plans that satisfy the semantics of Windowed Continuous Datalog.
   We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.

### 2015-11-17 article.tex
fixed left/mid/right sizing of braces and "givens"

@@
 \begin{align*}
-  cons(\Pi) &= \{ e \in \hat{H} \mid \Pi \models e \} \\
+  cons(\Pi) &= \left\{ e \in \hat{H} \,\middle\vert\, \Pi \models e \right\} \\
   &= \bigcap_\mathcal{I} \models_\mathcal{I} \Pi
@@
 For convenience, we also write $strip(F^T) = F$, where $F^T$ is a set
-of fact instances and $F = \{ strip(f) \mid f \in F^T \}$.
+of fact instances and $F = \left\{ strip(f) \,\middle\vert\, f \in F^T \right\}$.
 
@@
 The {\em entailment of the TA-Datalog program} $\Pi^T$ is therefore:
-\[ cons^T(\Pi^T) = \{ \langle e, t_e \rangle \in \hat{H}^T \mid
-\Pi^T \models \langle e, t_e \rangle \} \]
+\[ cons^T(\Pi^T) = \left\{ \langle e, t_e \rangle \in \hat{H}^T \,\middle\vert\,
+\Pi^T \models \langle e, t_e \rangle \right\} \]
 
@@
 became true:
-\[ t_e = max(\{ t \mid \langle j, t \rangle \in J^T \}) \]
+\[ t_e = max\left(\left\{ t \,\middle\vert\, \langle j, t \rangle \in J^T \right\}\right) \]
 
@@
 \[
-W^{S,r}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \}
+W^{S,r}_i = \left\{ \langle f, t \rangle \,\middle\vert\, \langle f, t \rangle \in S \land (i - r) < S(\langle f, t\rangle) \leqslant i \right\}
 \]  
@@
 are each mapped by $S$ to any instant in the interval ${(i,i]}$.  This
-interval is defined as the set of numbers ${\{ x \mid i < x \leqslant i \}}$,
+interval is defined as the set of numbers ${\left\{ x \,\middle\vert\, i < x \leqslant i \right\}}$,
 which is the empty set for any $i$.
@@
 \[
-\{ e \mid \langle R , W^{S,x}_i \rangle \models e \} \cap \{ e \mid \langle R , W^{S,y}_{i-z} \rangle \models e \} = \varnothing
+\left\{ e \,\middle\vert\, \langle R , W^{S,x}_i \rangle \models e \right\} \cap \left\{ e \,\middle\vert\, \langle R , W^{S,y}_{i-z} \rangle \models e \right\} = \varnothing
 \]
@@
 is the set $E^T_{i}$ of entailed fact instances whose justifications are at
-least partially contained in $W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i$.
+least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i$.
 {\nobreak We define $E^T_i$ as:
 \[
-E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cap \left( W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i \right) \neq \varnothing \}
+E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_j,r_j}_i \neq \varnothing \right) \right\}
 \]
-where $J^T$ is a justification for the entailment of $x$ by
+where $J^T$ is a justification for the entailment of $e$ by
 $\Pi^{C,r}_i$.} By this definition and Axiom~\ref{axiom:continuous

### 2015-11-17 article.tex
fixed definitions to include stream union and cascaded that fix, and to be based on (sliding) windows rather than pairings of streams and ranges (allows incorporation of valid windows into definitions at the end)

@@
 the instant records the time at which the fact instance appears on the
-stream. This has the consequence that a particular fact instance
-(i.e. a fact accompanied by a particular instant that is its time
-annotation) may appear only once in a given stream. For convenience,
+stream. This makes the stream a functional mapping from fact instances
+to instants, with the consequence that a particular fact instance
+% (i.e. a fact accompanied by a particular instant that is its time annotation)
+may appear only once in a given stream. For convenience,
 we indicate that a fact instance $\langle f, t\rangle$ appears in $S$
@@
 are not ordered by their instants as {\em out-of-order streams}.
+
+A \emph{union} of the mappings of two streams, written $S \cup S^\prime$,
+may itself be a stream, but only in the case that each fact instance occurs
+in only one of the two streams, or is mapped to the same instant in both,
+thereby preserving the functionality of the mapping.  It should be noted that
+the union of two in-order streams is always a stream, as each fact instance
+is mapped \emph{in each in-order stream} to its inherent instant.
 
@@
 of fact instances $S$.  It should be noted that, in practice, data may be
-drawn from more than one source, and so $S$ may be a union of
-the mappings represented by $n$ in-order streams $S_0,\dots,S_n$ without altering
-the semantics presented here.
+drawn from more than one source, and so $S$ may be a union of $n$ in-order
+streams $S_1,\dots,S_n$ without altering the semantics presented here.
 
@@
 from then on.  As such, we define a {\em Windowed Continuous Datalog program}
-$\Pi^{C, r} = \langle R, F^0, \{\langle S_0, r_0 \rangle,\dots,\langle S_n, r_n \rangle\} \rangle$
-(where $r = max(r_0,\dots,r_n)$) to be a tuple with components $R$ and $F^0$ as defined above for a
-C-Datalog program, but with each part of $S = S_0 \cup \dots \cup S_n$
+$\Pi^{C, r} = \langle R, F^0, \{ W^{S_1, r_1},\dots, W^{S_n, r_n} \rangle\} \rangle$
+(where $r = max(r_1,\dots,r_n)$) to be a tuple with components $R$ and $F^0$ as defined above for a
+C-Datalog program, but with each part of $S = S_1 \cup \dots \cup S_n$
 associated with the range $r_x \in \mathbb{N}$ of a sliding window over
-$S_x$, for $x \in [0,n]$. This specifies that the fact instances
+$S_x$, for $x \in [1,n]$. This specifies that the fact instances
 from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
@@
 We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
-\{ \langle S_0, r_0 \rangle,\dots,\langle S_n, r_n \rangle \} \rangle$
+\{ W^{S_1, r_1},\dots, W^{S_n, r_n} \} \rangle$
 at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
@@
 finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
-$r_x \in \mathbb{N}$, for all $x \in [0,n]$; we refer to this as the {\em instantaneous state} of
+$r_x \in \mathbb{N}$, for all $x \in [1,n]$; we refer to this as the {\em instantaneous state} of
 the WC-Datalog program, written
-$\Pi^{C,r}_i = \langle R, F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i \rangle$.
+$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.
 As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
 reduced to the Datalog program
-$\Pi = \langle R, strip(F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i) \rangle$.
+$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.
 
@@
 
-The decomposition of streams into more sparse streams without loss of
-information allows us to define the state of a valid window
-$W^{P \setminus N}$ as a union of the states of sliding windows over
-each of the singleton streams $S_{\langle f, t \rangle} =
-(\langle f, t \rangle \mapsto t^\prime)$, where
-$S(\langle f, t \rangle) = t^\prime$.  The range of the sliding
-window over each $S_{\langle f, t \rangle}$ is then the difference between
-the instant at which $\langle f, t \rangle$ occurs in $P$ and that
-at which it occurs in $N$:
+The decomposition of any stream into any number of more
+sparse streams without loss of information allows us to define the
+state of a valid window $W^{P \setminus N}$ as a union of the states
+of sliding windows over each of the singleton streams
+$(\langle f, t \rangle \mapsto t)$, the union of which is equal to $P$,
+where $P(\langle f, t \rangle) = t$ means that $P$ maps
+$\langle f, t \rangle$ to $t$.  The range of the sliding
+window over each singleton stream is the difference between
+the instant at which $\langle f, t \rangle$ occurs in $P$ ($t$, as $P$
+is in-order) and that at which it occurs in $N$:
 \[
-  W^{P \setminus N}_i = \{ \langle f, t \rangle \mid \langle f, t \rangle \in W^{S_{\langle f, t \rangle},N(\langle f, t \rangle) - P(\langle f, t \rangle)}_i\}
+  W^{P \setminus N}_i = \left\{ \langle f, t \rangle \,\middle\vert\, \begin{split}
+    P(\langle f, t \rangle) = t \land \\
+    \langle f, t \rangle \in W^{\left( \langle f, t \rangle \mapsto t \right),N(\langle f, t \rangle) - t}_i
+  \end{split}\right\}
 \]
@@
 conventional Datalog program $\Pi = \langle R, F \rangle$. The
-relationship between these \ast-Datalog programs and their
+relationship between these $\ast$-Datalog programs and their
 entailments is expressed in the commutative diagram in

### 2015-11-17 article.tex
directly address points regarding count-based windows, persistent facts being annotated with the instant 0 and contradictions invalidating an entire window state.

@@
 
-Based on this definition of streams, we go on to define the notion of
-sliding windows over streams, as well as states of those windows at
-specific instants.  We do not consider count-based window functions, as
-these serve a semantic purpose at high specificity, e.g. \emph{retain
-the last ten buses to stop at a specific stop}.  In contrast, as will be
-formally described in \refdef{continuous datalog: CDP}, we consider only
-the application of the \emph{entirety of a rule set} within the
-consecutive states of one (or more) window(s); in this context, an
-example application of a count-based window might be to \emph{determine
-the implications of the last hundred facts reported by a sensor network,
-given some background knowledge regarding that network}.
+Based on this definition of streams, we go on to define a notion of
+sliding time-based windows over streams, as well as states of those windows at
+specific instants.  Every fact instantiated in the state of such a window at a
+given instant is considered to be true at that instant.  We do not consider count-based
+window functions, as the state of such windows (and therefore the truth of the
+facts instantiated therein) is determined by the contents of the stream as a
+whole, which is a form of aggregation that contrasts with that of time-based
+windows: where the states of the latter are determined by some
+function of each element of the stream (the time from the instant of its
+appearance thereon), the states of the former are determined by a function of
+the stream itself.  This is more similar to the aggregation functions of SQL,
+such as $count()$, $max()$, etc., which are not part of the semantics of
+positive Datalog or (outside of specific heavily restricted cases) OWL 2 RL
+\citep{w3cowl2profiles}.
+
+% serve a semantic purpose at high specificity, e.g. \emph{retain
+% the last ten buses to stop at a specific stop}.  In contrast, as will be
+% formally described in \refdef{continuous datalog: CDP}, we consider only
+% the application of the \emph{entirety of a rule set} within the
+% consecutive states of one (or more) window(s); in this context, an
+% example application of a count-based window might be to \emph{determine
+% the implications of the last hundred facts reported by a sensor network,
+% given some background knowledge regarding that network}.
 
@@
 C-Datalog program) $\Pi^C = \langle R, F^0, S \rangle$ to be a tuple
-which comprises a finite set of rules $R$, a finite set of persistent/static
-fact instances $F^0$ (all of which have the instant $0$, meaning that they
-cannot appear in any state of any window, by
-\refdef{continuous datalog: window} and Axioms~\ref{axiom:continuous datalog: window range leq 0} and
+which comprises a finite set of rules $R$, a finite set of persistent/\emph{static}
+fact instances $F^0$ (all of which have the instant $0$, meaning that the facts
+instantiated are recognised to be true from the beginning of time and cannot
+appear in any state of any window, by \refdef{continuous datalog: window} and
+Axioms~\ref{axiom:continuous datalog: window range leq 0} and
 \ref{axiom:continuous datalog: window start leq 0}) and an in-order stream
@@
 state $\Pi^{C,r}_i$ that is unsatisfiable, i.e. that contains at least
-one contradiction.  Such a state has no entailment.
+one contradiction.  Such a state has no entailment.  As this is not often
+desirable in real-world applications, policies may feasibly be employed in
+practise to eliminate facts from a window prematurely, with a view to making
+an inconsistent state satisfiable. However, such policies are not discussed
+here as the policy implemented (if any) depends upon the nature of the
+application.
 

### 2015-11-17 article.tex
small changes no time aaaaaaah

@@
   The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the \emph{iterative/recursive execution} of queries over the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}, in which each query expresses a rule in a logic program.
-  This reflects precisely the approach to reasoning using Datalog in deductive databases, which has been characterised as the ``QUOTE'' \citep{DATALOG CHARACTERISATION}.
+  This reflects precisely the approach to reasoning using Datalog in deductive databases, in which queries may be executed on intensional relations whose definitions must be used in turn to query the database, recursing in this manner until base only extensional relations are queried.
   Meanwhile, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
+  %%%FIXME
+  NEED SOMETHING ON TIME-DECAY MODEL
 
@@
 of entailments from this sequence of instantaneous states as a stream
-of facts; the data model of these streams is subsumed by that currently proposed
+of time-annotated facts. The data model of these streams is subsumed by that currently proposed
 by the RDF stream processing community\footnote{As presented at
 \url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}
-on \mbox{03/11/2015}.}, thereby allowing the integration of WC-Datalog programs
-into stream processing pipelines. Finally, we show that these streams may be
+on \mbox{03/11/2015}.}, in that each time-annotated fact may be expressed as a
+RDF graph containing a single triple that is annotated with one or more
+predicated timestamps; therefore integration of WC-Datalog programs
+into stream processing pipelines is inherently supported. Finally, we show that these streams may be
 manipulated to retrieve the instantaneous sets of entailments produced

### 2015-11-24 article.tex
sooooomany updates

@@
 \usepackage{amscd}
+\usepackage{mathtools}
 
@@
 $\bar{s}\left( \alpha_0 \right)$ in $\mathcal{I}^T$ is better described in terms
-of an entailment and its time-annotated justification, as per
+of the time-annotated justification of an entailed fact instance, as per
 \refdef{continuous datalog: instance justification}.
@@
 
+From the definition of entailment for TA-Datalog in \refdef{continuous
+  datalog: ta datalog program} and the definition of the states of sliding windows in
+\refdef{continuous datalog: window}, we can derive the following:
+
+{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
+If two states of windows over the same in-order stream $S$ cover
+non-overlapping time periods, then the set of fact instances entailed
+by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
+set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
+where ${z \geqslant x}$.
+\[
+\left\{ e \,\middle\vert\, \langle R , W^{S,x}_i \rangle \models e \right\} \cap \left\{ e \,\middle\vert\, \langle R , W^{S,y}_{i-z} \rangle \models e \right\} = \varnothing
+\]
+\end{axiom}}
+
 Given these definitions, a Continuous Datalog program may be formally
@@
 of fact instances $S$.  It should be noted that, in practice, data may be
-drawn from more than one source, and so $S$ may be a union of $n$ in-order
-streams $S_1,\dots,S_n$ without altering the semantics presented here.
+drawn from more than one source, and so $S$ may be a union of $s$ in-order
+streams $S_1,\dots,S_s$ without altering the semantics presented here.
 
@@
 \refdef{continuous datalog: instance justification}), and $E^S$ is an
-in-order stream, $E^S$ expresses the times at which each entailed fact
-instance was first justified by $\Pi^C$.
+in-order stream, $E^S$ expresses the times at and order in which each
+entailed fact instance was first justified by $\Pi^C$.
 
@@
 from then on.  As such, we define a {\em Windowed Continuous Datalog program}
-$\Pi^{C, r} = \langle R, F^0, \{ W^{S_1, r_1},\dots, W^{S_n, r_n} \rangle\} \rangle$
-(where $r = max(r_1,\dots,r_n)$) to be a tuple with components $R$ and $F^0$ as defined above for a
-C-Datalog program, but with each part of $S = S_1 \cup \dots \cup S_n$
+$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle\} \rangle$
+(where $r = max(r_1,\dots,r_s)$) to be a tuple with components $R$ and $F^0$ as defined above for a
+C-Datalog program, but with each part of $S = \mathop{\cup}^s_{x=1}S_x$
 associated with the range $r_x \in \mathbb{N}$ of a sliding window over
-$S_x$, for $x \in [1,n]$. This specifies that the fact instances
+$S_x$, for $x \in (1,s)$. This specifies that the fact instances
 from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
@@
 
-From the definition of entailment for TA-Datalog in \refdef{continuous
-  datalog: ta datalog program} and the definition of windows in
-\refdef{continuous datalog: window}, we can derive the following:
-
-{\nobreak\begin{axiom}\label{axiom:continuous datalog: entailment disjointness}
-If two states of windows over the same in-order stream $S$ cover
-non-overlapping time periods, then the set of fact instances entailed
-by ${\langle R, W^{S,x}_i \rangle}$ is necessarily disjoint from the
-set of fact instances entailed by $\langle R, W^{S,y}_{i-z} \rangle$,
-where ${z \geqslant x}$.
-\[
-\left\{ e \,\middle\vert\, \langle R , W^{S,x}_i \rangle \models e \right\} \cap \left\{ e \,\middle\vert\, \langle R , W^{S,y}_{i-z} \rangle \models e \right\} = \varnothing
-\]
-\end{axiom}}
-
 \begin{definition}[Instanteous State of a WC-Datalog program]
@@
 We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
-\{ W^{S_1, r_1},\dots, W^{S_n, r_n} \} \rangle$
+\{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle$
 at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
@@
 finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
-$r_x \in \mathbb{N}$, for all $x \in [1,n]$; we refer to this as the {\em instantaneous state} of
+$r_x \in \mathbb{N}$, for all $x \in (1,s)$; we refer to this as the {\em instantaneous state} of
 the WC-Datalog program, written
-$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.
+$\Pi^{C,r}_i = \langle R, F^0 \cup \mathop{\cup}\limits_{x=1}^s W^{S_x,r_x}_i \rangle$.
 As previously shown in \refdef{continuous datalog: ta datalog program}, this may then be further
@@
 is the set $E^T_{i}$ of entailed fact instances whose justifications are at
-least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i$.
+least partially contained in $\mathop{\cup}\limits_{x=1}^n W^{S_x,r_x}_i$.
 {\nobreak We define $E^T_i$ as:
 \[
-E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_j,r_j}_i \neq \varnothing \right) \right\}
+E^T_i = \left\{ e \,\middle\vert\, \Pi^{C,r}_i \models e \land \left( J^T \cap \bigcup_{x=1}^n W^{S_x,r_x}_i \neq \varnothing \right) \right\}
 \]
@@
 $\Pi^{C,r} = \langle R, F^0, \{ \langle S, r \rangle \} \rangle$ may be reconstructed from
-windows of sufficient range {\nobreak over its entailed streams $P$ and $N$ at
-$i$:
-\[ E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i} \]
+windows of sufficient range over its entailed streams $P$ and $N$ at
+$i$, as expressed in \refeqn{continuous datalog: entailment equivalence}.
+{\nobreak\begin{equation} \labeleqn{continuous datalog: entailment equivalence}
+  E^T_{i} = W^{P,x}_{i} \setminus W^{N,y}_{i}
+\end{equation}
 where ${y \geqslant x \geqslant r}$}. The minimum range $x$ is such to achieve
@@
 \emph{sound} set $E^T_{i}$.  It should be noted that the proof holds in the
-case of $n$ windows of varying range, with $x \geqslant max(r_0,\dots,r_n)$.
+case of $n$ windows of varying range, with $x \geqslant max(r_1,\dots,r_n)$.
 We capture this derivation in the form of a \emph{valid window}, as per
@@
 The decomposition of any stream into any number of more
-sparse streams without loss of information allows us to define the
-state of a valid window $W^{P \setminus N}$ as a union of the states
-of sliding windows over each of the singleton streams
-$(\langle f, t \rangle \mapsto t)$, the union of which is equal to $P$,
-where $P(\langle f, t \rangle) = t$ means that $P$ maps
-$\langle f, t \rangle$ to $t$.  The range of the sliding
-window over each singleton stream is the difference between
-the instant at which $\langle f, t \rangle$ occurs in $P$ ($t$, as $P$
-is in-order) and that at which it occurs in $N$:
-\[
-  W^{P \setminus N}_i = \left\{ \langle f, t \rangle \,\middle\vert\, \begin{split}
-    P(\langle f, t \rangle) = t \land \\
-    \langle f, t \rangle \in W^{\left( \langle f, t \rangle \mapsto t \right),N(\langle f, t \rangle) - t}_i
+sparse streams without loss of information allows us to alternatively define the
+state of the valid window $W^{P \setminus N}$ at any instant $i \in \mathbb{T}$
+as the union of the states of the set of sliding windows $\mathcal{W}^{P \setminus N}$
+over each of the singleton streams $(\langle f, t \rangle \mapsto t^\prime)$
+at $i$, the union of these streams being equal to $P$.  Remembering that
+$P(\langle f, t \rangle) = t^\prime$ means that $\langle f, t \rangle$ occurs
+in $P$ at $t^\prime$, the range of each sliding window in $\mathcal{W}^{P \setminus N}$
+is the difference $d$ between $P(\langle f, t \rangle)$ (which equals $t$, as $P$ is
+in-order) and $N(\langle f, t \rangle)$.  This difference is always positive by Axiom
+\ref{axiom:continuous datalog: entailment precedes negation}, and bounded at the
+upper limit $r$ of the program $\Pi^{C,r} \models \langle E^0, P, N \rangle$ by
+\refdef{continuous datalog: CDP}.
+\begin{gather}
+  W^{P \setminus N}_i = \bigcup_{W^{p,d} \in \mathcal{W}^{P \setminus N}} W^{p,d}_i \\
+  \mathcal{W}^{P \setminus N} = \left\{ W^{p,d} \,\middle\vert\, \begin{split}
+    \langle f, t \rangle \in P \land \\
+    p = (\langle f, t \rangle \mapsto t) \land \\
+    d = N(\langle f, t \rangle) - t
   \end{split}\right\}
-\]
+\end{gather}
 Treated this way, the entailment of one WC-Datalog program, e.g. that in
-\refeqn{continuous datalog: example WC-Datalog program},
-may then be taken as a part or all of the data in a successive WC-Datalog
-program, without modification of the existing semantics or loss of information
-regarding the validity of the entailments.  However, in the interest
-of being concise, such a successive WC-Datalog program may express valid
-windows alongside any sliding windows. For example, 
-a program composed of the set of rules $R^\prime \neq R$, the set of base facts
-$F^{0\prime} \neq F^0$, a sliding window of range $x \geqslant r$ over the
-in-order stream $S^\prime \neq S$, and the entailment of
+\refeqn{continuous datalog: example WC-Datalog program}, may then be taken as a
+part or all of the data in a successive WC-Datalog program, without modification
+of the existing semantics or loss of information regarding the validity of the
+entailments.
+% However, in the interest
+% of being concise, such a successive WC-Datalog program may express valid
+% windows alongside any sliding windows.
+For example, 
+a program composed of the set of rules $R^\prime$, the set of base facts
+$F^{0\prime}$, a sliding window of range $x \geqslant r$ over the
+in-order stream $S^\prime$, and the entailment of
 $\Pi^{C,r} \models \langle E^0, P, N \rangle$, would be expressed as in \refeqn{continuous datalog: successive WC-Datalog program}.
@@
   \labeleqn{continuous datalog: successive WC-Datalog program}
-    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{P \setminus N}, W^{S^\prime,x} \} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
+    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{S^\prime,x} \} \cup \mathcal{W}^{P \setminus N} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
 \end{gather}
@@
 In conclusion, given a Continuous Datalog program $\Pi^C = \langle R,
-F^0, S_0 \cup \dots \cup S_n \rangle$, we can apply to it a set of
-windows of range $r_0,\dots,r_n$ to yield a Windowed Continuous Datalog
-program $\Pi^{C,r} = \langle R, F^0, \{ W^{S_0, r_0}, \dots,
-W^{S_n, r_n} \} \rangle$. We can evaluate these windows at
-an instant $i \in \mathbb{T}$ to yield a Time-Annotated Datalog program
-$\Pi^T = \langle R, F^0 \cup W^{S_0,r_0}_i \cup \dots \cup W^{S_n,r_n}_i
-\rangle$, from we which can then strip the time annotations to yield a
-conventional Datalog program $\Pi = \langle R, F \rangle$. The
-relationship between these $\ast$-Datalog programs and their
-entailments is expressed in the commutative diagram in
-\reffig{CD-TD-D}.
-
+F^0, S_1 \cup \dots \cup S_n \rangle$, we can apply to it a set of
+windows of range $r_1,\dots,r_n$ to yield a Windowed Continuous Datalog
+program $\Pi^{C,r} = \langle R, F^0, \{ W^{S_1, r_1}, \dots,
+W^{S_n, r_n} \} \rangle$. We can evaluate these windows and a valid window
+over the streamed entailment at an instant $i \in \mathbb{T}$ to yield a
+Time-Annotated Datalog program $\Pi^T = \langle R, F^0 \cup W^{S_1,r_1}_i
+\cup \dots \cup W^{S_n,r_n}_i \rangle$.  From this we can then strip the
+time annotations to yield a conventional Datalog program $\Pi = \langle
+R, F \rangle$, where $F = strip(F^0 \cup W^{S_1,r_1}_i \cup \dots \cup
+W^{S_n,r_n}_i)$. The relationship between these $\ast$-Datalog programs,
+mirrored in the relationship between their entailments, is expressed in
+the commutative diagram in \reffig{CD-TD-D}.
+
+\begin{nestedsection}{In the Semantic Web: Windowed Continuous RIF-Core}{continuous datalog: rif}
+  It has been noted that RIF-Core \citep{w3crifcore} shares the semantics
+  of Positive Datalog, extended solely with a set of built-in predicates
+  describing the persistent characteristics of and relationships between
+  all literal values. These predicates require an updated notion of rule
+  \emph{safety}, but this is only relevant to the formulation of rules,
+  at the semantic level; as this is unchanged in any $\ast$-Datalog from
+  that of positive Datalog, the extended form of rule safety does not
+  impact the semantics of satisfiability and entailment of $\ast$-RIF-Core
+  programs, as an extension of $\ast$-Datalog programs.
+
+  With regards to the notation of WC-Datalog, these built-in predicates
+  may be represented by the set of facts $B$, or the set of persistent
+  fact instances ${B^0 = \{ \langle b, 0 \rangle \mid b \in B \}}$, that
+  are the collected extension of those predicates. A RIF-Core program may
+  therefore be represented as a Datalog program
+  ${\Pi = \langle R, F \cup B \rangle}$, a Time-Annotated RIF-Core program
+  as a TA-Datalog program ${\Pi^{T} = \langle R, F^T \cup B^0 \rangle}$, a
+  Continuous RIF-Core program as a C-Datalog program
+  ${\Pi^{C} = \langle R, F^0 \cup B^0, S_1 \cup \dots \cup S_n \rangle}$
+  and a Windowed Continuous RIF-Core program as a WC-Datalog program
+  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{W^{S_1,r_1}, \dots, W^{S_n,r_n}\} \rangle}$.
+  Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
+  ${\langle R, B^0 \rangle \models \varnothing}$, for any set $R$ of
+  \emph{safe} RIF-Core rules. As such, the form of entailment for each
+  type of RIF-Core program is the same as that for the equivalent variety
+  of Datalog program:
+  \begin{align*}
+    \Pi =&& \langle R, F \cup B \rangle &\models E \\
+    \Pi^{T} =&& \langle R, F^T \cup B^0 \rangle &\models E^T \\
+    \Pi^{C} =&& \langle R, F^0 \cup B^0, S \rangle &\models E^S \\
+    \Pi^{C,r} =&& \langle R, F^0 \cup B^0, S, r \rangle &\models \langle E^0, P, N \rangle \\
+    \Pi^{C,r}_i =&& \langle R, F^0 \cup B^0 \cup W^{S,r}_i \rangle &\models E^0 \cup E^T_i
+  \end{align*}
+
+  The reason for maintaining the distinction between $F$/$F^T$/$F^0$ and
+  $B$/$B^0$ is the fact that the sets $F$/$F^T$/$F^0$ are strictly finite,
+  whereas many of the extensions of the built-in predicates in $B$/$B^0$ are
+  strictly infinite. As such, these predicates are not expected to be
+  extended in any real collection of data, and so propositions based on
+  built-in predicates cannot be satisfied through an inspection of collected
+  facts, as is the case for both the facts in any $F$ and fact instances in
+  any $F^T$, $F^0$ and $S$; instead, propositions that extend built-in
+  predicates can be satisfied by a given interpretation by performing
+  calculations on arguments substituted by a variable substitution $\bar{s}$,
+  where the propositions in the same rule body that do not extend built-in
+  predicates are satisfied by the interpretation and the substitution
+  $\bar{s}$. This is the reason that propositions that extend built-in
+  predicates do not contribute to variable safety in RIF-Core, in the majority
+  of cases.
+
+  As can be seen, the built-in predicates of RIF-Core do not alter the
+  semantics of entailment in positive Datalog, nor do they alter these
+  semantics in its time-annotated and continuous extensions here presented.
+  Rather, they provide an additional set of persistent facts that may be
+  utilised within the set $R$ of rules for any *-Datalog program.
 \end{nestedsection}
 
+\end{nestedsection}
+
 \begin{nestedsection}{CoDeR Model for Continuous Deductive Reasoning}{model}
-  CoDeR is an abstract model for continuous deductive reasoning over streamed RDF data that expresses the semantics of Continuous Datalog, thus supporting continuous reasoning expressed in both RIF-Core and OWL 2 RL.
-  It is composed of a stream-based data model for expressing the streams $S$, $P$ and $N$ of each ${\Pi^{C,r} = \langle R,F^0,S,r \rangle \models \langle E^0,P,N \rangle}$, and a minimal algebra of stream-to-stream operators for expressing the persistent set of rules $R$.
+  \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
+  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_i$, $P$ and $N$ and the window ranges $r_i$ for $i \in (1,s)$ of each ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle \models \langle E^0,P,N \rangle}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window range $r$.
   It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
@@
   \begin{nestedsection}{Data Model}{model: data}
-    The data model of CoDeR is composed of multiple-source multiple-sink streams.
-    These streams are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
-
-    In the first case, input/output streams provide the semantics of the streams $S$, $P$ and $N$ for any $\Pi^{C,r}$ expressed by a CoDeR-based system.
-    These streams are sequences of instances of heterogeneous RDF graphs compatible with the ongoing work on query semantics by the RSP working group\footnote{\url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}};
-    an instance of an RDF graph is an RDF graph annotated with a time interval ${(e,n]}$: ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
-    As the atomic units of Continuous Datalog are instances of facts, being triples in RDF, graph instances arriving on the input stream must be translated into sequences of fact instances using the graph-to-triples operator described in \refsec{model: algebra}, where each fact instance inherits the time interval of the graph instance in which it originated.
-
-    The interval ${(e,n]}$ for any graph or triple instance is the the period of time for which the data is \emph{valid} \citep{SemanticStreamingManagement,sparkwave} according to the ${\Pi^{C,r} = \langle R, F^0, S, r \rangle}$ expressed.
-    For graphs on the input stream, $e$ is the instant at which a given graph was asserted, and ${n = e + r}$ is the instant at which it will leave the semantic sliding window over $S$ of $\Pi^{C,r}$ (or become ``negated'').
-    For graphs on the output stream, $e$ is the latest instant at which any part of the justification of a given graph was asserted, and $n$ is the earliest instant at which any part of its justification will leave the sliding window.
-    \reffig{intersected-intervals} provides an example of this for some ${\Pi^{C,r} = \langle \{c \Leftarrow a \land b\},\{\},[\langle a,1 \rangle,\langle b,2 \rangle ], 3\rangle}$ over instants ${i = 1 \dots 4}$.
+    The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
+  The data model of CoDeR is composed of multiple-source multiple-sink data streams, as shown in \reffig{CoDeR: stream}.
+  These streams are based on the notion of in-order streams presented in \refdef{continuous datalog: stream}, but are divided into two classes: \emph{input}/\emph{output} streams and \emph{inter-operator} streams.
+  In addition, the notion of an \emph{instance} is extended in \refdef{CoDeR: graph instance} from that of an instance of a \emph{fact} that became true at a specific instant, as presented in \refdef{continuous datalog: fact instance}.
+  The streams of CoDeR are streams of this extended form of instance, as per \refdef{CoDeR: stream}, which are compatible with the ongoing work on continuous query semantics by the RSP working group\footnote{According to \url{https://github.com/streamreasoning/RSP-QL/blob/master/Semantics.md}, 4th Nov 2015, a CoDeR graph instance may be expressed as an RSP graph with two timestamps predicated as `became valid' and `ceased to be valid'.}.
+  \begin{definition}[Graph Instance]\labeldef{CoDeR: graph instance}
+    A \emph{graph instance} is an RDF graph with some additional temporal annotation regarding both the instant from which it became valid and the instant at which it will cease to be valid;
+    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${(e,n]}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
+    The names $e$ and $n$ are so named because they relate to the instants of the instance's {$e$ntailment} and {$n$egation} by the program.
+    It should be noted that validity is a stronger concept than truth, as all facts within a graph that is valid in a system are understood to be true, but facts may be true despite not being represented in a graph that is valid within the system.
+    Two graph instances containing the same facts and annotation for $e$ and $n$ are considered identical;
+    two graph instances containing the same facts but different annotations for $e$ and/or $n$ are considered to be distinct.
+  \end{definition}
+  \begin{definition}[CoDeR Stream]\labeldef{CoDeR: stream}
+    All \emph{streams} of CoDeR are sequences of graph instances, using the definition of \emph{sequence} used in \refdef{continuous datalog: stream} of streams in WC-Datalog.
+    They are also \emph{in-order} with regards to their temporal annotation $e$, but potentially \emph{out-of-order} with regards to their temporal annotation $n$.
+  \end{definition}
+  \begin{definition}[Empty Stream]\labeldef{CoDeR: empty stream}
+    The \emph{empty stream} is a CoDeR stream that contains no graph instances, and is denoted by the symbol $\varnothing$.
+    This means that the stream will contain no graph instances that map to instant $i$, for all instants ${i \in \mathbb{T}}$.
+    This stream may occur in specific expressions of any program, as described in \refsec{CoDeR: transformations}.
+  \end{definition}
+
+  It should be noted that the atomic units of WC-RIF-Core are instances of facts, being \emph{triples} in RDF, and so graph instances must be translated into sequences of fact instances in order to truly express the stream $S$ of a program $\Pi^{C,r}$.
+  This is achieved by the graph-to-triples operator ${{}_g\pi_t}$ detailed in \refsec{CoDeR: graph-to-triples}, which produces a stream of graph instances in which each instantiated graph contains a single triple, thereby expressing a single fact.
+
   \begin{figure}
     \centering
-      \includegraphics[width=0.3\textwidth]{intersected-intervals}
-      \caption{Derivation of valid-time for entailed fact instance from those of its justification.}
-      \labelfig{intersected-intervals}
+    \includegraphics[width=0.45\textwidth]{logical-stream}
+    \caption[CoDeR Data Stream]{A stream of the CoDeR data model, contributed to by two sources and consumed from by two sinks.}
+    \labelfig{CoDeR: stream}
   \end{figure}
+
+  As has been noted in \refdef{CoDeR: graph instance}, the interval ${(e,n]}$ for any graph instance is the the period of time for which that graph instance is \emph{valid} \citep{SemanticStreamingManagement,sparkwave}.
+  The notion of validity within WC-RIF-Core programs relates to the period of time for which a given fact instance ${\langle f, t \rangle}$ is justified by a state ${\Pi^{C,r}_i = \langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_i \rangle \models \langle f, t \rangle}$ of a program $\Pi^{C,r}$.
+  As such, the interval ${(e,n]}$ is used to express the window ranges $r_x$ for $x \in (1,s)$ of the ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle}$:
+  for graph instances entering a CoDeR-based system, $e$ is the instant at which every fact in the given graph was asserted, and $n$ is the instant at which they will leave the semantic window over $S_x$ described by $\Pi^{C,r}$ (a.k.a.
+being ``negated''), where ${n = e + r_x}$ in the general case;
+  for graph instances produced by a CoDeR-based system, $e$ is the latest instant at which any fact within the given graph was asserted, and $n$ is the earliest instant at which any fact within the graph will be negated.
+  \reffig{CoDeR: intersected-intervals,CoDeR: consecutive states} provide an example of this for the ${\Pi^{C,3} = \langle \{c \Leftarrow a \land b\},\{\},\{ W^{S,3} \}, 3\rangle}$, where ${S = \langle a,1 \rangle,\langle b,2 \rangle}$ over instants ${i = 1 \dots 4}$.
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{intersected-intervals-example}
+    \caption[Derivation of valid-time for entailed fact instance from those of its justification]{Derivation of valid-time for entailed fact instance \emph{c} from those of its justification \emph{a} and \emph{b}.}
+    \labelfig{CoDeR: intersected-intervals}
+  \end{figure}
+  \begin{figure}[h!]
+    \centering
     \begin{align*}
-      \Pi^{C,r}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
-      \Pi^{C,r}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
-      \Pi^{C,r}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
-      \Pi^{C,r}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
+      \Pi^{C,3}_{1} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle \} \rangle & \models & \{ \} \\
+      \Pi^{C,3}_{2} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+      \Pi^{C,3}_{3} & = \langle \{c \Leftarrow a \land b\}, \{ \langle a, 1 \rangle, \langle b, 2 \rangle \} \rangle & \models & \{ \langle c, 2 \rangle \} \\
+      \Pi^{C,3}_{4} & = \langle \{c \Leftarrow a \land b\}, \{ \langle b, 2 \rangle \} & \models & \{ \}
     \end{align*}
-    It is clear that a CoDeR output stream expresses the semantic stream $P$ entailed by some $\Pi^{C,r}$.
-    However, as the semantic stream $N$ is defined as the sequence of entailments of some $\Pi^{C,r}$ ordered by the instant $n$ at which $W^{S,r}_{n}$ \emph{ceased} to justify them, the output stream of a CoDeR-based system also expresses the stream $N$.
-    As such, the annotation of streamed RDF graphs with the interval $(e,n]$ gives the streams of CoDeR a dual semantics with regard to Continuous Datalog.
-
-    In the second case, inter-operator streams are produced and consumed by the stream-to-stream operators of CoDeR's minimal algebra.
+    \caption{Consecutive states of a WC-RIF-Core program and their entailments}
+    \labelfig{CoDeR: consecutive states}
+  \end{figure}
+
+  The first class of CoDeR streams, input/output streams, provide the semantics of the streams and sets $S_x$, $P$, $N$, $F^0$ and $E^0$, for any ${\Pi^{C,r} = \langle R,F^0,\{W^{S_x,r_x} \mid x \in (1,s)\} \rangle \models \langle E^0, P, N \rangle}$ expressed by a CoDeR-based system.
+  These streams and sets are represented as CoDeR streams, as per \refdef{CoDeR: stream}, that are, in the case of input streams, produced elsewhere in the Semantic Web or, in the case of output streams, exposed on the Semantic Web as a resource themselves, as per \refdef{CoDeR: input/output}.
+  \begin{definition}[Input/Output Stream]\labeldef{CoDeR: input/output}
+    Both \emph{input} and \emph{output} streams of CoDeR are CoDeR streams that are identified on the Semantic Web by IRIs.
+    The graphs of which there are instances in an input/output stream are heterogeneous, in so much as each graph may contain any number of triples, each in turn being composed of any values.
+  \end{definition}
+  In each CoDeR-based system the input streams express the logical streams $S_x$ for $x \in (1,s)$ of the given $\Pi^{C,r}$, and a single output stream expresses the logical stream $P$ entailed by the $\Pi^{C,r}$.
+  This single output stream also expresses the stream $N$ entailed by the $\Pi^{C,r}$, as that logical stream is defined as the set of instances in the stream $P$ ordered in a sequence by the instant $n$ at which ${\langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_{n}}$ \emph{ceased} to justify them;
+  this instant $n$ is, therefore, the same $n$ with which each graph is annotated in CoDeR, being the instant at which the graph instance will cease to be valid.
+  As such, the annotation of streamed RDF graphs with the interval ${(e,n]}$ gives \emph{all} streams of CoDeR a dual semantics with regard to WC-RIF-Core.
+
+  In addition, the fact instances of $F^0$ are expressed as another input stream $S_0$ over which there is an infinite-range sliding window, where each graph instance is annotated with the instants ${e = 0}, {n = \infty}$ to indicate that it is valid indefinitely;
+  likewise, the fact instances $E^0$ appear at the start of the output stream, naturally annotated with the instants ${e = 0}, {n = \infty}$.
+
+  The second class of streams, inter-operator streams, are CoDeR streams that are both produced and consumed exclusively by the stream-to-stream operators of CoDeR's minimal algebra.
   These streams are sequences of the intermediate results of processing between steps in the production of entailments.
-    As the production of entailments in CoDeR is cast as an iterative pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs;
-    an inter-operator stream is a sequence of instances of homogeneous RDF graphs, each matching the same graph pattern, that being a sub-pattern of the body of one of the rules $R$ of the expressed $\Pi^{C,r}$.
-    These instances of RDF graphs could be considered to be ``entailed'' by the sub-pattern of the stream on which they appear and the presence of valid instances of their constituent RDF triples within the system;
-    as such, the same dual semantics of valid-times may be applied to these intermediate matches as to the true entailments of the model found on output streams.
-    It follows that inter-operator streams are functionally the same as input/output streams, only being distinct in their semantics:
-    the former are sequences of homogeneous RDF graphs that hold no value outside of a system, while the latter contain heterogeneous graphs that are either the entailments of the system or the assertions of its predecessor in the stream pipeline.
-
-    In addition, each stream in the model is semantically distinct:
-    each input/output stream is from a distinct source and uniquely identifiable by the source's IRI;
-    each inter-operator stream contains graphs matching a specific graph pattern and are uniquely identifiable by that pattern.
-    As the same graph pattern may be a sub-pattern of many rule bodies, and so appear in multiple places in a processing plan, one inter-operator stream may be the input to multiple operators.
-    Furthermore, two or more operators may contribute to the same stream within the model, where the pattern of the stream in question is the union of the patterns of the streams produced by each of the operators in isolation.
-    In this case, each operator contributes results to the common stream as they are produced, passively interleaving the streams produced by the individual operators such that the resultant stream is still ordered by the entailment instant $e$ of the constituent instances.
-    As this is the active behaviour of a stream union operator, such an operator is not needed in CoDeR.
-    However, within the algebra of CoDeR, such a stream is still denoted ${op_{application} \cup op_{application}}$.
+  As the production of entailments in CoDeR is cast as an iterative pattern matching problem, all intermediate results of this process will also take the form of instances of RDF graphs.
+  However, unlike the graph instances in input and output streams, all instances in a given inter-operator stream will be of graphs that match the same sequenced graph pattern, as per \refdef{CoDeR: sequenced pattern}, yielding \refdef{CoDeR: inter-operator} for inter-operator streams.
+  \begin{definition}[Sequential Conjunction]\labeldef{CoDeR: sequential conjunction}
+    \emph{Sequential conjunction} is the restriction of conventional conjunction to express not only that two facts are true, but that one precedes the other, for some notion of weak precedence.
+    In the context of CoDeR, where the validity of facts changes over time, the sequential conjunction of one fact $a$ with another $b$ is definable as in \refeqn{CoDeR: seq}.
+    \begin{equation}
+      \labeleqn{CoDeR: seq}
+      a\ seq\ b \equiv
+        \begin{split}
+          a \in g_a \land \\
+          b \in g_b \land \\
+          \langle g_a, e_a, n_a \rangle \land \\
+          \langle g_b, e_b, n_b \rangle \land \\
+          e_a \leqslant e_b < min\left( n_a , n_b \right)
+        \end{split}
+    \end{equation}
+    This means that not only are $a$ and $b$ valid at the same instant, but $a$ became valid during an instant prior to (or the same as) that at which $b$ became valid.
+    For example, if we replace the rule ${c \Leftarrow a \land b}$ in the continuous program \reffig{CoDeR: consecutive states} with ${c \Leftarrow a\ seq\ b}$, then the entailment of each state (and so the streamed entailment of the program) would remain unchanged.
+    However, if the rule that uses the complementary sequential conjunction ${c \Leftarrow b\ seq\ a}$ were substituted into that program, then its entailment would be empty, because $a$ strictly preceded $b$.
+  \end{definition}
+  \begin{definition}[Sequenced Pattern]\labeldef{CoDeR: sequenced pattern}
+    A \emph{sequenced pattern} is a SPARQL advanced graph pattern \citep{w3csparql}, extended with the notion of ordering between the components of the graph pattern.
+    Where SPARQL graph patterns may be composed using explicit disjunction (UNION and OPTIONAL clauses) or implicit conventional conjunction, sequenced graphs may also make use of explicit \emph{sequential conjunction} between its constituent components, as per \refdef{CoDeR: sequential conjunction}, where each component may be a triple pattern or another sequenced graph pattern.
+    As such, the set of all sequenced graph patterns completely subsumes the set of advanced graph patterns of SPARQL, those being sequenced graph patterns that make no use of sequential conjunction.
+
+    Sequential conjunction in graph patterns is expressed as a binary clause, SEQ, similar to ${seq}$ in EP-SPARQL \citep{EP-SPARQL}:
+    given a sequenced pattern ${\{x\ SEQ\ y\}}$ composed of the sequential conjunction of two subgraph patterns $x$ and $y$, any graph that matches the pattern must contain one subgraph $a$ that matches the pattern $x$ and another subgraph $b$ that matches the pattern $y$, where $a$ must have become valid before (or simultaneously with) $b$.
+  \end{definition}
+  \begin{definition}[Inter-Operator Stream]\labeldef{CoDeR: inter-operator}
+    The \emph{inter-operator} streams of CoDeR are CoDeR streams that are each associated with a specific \emph{sequenced graph pattern} that is a sub-pattern of the body of one of the rules $R$.
+    The graphs of which there are instances in an inter-operator stream are homogeneous with respect to the graph pattern associated with that stream, in so much as every such graph is a ``match'' for that graph pattern, as per \refdef{CoDeR: sequenced pattern}.
+    While the ordering in a sequenced graph pattern is not represented directly in the graph instances matched, those being temporally-annotated RDF graphs, the nature of the processing that produces a given graph instance may dictate a \emph{necessary ordering} between the components of every graph instance that it produces (see \refsec{CoDeR: SteM}).
+    It is this \emph{necessary ordering} that is represented by the ordering in a sequenced graph pattern.
+  \end{definition}
+  It should be noted that, by \refdef{CoDeR: sequenced pattern} of sequenced patterns, the conventional conjunction between two patterns $a$ and $b$ is semantically equivalent to the disjunction of two graph patterns that are complementary sequential conjunctions between the two component patterns, as shown in \refeqn{CoDeR: disjunction of sequential conjunctions}.
+  As such, any SPARQL graph pattern may be expressed without the use of conventional conjunction.
+  \begin{equation}
+    \labeleqn{CoDeR: disjunction of sequential conjunctions}
+    \{a\ SEQ\ b\}\ UNION\ \{b\ SEQ\ a\} \equiv \{a\ .\ b\ .\}
+  \end{equation}
+
+  It follows that the distinction between input/output streams and inter-operator streams is one of semantics, rather than functionality.
+  Both are, syntactically, CoDeR streams as per \refdef{CoDeR: stream}.
+  Semantically, the contents of input streams are the assertions of the predecessor of a CoDeR-based system in a wider stream processing pipeline, as identified on the Semantic Web by an IRI, and cannot be expected to match any given pattern.
+  It follows that the same is true of the output stream, as the entire contents of all input streams is part of the entailment of a WC-RIF-Core program.
+  In contrast, the contents of inter-operator streams are not expected, in the general case, to hold any intrinsic value outside of a system (though there is no reason it should not be exposed on the Semantic Web);
+  instead, they contain the intermediate result of some highly specific part of the inference over the program expressed by the system, represented by a sequenced graph pattern.
   \end{nestedsection}
@@
   % Minimal Algebra
-  \begin{nestedsection}{Minimal Algebra}{model: algebra}
-    The minimal algebra of CoDeR is based on that of Positive Datalog, that being ${\{\bot, \text{:-}, \wedge\}}$.
-    Of these, implication $\text{:-}$ and conjunction $\wedge$ may be expressed as stream-to-stream operators, the former being extended to apply to instances as $\text{:-}_{head}$ and the latter being the window-join $\Join$ \citep{niagaraCQ}, in order to support forward-chaining processing of fixed rules;
-    $\bot$ is not an operator, but the special class of values that should not exist, represented in CoDeR by the ${rif\text{:}error()}$ class from RIF \citep{w3crif}.
-    % Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, algebraically represented by ${{}_b\Join_p}$.
-    Further, in order to support both fixed-plan pattern-matching algorithms such as the Rete algorithm \citep{forgy79} and adaptive algorithms such as the Eddies algorithm \citep{eddies} and its derivatives \citep{CACQ,TCQ}, CoDeR replaces the window-join operator with the State Module of Eddies, which we represent algebraically by $\rstreamjoin$.
-    % In addition to $\text{:-}_{head}$ and ${{}_b\Join_p}$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
-    In addition to $\text{:-}_{head}$ and $\rstreamjoin$, CoDeR utilises a pair of stream-filter operators, $\sigma_{triple}$ and $\sigma_{external}$, for identifying \emph{basic patterns} as defined in SPARQL \citep{w3csparql} and expressing the datatype functions of RIF-Core \citep{w3crifcore}, respectively.
-    Finally, as basic patterns match individual triples and CoDeR's streams are composed of graphs, the CoDeR algebra includes a graph-to-triples operator ${{}_g\pi_t}$ to translate input streams to the streams of triples on which $\sigma_{triple}$ operates.
-    % In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, {}_b{\Join_p}, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
-    In summary, the minimal algebra of CoDeR consists of the set of operators ${\{\text{:-}_{head}, \rstreamjoin, \sigma_{triple}, \sigma_{external}, {}_g{\pi_t}\}}$.
-
-    As all operators are stream-to-stream in nature, they each take one or two streams as input and produce a single stream (though they may contribute to any number of union streams).
-    As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be used as input to another operator.
-    For example, \refeqn{basic-example-plan} is a plan expressed in the CoDeR algebra for the ${\Pi^{C,r} = \langle \{b \Leftarrow a\}, \{\}, S, r \rangle}$, where $a$ and $b$ are RDF triples.
-    \begin{equation}\labeleqn{basic-example-plan}
-      \text{:-}_b \left( \sigma_a \left( {}_g{\pi_t}\,S \right) \right)
-    \end{equation}
-
-    \begin{description}
-      \item[$\sigma_{triple}\,X$] \hfill \\
-        \begin{figure}[t]
+  \begin{nestedsection}{An Algebra for CoDeR}{model: algebra}
+    An algebra for CoDeR is any set of operators fit for the purpose of describing the process of continuous inference, as expressed in any WC-RIF-Core program, over CoDeR streams.
+    In this section we present such a set of operators, detailed individually with regards to the nature of the CoDeR streams that each produces and may be applied to, as well as the specific actions that each takes with respect to each graph instance it receives.
+    We go on to present the transformations of expressions in this algebra that support the optimisation of processing WC-RIF-Core programs.
+
+    The algebra fundamentally consists of the following set of stream-to-stream operators: ${\{{}_g\pi_t, \sigma_{triple}, \sigma_{external}, \coloneq_{head}, \cup, \rstreamjoin\}}$.
+    These operators encapsulate the practical expression of the set of functions required to derive the entailment of WC-RIF-Core programs in the context of CoDeR input/output streams.
+    These functions are those required for WC-Datalog programs, with the addition of two functions:
+    a one-pass unary evaluation function $\sigma_{external}$ for the built-in predicates of RIF-Core;
+    and a one-pass unary graph decomposition function ${{}_g\pi_t}$ to convert the CoDeR input stream of heterogeneous RDF graph instances into a stream of homogeneous RDF triple instances, the latter being the form of a stream $S_x$ of a program ${\Pi^{C,r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,n) \} \rangle}$.
+    The continuous derivation of the entailment of WC-Datalog programs requires, in turn, the set of functions required for querying deductive databases using Datalog, with the addition of another two functions:
+    a two-pass unary valid-window maintenance function $w$ to enforce the windows of range $r_x$, for $x \in (1,n)$, over the CoDeR input streams;
+    and a two-pass unary window-to-stream function $s$ to extract the stream of novel graph instances from a sequence of consecutive valid-windows, closely related to the ${IStream()}$ function of CQL \citep{StanfordStream}.
+    Finally, the functions required for querying deductive databases using Datalog are conventionally the complete set required of the relational algebra\footnote{This refers to the complete relational algebra less the functionality of set minus $\setminus$, which is analogous to the NaF operator $\sim$ that is not supported by positive Datalog.}, being ${\{\sigma,\pi,\Join,\cup\}}$ as described by \citet{codd70relationalModel}.
+    However, three of these may be restricted when processing semantic data, as discussed in \refsec{background: semantic queries}, becoming:
+    the one-pass unary equality-based triple-select $\sigma_{triple}$;
+    the two-pass binary natural join $\Join^\ast$ between \emph{sets} of variable bindings for graph patterns (requiring valid-windows over streams as input, and producing valid-windows);
+    and the one-pass unary template-project $\pi_{triple}$ that projects constant values and bindings for variables in graph patterns from graph instances into new triple instances.
+    The one-pass binary set operator $\cup$ utilised by the relational algebra need only be modified to preserve the mapping of instances to instants when applied to CoDeR streams.
+
+    All one-pass functions required to derive the entailment of a WC-RIF-Core program expressed by an equivalent stream-to-stream operator in CoDeR, as detailed in \reftab{WC-RIF-Core-functions-to-CoDeR-operators}.
+    \begin{table}
+        \centering
+        \begin{tabular}{r || c | c | c | c | c | c | c | c}
+          WC-RIF-Core & $\sigma_{triple}$ & $\sigma_{external}$ & ${}_g\pi_t$ & $\pi_{triple}$ & $\cup$ & $w$ & $s$ & $\Join^\ast$ \\ \hline
+          CoDeR & $\sigma_{triple}$ & $\sigma_{external}$ & ${}_g\pi_t$ & $\coloneq_{head}$ & $\cup$ & \multicolumn{3}{c}{$\rstreamjoin \cup \rstreamjoin$}
+        \end{tabular}
+        \caption[WC-RIF-Core Functions to CoDeR Operators]{The mapping of each function of WC-RIF-Core to their expression using the operators of the CoDeR algebra.}
+        \labeltab{WC-RIF-Core-functions-to-CoDeR-operators}
+      \end{table}
+      The two-pass functions $\{w,s,\Join^\ast\}$, as shown in the same table, are combined into a single stream-to-stream operator in CoDeR in order to make the algebra universally stream-to-stream in nature.
+      This is possible because $w$ is solely used to produce a set from a stream, $s$ to produce a stream from a set, and $\Join^\ast$ is the only function that must specifically act between sets;
+      it follows that the former two will always, and need only, be used in direct conjunction with the latter in the form $s(w(X) \Join^\ast w(Y))$, where $X$ and $Y$ are inter-operator streams.
+
+    As all the operators of this algebra are stream-to-stream in nature, they each receive one or two streams and produce a single stream.
+    As such, the notation for applying an operator to a stream is also the definition of a stream, and may, therefore, be the argument for another operator.
+    % , termed the \emph{successor} operator, as per \refdef{CoDeR: successor operator}.
+    % For example, \refeqn{basic-example-plan} is a plan for the ${\Pi^{C,r} = \langle \{b \Leftarrow a\}, \{\}, S, r \rangle}$ expressed in the CoDeR algebra, where $a$ and $b$ are RDF triples.
+    % \begin{equation}\labeleqn{basic-example-plan}
+    %   \coloneq_b \left( \sigma_a \left( {}_g{\pi_t}\ S \right) \right)
+    % \end{equation}
+    % In this case, $\sigma_a$ is the successor operator of ${{}_g\pi_t}$, and is succeeded by $\coloneq_b$ in turn.
+
+    % \begin{definition}[Successor Operator]\labeldef{CoDeR: successor operator}
+    %   A \emph{successor} of an operator $a$ is any operator $b$ that receives the stream produced by $a$, i.e. $b \left( a \left( S \right) \right)$ for some suitable stream $S$.
+    %   This relationship is denoted ${a \circ b}$, as it is analogous to operator composition, and may be chained to represent extended operator pipelines, e.g. ${a \circ b \circ c}$.
+    %   Where an asymmetric binary operator is a successor, the succeeding function of the operator is noted above the successor symbol, e.g.
+${{}_g\pi_t \circ \sigma_p \overset{build}{\circ} \rstreamjoin}$, where $p$ is some triple pattern.
+    % \end{definition}
+
+    The following descriptions present a thorough treatment of the individual operators of the algebra.
+    Each provides a description of its behaviour on a triple-by-triple basis, and the formal definition of the stream produced in terms of the stream(s) to which the operator is applied.
+    Because many of the operators' behaviours are defined in terms of bindings within graph instances for variables in the sequenced pattern associated with a received stream, the functions $vars(stream)$ and $binding(graph, variable)$ are defined:
+    the former maps inter-operator streams to the sets of variables in their associated patterns;
+    the latter maps graphs and variables from any inter-operator stream and associated pattern to the value bound to that variable in that graph.
+
+    \begin{nestedsection}{Graph-to-Triples}{CoDeR: graph-to-triples}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.3\textwidth]{graph-to-triple}
+        \caption{The graph-to-triple operator.}
+        \labelfig{CoDeR: graph-to-triple}
+      \end{figure}
+      The graph-to-triples operator ${{}_g\pi_t}$ is a one-pass unary operator that deconstructs graph instances into sequences of triple instances, the union of those triples instantiated being equal to the contents of the original graph.
+      It takes an \emph{input} stream $X$ and produces an \emph{inter-operator} stream ${{}_g\pi_t\ X}$, whose associated pattern is always the singleton pattern ${\{?subj\ ?pred\ ?obj\ .\}}$.
+
+      Each singleton graph instance on the produced stream is composed of a triple drawn from a received graph instance and the same valid-time interval as that graph instance.
+      Consider, for example, the input stream containing an instance of a graph containing the RDF triples $a$ and $b$ and valid from instant $e$ to instant $n$:
+      \[
+        {}_g{\pi_t} (\dots,\langle \{a\ .\ b\ .\},e,n \rangle,\dots) =
+          (\dots,\langle \{a\ .\},e,n \rangle,\langle \{b\ .\},e,n \rangle,\dots)
+      \]
+      More generally:
+      \begin{equation*}
+        {}_g{\pi_t} \left( X \right) =
+        \left(
+          \langle \{t .\}, e, n \rangle \:\middle\vert\: \langle g, e, n \rangle \in X \land t \in g
+        \right)
+      \end{equation*}
+    \end{nestedsection}
+    \begin{nestedsection}{Triple-Select}{CoDeR: triple-select}
+      The triple-select operator identifies instances of triples in a stream $X$ that match some \emph{triple pattern} \citep{w3csparql}.
+      This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream of triple instances, rather than a set of triples, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to temporally annotated semantic data.
+      Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the triple pattern to be matched.
+      \begin{figure}
         \centering
@@
         \caption{The triple-select operator.}
-          \labelfig{continuous datalog: basic pattern match}
+        \labelfig{CoDeR: basic pattern match}
       \end{figure}
-        The triple-select operator identifies instances of triples in a stream $X$ that match some \emph{basic pattern} \citep{w3csparql}.
-        This operator is semantically equivalent to the \emph{basic pattern match} of SPARQL applied to a stream, and to the \emph{alpha nodes} of the Rete pattern matching algorithm \citep{forgy79} applied to semantic data.
-        Like these operators, the triple-select is one-pass, and so is trivially interpreted as the unary stream-to-stream operator $\sigma_{triple}$, where ${triple}$ is the basic pattern to be matched.
-        It takes an inter-operator stream $X$ of triple instances as input and produces an inter-operator stream of triple instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
-        The contents of the produced stream are characterised as those triple instances from the input stream whose triples match the specified basic pattern.
-      \item[$\sigma_{external}\,X$] \hfill \\
-        \begin{figure}[b]
+
+      The triple-select takes an inter-operator stream $X$ of singleton graph instances and produces an inter-operator stream, where the produced stream contains a subset of those instances in the received stream and maintains the ordering of those instances.
+      The set of instances produced is defined as the subset of those received that contains only those singleton graph instances for whom the single triple matches the specified triple pattern.
+
+      More formally, where ${matches(graph,pattern)}$ is true iff the $graph$ matches the $pattern$ as per \refdef{CoDeR: sequenced pattern}:
+      \begin{equation*}
+        \sigma_{triple} \left( X \right) =
+        \left(
+          \langle g, e, n \rangle \:\middle\vert\: \langle g, e, n \rangle \in X \land matches\left(g,\{triple\ .\}\right)
+        \right)
+      \end{equation*}
+    \end{nestedsection}
+    \begin{nestedsection}{External-Select}{CoDeR: external-select}
+      The external-select operator encompasses all of the one-pass built-in datatype predicates and functions of RIF-Core.
+      Like the triple-select, it is trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is one of the built-in predicates of RIF-Core and each $a_i$, ${1 \leq i \leq n}$, is either a variable, a constant value or a built-in function of RIF-Core to which $f$ is to be applied.
+      \begin{figure}
         \centering
@@
         \caption{The operator encompassing all datatype functions of RIF-Core.}
-          \labelfig{continuous datalog: datatype function}
+        \labelfig{CoDeR: datatype function}
       \end{figure}
-        This operator encompasses all of the one-pass built-in datatype functions of RIF-Core, and is also trivially interpreted as a unary stream-to-stream operator $\sigma_{external}$, where ${external = f(a_1,\dots,a_n)}$, $f$ is the datatype function to apply and each $a_i$, ${1 \leq i \leq n}$ is either a variable or a constant value to which $f$ is to be applied.
-        It filters elements of its input stream $X$ based not on the semantic structure of the data but directly on the values of variables, which are consistent throughout any given graph pattern;
-        thus the $\sigma_{external}$ operator expresses an \emph{advanced} component of an \emph{advanced graph pattern}, as specified in SPARQL.
-        As such, it takes an inter-operator stream $X$ of graph instances as input and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the input stream, maintaining the ordering of those instances.
-        The contents of the produced stream are characterised as those graph instances from the input stream whose graphs match a specific advanced graph pattern.
-      \item[${{}_g{\pi_t}\,X}$] \hfill \\
-        \begin{figure}[t]
-          \centering
-          \includegraphics[width=0.3\textwidth]{graph-to-triple}
-          \caption{The graph-to-triple operator.}
-          \labelfig{continuous datalog: graph-to-triple}
-        \end{figure}
-        The graph-to-triples operator ${{}_g{\pi_t}}$ is a unary operator that deconstructs graph instances into sequences of triple instances, the union of the triples instantiated therein being equal to the contents of the original graph.
-        It takes a heterogeneous (\emph{input/output}) stream $X$ of RDF graph instances and produces a homogeneous (\emph{inter-operator}) stream of RDF triple instances.
-        Each triple instance produced is assigned the same valid-time interval as the graph instance in whose graph the instantiated triple was contained.
-        Consider, for example, the input/output stream containing an instance of a graph containing the RDF triples $a$ and $b$ and valid from instant $t_e$ to instant $t_n$:
+      It should be noted that the built-in functions of RIF-Core resemble the built-in predicates, but evaluate to a value rather than either ${true}$ or ${false}$, and so are only valid as arguments to another built-in predicate or function.
+      The $\sigma_{external}$ filters elements of its received stream $X$ based not on the semantic structure of the data but directly on the values bound within each graph instance for variables of the graph pattern of $X$;
+      thus the $\sigma_{external}$ operator not only expresses the built-in predicates and functions of RIF-Core but is analogous to the non-NaF FILTER clauses of \emph{advanced graph patterns}, as specified in SPARQL.
+
+      The external-select takes an inter-operator stream $X$ and produces an inter-operator stream of graph instances, where the produced stream contains a subset of those instances in the received stream and maintains the ordering of those instances.
+      The contents of the produced stream are characterised as those graph instances from the received stream whose graphs match the sequenced pattern that is the conventional conjunction of the pattern of $X$ and the proposition ${external}$.
+      The sequenced pattern associated with stream $X$ must contain all the variables referenced as arguments that must be bound in the proposition ${external}$ \citep{w3crifcoreBuiltins}, as the proposition will always evaluate to false where not all necessary arguments are bound.
+      As such, the produced stream will necessarily be the empty stream $\varnothing$ (as per \refeqn{CoDeR: empty result of external select over unbound variables} in \refsec{CoDeR: transformations}).
+
+      More formally:
+      % \begin{equation*}
       \begin{multline*}
-          {}_g{\pi_t} (\dots,\langle \{a,b\},t_e,t_n \rangle,\dots) = \\
-            (\dots,\langle a,t_e,t_n \rangle,\langle b,t_e,t_n \rangle,\dots)
+        \sigma_{f(a_0,\dots,a_n)} \left( X \right) =
+        \left(
+          \langle g, e, n \rangle
+        \:\middle\vert\:
+          \langle g, e, n \rangle \in X \land \right. \\
+          \begin{split}
+            \forall x \in (0,n) v_x = \left\{
+              \begin{split}
+                & binding\left(g, a_x\right) && \text{, if } a_x \in vars(X) \\
+                & a_x && \text{, otherwise}
+              \end{split}\right. & \land \\
+            f(v_{0},\dots,v_{n}) &
+        \left.\right)
+          \end{split}
+      % \end{equation*}
       \end{multline*}
-      \item[$\text{:-}_{head}\,X$] \hfill \\
-        \begin{figure}[b]
+    \end{nestedsection}
+    \begin{nestedsection}{Instance Implication}{CoDeR: instance implication}
+      \begin{figure}
         \centering
@@
         \caption{The instance implication operator.}
-          \labelfig{continuous datalog: instance implication}
+        \labelfig{CoDeR: instance implication}
       \end{figure}
-        Instance implication is a one-pass operator that takes as input an inter-operator stream $X$ whose graph pattern is the body of one of the rules $R$ of the expressed $\Pi^{C,r}$, and produces an inter-operator stream whose graph pattern is the head of that axiomatic rule.
-        Each graph instance received on the input stream causes the production of a single graph instance that matches the pattern of the head clause and is populated with the bindings of the received instance.
+      Instance implication $\coloneq_{head}$ is a one-pass unary operator that takes an inter-operator stream $X$ whose associated pattern is the body of one of the rules $R$ of the expressed $\Pi^{C,r}$, and produces an inter-operator stream whose associated pattern is the head of that axiomatic rule, ${head}$.
+      Each graph instance received on $X$ causes the production of a single graph instance that matches the pattern of the head clause and is populated with the values bound to the pertinent variables within the received instance.
       This is similar to the CONSTRUCT clause of SPARQL, or the final step of \emph{conflict set resolution} within the Rete algorithm \citep{forgy79}.
-        In addition, each produced instance inherits its valid-time interval from the graph instance that implied it.
-      % \item[${S_1\,{}_b{\Join_p}\,S_2}$] \hfill \\
-      \item[${X\,\rstreamjoin\,Y}$] \hfill \\
-        \begin{figure}[t]
+      In addition, each produced instance inherits its valid-time interval from the graph instance from which it is derived.
+
+      More formally, using the example case of a singleton graph pattern ${head = \{ a_0\ a_1\ a_2\ .\}}$ for readability:
+      % \begin{equation*}
+      \begin{multline*}
+        \coloneq_{\{ a_0\ a_1\ a_2\ .\}} \left( X \right) =
+        \left(
+          \langle \{v_{0}\ v_{1}\ v_{2}\ .\}, e, n \rangle
+        \:\middle\vert\:
+          \langle g, e, n \rangle \in X \land \right. \\
+            \forall x \in (0,2) v_x = \left\{
+              \begin{split}
+                & binding\left(g, a_x\right) && \text{, if } a_x \in vars(X) \\
+                & a_x && \text{, otherwise}
+              \end{split}
+        \right)
+      % \end{equation*}
+      \end{multline*}
+    \end{nestedsection}
+    \begin{nestedsection}{Stream Union}{CoDeR: stream union}
+      Stream union is a one-pass binary operator $\cup$, meaning that it takes two streams and produces a single stream that is composed of all the mappings of instances to times from both received streams.
+      No new instances are created by the stream union operator, unlike the projection-like operators ${{}_g\pi_t}$ and $\coloneq_{head}$, nor are any instances filtered from the received stream, unlike the select-based operators $\sigma_{triple}$ and $\sigma_{external}$.
+      Both the streams received and the stream produced may be either input/output streams or inter-operator streams, but the produced stream may only be an inter-operator stream if \emph{both} of its received streams are as well:
+      if one of the received streams is not an inter-operator stream then there is no way to reliably define the pattern matched by each instance in the produced stream, as is necessary for an inter-operator stream.
+
+      More formally:
+      \begin{equation*}
+        \left( X \right) \cup \left( Y \right) =
+        \left(
+          \langle g_i, e_i, n_i \rangle
+        \:\middle\vert\:
+            \langle g, e, n \rangle \in X \lor \langle g, e, n \rangle \in Y
+        \right)
+      \end{equation*}
+
+      As shown in \reffig{CoDeR: stream union}, this algebraic operator does not have an operator node to represent it in dataflow representations of a processing plan.
+      \begin{figure}
         \centering
-          \begin{minipage}[b]{0.45\textwidth}\includegraphics[width=\textwidth]{SteM}\end{minipage}\\\hfill\\
-          \begin{minipage}[b]{0.45\textwidth}\includegraphics[width=\textwidth]{window-join}\end{minipage}
-          \caption{Exploded views of two-pass joining operators.\hspace{\textwidth}Top: the SteM operator. Bottom: a typical window-join.}
-          \labelfig{continuous datalog: SteM}
+        \includegraphics[width=0.45\textwidth]{stream-union}
+        \caption{The expression of stream union in CoDeR as the contribution of multiple operators to the same stream.}
+        \labelfig{CoDeR: stream union}
       \end{figure}
-        State Modules (\emph{SteM}s) are asymmetrical binary operators that each perform the dual function of maintaining a window of graph instances and ``joining'' each graph instance on a stream against those within the maintained window (known as ``probing'' the window).
-        These are the same two functions performed for each window in a symmetric window-join, extended from Rete \emph{beta nodes} in \citep{ReteDBMS}, with results of each probe being produced to the same result stream as they occur.
-        Window-joins are, therefore, functionally identical to the union of two complementary SteMs:
-        % ${A \Join B = \left( A\,{}_b{\Join_p}\,B \right) \cup \left( B\,{}_b{\Join_p}\,A \right)}$, where $A$ and $B$ are inter-operator streams.
-        ${A \Join B = \left( A\,\rstreamjoin\,B \right) \cup \left( B\,\rstreamjoin\,A \right)}$, where $A$ and $B$ are inter-operator streams.
-        In addition, it should be noted that SteMs express the semantics of the sequential intersection operator ${seq}$ proposed in EP-SPARQL \citep{EP-SPARQL}, in that they join new data from one stream with the prior (valid) contents of another, thereby supporting rudimentary event processing.
-
-        The implementations of SteMs in the literature show a great variety of behaviour regarding window maintenance \citep{SteMs}, in the most extreme case ceasing to be a wholly stream-to-stream operator and supporting the incorporation of ``static'' data into stream processing systems.
-        In the minimal algebra of CoDeR, however, SteMs simply support the building of a \emph{valid window} (as defined in \refsec{semantics}) from its first inter-operator stream $X$, and the probing of that window with data from its second inter-operator stream $Y$ to produce an inter-operator stream of the homogeneously joined graph instances.
-        Furthermore, the nature of probing in CoDeR is strictly limited to that of the natural join between the variables in the patterns for streams $X$ and $Y$.
-        Finally, in contrast to the direct inheritance of valid-times in unary operators of CoDeR, the valid-time interval assigned to each graph produced by a SteM is the intersection of the valid-times of the two joined graphs.
-    \end{description}
+      This is because the multi-source and in-order natures of CoDeR data streams allows for the behaviour of $\cup$ to be expressed passively through those streams, rather than within a distinct data processing operator.
+    \end{nestedsection}
+    \begin{nestedsection}{SteM}{CoDeR: SteM}
+      State Modules were originally proposed in the adaptive continuous processing literature, and their potential implementations for static and sliding-window datasets are thoroughly detailed by \citet{SteMs}.
+      In CoDeR they are asymmetrical binary operators, represented algebraically by the binary operator $\rstreamjoin$, that simply supports the maintenance (or ``building'') of a \emph{valid window}, as per \refdef{continuous datalog: valid window}, from one inter-operator stream and naturally joining each graph instance on a second inter-operator stream against those within the maintained window (known as ``probing'' the window).
+      Since the persistent fact instances of $F^0$ are simply included as the CoDeR input stream $S_0$ and are annotated as being perpetually valid, as discussed in \refsec{CoDeR: data}, the applicable part of $F^0$ (and its entailment with respect to the rules $R$) will be perpetuated indefinitely in each SteM.
+      The nature of probing in CoDeR is the same as that of joining the sets of matches for two graph patterns in SPARQL.
+      As such, ${X \rstreamjoin Y}$ produces an inter-operator stream of graph instances rather than a stream of tuple instances that would be produced by a conventional natural window-join.
+
+      Building and probing a window are the same two functions performed for each window in a symmetric natural window-join (that shall be noted as the operator ${X \Join^\ast_w Y}$ in this section), as discussed by \citet{niagaraCQ} and extended from Rete \emph{beta nodes} in \citep{ReteDBMS}, with results of probing each window being produced to the same result stream.
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.45\textwidth]{SteM}\\
+        \includegraphics[width=0.45\textwidth]{window-join}
+        \caption{Exploded views of two-pass joining stream operators.}
+        \labelfig{CoDeR: SteM}
+      \end{figure}
+      Therefore, natural window-joins are functionally identical to the stream union of two \emph{complementary} SteMs (i.e.
+${X \rstreamjoin Y}$ and ${Y \rstreamjoin X}$) applied to inter-operator streams, as in \refeqn{window-join-from-union-of-SteMs}.
+      \begin{equation}\labeleqn{window-join-from-union-of-SteMs}
+        X \Join^\ast_w Y = \left( X \rstreamjoin Y \right) \cup \left( Y \rstreamjoin X \right)
+      \end{equation}
+      In addition, it should be noted that SteMs express the semantics of the sequential join operator ${seq}$ proposed in EP-SPARQL \citep{EP-SPARQL}, in that they join new data from one stream with the prior (valid) contents of another, thereby supporting rudimentary event processing.
+      As such, the sequenced pattern associated with the stream produced by ${X \rstreamjoin Y}$ would be ${\{x\ SEQ\ y\}}$, as per \refdef{CoDeR: sequenced pattern}, where $x$ is the pattern associated with stream $X$, and $y$ is the pattern associated with $Y$.
+      This leads to the isomorphism between \refeqn{CoDeR: disjunction of sequential conjunctions,window-join-from-union-of-SteMs}.
+
+      Finally, in contrast to the direct inheritance of valid-times in ${{}_g\pi_t}$ and $\coloneq_{head}$, the valid-time interval assigned to each graph produced by a SteM is the intersection of the valid-times of the two unioned graphs.
+More formally:
+      \begin{equation*}
+        \left( X \right) \rstreamjoin \left( Y \right) =
+        \left(
+          \langle g, e_Y, min \left( n_{X}, n_{Y} \right) \rangle \:\middle\vert\:
+          \begin{split}
+            \langle g_X, e_{X}, n_{X} \rangle \in X & \land \\
+            \langle g_Y, e_{Y}, n_{Y} \rangle \in Y & \land \\
+            e_{X} \leqslant e_{Y} < n_{X} & \land \\
+            \left( \{g_X\} \Join^\ast \{g_Y\} = \{g\} \right) &
+          \end{split}
+        \right)
+      \end{equation*}
+    \end{nestedsection}
+
+    %%% TRANSFORMATIONS
+
+    As is the case with relations in the relational algebra, the same stream may be defined in a variety of ways using the set of operators.
+    As each inter-operator stream is associated with a sequenced pattern, each definition of a stream may also be called a \emph{plan} for matching that stream's pattern.
+    Such equivalences between plans rely on a set of fundamental equivalences known as \emph{transformations}.
+    The most fundamental of the transformations of the relational algebra are traditionally classified as \emph{identity}, \emph{commutative}, \emph{associative} or \emph{distributive}, all of which apply equally well to the operators of CoDeR due to their similarity to those of SPARQL, which in turn have an equivalence to those of the relational algebra, as discussed by \citet{cyganiak05relationalSPARQL}.
+
+    In the remainder of this section: $a$ and $b$ are triple patterns; $g$ is a sequenced pattern; $p$ and $q$ are non-ground propositions using RIF-Core built-in predicates; $A$, $B$ and $C$ are CoDeR streams; and $X$, $Y$ and $Z$ are inter-operator streams.
+
+    Equations \ref{eqn:CoDeR: identity of graph-to-triples under succession} to \ref{eqn:CoDeR: identity of streams under union} enumerate the unconditional \emph{identity} transformations of the operators described in \refsec{CoDeR: graph-to-triples}.
+    \begin{align}
+      \labeleqn{CoDeR: identity of graph-to-triples under succession}
+      {}_g\pi_t\ {}_g\pi_t A &= {}_g\pi_t A \\
+      \labeleqn{CoDeR: identity of instance implication under succession}
+      \coloneq_g\ \coloneq_g X &=\ \coloneq_g X \\
+      \labeleqn{CoDeR: identity of streams under union}
+      A \cup A &= A
+    \end{align}
+    In addition, there are five identities that are conditional on the nature of the propositions $p$ and $q$ and the triple patterns $a$ and $b$.
+    The first of these are presented in \refeqn{CoDeR: identity of triple select under succession,CoDeR: identity of triple select under succession to instance implication}, where applying $\sigma_{a}$ to a stream associated with the singleton pattern ${\{b\ .\}}$ (as produced by $\sigma_{b}$ and $\coloneq_{\{b\ .\}}$) in the case where $b$ subsumes $a$ (or ${a \subseteq b}$) produces a stream that is identical to the original.
+    The same is true for \refeqn{CoDeR: identity of external select under succession}, only for propositions $p$ and $q$ where $q$ subsumes $p$ (or ${p \subset q}$ e.g.
+${p = greater{-}than(?x,5)}$ and ${q = greater{-}than(?x,4)}$).
+    Furthermore, in \refeqn{CoDeR: identity of graph-to-triples under succession to instance implication,CoDeR: identity of graph-to-triples under succession to triple-select}, applying ${{}_g\pi_t}$ to a stream associated with a singleton pattern (as produced by $\sigma_{triple}$ and $\coloneq_{\{a\ .\}}$) produces a stream that is identical to the original.
+    \begin{align}
+      \labeleqn{CoDeR: identity of triple select under succession}
+      \sigma_b\ \sigma_a X &= \sigma_a X && \text{where } a \subseteq b \\
+      \labeleqn{CoDeR: identity of triple select under succession to instance implication}
+      \sigma_b \coloneq_{\{a\ .\}} X &=\ \coloneq_{\{a\ .\}} X && \text{where } a \subseteq b \\
+      \labeleqn{CoDeR: identity of external select under succession}
+      \sigma_q\ \sigma_p X &= \sigma_p X && \text{where } p \subseteq q \\
+      \labeleqn{CoDeR: identity of graph-to-triples under succession to instance implication}
+      {}_g\pi_t \sigma_{a} X &=\ \sigma_{a} X \\
+      \labeleqn{CoDeR: identity of graph-to-triples under succession to instance implication}
+      {}_g\pi_t \coloneq_{\{a\ .\}} X &=\ \coloneq_{\{a\ .\}} X
+    \end{align}
+
+  \refeqn{CoDeR: commutativity of triple select,CoDeR: commutativity of external select,CoDeR: commutativity of union} enumerate the \emph{commutative} transformations of the operators described in \refsec{CoDeR: operator detail}.
+  These are the transformations that involve the re-ordering of unary operators of the same kind within a plan, as well as the swapping of inputs to binary operators, to produce another for the same sequenced pattern.
+  \begin{align}
+    \labeleqn{CoDeR: commutativity of triple select}
+    \sigma_a\ \sigma_b X &= \sigma_b\ \sigma_a X \\
+    \labeleqn{CoDeR: commutativity of external select}
+    \sigma_p\ \sigma_q X &= \sigma_q\ \sigma_p X \\
+    \labeleqn{CoDeR: commutativity of union}
+    A \cup B &= B \cup A
+  \end{align}
+  It should be noted that the SteM operator $\rstreamjoin$ is \emph{not} commutative, unlike the symmetric window-join.
+  However the definition of the window-join in terms of the commutative stream union and complementary SteMs in \refeqn{window-join-from-union-of-SteMs} does affirm the fact that the window-join \emph{is} commutative.
+
+  \refeqn{CoDeR: associativity of union,CoDeR: associativity of SteM} enumerate the associative transformations of the operators described in \refsec{CoDeR: operator detail}.
+  These are the transformations that involve the re-ordering of binary operators of the same kind within a plan to produce another for the same sequenced pattern.
+  \begin{align}
+    \labeleqn{CoDeR: associativity of union}
+    A \cup \left( B \cup C \right) &= \left( A \cup B \right) \cup C \\
+    \labeleqn{CoDeR: associativity of SteM}
+    X \rstreamjoin \left( Y \rstreamjoin Z \right) &= \left( X \rstreamjoin Y \right) \rstreamjoin Z
+  \end{align}
+
+  \refeqn{CoDeR: distributivity of selects,CoDeR: distributivity of triple select over union,CoDeR: distributivity of external select over union,CoDeR: distributivity of instance implication over union,CoDeR: distributivity of graph-to-triples over union} enumerate the distributive transformations of the operators described in \refsec{CoDeR: operator detail}.
+  These are the transformations that involve the re-ordering of operators of different kinds within a plan to produce another for the same sequenced pattern.
+  \begin{align}
+    \labeleqn{CoDeR: distributivity of selects}
+    \sigma_a\ \sigma_p X &= \sigma_p\ \sigma_a X \\
+    \labeleqn{CoDeR: distributivity of graph-to-triples over union}
+    {}_g\pi_t \left( A \cup B \right) &= \left( {}_g\pi_t A \right) \cup \left( {}_g\pi_t B \right) \\
+    \labeleqn{CoDeR: distributivity of triple select over union}
+    \sigma_a \left( X \cup Y \right) &= \left( \sigma_a X \right) \cup \left( \sigma_a Y \right) \\
+    \labeleqn{CoDeR: distributivity of external select over union}
+    \sigma_p \left( X \cup Y \right) &= \left( \sigma_p X \right) \cup \left( \sigma_p Y \right) \\
+    \labeleqn{CoDeR: distributivity of instance implication over union}
+    \coloneq_{g} \left( A \cup B \right) &= \left( \coloneq_{g} A \right) \cup \left( \coloneq_{g} B \right)
+  \end{align}
+  In addition, there are three conditional transformations, enumerated in \refeqn{CoDeR: distributivity of external select over SteM to left,CoDeR: distributivity of external select over SteM to right,CoDeR: distributivity of external select over SteM to both}, that distribute $\sigma_{external}$ over $\rstreamjoin$ and are analogous to those distributing selection $\sigma$ over join $\Join$ in the relational algebra.
+  Consider the non-ground propositions
+  \begin{align*}
+    p &= greater{-}than(?1, 5) & q &= less{-}than(?3, 5)
+  \end{align*}
+  and associations between inter-operator streams and sequenced patterns
+  \begin{align*}
+    X &\mapsto \{?1\ pred1\ ?2\ .\} & Y &\mapsto \{?2\ pred2\ ?3\ .\} & Z &\mapsto \{?1\ pred3\ obj\ .\}
+  \end{align*}
+  The conditional transformations are thus: 
+  % In the first case, non-ground proposition $p$ requires a set of variables ${\{?1\}}$, wholly subsumed by those present in the pattern associated with $X$ but not subsumed by the pattern associated with $Y$, to be bound.
+  % As such, $\sigma_{p}$ may be distributed over $\rstreamjoin$ to $X$ but not to $Y$ (due to \refeqn{CoDeR: empty result of external select over unbound variables,CoDeR: identity of empty stream under SteM right} making ${\left( \sigma_p X \right) \rstreamjoin \left( \sigma_p Y \right) = \varnothing}$).
+  \begin{equation}
+    \labeleqn{CoDeR: distributivity of external select over SteM to left}
+    \sigma_p \left( X \rstreamjoin Y \right) = \left( \sigma_p X \right) \rstreamjoin Y
+  \end{equation}
+  % In the second case, the non-ground proposition $q$ requires a set of variables ${\{?3\}}$, wholly subsumed in the pattern associated with $Y$ but not that associated with $X$, to be bound.
+  % As such, $\sigma_{q}$ may be distributed over $\rstreamjoin$ to $Y$ but not to $X$ (due to \refeqn{CoDeR: empty result of external select over unbound variables,CoDeR: identity of empty stream under SteM left}).
+  \begin{equation}
+    \labeleqn{CoDeR: distributivity of external select over SteM to right}
+    \sigma_q \left( X \rstreamjoin Y \right) = X \rstreamjoin \left( \sigma_q Y \right)
+  \end{equation}
+  % In the final case, the set of variables required by $p$ to be bound is wholly subsumed in the patterns associated with both $X$ and $Z$.
+  % As such, $\sigma_{p}$ may be distributed over $\rstreamjoin$ to both $X$ and $Z$.
+  \begin{equation}
+    \labeleqn{CoDeR: distributivity of external select over SteM to both}
+    \sigma_p \left( X \rstreamjoin Z \right) = \left( \sigma_p X \right) \rstreamjoin \left( \sigma_p Z \right)
+  \end{equation}
+  % It should be noted that, in the final case, $\sigma_p$ may alternatively be distributed to only one or the other of streams $X$ and $Y$ as in \refeqn{CoDeR: distributivity of external select over SteM to left,CoDeR: distributivity of external select over SteM to right};
+  % this is because filtering certain bindings for, for example, ${?1}$ from $X$ will prevent $\rstreamjoin$ from finding matches for those bindings for ${?1}$ if they should occur in $Z$, and vice versa.
+
+  Finally, the \emph{empty stream}, as per \refdef{CoDeR: empty stream}, may be created in a number of ways and creates a set of special identities, the latter enumerated in \refeqn{CoDeR: identity of empty stream under graph-to-triples,CoDeR: identity of empty stream under triple select,CoDeR: identity of empty stream under external select,CoDeR: identity of empty stream under instance implication,CoDeR: identity of empty stream under SteM right,CoDeR: identity of empty stream under SteM left,CoDeR: identity of stream under union with empty stream}.
+  \begin{align}
+    \labeleqn{CoDeR: identity of empty stream under graph-to-triples}
+    {}_g\pi_t \varnothing &= \varnothing \\
+    \labeleqn{CoDeR: identity of empty stream under triple select}
+    \sigma_a \varnothing &= \varnothing \\
+    \labeleqn{CoDeR: identity of empty stream under external select}
+    \sigma_p \varnothing &= \varnothing \\
+    \labeleqn{CoDeR: identity of empty stream under instance implication}
+    \coloneq_g \varnothing &= \varnothing \\
+    \labeleqn{CoDeR: identity of empty stream under SteM right}
+    X \rstreamjoin \varnothing &= \varnothing \\
+    \labeleqn{CoDeR: identity of empty stream under SteM left}
+    \varnothing \rstreamjoin X &= \varnothing \\
+    \labeleqn{CoDeR: identity of stream under union with empty stream}
+    \varnothing \cup A &= A
+  \end{align}
+  The ways in which the empty stream is created all stem from the possibility that a successor $\sigma_\ast$ operator might filter the stream that it receives based on a pattern that will never match any graph instance produced by the operator that it succeeds.
+  \refeqn{CoDeR: empty result of triple select under succession,CoDeR: empty result of triple select under succession from instance implication,CoDeR: empty result of external select under succession,CoDeR: empty result of external select over unbound variables} provide an example of each possible case where this occurs.
+  \begin{align}
+    \labeleqn{CoDeR: empty result of triple select under succession}
+    \text{Conflicting patterns:} && \sigma_{?1\ pred1\ ?2}\ \sigma_{?1\ pred2\ ?2} X &= \varnothing \\
+    \labeleqn{CoDeR: empty result of triple select under succession from instance implication}
+    \text{Conflicting patterns:} && \sigma_{?1\ pred1\ ?2} \coloneq_{\{?1\ pred2\ ?2\ .\}} X &= \varnothing \\
+    \labeleqn{CoDeR: empty result of external select under succession}
+    \text{Mutually exclusive propositions:} && \sigma_{greater{-}than(?1,5)}\ \sigma_{less{-}than(?1,5)} X &= \varnothing \\
+    \labeleqn{CoDeR: empty result of external select over unbound variables}
+    \text{Unbound variables in proposition:} && \sigma_{greater{-}than(?3,5)}\ \sigma_{?1\ pred1\ ?2} X &= \varnothing
+  \end{align}
+  \refeqn{CoDeR: empty result of external select over unbound variables} expresses the extension of rule safety in RIF-Core relating to the required binding elsewhere in a pattern of any variables that are necessary arguments to a built-in predicate or function in that pattern.
+  The example given assumes that both arguments of the concisely-named predicate ${greater{-}than()}$ must be bound, as they must in the actual built-in predicate of RIF-Core that it emulates:
+  \[
+    {<}http://www.w3.org/2007/rif{-}builtin{-}predicate\#numeric{-}greater{-}than{>}
+  \]
+  However, there are a few cases where specific arguments of specific predicates need not be bound \citep{w3crifcoreBuiltins}; the transformation in \refeqn{CoDeR: empty result of external select over unbound variables} does \emph{not} apply in the cases where unbound variables appear as those specific arguments.
+\end{nestedsection}
+
+\begin{nestedsection}{CoDeR Streams Notation}{model: stream notation}
+  In compliment to the set of operators of the CoDeR algebra, we propose an extended notation that allows the accurate description a system built according to the CoDeR model, taking into account the iterative nature of inference and the uniqueness of streams within the model.
+  This notation is needed to express two of the fundamental characteristics of the CoDeR data model, those being the intention of performing inference and the uniqueness of streams within the data model.
+
+  Where many stream processing tasks (including continuous querying) are one-pass, the process of reasoning and inference is iterative/recursive.
+  This theoretically necessitates the union of the output stream with the input stream ``upstream'' of the remaining operators of any processing plan in the form of a \emph{working stream}, as per \refdef{CoDeR: working stream} and as shown in \reffig{CoDeR: re-entrant stream}.
+  \begin{definition}[Working Stream]\labeldef{CoDeR: working stream}
+    The \emph{working stream} of a CoDeR-based system is a CoDeR stream that is the union of all the system's input streams and its output stream.
+    It facilitates the \emph{iterative} application of the rules of the system to the streamed data, in the same manner as the \emph{working memory} of the Rete algorithm \citep{forgy79}.
+  \end{definition}
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.9\textwidth]{CoDeR-streams}
+    \caption[Streams in CoDeR System]{An abstract CoDeR-based system, demonstrating the purview of each class of stream: Input and Output streams to receive data from and publish data to external sources; Inter-operator streams facilitate operator nesting/pipelines; and the re-entrant Working Stream facilitates iterative processing.}
+    \labelfig{CoDeR: re-entrant stream}
+  \end{figure}
+  This feedback of the output stream into the \emph{working stream} may cause cause the definition of the output stream of a system to be circular, thus impossible to express purely in terms of the operators of the algebra described in \refsec{CoDeR: algebra}.
+  This problem of inexpressibility also applies to the sharing of streams between potentially many sinks and sources:
+  when relying solely on the nested-operator notation of the relational algebra, the same stream must be defined multiple times throughout a plan if it occurs in multiple sub-plans, though in reality that stream need only occur once in the modelled system.
+
+  As such, the complete expression of the processing required for inference over WC-RIF-Core programs necessitates that streams may be named and that, once named, those streams may be referenced by those names within the definitions of other streams.
+  Further, it must be possible to provide multiple contributing definitions for each named stream, this being an expression of \emph{stream union} derived from the manner in which the same predicate may have multiple disjunctive intensions within a Datalog program.
+
+  Finally, it must be possible to project the variables in a sequenced pattern associated with an inter-operator stream into a variant of that pattern, in order for that stream to be referenced within other plans that may use conflicting variable names.
+  However, this ``projection'' of variable names is not an operator of the algebra, those being related to pattern matching, which is variable-name agnostic\footnote{
+    By this I mean that a pattern is matched based on the existence of components that satisfy certain comparisons against given constant values and against specific other components.
+    How each of those components is labelled is irrelevant to the process, as long as the nature of those comparisons is maintained.
+    This is distinct from interpretations of the relational algebra such as SQL, in which the naming of the attributes adds semantic value to the data.}.
+  Instead, it is simply a requirement of the expression of plans within a system, met in the convention of naming inter-operator streams using variants of their pattern, i.e. ${S_{pattern}}$.
+  When naming or referencing a stream, the variables may therefore be renamed as appropriate for the context in which the stream is referenced.
+
+  These requirements are met in CoDeR using the notation shown in \refeqn{CoDeR: stream assignment}, an example of which for expressing the plan for some $\Pi^{C,r}$ is given in \refeqn{CoDeR: stream assignment example}.
+  \begin{align}
+    [stream\ name] &= [stream\ definition] \labeleqn{CoDeR: stream assignment} \\
+    \begin{split}
+      working\_stream &= {<}\text{http://example.org/test\_stream}{>}\\
+      S_{\{?s\ ?p\ ?o\ .\}} &= {}_g\pi_t\ working\_stream \\
+      S_{\{?s\ inside\ ?o\ .\}} &= \sigma_{?s\ inside\ ?o}\ S_{\{?s\ ?p\ ?o\ .\}} \\
+      S_{\{?1\ inside\ ?2\ .\ ?2\ inside\ ?3\ .\}} &=\ S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \\
+      S_{\{?1\ inside\ ?2\ .\}} &=\ \coloneq_{\{?1\ inside\ ?2\ .\}}\ S_{\{?1\ inside\ ?3\ .\ ?3\ inside\ ?2\ .\}} \\
+      & \vdots \\
+      inference\_stream &= S_{\{?s\ inside\ ?p\ .\}} \cup \dots \cup \dots \\
+      working\_stream &= inference\_stream
+    \end{split} \labeleqn{CoDeR: stream assignment example}
+  \end{align}
+  The first line of \refeqn{CoDeR: stream assignment example} demonstrates the definition of the plan's input stream by its IRI, and its being named ${working\_stream}$.
+  The third through fifth lines demonstrate the dual definition of a single inter-operator named stream using variant graph patterns, in this case expressing the transitive property of the example predicate ${inside}$ via the definition of of an inter-operator named stream (using \refeqn{window-join-from-union-of-SteMs,CoDeR: identity of streams under union}).
+  The penultimate line demonstrates the definition of the plan's inference stream, this being the \emph{stream union} of all streams produced by processing the separate rules of $R$.
+  The last line asserts that the inference stream contributes through \emph{stream union} with the input stream to the previously named ${working\_stream}$, thereby making the plan iterative.
+
+  In the plan presented in \refeqn{CoDeR: stream assignment example}, the stream defined on the fourth line is defined once and used once.
+  As such, the stream definition on the fifth line may be rewritten, replacing the stream reference $S_{\{?1\ inside\ ?3\ .\ ?3\ inside\ ?2\ .\}}$ with its definition ${\left( S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \right)}$ from the fourth line, renaming variables as appropriate.
+  This results in the plan presented in \refeqn{CoDeR: stream assignment example rewritten}.
+  \begin{equation}
+    \begin{split}
+      working\_stream &= {<}\text{http://example.org/test\_stream}{>}\\
+      S_{\{?s\ ?p\ ?o\ .\}} &= {}_g\pi_t\ working\_stream \\
+      S_{\{?s\ inside\ ?o\ .\}} &= \sigma_{?s\ inside\ ?o}\ S_{\{?s\ ?p\ ?o\ .\}} \\
+      S_{\{?1\ inside\ ?3\ .\}} &=\ \coloneq_{\{?1\ inside\ ?3\ .\}}\ \left( S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \right) \\
+      & \vdots \\
+      output\_stream &= S_{\{?s\ inside\ ?p\ .\}} \cup \dots \cup \dots \\
+      working\_stream &= output\_stream
+    \end{split} \labeleqn{CoDeR: stream assignment example rewritten}
+  \end{equation}
+  This method of rewriting may also be used to isolate a part of the plan as a single algebraic definition, to which the transformations presented in \refsec{CoDeR: transformations} may be applied.
+  This rewriting process may then be carried out in reverse, assigning names to streams produced by sub-definitions, then replacing those sub-definitions with references to the corresponding named streams in the original definition.
+
+  The example plans presented in \refeqn{CoDeR: stream assignment example,CoDeR: stream assignment example rewritten} also demonstrate that iteration within a plan follows naturally from the uniqueness of inter-operator streams by their associated sequenced pattern, discussed in \refsec{CoDeR: data}:
+  in each case, the second definition of the stream $S_{\{?1\ inside\ ?3\ .\}}$ is recursive.
+  The generalised case is of a rule whose head is a singleton graph pattern ${\{h\ .\}}$ that is a variant of the pattern of some triple-select operator $\sigma_{h}$;
+  both the triple-select and instance implication contribute to the same stream, resulting in an iterative feedback of entailed graph instances into the plan that bypasses the \emph{working stream}, as shown in \reffig{CoDeR: unique stream feedback}.
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.9\textwidth]{CoDeR-unique-stream-feedback}
+    \caption[Iteration by Uniqueness of Streams in CoDeR]{An abstract CoDeR-based system, demonstrating the natural iterative feedback caused by the uniqueness of streams in the model.}
+    \labelfig{CoDeR: unique stream feedback}
+  \end{figure}
+  To support the correctness of such behaviour, it can be shown that any such iterative feedback may be derived algebraically from the definition and reference of the working stream in any plan that expresses it precisely as in \refdef{CoDeR: working stream}.
+  %
+  % In order to derive this proof, the selection of triple instances matching the pattern $h$ from the \emph{working stream}, i.e.
+${\sigma_{h}\ {}_{g}\pi_{t}\ working\_stream}$, must be rewritten (as already described) until all instance implication operators of the plan are present in a single algebraic expression.
+  % This rewriting is demonstrated in \refeqn{CoDeR: named stream rewriting}, where:
+  % \begin{itemize}
+  %   \item
+  %     $S_{g}$ is any inter-operator stream the precise nature of whose associated sequenced pattern is not relevant to the proof;
+  %   \item
+  %     $f$ is any sequenced pattern that is strictly incompatible with $h$ as per the transformation expressed in \refeqn{CoDeR: empty result of triple select under succession from instance implication};
+  %   \item
+  %     `$\dots$' is the remainder of the inter-operator streams that contribute to the CoDeR output stream, and whose associated sequenced pattern is neither $h$ nor incompatible with it.
+  % \end{itemize}
+  % \begin{equation} \labeleqn{CoDeR: named stream rewriting}
+  %   \begin{split}
+  %     \sigma_{h}\ {}_g\pi_t\ working\_stream &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup output\_stream \right) \\
+  %     &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \left( S_{\{h\ .\}} \cup \dots \cup S_{\{f\ .\}} \right) \right) \\
+  %     &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \left( \left( \coloneq_{\{h\ .\}} S_{g} \right) \cup \dots \cup \left( \coloneq_{\{f\ .\}}\ S_{g} \right) \right) \right)
+  %   \end{split}
+  % \end{equation}
+  % The proof may then be derived by showing that the filtering of the entire working stream by $h$ is equal to the union of the entailed stream matching pattern ${\{h\ .\}}$ with the remainder of the working stream filtered by $h$, less those entailed streams that match patterns $f$ that are incompatible with $h$.
+  % This result is presented in \refeqn{CoDeR: transformations for closer iteration}, and may be proven by applying the following transformations to the right hand side:
+  % \ref{eqn:CoDeR: associativity of union} and \ref{eqn:CoDeR: commutativity of union}%;
+  % to re-arrange the union;
+  % \ref{eqn:CoDeR: distributivity of graph-to-triples over union} and \ref{eqn:CoDeR: distributivity of triple select over union}%;
+  % to distribute the selection by $h$;
+  % \ref{eqn:CoDeR: identity of graph-to-triples under succession to instance implication}%;
+  % to simplify each selection;
+  % \ref{eqn:CoDeR: identity of triple select under succession to instance implication}%;
+  % to eliminate selection over the entailed stream matching ${\{h\ .\}}$;
+  % and \ref{eqn:CoDeR: empty result of triple select under succession from instance implication} and \ref{eqn:CoDeR: identity of stream under union with empty stream}%.
+  % to eliminate from consideration the entailed streams matching ${\{f\ .\}}$.
+  % \begin{equation} \labeleqn{CoDeR: transformations for closer iteration}
+  %   \left( \begin{split}
+  %     \left( \coloneq_{\{h\ .\}}\ S_{g} \right) \cup \\
+  %     \left( \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \dots \right) \right)
+  %   \end{split} \right)
+  %   &
+  %     =\ \sigma_{h}\ {}_g\pi_t\ \left( \begin{split}
+  %       input\_stream \cup \\
+  %       \left( \left( \coloneq_{\{h\ .\}} S_{g} \right) \cup \dots \cup \left( \coloneq_{\{f\ .\}} S_{g} \right) \right)
+  %     \end{split} \right)
+  % \end{equation}
+  This reflects the use of \emph{variant tabling} for optimising Prolog programs as discussed by \citet{DS4SbsmptvTbls}.
+\end{nestedsection}
 
@@
   Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
-    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the contents of the semantic sliding window of range $r$ over the system input streams $S$ defined in a given ${\Pi^{C,r} = \langle R,F^0,S,r\rangle}$.
-    It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of Continuous Datalog.
+  as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n) \} \rangle}$.
+  It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of WC-RIF-Core.
   It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
-    i.e. such a CoDeR-based system is guaranteed to produce the complete set of results according to the semantics of Continuous Datalog.
-  \end{nestedsection}
+  i.e. such a CoDeR-based system is guaranteed to produce the complete stream of results according to the semantics of WC-RIF-Core.
 \end{nestedsection}

### 2015-12-01 article.tex
Complete rewrite of most of the introduction to neaten and bring more to the point.  May include HORRORS

@@
   \begin{abstract}
-    In this paper, we introduce Continuous Datalog, an extension of
+    In this paper, we introduce Windowed Continuous Datalog, an extension of
     conventional Datalog that adds the concept of streams of
@@
     R4 stream reasoner, which implements the CoDeR model as a Rete-based
-    dataflow network and provide a brief comparative evaluation of R4
+    dataflow network and provide a comparative evaluation of R4
     against ETALIS and Sparkwave.
@@
   With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
-  In the most recent literature, this has been characterised as a layer of higher complexity processing on top of Complex Event Processing (CEP) \citep{mileo15webSR}, deeming it suitable only for application to low-throughput macro-event streams.
-  However, Semantic Web reasoning languages provide many varying degrees of expressivity corresponding to to varying levels of complexity.
-  Furthermore, in many cases static ontologies are applicable to streamed data (such as the description and classification of readings from sensor networks, for example), and therefore do not require the features unique to CEP (and temporal reasoning \citep{OrgunWadge92,Tuzhilin93}), rather requiring a concept of data becoming less relevant over time.
+
+  In the most recent literature, stream reasoning has been characterised as a layer of higher complexity processing on top of Complex Event Processing (CEP) \citep{mileo15webSR}, deeming it suitable only for application to low-throughput macro-event streams.
+  However, this expression of reasoning on top of CEP introduces semantics of temporal reasoning \citep{OrgunWadge92,Tuzhilin93}, whereas in many cases \emph{temporally-agnostic} ontologies, which do \emph{not} make use of temporal reasoning, are applicable to \emph{temporally-sensitive} streamed data, which is relevant for an inherently limited amount of time.
+  For example, the descriptions of sensors in a sensor network and the various types of reading they produce are often temporally agnostic, while the readings themselves are accurate (and so relevant) for only a short time after they are produced.
+  Furthermore, the inferences gleaned from applying the former to the latter are equally temporally sensitive, and so may most accurately be represented as a stream themselves.
   A prime expression of low-complexity reasoning that is capable of such tasks in a static context is the rule language profile of OWL 2 (OWL 2 RL) \citep{w3cowl2profiles}.
   The continuous application of reasoning of this nature to streamed RDF data may obviously be characterised as `stream reasoning', but occupies the same general complexity bracket as stream querying in the hierarchy presented by \citep{mileo15webSR}, making it applicable to high-throughput streams.
-  It is this low-complexity continuous reasoning that is the focus of this paper.
-
-  The value of such processing is the same as that of reasoning over comparatively static data, that being that the division of knowledge into particular knowledge and general knowledge allows for the expression of a greater volume of information in a lesser volume of data.
-  In addition, this division allows the application of a single domain of general knowledge to a variety of distinct sets of particular knowledge, and the application of many distinct domains of general knowledge to a single set of particular knowledge.
-  It should be noted that this division of general and particular knowledge in the expression of a reasoning problem depends on the generalisation of that problem:
-  in Description Logics, it may be considered that the TBox of a knowledge base describes general knowledge and the ABox describes particular knowledge;
-  however, the fundamental rules by which entailments are derived from knowledge bases by Description Logics constitute a more general knowledge than that of the TBox of some specific knowledge base, by which the entirety of a knowledge base may be considered particular knowledge.
-  In the latter case, a distinction may still be made between a TBox and ABox in the context of streaming, with the former being expressed in static data and the latter being streamed or vice versa, though this is dependent on the rate at which the TBox and ABox are expected to change.
-  In the existing stream processing literature, the vast majority of current use cases consider the application of some \emph{specific} static domain of general knowledge (e.g.
-a specific TBox) to some stream of particular knowledge;
-  thus, with stream reasoning the volume of streamed data may be dramatically reduced by extracting the static domain of general knowledge that underpins the data and providing that domain as a static ontology, then streaming only the minimal particular knowledge that would be necessary to deduce the full data set.
+  It is this form of stream-based low-complexity reasoning that is the focus of this paper.
+
+  Most stream processing makes use of the concept of \emph{sliding windows} over streams.
+  In stream querying, these windows are \emph{global}, in so much as streamed data is stored in windows, then that windowed data \emph{as a whole} is queried \citep{C-SPARQL,EP-SPARQL,CQELS}.
+  This may be seen as each datum in a sliding windows being considered to be \emph{true} by a temporally-agnostic continuous query, for the interval from its arrival for the duration of the window.
+  It is this notion of windowing that best expresses the form of stream processing we have stated as our focus.
+
+  In contrast, in the recent temporal stream reasoning/CEP literature \citep{LARS}, windows are \emph{localised within} the ``general knowledge'' (e.g.
+rules in a logic program, the TBox in an ontology), meaning that individual rules/TBox axioms are applied to sliding windows over both streamed data \emph{and} temporally sensitive inferences resulting from other rules/TBox axioms;
+  the data is considered to \emph{have been true} at the moment it appeared on a stream/was inferred, and sliding windows retain prior (and future) data according to temporal constraints in the reasoning.
+  Localised sliding windows may be used to express a globally windowed \emph{query} (as a single non-recursive rule), and a set of such queries might therefore be used to (partially) express CEP \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb,LARS}.
+
+  Localised sliding windows are inherently incapable of expressing temporally-agnostic reasoning over globally windowed data, due to the maintenance of prior inferences within localised sliding windows.
+  Inferences from globally windowed streamed data will become untrue as soon as they are no longer implied by the contents of the global window, whereas those inferences would be retained in a localised window for the same duration as any streamed data also retained.
+  They may, therefore, be retained \emph{after} the global window would have ceased to imply them, and contribute to further inferences that may, in turn, ``outlive'' them.
+  As such, stream reasoning semantics that rely on localised sliding windows (such as that which underpins \citep{LARS}) are unable to express the stream-based reasoning that is the focus of this paper.
+
+  On the other hand, much research has been conducted into the notion of \emph{continuous reasoning}, which has varied from truth-maintenance in ontologies \citep{stream-truth-maintenance}, to the efficient update/deletion of facts \citep{dred,inc-materialisation,ahmeti14rdfsUpdate}, to the addition of time-decaying semantics to Answer Set Programming (ASP) \citep{gebser12streamASP}.
+  While these investigate the challenge of changing data in temporally-agnostic (as well as temporal) knowledge-bases, none explicitly provide a semantics in which the entailment of the knowledge base is itself a stream.
+  While it is possible to produce streams of data from the changes in a knowledge base, the semantics of such streams have not been explicitly defined or, therefore, thoroughly investigated.
 
   % motivating scenario
-  Take, for example, the case of extensive sensor networks with many and varied sensors, such as those served by the SemsorGrid4Env project\footnote{\url{http://www.semsorgrid4env.eu/}}.
-  Any stand-alone semantic representation of this data may contain classification information for each sensor, according to what they measure, as well as how frequently and how accurately, as well as other factors.
-  It may also contain explicit data concerning events that are reported by no one sensor, and are instead extrapolated from the base sensor data.
-  In both of these cases, the classification criteria and extrapolation process for macro-events may be expressed as general knowledge that is applicable to all sensor readings.
-  As such, a consumer of the streams produced using SemsorGrid4Env could download the appropriate ontologies once, then use them to reason over a reduced set of streamed data containing only the base readings from the sensor network, thus obtaining the classification and macro-event information desired without having to stream it explicitly.
-  Furthermore, sensor readings are often superseded by later readings of the same type from the same sensor, and so older readings may be discarded as time progresses.
-
-  % *brief* summary of state of the art (to be expanded in related work)
-  Recent research has approached the task of stream reasoning from two different directions.
-  %   RDF stream processing and reasoning
-  The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the \emph{iterative/recursive execution} of queries over the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}, in which each query expresses a rule in a logic program.
-  This reflects precisely the approach to reasoning using Datalog in deductive databases, in which queries may be executed on intensional relations whose definitions must be used in turn to query the database, recursing in this manner until base only extensional relations are queried.
-  Meanwhile, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{dred,inc-materialisation,stream-truth-maintenance}.
-  %%%FIXME
-  NEED SOMETHING ON TIME-DECAY MODEL
-
-  %   issues with existing art: no explicit semantics for reasoning
-  While the iterative execution of queries as rules is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, as practised in deductive databases, the query-by-query windowing semantics of most stream-querying systems do not translate to a coherent semantics for multi-rule stream reasoning of the kind investigated in this paper.
-  This is because the windowing semantics proposed for stream querying involve applying the entirety of a ``problem'' (be it a single query or a unified body of general knowledge, as we propose) to a bounded subset of the data in a stream (a window).
-  This semantics is henceforth referred to as \emph{global windowing}.
-  The iterative execution of individually windowed queries, however, both introduces and \emph{fails to properly consider} the semantics of Complex Event Processing, leading to inconsistencies between the results of identical rule sets expressed in the cited systems \citep{LARS}.
-
-  Truth-maintenance algorithms for continuously changing knowledge bases and logic programs, on the other hand, do support coherent semantics for continuous reasoning.
-  However, these semantics cast the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams.
-  Without research to extend these semantics to entail concise streams that maintain in their own right the coherence of the existing semantics, these algorithms preclude the further processing of entailments in a stream-processing manner.
-
-  As such, until recently, stream reasoning either lacked a coherent semantics for continuous truth and entailment, or was necessarily a terminus rather than intermediate process in any streamed data pipeline \citep{valle09streamReasoningSummary}.
-  However, the work by \citet{LARS} has produced the LARS framework, by which the reasoning semantics of stream processing systems may be expressed and formally analysed.
-  LARS extends the semantics of Answer Set Programming (itself an extension of Datalog semantics with features such as Negation as Failure and aggregation) with the notion of streamed data with a point-in-time semantics.
-  Paired with this is the notion of a logical windowing operator that supports the expression of windows as part of the general knowledge of a reasoning problem, as is necessary to express reasoning as performed by the iterative execution of continuous queries in the manner already discussed.
-  Though this form of windowing also allows the expression of temporally-aware rules in the manner of Complex Event Processing, it does \emph{not} align with or support the expression of \emph{global windowing} derived from stream \emph{querying}.
-  Furthermore, the expression of granular windowing such as that expressible in LARS has not been defined in existing Semantic Web reasoning languages such as RIF \citep{w3crif} and OWL \citep{w3cowl2}.
-  Finally, the complexity of entailment over programs as defined in LARS makes it unsuitable as a semantics on which to \emph{base} the expression of reasoning in real-time:
-  the complexity of entailment in LARS is PSPACE, or co-NP given restrictions on window-nesting, which is generally considered too high a complexity for near-real-time entailment over high-throughput streams \citep{mileo15webSR}.
+  % Take, for example, the case of extensive sensor networks with many and varied sensors, such as those served by the SemsorGrid4Env project\footnote{\url{http://www.semsorgrid4env.eu/}}.
+  % Any stand-alone semantic representation of this data may contain classification information for each sensor, according to what they measure, as well as how frequently and how accurately, as well as other factors.
+  % It may also contain explicit data concerning events that are reported by no one sensor, and are instead extrapolated from the base sensor data.
+  % In both of these cases, the classification criteria and extrapolation process for macro-events may be expressed as general knowledge that is applicable to all sensor readings.
+  % As such, a consumer of the streams produced using SemsorGrid4Env could download the appropriate ontologies once, then use them to reason over a reduced set of streamed data containing only the base readings from the sensor network, thus obtaining the classification and macro-event information desired without having to stream it explicitly.
+  % Furthermore, sensor readings are often superseded by later readings of the same type from the same sensor, and so older readings may be discarded as time progresses.
+
+  % % *brief* summary of state of the art (to be expanded in related work)
+  % Recent research has approached the task of stream reasoning from two different directions, though each is flawed with respect to .
+  % %   RDF stream processing and reasoning
+  % The RDF stream processing community has proposed extensions to their continuous semantic query languages to support the \emph{iterative/recursive execution} of queries over the results of other queries \citep{C-SPARQL,EP-SPARQL,walavalkar08streamingkb}, in which each query expresses a rule in a logic program.
+  % This reflects precisely the approach to reasoning using Datalog in deductive databases, in which queries may be executed on intensional relations whose definitions must be used in turn to query the database, recursing in this manner until base only extensional relations are queried.
+  % Meanwhile, the reasoning community has explored the concept of continuously reasoning over streamed data in a variety of forms, primarily through a range of truth maintenance algorithms \citep{inc-materialisation,stream-truth-maintenance}.
+  % %%%FIXME
+  % NEED SOMETHING ON TIME-DECAY MODEL
+
+  % %   issues with existing art: no explicit semantics for reasoning
+  % While the iterative execution of queries as rules is sufficient for performing reasoning to the expressivity of Datalog when applying one-shot queries to static data, as practised in deductive databases, the query-by-query windowing semantics of most stream-querying systems do not translate to a semantics for multi-rule stream reasoning of the kind investigated in this paper.
+  % This is because the windowing semantics proposed for stream querying involve applying the entirety of a ``problem'' (be it a single query or a unified body of general knowledge, as we propose) to a bounded subset of the data in a stream (a window).
+  % This semantics is henceforth referred to as \emph{global windowing}.
+  % The iterative execution of individually windowed queries, however, both introduces and \emph{fails to properly consider} the semantics of Complex Event Processing, leading to inconsistencies between the results of identical rule sets expressed in the cited systems \citep{LARS}.
+
+  % Truth-maintenance algorithms for continuously changing knowledge bases and logic programs, on the other hand, do support coherent semantics for continuous reasoning.
+  % However, these semantics cast the problem as that of maintaining a knowledge base and does not support the expression of entailments in streams.
+  % Without research to extend these semantics to entail concise streams that maintain in their own right the coherence of the existing semantics, these algorithms preclude the further processing of entailments in a stream-processing manner.
+
+  % As such, until recently, stream reasoning either lacked a coherent semantics for continuous truth and entailment, or was necessarily a terminus rather than intermediate process in any streamed data pipeline \citep{valle09streamReasoningSummary}.
+  % However, the work by \citet{LARS} has produced the LARS framework, by which the reasoning semantics of existing stream processing systems may be expressed and formally contrasted.
+  % LARS extends the semantics of Answer Set Programming (itself an extension of Datalog semantics with features such as Negation as Failure and aggregation) with the notion of streamed data with a point-in-time semantics.
+  % Paired with this is the notion of a logical windowing operator that supports the expression of windows as part of the general knowledge of a reasoning problem, as is necessary to express reasoning as performed by the iterative execution of continuous queries in the manner already discussed.
+  % Though this form of windowing also allows the expression of temporally-aware rules in the manner of Complex Event Processing, it does \emph{not} align with or support the expression of \emph{global windowing} derived from stream \emph{querying}.
+  % Furthermore, a standard for the expression of \emph{local} windowing such as that expressible in LARS has not been defined in existing Semantic Web reasoning languages such as RIF \citep{w3crif} and OWL \citep{w3cowl2}.
+  % Finally, the complexity of entailment over programs as defined in LARS makes it unsuitable as a semantics on which to \emph{base} the expression of reasoning in real-time:
+  % the complexity of entailment in LARS is PSPACE, or co-NP given restrictions on window-nesting, which is generally considered too high a complexity for near-real-time entailment over high-throughput streams \citep{mileo15webSR}.
 
   % Rete as non-streaming implementationexample (application to dataflow networks)
-  Progressing from reasoning semantics to data processing, Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
+  Progressing from stream-based reasoning semantics to data processing, Rete networks \citep{forgy79} are an example of production systems (a form of rule-based reasoning system) that are particularly suitable for extension with streaming semantics, being modelled as dataflow networks composed of incremental operators.
   These systems send data from operator to operator, with `memories' compiled for two-pass operators in which to retain the specific data received, in much the same way as sliding windows are used in stream processing.
@@
   However, \citet{forgy79} did not incorporate the semantics of stream processing into the model underlying the Rete algorithm.
-  In addition, more recent data-flow-based models either don't consider ``reasoning'' more expressive than RDFS \citep{sparkwave}, or are based on rule/query-level windowing \citep{C-SPARQL,walavalkar08streamingkb}, with problems as already discussed.
-  Despite this, data-flow-based models and the vast body of associated literature present a firm foundation for the processing of reasoning tasks in a flexible and scalable manner.
+  In addition, more recent data-flow-based models either don't consider ``reasoning'' more expressive than RDFS \citep{sparkwave}, or are based on rule/query-level windowing \citep{C-SPARQL,walavalkar08streamingkb}, which has already been shown to be incompatible with globally-windowed stream-based reasoning.
+  Despite this, data-flow-based models and the vast body of associated literature present a firm foundation for the processing of stream-based reasoning tasks in a flexible and scalable manner.
 
   % overview
-  In this paper we present, first, a simple semantics extended from those of positive Datalog that provide a coherent expression of entailments as streams, which we dub Windowed Continuous Datalog.
-  The scope of this semantics is to be an expression of inference that extends the static form currently described by Semantic Web reasoning languages with global reasoning semantics, thereby incorporating potentially high-throughput streams of semantic data but resulting in a low enough computational time-complexity that entailments be feasibly computed continuously in real-time.
+  In this paper we present, first, a simple semantics that extends those of positive Datalog to both incorporate streamed data and provide an expression of entailments over time as streams.
+  The scope of this semantics, which we dub Windowed Continuous Datalog, is to be an expression of inference that extends the static form currently described by Semantic Web reasoning languages with global windowing semantics, thereby incorporating potentially high-throughput streams of semantic data while retaining a low enough computational time-complexity that streams of entailments may feasibly be computed continuously in real-time.
   As entailment in positive Datalog has polynomial-time complexity and is the equivalent of entailment in RIF-Core \citep{w3crifcore}, which in turn may be used to express the entailment rules of the rule language profile of OWL 2 \citep{w3cowl2profiles}, the continuous semantics of Windowed Continuous Datalog may be used to apply OWL 2 ontologies to streams of RDF data.
-  Next, we present the abstract processing model CoDeR that provides for the expression of data-flow processing plans that satisfy the semantics of Windowed Continuous Datalog.
-  We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically coherent stream reasoning over sets of rules expressed in RIF-Core.
-  Due to the lack of benchmarks for stream \emph{reasoning} use cases, we use single-rule rule-sets based on the individual queries of two stream-\emph{querying} benchmarks to contrast the performance of R4 against that of contemporary systems Etalis and Sparkwave.
-  Finally, we discuss differences between the three systems, as well as improvements that could be made to the efficiency of R4 without breaking from the processing model of CoDeR and, therefore, the semantics of Windowed Continuous Datalog.
+
+  Next, we present the abstract processing model CoDeR that provides for the expression of data-flow processing plans that satisfy the semantics of Windowed Continuous RIF-Core.
+  We also present the system R4 that utilises the Rete pattern-matching algorithm to generate processing plans expressed in the CoDeR model to achieve semantically sound stream-based reasoning over sets of rules expressed in RIF-Core.
+  Due to the lack of benchmarks for expressive stream-based \emph{reasoning} use cases, we use the modified big-data/RDFS evaluation described by \citet{sparkwave} to contrast the performance of R4 against that of contemporary systems Etalis \citep{EP-SPARQL} and Sparkwave \citep{sparkwave}.
+  From this, we assert that R4 is empirically faster than these contemporaries, despite having a more expressive underlying semantics.
+  Finally, we briefly discuss possible extensions to R4 to further improve its quality of service characteristics.
 \end{nestedsection}

### 2015-12-01 article.tex
many minor formatting edits and points of clarification. also slightly revised wording of conclusion to match introduction and new conclusions

@@
 
+It should be noted that a fact instance only indicates that the
+fact is recognised as true at that instant.  This does \emph{not}
+imply that the fact is true for \emph{only} that instant.  As such,
+a fact instance denotes an interval of indefinite length from a
+definite start for which the fact is to be considered true.
+
 In \refdef{continuous datalog: datalog program}, we took a Datalog
@@
 appearance thereon), the states of the former are determined by a function of
-the stream itself.  This is more similar to the aggregation functions of SQL,
+the stream itself.  This is more similar to the aggregation functions of SQL and ASP,
 such as $count()$, $max()$, etc., which are not part of the semantics of
@@
 reduced to the Datalog program
-$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_j,r_j}_i \rangle$.
+$\Pi = \langle R, strip(F^0 \cup \mathop{\cup}\limits_{x=1}^n W^{S_x,r_x}_i \rangle$.
 
@@
 $E^T_{i}$ of instances entailed at any given instant $i$ by any single-window program
-$\Pi^{C,r} = \langle R, F^0, \{ \langle S, r \rangle \} \rangle$ may be reconstructed from
+$\Pi^{C,r} = \langle R, F^0, \{ W^{S, r} \} \rangle$ may be reconstructed from
 windows of sufficient range over its entailed streams $P$ and $N$ at
@@
   \labeleqn{continuous datalog: successive WC-Datalog program}
-    \Pi^{C,x} = \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{S^\prime,x} \} \cup \mathcal{W}^{P \setminus N} \rangle \models \langle E^{0\prime}, P^\prime, N^\prime \rangle
+    \begin{split}
+      \Pi^{C,x} &= \langle R^\prime, F^{0\prime} \cup E^0, \{ W^{S^\prime,x} \} \cup \mathcal{W}^{P \setminus N} \rangle \\
+      &\models \langle E^{0\prime}, P^\prime, N^\prime \rangle
+    \end{split}
 \end{gather}
@@
   and a Windowed Continuous RIF-Core program as a WC-Datalog program
-  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{W^{S_1,r_1}, \dots, W^{S_n,r_n}\} \rangle}$.
+  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n)\} \rangle}$.
   Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
@@
   \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
-  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_i$, $P$ and $N$ and the window ranges $r_i$ for $i \in (1,s)$ of each ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle \models \langle E^0,P,N \rangle}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window range $r$.
+  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_x$, $P$ and $N$ and the window ranges $r_x$ for $x \in (1,s)$ of each $\Pi^{C,r}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window ranges $r_x$.
   It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
@@
     They are also \emph{in-order} with regards to their temporal annotation $e$, but potentially \emph{out-of-order} with regards to their temporal annotation $n$.
+    In practise, where a stream received from some source external to the model is not annotated with one or either, these instants are to be assigned by some trivial pre-processing;
+    this will be based on either the existing annotation, if it exists, or the current time, and the sliding window definitions in the modelled $\Pi^{C,r}$.
   \end{definition}
@@
     This means that the stream will contain no graph instances that map to instant $i$, for all instants ${i \in \mathbb{T}}$.
-    This stream may occur in specific expressions of any program, as described in \refsec{CoDeR: transformations}.
+    This stream may occur in specific expressions of any program when the range of an operator is mutually exclusive to that of the operator that succeeds it in a dataflow.
   \end{definition}
@@
 
+  Where many stream processing tasks (including continuous querying) are one-pass, the process of reasoning and inference is iterative/recursive.
+  This theoretically necessitates the union of the output stream with the input stream ``upstream'' of the remaining operators of any processing plan in the form of a \emph{working stream}, as per \refdef{CoDeR: working stream}.
+  \begin{definition}[Working Stream]\labeldef{CoDeR: working stream}
+    The \emph{working stream} of a CoDeR-based system is a CoDeR stream that is the union of all the system's input streams and its output stream.
+    It facilitates the \emph{iterative} application of the rules of the system to the streamed data, in the same manner as the \emph{working memory} of the Rete algorithm \citep{forgy79}.
+  \end{definition}
+
   The second class of streams, inter-operator streams, are CoDeR streams that are both produced and consumed exclusively by the stream-to-stream operators of CoDeR's minimal algebra.
@@
         \begin{split}
-          a \in g_a \land \\
-          b \in g_b \land \\
-          \langle g_a, e_a, n_a \rangle \land \\
-          \langle g_b, e_b, n_b \rangle \land \\
-          e_a \leqslant e_b < min\left( n_a , n_b \right)
+          a \in g_a & \land \\
+          b \in g_b & \land \\
+          \langle g_a, e_a, n_a \rangle & \land \\
+          \langle g_b, e_b, n_b \rangle & \land \\
+          e_a \leqslant e_b < min\left( n_a , n_b \right) &
         \end{split}
@@
     Finally, the functions required for querying deductive databases using Datalog are conventionally the complete set required of the relational algebra\footnote{This refers to the complete relational algebra less the functionality of set minus $\setminus$, which is analogous to the NaF operator $\sim$ that is not supported by positive Datalog.}, being ${\{\sigma,\pi,\Join,\cup\}}$ as described by \citet{codd70relationalModel}.
-    However, three of these may be restricted when processing semantic data, as discussed in \refsec{background: semantic queries}, becoming:
+    However, three of these may be restricted when processing semantic data, becoming:
     the one-pass unary equality-based triple-select $\sigma_{triple}$;
@@
     All one-pass functions required to derive the entailment of a WC-RIF-Core program expressed by an equivalent stream-to-stream operator in CoDeR, as detailed in \reftab{WC-RIF-Core-functions-to-CoDeR-operators}.
-    \begin{table}
+    \begin{table*}
         \centering
@@
         \labeltab{WC-RIF-Core-functions-to-CoDeR-operators}
-      \end{table}
+      \end{table*}
       The two-pass functions $\{w,s,\Join^\ast\}$, as shown in the same table, are combined into a single stream-to-stream operator in CoDeR in order to make the algebra universally stream-to-stream in nature.
@@
     Each provides a description of its behaviour on a triple-by-triple basis, and the formal definition of the stream produced in terms of the stream(s) to which the operator is applied.
-    Because many of the operators' behaviours are defined in terms of bindings within graph instances for variables in the sequenced pattern associated with a received stream, the functions $vars(stream)$ and $binding(graph, variable)$ are defined:
-    the former maps inter-operator streams to the sets of variables in their associated patterns;
-    the latter maps graphs and variables from any inter-operator stream and associated pattern to the value bound to that variable in that graph.
+    Many of the operators' behaviours are defined in terms of bindings within graph instances for variables in the sequenced pattern associated with a received stream;
+    to represent this relationship between streams/patterns and the graphs/propositions that match them we use the same notation of an \emph{extended variable substitution} $\bar{s}$ used in the satisfaction of $\ast$-Datalog programs.
 
@@
       Consider, for example, the input stream containing an instance of a graph containing the RDF triples $a$ and $b$ and valid from instant $e$ to instant $n$:
-      \[
-        {}_g{\pi_t} (\dots,\langle \{a\ .\ b\ .\},e,n \rangle,\dots) =
+      \begin{multline*}
+        {}_g{\pi_t} (\dots,\langle \{a\ .\ b\ .\},e,n \rangle,\dots) = \\
           (\dots,\langle \{a\ .\},e,n \rangle,\langle \{b\ .\},e,n \rangle,\dots)
-      \]
+      \end{multline*}
       More generally:
@@
         \left(
-          \langle g, e, n \rangle \:\middle\vert\: \langle g, e, n \rangle \in X \land matches\left(g,\{triple\ .\}\right)
+          \langle g, e, n \rangle
+        \:\middle\vert\:
+          \begin{split}
+            \langle g, e, n \rangle \in X & \land \\
+            \exists \bar{s} \left(
+              \bar{s}\left( \{ triple\ .\} \right) = g
+            \right)
+          \end{split}
         \right)
@@
       The sequenced pattern associated with stream $X$ must contain all the variables referenced as arguments that must be bound in the proposition ${external}$ \citep{w3crifcoreBuiltins}, as the proposition will always evaluate to false where not all necessary arguments are bound.
-      As such, the produced stream will necessarily be the empty stream $\varnothing$ (as per \refeqn{CoDeR: empty result of external select over unbound variables} in \refsec{CoDeR: transformations}).
+      As such, the produced stream will necessarily be the empty stream $\varnothing$, as per \refdef{CoDeR: empty stream}.
 
       More formally:
-      % \begin{equation*}
-      \begin{multline*}
-        \sigma_{f(a_0,\dots,a_n)} \left( X \right) =
+      \begin{equation*}
+        \sigma_{p} \left( X \right) =
         \left(
@@
         \:\middle\vert\:
-          \langle g, e, n \rangle \in X \land \right. \\
           \begin{split}
-            \forall x \in (0,n) v_x = \left\{
+            \langle g, e, n \rangle \in X & \land \\
+            \exists \bar{s} \left(
               \begin{split}
-                & binding\left(g, a_x\right) && \text{, if } a_x \in vars(X) \\
-                & a_x && \text{, otherwise}
-              \end{split}\right. & \land \\
-            f(v_{0},\dots,v_{n}) &
-        \left.\right)
+                \bar{s}\left( X \right) = g & \land \\
+                \bar{s}\left( p \right) &
               \end{split}
-      % \end{equation*}
-      \end{multline*}
+            \right)
+          \end{split}
+        \right)
+      \end{equation*}
     \end{nestedsection}
@@
       More formally, using the example case of a singleton graph pattern ${head = \{ a_0\ a_1\ a_2\ .\}}$ for readability:
-      % \begin{equation*}
-      \begin{multline*}
-        \coloneq_{\{ a_0\ a_1\ a_2\ .\}} \left( X \right) =
+      \begin{equation*}
+        \coloneq_{head} \left( X \right) =
         \left(
-          \langle \{v_{0}\ v_{1}\ v_{2}\ .\}, e, n \rangle
+          \langle h, e, n \rangle
         \:\middle\vert\:
-          \langle g, e, n \rangle \in X \land \right. \\
-            \forall x \in (0,2) v_x = \left\{
           \begin{split}
-                & binding\left(g, a_x\right) && \text{, if } a_x \in vars(X) \\
-                & a_x && \text{, otherwise}
+            \langle g, e, n \rangle \in X & \land \\
+            \exists \bar{s} \left(
+              \begin{split}
+                \bar{s}\left( X \right) = g & \land \\
+                \bar{s}\left( head \right) = h &
               \end{split}
             \right)
-      % \end{equation*}
-      \end{multline*}
+          \end{split}
+        \right)
+      \end{equation*}
     \end{nestedsection}
@@
         \left(
-          \langle g_i, e_i, n_i \rangle
+          \langle g, e, n \rangle
         \:\middle\vert\:
-            \langle g, e, n \rangle \in X \lor \langle g, e, n \rangle \in Y
+            \begin{split}
+              \langle g, e, n \rangle \in X & \lor \\
+              \langle g, e, n \rangle \in Y
+            \end{split}
         \right)
@@
       In CoDeR they are asymmetrical binary operators, represented algebraically by the binary operator $\rstreamjoin$, that simply supports the maintenance (or ``building'') of a \emph{valid window}, as per \refdef{continuous datalog: valid window}, from one inter-operator stream and naturally joining each graph instance on a second inter-operator stream against those within the maintained window (known as ``probing'' the window).
-      Since the persistent fact instances of $F^0$ are simply included as the CoDeR input stream $S_0$ and are annotated as being perpetually valid, as discussed in \refsec{CoDeR: data}, the applicable part of $F^0$ (and its entailment with respect to the rules $R$) will be perpetuated indefinitely in each SteM.
+      Since the persistent fact instances of $F^0$ are simply included as the CoDeR input stream $S_0$ and are annotated as being perpetually valid, as discussed in \refsec{model: data}, the applicable part of $F^0$ (and its entailment with respect to the rules $R$) will be perpetuated indefinitely in each SteM.
       The nature of probing in CoDeR is the same as that of joining the sets of matches for two graph patterns in SPARQL.
@@
         \left(
-          \langle g, e_Y, min \left( n_{X}, n_{Y} \right) \rangle \:\middle\vert\:
+          \langle g, e_Y, n_m \rangle \:\middle\vert\:
           \begin{split}
@@
             e_{X} \leqslant e_{Y} < n_{X} & \land \\
-            \left( \{g_X\} \Join^\ast \{g_Y\} = \{g\} \right) &
+            \{g_X\} \Join^\ast \{g_Y\} = \{g\} & \land \\
+            n_m = min \left( n_{X}, n_{Y} \right) &
           \end{split}
@@
 
-    %%% TRANSFORMATIONS
-
-    As is the case with relations in the relational algebra, the same stream may be defined in a variety of ways using the set of operators.
-    As each inter-operator stream is associated with a sequenced pattern, each definition of a stream may also be called a \emph{plan} for matching that stream's pattern.
-    Such equivalences between plans rely on a set of fundamental equivalences known as \emph{transformations}.
-    The most fundamental of the transformations of the relational algebra are traditionally classified as \emph{identity}, \emph{commutative}, \emph{associative} or \emph{distributive}, all of which apply equally well to the operators of CoDeR due to their similarity to those of SPARQL, which in turn have an equivalence to those of the relational algebra, as discussed by \citet{cyganiak05relationalSPARQL}.
-
-    In the remainder of this section: $a$ and $b$ are triple patterns; $g$ is a sequenced pattern; $p$ and $q$ are non-ground propositions using RIF-Core built-in predicates; $A$, $B$ and $C$ are CoDeR streams; and $X$, $Y$ and $Z$ are inter-operator streams.
-
-    Equations \ref{eqn:CoDeR: identity of graph-to-triples under succession} to \ref{eqn:CoDeR: identity of streams under union} enumerate the unconditional \emph{identity} transformations of the operators described in \refsec{CoDeR: graph-to-triples}.
-    \begin{align}
-      \labeleqn{CoDeR: identity of graph-to-triples under succession}
-      {}_g\pi_t\ {}_g\pi_t A &= {}_g\pi_t A \\
-      \labeleqn{CoDeR: identity of instance implication under succession}
-      \coloneq_g\ \coloneq_g X &=\ \coloneq_g X \\
-      \labeleqn{CoDeR: identity of streams under union}
-      A \cup A &= A
-    \end{align}
-    In addition, there are five identities that are conditional on the nature of the propositions $p$ and $q$ and the triple patterns $a$ and $b$.
-    The first of these are presented in \refeqn{CoDeR: identity of triple select under succession,CoDeR: identity of triple select under succession to instance implication}, where applying $\sigma_{a}$ to a stream associated with the singleton pattern ${\{b\ .\}}$ (as produced by $\sigma_{b}$ and $\coloneq_{\{b\ .\}}$) in the case where $b$ subsumes $a$ (or ${a \subseteq b}$) produces a stream that is identical to the original.
-    The same is true for \refeqn{CoDeR: identity of external select under succession}, only for propositions $p$ and $q$ where $q$ subsumes $p$ (or ${p \subset q}$ e.g.
-${p = greater{-}than(?x,5)}$ and ${q = greater{-}than(?x,4)}$).
-    Furthermore, in \refeqn{CoDeR: identity of graph-to-triples under succession to instance implication,CoDeR: identity of graph-to-triples under succession to triple-select}, applying ${{}_g\pi_t}$ to a stream associated with a singleton pattern (as produced by $\sigma_{triple}$ and $\coloneq_{\{a\ .\}}$) produces a stream that is identical to the original.
-    \begin{align}
-      \labeleqn{CoDeR: identity of triple select under succession}
-      \sigma_b\ \sigma_a X &= \sigma_a X && \text{where } a \subseteq b \\
-      \labeleqn{CoDeR: identity of triple select under succession to instance implication}
-      \sigma_b \coloneq_{\{a\ .\}} X &=\ \coloneq_{\{a\ .\}} X && \text{where } a \subseteq b \\
-      \labeleqn{CoDeR: identity of external select under succession}
-      \sigma_q\ \sigma_p X &= \sigma_p X && \text{where } p \subseteq q \\
-      \labeleqn{CoDeR: identity of graph-to-triples under succession to instance implication}
-      {}_g\pi_t \sigma_{a} X &=\ \sigma_{a} X \\
-      \labeleqn{CoDeR: identity of graph-to-triples under succession to instance implication}
-      {}_g\pi_t \coloneq_{\{a\ .\}} X &=\ \coloneq_{\{a\ .\}} X
-    \end{align}
-
-  \refeqn{CoDeR: commutativity of triple select,CoDeR: commutativity of external select,CoDeR: commutativity of union} enumerate the \emph{commutative} transformations of the operators described in \refsec{CoDeR: operator detail}.
-  These are the transformations that involve the re-ordering of unary operators of the same kind within a plan, as well as the swapping of inputs to binary operators, to produce another for the same sequenced pattern.
-  \begin{align}
-    \labeleqn{CoDeR: commutativity of triple select}
-    \sigma_a\ \sigma_b X &= \sigma_b\ \sigma_a X \\
-    \labeleqn{CoDeR: commutativity of external select}
-    \sigma_p\ \sigma_q X &= \sigma_q\ \sigma_p X \\
-    \labeleqn{CoDeR: commutativity of union}
-    A \cup B &= B \cup A
-  \end{align}
-  It should be noted that the SteM operator $\rstreamjoin$ is \emph{not} commutative, unlike the symmetric window-join.
-  However the definition of the window-join in terms of the commutative stream union and complementary SteMs in \refeqn{window-join-from-union-of-SteMs} does affirm the fact that the window-join \emph{is} commutative.
-
-  \refeqn{CoDeR: associativity of union,CoDeR: associativity of SteM} enumerate the associative transformations of the operators described in \refsec{CoDeR: operator detail}.
-  These are the transformations that involve the re-ordering of binary operators of the same kind within a plan to produce another for the same sequenced pattern.
-  \begin{align}
-    \labeleqn{CoDeR: associativity of union}
-    A \cup \left( B \cup C \right) &= \left( A \cup B \right) \cup C \\
-    \labeleqn{CoDeR: associativity of SteM}
-    X \rstreamjoin \left( Y \rstreamjoin Z \right) &= \left( X \rstreamjoin Y \right) \rstreamjoin Z
-  \end{align}
-
-  \refeqn{CoDeR: distributivity of selects,CoDeR: distributivity of triple select over union,CoDeR: distributivity of external select over union,CoDeR: distributivity of instance implication over union,CoDeR: distributivity of graph-to-triples over union} enumerate the distributive transformations of the operators described in \refsec{CoDeR: operator detail}.
-  These are the transformations that involve the re-ordering of operators of different kinds within a plan to produce another for the same sequenced pattern.
-  \begin{align}
-    \labeleqn{CoDeR: distributivity of selects}
-    \sigma_a\ \sigma_p X &= \sigma_p\ \sigma_a X \\
-    \labeleqn{CoDeR: distributivity of graph-to-triples over union}
-    {}_g\pi_t \left( A \cup B \right) &= \left( {}_g\pi_t A \right) \cup \left( {}_g\pi_t B \right) \\
-    \labeleqn{CoDeR: distributivity of triple select over union}
-    \sigma_a \left( X \cup Y \right) &= \left( \sigma_a X \right) \cup \left( \sigma_a Y \right) \\
-    \labeleqn{CoDeR: distributivity of external select over union}
-    \sigma_p \left( X \cup Y \right) &= \left( \sigma_p X \right) \cup \left( \sigma_p Y \right) \\
-    \labeleqn{CoDeR: distributivity of instance implication over union}
-    \coloneq_{g} \left( A \cup B \right) &= \left( \coloneq_{g} A \right) \cup \left( \coloneq_{g} B \right)
-  \end{align}
-  In addition, there are three conditional transformations, enumerated in \refeqn{CoDeR: distributivity of external select over SteM to left,CoDeR: distributivity of external select over SteM to right,CoDeR: distributivity of external select over SteM to both}, that distribute $\sigma_{external}$ over $\rstreamjoin$ and are analogous to those distributing selection $\sigma$ over join $\Join$ in the relational algebra.
-  Consider the non-ground propositions
-  \begin{align*}
-    p &= greater{-}than(?1, 5) & q &= less{-}than(?3, 5)
-  \end{align*}
-  and associations between inter-operator streams and sequenced patterns
-  \begin{align*}
-    X &\mapsto \{?1\ pred1\ ?2\ .\} & Y &\mapsto \{?2\ pred2\ ?3\ .\} & Z &\mapsto \{?1\ pred3\ obj\ .\}
-  \end{align*}
-  The conditional transformations are thus: 
-  % In the first case, non-ground proposition $p$ requires a set of variables ${\{?1\}}$, wholly subsumed by those present in the pattern associated with $X$ but not subsumed by the pattern associated with $Y$, to be bound.
-  % As such, $\sigma_{p}$ may be distributed over $\rstreamjoin$ to $X$ but not to $Y$ (due to \refeqn{CoDeR: empty result of external select over unbound variables,CoDeR: identity of empty stream under SteM right} making ${\left( \sigma_p X \right) \rstreamjoin \left( \sigma_p Y \right) = \varnothing}$).
-  \begin{equation}
-    \labeleqn{CoDeR: distributivity of external select over SteM to left}
-    \sigma_p \left( X \rstreamjoin Y \right) = \left( \sigma_p X \right) \rstreamjoin Y
-  \end{equation}
-  % In the second case, the non-ground proposition $q$ requires a set of variables ${\{?3\}}$, wholly subsumed in the pattern associated with $Y$ but not that associated with $X$, to be bound.
-  % As such, $\sigma_{q}$ may be distributed over $\rstreamjoin$ to $Y$ but not to $X$ (due to \refeqn{CoDeR: empty result of external select over unbound variables,CoDeR: identity of empty stream under SteM left}).
-  \begin{equation}
-    \labeleqn{CoDeR: distributivity of external select over SteM to right}
-    \sigma_q \left( X \rstreamjoin Y \right) = X \rstreamjoin \left( \sigma_q Y \right)
-  \end{equation}
-  % In the final case, the set of variables required by $p$ to be bound is wholly subsumed in the patterns associated with both $X$ and $Z$.
-  % As such, $\sigma_{p}$ may be distributed over $\rstreamjoin$ to both $X$ and $Z$.
-  \begin{equation}
-    \labeleqn{CoDeR: distributivity of external select over SteM to both}
-    \sigma_p \left( X \rstreamjoin Z \right) = \left( \sigma_p X \right) \rstreamjoin \left( \sigma_p Z \right)
-  \end{equation}
-  % It should be noted that, in the final case, $\sigma_p$ may alternatively be distributed to only one or the other of streams $X$ and $Y$ as in \refeqn{CoDeR: distributivity of external select over SteM to left,CoDeR: distributivity of external select over SteM to right};
-  % this is because filtering certain bindings for, for example, ${?1}$ from $X$ will prevent $\rstreamjoin$ from finding matches for those bindings for ${?1}$ if they should occur in $Z$, and vice versa.
-
-  Finally, the \emph{empty stream}, as per \refdef{CoDeR: empty stream}, may be created in a number of ways and creates a set of special identities, the latter enumerated in \refeqn{CoDeR: identity of empty stream under graph-to-triples,CoDeR: identity of empty stream under triple select,CoDeR: identity of empty stream under external select,CoDeR: identity of empty stream under instance implication,CoDeR: identity of empty stream under SteM right,CoDeR: identity of empty stream under SteM left,CoDeR: identity of stream under union with empty stream}.
-  \begin{align}
-    \labeleqn{CoDeR: identity of empty stream under graph-to-triples}
-    {}_g\pi_t \varnothing &= \varnothing \\
-    \labeleqn{CoDeR: identity of empty stream under triple select}
-    \sigma_a \varnothing &= \varnothing \\
-    \labeleqn{CoDeR: identity of empty stream under external select}
-    \sigma_p \varnothing &= \varnothing \\
-    \labeleqn{CoDeR: identity of empty stream under instance implication}
-    \coloneq_g \varnothing &= \varnothing \\
-    \labeleqn{CoDeR: identity of empty stream under SteM right}
-    X \rstreamjoin \varnothing &= \varnothing \\
-    \labeleqn{CoDeR: identity of empty stream under SteM left}
-    \varnothing \rstreamjoin X &= \varnothing \\
-    \labeleqn{CoDeR: identity of stream under union with empty stream}
-    \varnothing \cup A &= A
-  \end{align}
-  The ways in which the empty stream is created all stem from the possibility that a successor $\sigma_\ast$ operator might filter the stream that it receives based on a pattern that will never match any graph instance produced by the operator that it succeeds.
-  \refeqn{CoDeR: empty result of triple select under succession,CoDeR: empty result of triple select under succession from instance implication,CoDeR: empty result of external select under succession,CoDeR: empty result of external select over unbound variables} provide an example of each possible case where this occurs.
-  \begin{align}
-    \labeleqn{CoDeR: empty result of triple select under succession}
-    \text{Conflicting patterns:} && \sigma_{?1\ pred1\ ?2}\ \sigma_{?1\ pred2\ ?2} X &= \varnothing \\
-    \labeleqn{CoDeR: empty result of triple select under succession from instance implication}
-    \text{Conflicting patterns:} && \sigma_{?1\ pred1\ ?2} \coloneq_{\{?1\ pred2\ ?2\ .\}} X &= \varnothing \\
-    \labeleqn{CoDeR: empty result of external select under succession}
-    \text{Mutually exclusive propositions:} && \sigma_{greater{-}than(?1,5)}\ \sigma_{less{-}than(?1,5)} X &= \varnothing \\
-    \labeleqn{CoDeR: empty result of external select over unbound variables}
-    \text{Unbound variables in proposition:} && \sigma_{greater{-}than(?3,5)}\ \sigma_{?1\ pred1\ ?2} X &= \varnothing
-  \end{align}
-  \refeqn{CoDeR: empty result of external select over unbound variables} expresses the extension of rule safety in RIF-Core relating to the required binding elsewhere in a pattern of any variables that are necessary arguments to a built-in predicate or function in that pattern.
-  The example given assumes that both arguments of the concisely-named predicate ${greater{-}than()}$ must be bound, as they must in the actual built-in predicate of RIF-Core that it emulates:
-  \[
-    {<}http://www.w3.org/2007/rif{-}builtin{-}predicate\#numeric{-}greater{-}than{>}
-  \]
-  However, there are a few cases where specific arguments of specific predicates need not be bound \citep{w3crifcoreBuiltins}; the transformation in \refeqn{CoDeR: empty result of external select over unbound variables} does \emph{not} apply in the cases where unbound variables appear as those specific arguments.
+  %   \begin{nestedsection}{Transformations}{CoDer: transformations}
+  %     As is the case with relations in the relational algebra, the same stream may be defined in a variety of ways using the set of operators.
+  %     As each inter-operator stream is associated with a sequenced pattern, each definition of a stream may also be called a \emph{plan} for matching that stream's pattern.
+  %     Such equivalences between plans rely on a set of fundamental equivalences known as \emph{transformations}.
+  %     The most fundamental of the transformations of the relational algebra are traditionally classified as \emph{identity}, \emph{commutative}, \emph{associative} or \emph{distributive}, all of which apply equally well to the operators of CoDeR due to their similarity to those of SPARQL, which in turn have an equivalence to those of the relational algebra, as discussed by \citet{cyganiak05relationalSPARQL}.
+
+  %     Many of the transformations in this algebra are directly related to those of the relational algebra.
+  %     Those applied to selections $\sigma$ and joins $\Join$ are mirrored exactly by those applied to external-selects $\sigma_{external}$ and SteMs $\rstreamjoin$, except for the fact that SteMs are not commutative.
+  %     Also, all operators are distributive over stream union, which itself inherits all of the transformations of set union with respect to itself (commutative, associative and identity).
+  %     Furthermore, both $\sigma_\ast$ operators of CoDeR are commutative with respect to themselves and each other.
+  %     On the other hand, unlike in the relational algebra, triple-selects $\sigma_{triple}$ are not distributive over SteMs, and neither the graph-to-triples ${{}_g\pi_t}$ nor instance implication $\coloneq_{head}$ operators are commutative or distributive over other operators.
+
+  %     In addition, there are a number of identities that allow the elimination of redundant operators from a plan.
+  %     These are mostly based on the subsumption of the pattern associated with an inter-operator stream by a $\sigma_\ast$ operator.
+  %     For example, if a stream $s$ is produced on which a given variable $?v > x$, then any stream produced by an external-select on the proposition $?v > y \leqslant x$ applied to $s$ will be equal to $s$, making the external-select redundant.
+  %     Equally, if a stream $s$ of singleton graphs matches the pattern ${\{ s\ p\ ?o\ .\}}$, then any stream produced by a triple-select $\sigma_{?s\ p\ ?o}$ applied to $s$ will be equal to $s$, making the triple-select redundant.
+  %     Finally, a graph-to-triples operator is obviously redundant when applied to an inter-operator stream of singleton-graph instances.
+  % \end{nestedsection}
 \end{nestedsection}
 
-\begin{nestedsection}{CoDeR Streams Notation}{model: stream notation}
-  In compliment to the set of operators of the CoDeR algebra, we propose an extended notation that allows the accurate description a system built according to the CoDeR model, taking into account the iterative nature of inference and the uniqueness of streams within the model.
-  This notation is needed to express two of the fundamental characteristics of the CoDeR data model, those being the intention of performing inference and the uniqueness of streams within the data model.
-
-  Where many stream processing tasks (including continuous querying) are one-pass, the process of reasoning and inference is iterative/recursive.
-  This theoretically necessitates the union of the output stream with the input stream ``upstream'' of the remaining operators of any processing plan in the form of a \emph{working stream}, as per \refdef{CoDeR: working stream} and as shown in \reffig{CoDeR: re-entrant stream}.
-  \begin{definition}[Working Stream]\labeldef{CoDeR: working stream}
-    The \emph{working stream} of a CoDeR-based system is a CoDeR stream that is the union of all the system's input streams and its output stream.
-    It facilitates the \emph{iterative} application of the rules of the system to the streamed data, in the same manner as the \emph{working memory} of the Rete algorithm \citep{forgy79}.
-  \end{definition}
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.9\textwidth]{CoDeR-streams}
-    \caption[Streams in CoDeR System]{An abstract CoDeR-based system, demonstrating the purview of each class of stream: Input and Output streams to receive data from and publish data to external sources; Inter-operator streams facilitate operator nesting/pipelines; and the re-entrant Working Stream facilitates iterative processing.}
-    \labelfig{CoDeR: re-entrant stream}
-  \end{figure}
-  This feedback of the output stream into the \emph{working stream} may cause cause the definition of the output stream of a system to be circular, thus impossible to express purely in terms of the operators of the algebra described in \refsec{CoDeR: algebra}.
-  This problem of inexpressibility also applies to the sharing of streams between potentially many sinks and sources:
-  when relying solely on the nested-operator notation of the relational algebra, the same stream must be defined multiple times throughout a plan if it occurs in multiple sub-plans, though in reality that stream need only occur once in the modelled system.
-
-  As such, the complete expression of the processing required for inference over WC-RIF-Core programs necessitates that streams may be named and that, once named, those streams may be referenced by those names within the definitions of other streams.
-  Further, it must be possible to provide multiple contributing definitions for each named stream, this being an expression of \emph{stream union} derived from the manner in which the same predicate may have multiple disjunctive intensions within a Datalog program.
-
-  Finally, it must be possible to project the variables in a sequenced pattern associated with an inter-operator stream into a variant of that pattern, in order for that stream to be referenced within other plans that may use conflicting variable names.
-  However, this ``projection'' of variable names is not an operator of the algebra, those being related to pattern matching, which is variable-name agnostic\footnote{
-    By this I mean that a pattern is matched based on the existence of components that satisfy certain comparisons against given constant values and against specific other components.
-    How each of those components is labelled is irrelevant to the process, as long as the nature of those comparisons is maintained.
-    This is distinct from interpretations of the relational algebra such as SQL, in which the naming of the attributes adds semantic value to the data.}.
-  Instead, it is simply a requirement of the expression of plans within a system, met in the convention of naming inter-operator streams using variants of their pattern, i.e. ${S_{pattern}}$.
-  When naming or referencing a stream, the variables may therefore be renamed as appropriate for the context in which the stream is referenced.
-
-  These requirements are met in CoDeR using the notation shown in \refeqn{CoDeR: stream assignment}, an example of which for expressing the plan for some $\Pi^{C,r}$ is given in \refeqn{CoDeR: stream assignment example}.
-  \begin{align}
-    [stream\ name] &= [stream\ definition] \labeleqn{CoDeR: stream assignment} \\
-    \begin{split}
-      working\_stream &= {<}\text{http://example.org/test\_stream}{>}\\
-      S_{\{?s\ ?p\ ?o\ .\}} &= {}_g\pi_t\ working\_stream \\
-      S_{\{?s\ inside\ ?o\ .\}} &= \sigma_{?s\ inside\ ?o}\ S_{\{?s\ ?p\ ?o\ .\}} \\
-      S_{\{?1\ inside\ ?2\ .\ ?2\ inside\ ?3\ .\}} &=\ S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \\
-      S_{\{?1\ inside\ ?2\ .\}} &=\ \coloneq_{\{?1\ inside\ ?2\ .\}}\ S_{\{?1\ inside\ ?3\ .\ ?3\ inside\ ?2\ .\}} \\
-      & \vdots \\
-      inference\_stream &= S_{\{?s\ inside\ ?p\ .\}} \cup \dots \cup \dots \\
-      working\_stream &= inference\_stream
-    \end{split} \labeleqn{CoDeR: stream assignment example}
-  \end{align}
-  The first line of \refeqn{CoDeR: stream assignment example} demonstrates the definition of the plan's input stream by its IRI, and its being named ${working\_stream}$.
-  The third through fifth lines demonstrate the dual definition of a single inter-operator named stream using variant graph patterns, in this case expressing the transitive property of the example predicate ${inside}$ via the definition of of an inter-operator named stream (using \refeqn{window-join-from-union-of-SteMs,CoDeR: identity of streams under union}).
-  The penultimate line demonstrates the definition of the plan's inference stream, this being the \emph{stream union} of all streams produced by processing the separate rules of $R$.
-  The last line asserts that the inference stream contributes through \emph{stream union} with the input stream to the previously named ${working\_stream}$, thereby making the plan iterative.
-
-  In the plan presented in \refeqn{CoDeR: stream assignment example}, the stream defined on the fourth line is defined once and used once.
-  As such, the stream definition on the fifth line may be rewritten, replacing the stream reference $S_{\{?1\ inside\ ?3\ .\ ?3\ inside\ ?2\ .\}}$ with its definition ${\left( S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \right)}$ from the fourth line, renaming variables as appropriate.
-  This results in the plan presented in \refeqn{CoDeR: stream assignment example rewritten}.
-  \begin{equation}
-    \begin{split}
-      working\_stream &= {<}\text{http://example.org/test\_stream}{>}\\
-      S_{\{?s\ ?p\ ?o\ .\}} &= {}_g\pi_t\ working\_stream \\
-      S_{\{?s\ inside\ ?o\ .\}} &= \sigma_{?s\ inside\ ?o}\ S_{\{?s\ ?p\ ?o\ .\}} \\
-      S_{\{?1\ inside\ ?3\ .\}} &=\ \coloneq_{\{?1\ inside\ ?3\ .\}}\ \left( S_{\{?1\ inside\ ?2\ .\}} \rstreamjoin S_{\{?2\ inside\ ?3\ .\}} \right) \\
-      & \vdots \\
-      output\_stream &= S_{\{?s\ inside\ ?p\ .\}} \cup \dots \cup \dots \\
-      working\_stream &= output\_stream
-    \end{split} \labeleqn{CoDeR: stream assignment example rewritten}
-  \end{equation}
-  This method of rewriting may also be used to isolate a part of the plan as a single algebraic definition, to which the transformations presented in \refsec{CoDeR: transformations} may be applied.
-  This rewriting process may then be carried out in reverse, assigning names to streams produced by sub-definitions, then replacing those sub-definitions with references to the corresponding named streams in the original definition.
-
-  The example plans presented in \refeqn{CoDeR: stream assignment example,CoDeR: stream assignment example rewritten} also demonstrate that iteration within a plan follows naturally from the uniqueness of inter-operator streams by their associated sequenced pattern, discussed in \refsec{CoDeR: data}:
-  in each case, the second definition of the stream $S_{\{?1\ inside\ ?3\ .\}}$ is recursive.
-  The generalised case is of a rule whose head is a singleton graph pattern ${\{h\ .\}}$ that is a variant of the pattern of some triple-select operator $\sigma_{h}$;
-  both the triple-select and instance implication contribute to the same stream, resulting in an iterative feedback of entailed graph instances into the plan that bypasses the \emph{working stream}, as shown in \reffig{CoDeR: unique stream feedback}.
-  \begin{figure}
-    \centering
-    \includegraphics[width=0.9\textwidth]{CoDeR-unique-stream-feedback}
-    \caption[Iteration by Uniqueness of Streams in CoDeR]{An abstract CoDeR-based system, demonstrating the natural iterative feedback caused by the uniqueness of streams in the model.}
-    \labelfig{CoDeR: unique stream feedback}
-  \end{figure}
-  To support the correctness of such behaviour, it can be shown that any such iterative feedback may be derived algebraically from the definition and reference of the working stream in any plan that expresses it precisely as in \refdef{CoDeR: working stream}.
-  %
-  % In order to derive this proof, the selection of triple instances matching the pattern $h$ from the \emph{working stream}, i.e.
-${\sigma_{h}\ {}_{g}\pi_{t}\ working\_stream}$, must be rewritten (as already described) until all instance implication operators of the plan are present in a single algebraic expression.
-  % This rewriting is demonstrated in \refeqn{CoDeR: named stream rewriting}, where:
-  % \begin{itemize}
-  %   \item
-  %     $S_{g}$ is any inter-operator stream the precise nature of whose associated sequenced pattern is not relevant to the proof;
-  %   \item
-  %     $f$ is any sequenced pattern that is strictly incompatible with $h$ as per the transformation expressed in \refeqn{CoDeR: empty result of triple select under succession from instance implication};
-  %   \item
-  %     `$\dots$' is the remainder of the inter-operator streams that contribute to the CoDeR output stream, and whose associated sequenced pattern is neither $h$ nor incompatible with it.
-  % \end{itemize}
-  % \begin{equation} \labeleqn{CoDeR: named stream rewriting}
-  %   \begin{split}
-  %     \sigma_{h}\ {}_g\pi_t\ working\_stream &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup output\_stream \right) \\
-  %     &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \left( S_{\{h\ .\}} \cup \dots \cup S_{\{f\ .\}} \right) \right) \\
-  %     &= \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \left( \left( \coloneq_{\{h\ .\}} S_{g} \right) \cup \dots \cup \left( \coloneq_{\{f\ .\}}\ S_{g} \right) \right) \right)
-  %   \end{split}
-  % \end{equation}
-  % The proof may then be derived by showing that the filtering of the entire working stream by $h$ is equal to the union of the entailed stream matching pattern ${\{h\ .\}}$ with the remainder of the working stream filtered by $h$, less those entailed streams that match patterns $f$ that are incompatible with $h$.
-  % This result is presented in \refeqn{CoDeR: transformations for closer iteration}, and may be proven by applying the following transformations to the right hand side:
-  % \ref{eqn:CoDeR: associativity of union} and \ref{eqn:CoDeR: commutativity of union}%;
-  % to re-arrange the union;
-  % \ref{eqn:CoDeR: distributivity of graph-to-triples over union} and \ref{eqn:CoDeR: distributivity of triple select over union}%;
-  % to distribute the selection by $h$;
-  % \ref{eqn:CoDeR: identity of graph-to-triples under succession to instance implication}%;
-  % to simplify each selection;
-  % \ref{eqn:CoDeR: identity of triple select under succession to instance implication}%;
-  % to eliminate selection over the entailed stream matching ${\{h\ .\}}$;
-  % and \ref{eqn:CoDeR: empty result of triple select under succession from instance implication} and \ref{eqn:CoDeR: identity of stream under union with empty stream}%.
-  % to eliminate from consideration the entailed streams matching ${\{f\ .\}}$.
-  % \begin{equation} \labeleqn{CoDeR: transformations for closer iteration}
-  %   \left( \begin{split}
-  %     \left( \coloneq_{\{h\ .\}}\ S_{g} \right) \cup \\
-  %     \left( \sigma_{h}\ {}_g\pi_t\ \left( input\_stream \cup \dots \right) \right)
-  %   \end{split} \right)
-  %   &
-  %     =\ \sigma_{h}\ {}_g\pi_t\ \left( \begin{split}
-  %       input\_stream \cup \\
-  %       \left( \left( \coloneq_{\{h\ .\}} S_{g} \right) \cup \dots \cup \left( \coloneq_{\{f\ .\}} S_{g} \right) \right)
-  %     \end{split} \right)
-  % \end{equation}
-  This reflects the use of \emph{variant tabling} for optimising Prolog programs as discussed by \citet{DS4SbsmptvTbls}.
-\end{nestedsection}
-
-  % As the SteM ${{}_b{\Join_p}}$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
+  \begin{nestedsection}{Summary of CoDeR}{CoDeR: summary}
+    Given the time-based nature of windowing and of the justification of entailments in WC-RIF-Core, it is possible to determine in advance the instant at which fact instances will be negated.
+    Furthermore, facts in RDF may be grouped into graphs, meaning that fact instances with the same annotation that will be negated at the same instant may be grouped into an annotated graph without loss of information.
+    As such, both streams $P$ and $N$ entailed by a $\Pi^{C,r}$ may be represented by streams of CoDeR's twice-annotated graphs.
     As the SteM $\rstreamjoin$ is the only CoDeR operator that maintains any form of window over its input, and SteMs maintain exclusively \emph{valid windows}, no data will persist within a CoDeR-based system when it is no longer valid according to its annotations.
@@
     as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n) \} \rangle}$.
-  It follows that no facts may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of WC-RIF-Core.
-  It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^C$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
+
+    It follows that no fact instance may persist in a CoDeR based system if it is no longer valid, and thus only entailments that are valid at a given moment may be produced in that moment, thereby guaranteeing correct results according to the semantics of WC-RIF-Core.
+    It also follows that, given a system with a processing plan that is expressible in the CoDeR algebra and accurately represents a set of rules $R$ of a $\Pi^{C,r}$, all fact instances that are logically entailed by that $\Pi^{C,r}$ at a given instant will be persisted within that system at that instant,
     i.e. such a CoDeR-based system is guaranteed to produce the complete stream of results according to the semantics of WC-RIF-Core.
+  \end{nestedsection}
 \end{nestedsection}
@@
 
-    When a join node is left activated, i.e.
+    When a join node is left-activated, i.e.
 it receives a token through its left input, it first adds the new token to its left window, then it prunes its right window before iterating over the remaining quads to find any matches.
@@
     The new token is annotated with the \emph{earliest} time with which its component quads were annotated, and then propagated to the next beta node, or to the terminal node if it is in the root node of a join tree.
-    Conversely, when the join node is right activated by receiving a quad from an alpha node, it adds the new quad to the right window, prunes the left window then matches the quad against the remaining tokens.
+    Conversely, when the join node is right-activated by receiving a quad from an alpha node, it adds the new quad to the right window, prunes the left window then matches the quad against the remaining tokens.
 
@@
     This leaves the annotation $a$ of a token in R4, in combination with the range $r$ of the semantic window associated with the rule set, able to express CoDeR's negation time ${n}$ of entailed instances by the same calculation as those of asserted instances: ${n = a + r}$.
-    As such, the interpretation of beta nodes in R4 is a valid implementation of the window-join operator characterised as a pair of CoDeR SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{continuous datalog: SteM}.
+    As such, the interpretation of beta nodes in R4 is a valid implementation of the window-join operator characterised as a pair of CoDeR SteMs $\rstreamjoin$ implicitly unioned by contribution to the same stream, as shown in \reffig{CoDeR: SteM}.
 
@@
 
-  \begin{nestedsection}{Motivating Case}{evaluation: motivating case}
-    \begin{figure*}[t]
-      \centering
-      \begin{verbatim}
-Forall ?obs ?sensor ?result ?value ?v ?hur ?hurWind (
-    If And( ?obs [ssn:observedProperty -> wind_speed]
-        ?obs [ssn:observedBy -> ?sensor]
-            ?obs [ssn:observationResult -> ?result] 
-            ?result [ssn:hasValue -> ?value]
-            ?value [ssnExt:hasQuantityValue -> ?v]
-            ?hur [rdf:type -> yago:Hurricane111467018]
-            ?hur [dbpprop:1MinWinds -> ?hurWind] 
-            External (pred:numeric-greater-than(?v ?hurWind)) )
-    Then ?obs [ssn:observedWeatherType -> yago:Hurricane111467018] )
-      \end{verbatim}
-      \caption{The RIF-Core rule inspired by SRBench.}
-      \labelfig{SRBench-rule}
-    \end{figure*}
-
-    We first used a 1.1 million triple real-world dataset from the SemSorGrid4Env project that contains weather observations over a period of ten days.
-    Inspired by a query in \citep{SRBench}, we formulated a rule that entails that a given observation is of a weather pattern classified as a hurricane if it is of a wind speed that exceeds that of a previously classified hurricane;
-    it also entails that the wind speed observed exceeds that of a given historically recorded hurricane (\reffig{SRBench-rule}).
-    This rule involves dealing with static data regarding past hurricanes, dynamic data regarding ongoing sensor readings, and also needs to perform reasoning against an ontology of background knowledge to identify the hurricanes that are instances of the specified class of hurricanes.
-    For the historical hurricane data, we prepared a small dataset taken from dbpedia, containing information about a number of hurricanes.
-
-    The experiment consisted of pushing the whole stream to the compiled R4 system and measuring the time taken to process all the data, repeated for a variety the window ranges.
-    We ran this rule successfully on R4 and Jtalis getting correct results.
-    However, Sparkwave was unable to read the dataset correctly as it does not parse blank nodes;
-    this is because it uses the hash-join algorithm to improve time efficiency, as we show in \refsec{evaluation: comparative results}, but builds the hashtables based on the URIs of the incoming data as which blank nodes are not valid.
-  \end{nestedsection}
   \begin{nestedsection}{Comparative Results}{evaluation: comparative results}
@@
 
-    In order to compare the efficiency of R4 against both Etalis \emph{and} Sparkwave, we formulated a second experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
-    We used another 1.1 million triples dataset, this time using the provided data generator, containing information about 100,000 limited-time offers made available by some fictional online marketplace.
+    In order to compare the efficiency of R4 against both Etalis and Sparkwave, we formulated an experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
+    We used another 1.1 million triples dataset using the provided data generator, containing information about 100,000 limited-time offers made available by some fictional online marketplace.
     We used a small schema that has 329 product types arranged in a 4-level hierarchy.
@@
 \begin{nestedsection}{Conclusions}{conclusions}
-  We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a well-founded semantics for continuous reasoning at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
-  In this way it differs from other temporal interpretations of Datalog \citep{OrgunWadge92,Tuzhilin93} in that it forgoes the concepts of conventional temporal reasoning, replacing them with the windowing semantics of stream processing and casting a continuous program as a series of discrete states.
-  WC-Datalog supports the derivation of sets of facts as the entailment of each discrete state of a continuous program and the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.
-  In addition to this, however, it also supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program via the application of windows to the concise streams, leading to the well-founded definition of a \emph{valid window} over the combined streams $P$ and $N$, being the outputs of the reasoning equivalents of the \emph{IStream} and \emph{DStream}.
-  Thus WC-Datalog represents a well-founded semantics not only for continuous reasoning, but for coherent and concise streamed entailments of continuous programs, suitable for consumption in further stream processing.
-
-  In order to express this semantics we have developed CoDeR, a processing model for Continuous Deductive Reasoning over RDF.
+  We have presented Windowed Continuous Datalog (\emph{WC-Datalog}), a semantics for stream-based reasoning and streamed entailments at the expressivity of OWL 2 RL, extended incrementally from that of positive Datalog with temporal annotation, streams and windowing semantics.
+  In this way it differs from temporal interpretations of Datalog \citep{OrgunWadge92,Tuzhilin93} in that it forgoes the concepts of conventional temporal reasoning, instead applying temporally-agnostic rules to global sliding windows over streams as in stream querying, and casting a continuous program as a series of discrete states.
+  WC-Datalog explicitly supports the derivation of sets of facts as the entailment of each discrete state of a continuous program \emph{and} the derivation of concise streams of facts as the entailment of a continuous program over time, in a similar manner to the \emph{RStream}, \emph{IStream} and \emph{DStream} operators of the CQL continuous query language \citep{CQL}.
+  In addition to this, however, it also supports the low-complexity derivation of the entailment of each discrete state from the streamed entailment of the continuous program via the application of sliding windows to the concise streams, leading to the definition of a \emph{valid window} over the combined streams $P$ and $N$.
+  Thus WC-Datalog represents a semantics not only for low-complexity continuous reasoning, but for \emph{concise} streamed entailments of continuous programs, suitable for consumption in further stream processing without loss of temporal information.
+
+  In order to express this semantics we have developed CoDeR, a processing model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over RDF.
   This processing model is composed of a stream-based data model with a dual semantics and a stream-to-stream extension of a semantic interpretation of the relational algebra.
-  The data model is composed of streams of RDF graphs annotated both with the time at which they became true and the time at which they will cease to be true, with the ordering of graphs by the former being the sequence of facts in $P$ and the ordering by the latter being the sequence of facts in $N$, hence providing a dual semantics for the stream.
-  As such, the window inherent in a continuous program may be represented in the annotation of each graph in the streamed input, rather than ranged window operators in the algebra.
+  The data model is composed of streams of RDF graphs annotated both with the time at which they became valid and the time at which they will cease to be valid, with the ordering of graphs by the former being the sequence of facts in $P$ and the ordering by the latter being the sequence of facts in $N$, hence providing a dual semantics for the stream.
+  As such, the windows inherent in a windowed continuous program may be represented in the annotation of each graph in the streamed input, rather than ranged window operators in the algebra.
   Instead, the algebra makes use of \emph{valid window} operators to maintain the set of graphs that are both considered to be true at a given moment \emph{and} applicable at a given point in the processing.
@@
   R4 utilises a variant of the Rete pattern matching algorithm \citep{forgy79} to produce processing plans for sets of rules expressed in RIF-Core \citep{w3crifcore}, the W3C recommendation for expressing sets of rules with the semantics of positive Datalog and the rule language used to express the set of entailment rules of OWL 2 RL \citep{w3cowl2profiles}.
-  R4 is sound and complete with respect to the semantics of WC-Datalog by virtue of implementing CoDeR.
-  Given that many state-of-the-art stream-querying and continuous-reasoning systems have varying semantics, and so the expected results of each system will vary for the same query definition, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and Etalis \citep{EP-SPARQL}.
-  We have demonstrated that R4 is comparable in both respects to these RDF stream-querying systems when applying low-range windows to high-throughput streams, despite Sparkwave and Etalis supporting more limited/less explicitly well-founded reasoning than that of R4.
-
-  Moving forward, we plan to improve the time-efficiency of R4 by investigating more efficient join algorithms within its implementations of SteMs.
-  At a higher level, we also plan to investigate alternate planing algorithms to improve the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
+  R4 is sound and complete with respect to the semantics of WC-RIF-Core by virtue of implementing CoDeR.
+  Given that there are no benchmarks for more expressive stream-based reasoning, we contrasted the time efficiency and memory consumption of R4 with those of Sparkwave \citep{sparkwave} and Etalis \citep{EP-SPARQL} using a modification of the big-data/RDFS benchmark BSBM \citep{BSBMresults}.
+  We have demonstrated that R4 is faster than these RDF stream-querying systems, without loss of comparative soundness or completeness, when applying low-range windows to high-throughput streams, despite Sparkwave and Etalis supporting more limited/less explicitly well-grounded reasoning than that of R4.
+
+  Moving forward, we plan to investigate alternate planing algorithms to improve the Quality of Service (\emph{QoS}) characteristics of the system, i.e.
 reduce the latency from receipt of some new graph to the production of entailments justified in part by that new graph.

### 2015-12-02 article.tex
fixed incorrect interval notations and inconsistent variable names

@@
 from then on.  As such, we define a {\em Windowed Continuous Datalog program}
-$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle\} \rangle$
+$\Pi^{C, r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle\} \rangle$
 (where $r = max(r_1,\dots,r_s)$) to be a tuple with components $R$ and $F^0$ as defined above for a
@@
 associated with the range $r_x \in \mathbb{N}$ of a sliding window over
-$S_x$, for $x \in (1,s)$. This specifies that the fact instances
+$S_x$, for $x \in [1,s]$. This specifies that the fact instances
 from $S_x$ are to contribute to the entailments of $\Pi^{C,r}$ from the instant
@@
 We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0,
-\{ W^{S_x, r_x} \mid x \in (1,s) \} \rangle$
+\{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle$
 at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
@@
 finite states $W^{S_x,r_x}_i$ of windows $W^{S_x,r_x}$ over streams $S_x$ of range
-$r_x \in \mathbb{N}$, for all $x \in (1,s)$; we refer to this as the {\em instantaneous state} of
+$r_x \in \mathbb{N}$, for all $x \in [1,s]$; we refer to this as the {\em instantaneous state} of
 the WC-Datalog program, written
@@
   and a Windowed Continuous RIF-Core program as a WC-Datalog program
-  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n)\} \rangle}$.
+  ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s]\} \rangle}$.
   Furthermore, ${\langle R, B \rangle \models \varnothing}$ and
@@
   \emph{CoDeR} is an abstract model for \emph{Co}ntinuous \emph{De}ductive \emph{R}easoning over streamed RDF data that expresses the semantics of Windowed Continuous RIF-Core, thus supporting continuous reasoning with general knowledge expressed in both RIF-Core and OWL 2 RL.
-  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_x$, $P$ and $N$ and the window ranges $r_x$ for $x \in (1,s)$ of each $\Pi^{C,r}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window ranges $r_x$.
+  It is composed of a stream-based semantic data model and a minimal algebra of stream-to-stream operators, the former for expressing the persistent fact instances $F^0$ and $E^0$, the streams $S_x$, $P$ and $N$ and the window ranges $r_x$ for $x \in [1,s]$ of each $\Pi^{C,r}$, the latter for expressing the persistent set of rules $R$, the built-in predicates of RIF-Core $B^0$ and enforcing the window ranges $r_x$.
   It casts the problem of applying those rules to streamed RDF data as that of iterative pattern matching, in a manner inspired by the Rete pattern matching algorithm \citep{forgy79}, though it is meant to support any such algorithm rather than prescribing one itself.
@@
     A \emph{graph instance} is an RDF graph with some additional temporal annotation regarding both the instant from which it became valid and the instant at which it will cease to be valid;
-    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${(e,n]}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
+    it is a triple composed of the graph ${\{(s,p,o),\dots,(s',p',o')\}}$ and the pair of instants $e$ and $n$ that bound the interval ${[e,n)}$ for which the graph is valid, written ${\langle \{(s,p,o),\dots,(s',p',o')\},e,n \rangle}$.
     The names $e$ and $n$ are so named because they relate to the instants of the instance's {$e$ntailment} and {$n$egation} by the program.
@@
 
-  As has been noted in \refdef{CoDeR: graph instance}, the interval ${(e,n]}$ for any graph instance is the the period of time for which that graph instance is \emph{valid} \citep{SemanticStreamingManagement,sparkwave}.
+  As has been noted in \refdef{CoDeR: graph instance}, the interval ${[e,n)}$ for any graph instance is the the period of time for which that graph instance is \emph{valid} \citep{SemanticStreamingManagement,sparkwave}.
   The notion of validity within WC-RIF-Core programs relates to the period of time for which a given fact instance ${\langle f, t \rangle}$ is justified by a state ${\Pi^{C,r}_i = \langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_i \rangle \models \langle f, t \rangle}$ of a program $\Pi^{C,r}$.
-  As such, the interval ${(e,n]}$ is used to express the window ranges $r_x$ for $x \in (1,s)$ of the ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,s) \} \rangle}$:
+  As such, the interval ${[e,n)}$ is used to express the window ranges $r_x$ for $x \in [1,s]$ of the ${\Pi^{C,r} = \langle R, F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s] \} \rangle}$:
   for graph instances entering a CoDeR-based system, $e$ is the instant at which every fact in the given graph was asserted, and $n$ is the instant at which they will leave the semantic window over $S_x$ described by $\Pi^{C,r}$ (a.k.a.
@@
 
-  The first class of CoDeR streams, input/output streams, provide the semantics of the streams and sets $S_x$, $P$, $N$, $F^0$ and $E^0$, for any ${\Pi^{C,r} = \langle R,F^0,\{W^{S_x,r_x} \mid x \in (1,s)\} \rangle \models \langle E^0, P, N \rangle}$ expressed by a CoDeR-based system.
+  The first class of CoDeR streams, input/output streams, provide the semantics of the streams and sets $S_x$, $P$, $N$, $F^0$ and $E^0$, for any ${\Pi^{C,r} = \langle R,F^0,\{W^{S_x,r_x} \mid x \in [1,s]\} \rangle \models \langle E^0, P, N \rangle}$ expressed by a CoDeR-based system.
   These streams and sets are represented as CoDeR streams, as per \refdef{CoDeR: stream}, that are, in the case of input streams, produced elsewhere in the Semantic Web or, in the case of output streams, exposed on the Semantic Web as a resource themselves, as per \refdef{CoDeR: input/output}.
@@
   \end{definition}
-  In each CoDeR-based system the input streams express the logical streams $S_x$ for $x \in (1,s)$ of the given $\Pi^{C,r}$, and a single output stream expresses the logical stream $P$ entailed by the $\Pi^{C,r}$.
+  In each CoDeR-based system the input streams express the logical streams $S_x$ for $x \in [1,s]$ of the given $\Pi^{C,r}$, and a single output stream expresses the logical stream $P$ entailed by the $\Pi^{C,r}$.
   This single output stream also expresses the stream $N$ entailed by the $\Pi^{C,r}$, as that logical stream is defined as the set of instances in the stream $P$ ordered in a sequence by the instant $n$ at which ${\langle R, F^0 \cup B^0 \cup \mathop{\cup}^s_{x=1} W^{S_x,r_x}_{n}}$ \emph{ceased} to justify them;
   this instant $n$ is, therefore, the same $n$ with which each graph is annotated in CoDeR, being the instant at which the graph instance will cease to be valid.
-  As such, the annotation of streamed RDF graphs with the interval ${(e,n]}$ gives \emph{all} streams of CoDeR a dual semantics with regard to WC-RIF-Core.
+  As such, the annotation of streamed RDF graphs with the interval ${[e,n)}$ gives \emph{all} streams of CoDeR a dual semantics with regard to WC-RIF-Core.
 
@@
     a one-pass unary evaluation function $\sigma_{external}$ for the built-in predicates of RIF-Core;
-    and a one-pass unary graph decomposition function ${{}_g\pi_t}$ to convert the CoDeR input stream of heterogeneous RDF graph instances into a stream of homogeneous RDF triple instances, the latter being the form of a stream $S_x$ of a program ${\Pi^{C,r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in (1,n) \} \rangle}$.
+    and a one-pass unary graph decomposition function ${{}_g\pi_t}$ to convert the CoDeR input stream of heterogeneous RDF graph instances into a stream of homogeneous RDF triple instances, the latter being the form of a stream $S_x$ of a program ${\Pi^{C,r} = \langle R, F^0, \{ W^{S_x, r_x} \mid x \in [1,s] \} \rangle}$.
     The continuous derivation of the entailment of WC-Datalog programs requires, in turn, the set of functions required for querying deductive databases using Datalog, with the addition of another two functions:
-    a two-pass unary valid-window maintenance function $w$ to enforce the windows of range $r_x$, for $x \in (1,n)$, over the CoDeR input streams;
+    a two-pass unary valid-window maintenance function $w$ to enforce the windows of range $r_x$, for $x \in [1,s]$, over the CoDeR input streams;
     and a two-pass unary window-to-stream function $s$ to extract the stream of novel graph instances from a sequence of consecutive valid-windows, closely related to the ${IStream()}$ function of CQL \citep{StanfordStream}.
@@
     Furthermore, the valid-time interval of any instance is either directly inherited from the interval of the input instance, in the case of one-pass operators, or derived from the intersections of the intervals of both contributing instances in the sole two-pass operator, the SteM;
-    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in (1,n) \} \rangle}$.
+    as such, it is guaranteed that all instance annotations accurately describe the intervals for which the given instance is justified by the states of a given WC-RIF-Core program ${\Pi^{C,r} = \langle R,F^0 \cup B^0, \{ W^{S_x,r_x} \mid x \in [1,s] \} \rangle}$.
 

### 2015-12-02 article.tex
very minor edits to fix typos etc.

@@
   This includes the growing number of use cases for the continuous processing of streams of semantic data, in which large volumes of time-sensitive data must be processed in near-real time, with the aim of minimising the volume of data transported, duplicated and stored across the Web whilst maximising the variety of problems that can be solved.
-  With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
+  With the focus on reasoning in the Semantic Web stack, varying in expressivity from the low complexity RDF entailment rules through to the high complexity of OWL Full, it becomes clear that the greatest challenge in semantic stream processing is that of stream reasoning.
 
@@
   However, this expression of reasoning on top of CEP introduces semantics of temporal reasoning \citep{OrgunWadge92,Tuzhilin93}, whereas in many cases \emph{temporally-agnostic} ontologies, which do \emph{not} make use of temporal reasoning, are applicable to \emph{temporally-sensitive} streamed data, which is relevant for an inherently limited amount of time.
-  For example, the descriptions of sensors in a sensor network and the various types of reading they produce are often temporally agnostic, while the readings themselves are accurate (and so relevant) for only a short time after they are produced.
+  An example taken from \citep{margara14streamWeb}, the descriptions of sensors in a sensor network and the various types of reading they produce are often temporally agnostic, while the readings themselves are accurate (and so relevant) for only a short time after they are produced.
   Furthermore, the inferences gleaned from applying the former to the latter are equally temporally sensitive, and so may most accurately be represented as a stream themselves.
@@
 instantaneous truth and entailments, in line with the work of the
-reasoning community.  However, it must also provide a continuous
+reasoning community.  However, it must also provide a concise streamed
 interpretation of those instantaneous entailments in order to
@@
 \begin{nestedsection}{Evaluation, Results and Discussion}{evaluation}
-  To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning to some extent on semantic streaming data.
+  To evaluate the performance of R4, we compare it with state-of-the-art systems that provide the capability of performing background reasoning to some extent on streamed semantic data.
   These include Etalis \citep{EP-SPARQL}, Sparkwave \citep{sparkwave}, Streaming knowledge bases \citep{walavalkar08streamingkb}, and the incremental reasoner presented in \citep{inc-reasoning-background-knowledge}.

### 2015-12-02 article.tex
altered R4 extension to RIF-Core in order to allow one window per streamed import.
Integrated 1st pass of Rehab's changes (without figures)

--- 2015-12-01 article.tex
@@
   A rule document containing any number of rules is submitted to the system.
-  We have chosen RIF-Core to be the language in which rules can be represented, for being compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
+  We have chosen RIF-Core to be the language in which rules can be represented, because it is compatible with RDF and other semantic web standards, as well as sharing the semantics of positive Datalog.
   The rule document specifies any domain-specific rules as well as an entailment regime \emph{if any} for background reasoning.
-  We also added a simple extension to RIF-Core in line with the extension of Datalog described in \refsec{semantics}, enabling users to specify a window range as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
+  We have also added a simple extension to RIF-Core (in line with the extension of Datalog described in \refsec{semantics}) that enables users to specify a window range as part of their rule set, as shown in \reffig{lst-S-RIF-Core}.
   Further, we propose the extension of the set of `PROFILE's that describe the `Import's of a rule set to include those definitions of RDF streams recognised by the W3C.
@@
   \begin{nestedsection}{Continuous Reasoning with Rete}{implementation: continuous rete}
-    The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, can be well fit into the stream reasoning model as it operates incrementally.
+    The Rete algorithm \citep{forgy79}, which was introduced as a solution for the many pattern/many object matching problem, is a natural fit for our stream reasoning model because it is a dataflow approach that operates incrementally.
     The RETE algorithm can process large data sets efficiently because it avoids iterating over both data elements (facts, or ``working memory'') and over the production rules.
@@
     Alpha nodes are single-entry nodes that form a discrimination network, as shown in \reffig{example-rete-network}.
-    In R4, alpha nodes can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
-    Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to its successor node(s) in the case of a match.
+    In R4, an alpha node can receive streamed quads from any number of nodes through its single input, reflecting the multiple-source nature of CoDeR's streams.
+    Each received quad is matched against some conditions, and is either dropped if there is no match, or propagated downstream to its successor node(s) in the event of a match.
     As such, R4's interpretation of alpha nodes are implementations of the triple-select operator $\sigma_{triple}$ of CoDeR, but can also filter based on the datatypes of the values in triples, thereby also implementing the other selection operator of CoDeR, $\sigma_{external}$.
@@
     In this context, each left (beta) and right (alpha) memory is implemented as a \emph{valid window}.
-    These windows are implemented as priorities queues in which quads are ordered according to their time annotations, and pruned by ``popping'' from the queue and deleting any quads whose annotation $a \leqslant i - r$, where $i$ is the current system time and $r$ is the range of the semantic window.
-
-    When a join node is left-activated, i.e.
-it receives a token through its left input, it first adds the new token to its left window, then it prunes its right window before iterating over the remaining quads to find any matches.
+    These windows are implemented as a combination of a priority queue (in which quads are ordered according to their time annotations) and a hashmap (in order to improve the speed of probing).
+Quads are removed from the priority queue when they expire (i.e.
+when their time annotations are less than the current time minus the window range); the corresponding entries in the hashmap are removed at the same time.
+
+    When a join node is left-activated (i.e.
+it receives a token through its left input), it first adds the new token to its left window, then it prepares its right window by removing any expired items (as above) so that n ooutdated results are matched.
+The right window is then searched according to the join node conditions (using the hashmap).
 When a match is found, a new token is created by duplicating the left token and adding the right quad to the new token's quad list.
 The new token is annotated with the \emph{earliest} time with which its component quads were annotated, and then propagated to the next beta node, or to the terminal node if it is in the root node of a join tree.
-    Conversely, when the join node is right-activated by receiving a quad from an alpha node, it adds the new quad to the right window, prunes the left window then matches the quad against the remaining tokens.
+Conversely, when the join node is right-activated by receiving a quad from an alpha node, it adds the new quad to the right window, prepares the left window then matches the quad against the remaining tokens.
 
@@
   To the best of our knowledge, the latter two implementations were never made public, so we compare R4 against Etalis and Sparkwave only.
-  The experiments are performed on an Intel Core i5 computer with 3.2 GHz and 8 GB memory.
+
+  \begin{nestedsection}{Methodological Considerations}
+    Following the development of semantic stream processors, a number of benchmarks have emerged to test their performance: SRBench\citep{SRBench}, LSBench\citep{linkedstreamproc}, CSRBench and CityBench\citep{ali15}.
+While these benchmarks provide rich datasets, they mostly do not consider reasoning; with the exception of SRBench, none of the other benchmarks' queries require inference capabilities.
+
+    One of the most important measures of any stream processing system is the maximum input throughput, defined in \citep{scharrenbach13} as the number of data elements in the input stream consumed by the system per time unit; we use this as the key performance indicator in our comparative evaluation of the three systems.
+  \end{nestedsection}
+
+  \begin{nestedsection}{Methodology}
+   The experiments were performed on an Intel Core i5 computer running at 3.2 GHz with 8 GB memory.
 We set up Jtalis (the Java wrapper of Etalis) over SWI-Prolog v7.2.1 and installed Sparkwave v0.5.1.
 
-  \begin{nestedsection}{Comparative Results}{evaluation: comparative results}
+We followed the experimentation strategy used by Sparkwave , using a synthetic dataset generated with the Berlin SPARQL Benchmark\citep{BSBMresults} containing 1.1 million triples (representing 100,000 limited-time offers made available by an online marketplace).
+For the background knowledge, we generated a small schema that described 329 product types arranged in a 4-level hierarchy.
+
+    The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
+    As offers may be associated with product super-types, some small degree of background reasoning is needed to determine the offer price for specific product sub-types.
+
     \begin{figure*}[t]
@@
 
-    In order to compare the efficiency of R4 against both Etalis and Sparkwave, we formulated an experiment using the synthetic dataset from the Berlin SPARQL Benchmark \citep{BSBMresults} which was used in the Sparkwave paper.
-    We used another 1.1 million triples dataset using the provided data generator, containing information about 100,000 limited-time offers made available by some fictional online marketplace.
-    We used a small schema that has 329 product types arranged in a 4-level hierarchy.
-
-    The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
-    As offers may be associated with product super-types, some small degree of background reasoning is needed to determine the offer price for specific product sub-types.
-
-    Before we move on to more detailed analysis of the results, we noticed that this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
+  \end{nestedsection}
+
+  \begin{nestedsection}{Comparative Results}{evaluation: comparative results}
+
+this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
     However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).

### 2015-12-02 article.tex

--- 2015-12-02 article.tex
@@
               \ZeroOrMore{\NTerm{Import}} 
-              \Optional{\NTerm{Window}}
               \Optional{\NTerm{Group}} 
@@
               \Optional{\NTerm{PROFILE}} 
+              \Optional{\NTerm{Window}} 
               \Token{)}

### 2015-12-02 article.tex
final draft, including re-working of evaluation to incorporate correctness/completeness and more clearly represent our comments and decisions regarding Etalis.  Also includes graphs for throughput and latency, based on Rehabs recent data.  Does not include table showing the results for Etalis using And, because Rehabs data for that was for a 500k dataset, not 1.1m

@@
     When a join node is left-activated (i.e.
-it receives a token through its left input), it first adds the new token to its left window, then it prepares its right window by removing any expired items (as above) so that n ooutdated results are matched.
+it receives a token through its left input), it first adds the new token to its left window, then it prepares its right window by removing any expired items (as above) so that no outdated results are matched.
 The right window is then searched according to the join node conditions (using the hashmap).
@@
   \begin{nestedsection}{Methodological Considerations}
-    Following the development of semantic stream processors, a number of benchmarks have emerged to test their performance: SRBench\citep{SRBench}, LSBench\citep{linkedstreamproc}, CSRBench and CityBench\citep{ali15}.
+    Following the development of semantic stream processors, a number of benchmarks have emerged to test their performance: SRBench \citep{SRBench}, LSBench \citep{linkedstreamproc}, CSRBench and CityBench \citep{ali15}.
 While these benchmarks provide rich datasets, they mostly do not consider reasoning; with the exception of SRBench, none of the other benchmarks' queries require inference capabilities.
@@
     One of the most important measures of any stream processing system is the maximum input throughput, defined in \citep{scharrenbach13} as the number of data elements in the input stream consumed by the system per time unit; we use this as the key performance indicator in our comparative evaluation of the three systems.
+    In addition, it is necessary to confirm that the data produced by a system is complete and correct according to its semantics.
+    It should be noted that, while correctness must be absolute, completeness may be expected to decline with decreased window size.
   \end{nestedsection}
@@
   \begin{nestedsection}{Methodology}
-   The experiments were performed on an Intel Core i5 computer running at 3.2 GHz with 8 GB memory.
+   The experiments were performed on an Intel \mbox{Core i5} computer running at 3.2 GHz with 8 GB memory.
 We set up Jtalis (the Java wrapper of Etalis) over SWI-Prolog v7.2.1 and installed Sparkwave v0.5.1.
 
-We followed the experimentation strategy used by Sparkwave , using a synthetic dataset generated with the Berlin SPARQL Benchmark\citep{BSBMresults} containing 1.1 million triples (representing 100,000 limited-time offers made available by an online marketplace).
+We followed the experimentation strategy used by Sparkwave because it requires reasoning over background knowledge to correctly answer the query, using a synthetic dataset generated with the Berlin SPARQL Benchmark (BSBM) \citep{BSBMresults} containing 1.1 million triples (representing 100,000 limited-time offers made available by an online marketplace).
 For the background knowledge, we generated a small schema that described 329 product types arranged in a 4-level hierarchy.
 
+    We configure each system with the generated schema, a rule inspired by a query from the BSBM and a window size (0.1, 1, 2, 5 and 10 second sliding windows are tested).
     The rule we chose to express in each system, shown in \reffig{BSBM-rule}, entails the offer price for all products that are on offer.
@@
 
+    We then measure the time taken for the system to consume all 1.1 million triples for each configuration of each system, from which we can calculate the maximum throughput and average latency of that system for each window size.
+    We also compare the contents of the streams of entailments produced by each system against the set that is produced by statically reasoning over the conventional logic program composed of the same 1.1 million triple dataset and background knowledge;
+    because of the tight localisation of offer IRIs in the stream, this provides a good benchmark of the correctness and completeness of each system for each window size.
+
   \end{nestedsection}
@@
 
-this rule takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${<10}$ minutes for the other two systems).
-    However, when we changed the `And' into a series of `seq's, the system runs 20x times faster (7.5 minutes for the same test).
-    Such a drastic difference is likely to be caused by the optimisation of Etalis for the \emph{seq} operator: being intended for \emph{event processing} rather than continuous querying/reasoning, the order of arrival of triples is often more relevant than their simple coincidence in the system.
-    In order to present results that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to modify the rule when evaluating Etalis to use \emph{seq} instead of \emph{And}.
-    It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time-efficiency and memory-consumption of the three systems.
+    Firstly, for every window size, the streams produced by each system are correct, both in their contents with regard to the entailment of the conventional program, and their ordering with regard to the entailment semantics of WC-Datalog.
+    Additionally, in all cases but that of Etalis with the smallest window size, the streams produced are complete with regard to the entailment of the conventional program.
+
+    Secondly, we note that the rule presented in \reffig{BSBM-rule} takes significantly longer to process in Etalis (3 hours for the 1.1 million dataset with 1 second time window compared to ${{<}10}$ minutes for the other two systems).
+    In order to further investigate this difference, we tried changing the \emph{And} into a series of \emph{seq}s \citep{EP-SPARQL}, which resulted in the system running 20x times faster (7.5 minutes for the same test).
+    Despite being semantically different, we would not expect such a drastic difference in running time unless Etalis were heavily optimised for the \emph{seq} operator;
+    this is likely, as Etalis is intended for \emph{event processing} rather than continuous querying/reasoning, making the order of arrival of triples more relevant than their simple coincidence in the system.
+
+    In order to present graphics that provide a meaningful comparison between the time efficiency of each system with regards to changing window ranges, as in \reffig{all-systems-varying-windows}, we chose to eliminate the results for Etalis using \emph{And} from our data set, but include those using the rule modified to use \emph{seq} when evaluating Etalis.
+    It should be noted that we recognise that this modified rule is not semantically equivalent to that by which we evaluate R4 and Sparkwave, which cannot express the \emph{seq} operator, but is sufficient to contrast the effect of window range on the time-efficiency of the three systems.
     \begin{figure}
       \centering
-      \includegraphics[width=0.45\textwidth]{all-systems-varying-windows}
-      \caption{A comparison of the times taken by the systems R4, Sparkwave and Etalis to process 1.1 million triples from a single input stream, given windows of various ranges over the input stream.}
+      \includegraphics[width=0.45\textwidth]{all-systems-varying-windows-throughput}\\
+      \includegraphics[width=0.45\textwidth]{all-systems-varying-windows-latency}
+      \caption{A comparison of the maximum throughputs and average processing latencies of the systems R4, Sparkwave and Etalis, given windows of various ranges over the input stream.}
       \labelfig{all-systems-varying-windows}
@@
 
-    As shown in \reffig{all-systems-varying-windows}, R4 is faster than Etalis for small windows but slower than Sparkwave, and, while the time-to-complete increases with window range in both Sparkwave and R4, that of Etalis appears to remain fairly constant for the window ranges tested.
-    The growing difference between R4 and Sparkwave as the range of the windows increases is likely due to the fact that Sparkwave uses a hash-join algorithm with garbage-collection window-maintenance while the current implementation of R4 uses loop-joins over lists maintained by indexing over time annotations;
-    as data remains valid for longer, the beta memories (i.e valid windows) grow bigger, and window-joins in R4 need significantly longer to iteratively check for matches compared to the hash-based look-up of those in Sparkwave.
-    More time-efficient, multi-index window-joins are to be considered for R4 in future work.
-
-    We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
-    Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
-    It shows that, while R4 and Jtalis follow similar patterns, Sparkwave's memory needs are significantly higher and continue to increase over time, a behaviour that should not occur when applying windows over streams.
-    Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to an interaction between the manual garbage collection performed by Sparkwave and the underlying JVM garbage collector.
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
-      \caption{A comparison of the memory consumed by the systems R4, Sparkwave and Etalis throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
-      \labelfig{all-systems-memory-consumption}
-    \end{figure}
+    As shown in \reffig{all-systems-varying-windows}, R4 is significantly faster than both Etalis and Sparkwave for the tested window sizes.
+    Interestingly, while the maximum throughput decreases and average processing latency increases (apparently asymptotically) with window range similarly in both Sparkwave and R4, those of Etalis using \emph{seq} appear to remain approximately constant for all window ranges tested.
+    It should be noted that this behaviour of Etalis remains in the case where the query uses \emph{And}.
+    However, despite its constancy, the throughput and processing latency of Etalis are so dramatically lower/higher than the other systems, respectively, that we project that R4 will remain faster than both Sparkwave and Etalis in cases where the window sizes are orders of magnitude larger than those tested.
+
+    % We also examined the memory consumption of R4, Sparkwave, and Jtalis throughout a run.
+    % Setting the sliding window range to five seconds, \reffig{all-systems-memory-consumption} shows the amount of memory used by the three engines during the life time of the BSBM rule.
+    % It shows that, while R4 and Jtalis follow similar patterns, Sparkwave's memory needs are significantly higher and continue to increase over time, a behaviour that should not occur when applying windows over streams.
+    % Though we cannot know the reason for this without a thorough inspection of the Sparkwave code-base, we suspect that this is related to an interaction between the manual garbage collection performed by Sparkwave and the underlying JVM garbage collector.
+    % \begin{figure}
+    %   \centering
+    %   \includegraphics[width=0.5\textwidth]{memoryConsumptionComparison}
+    %   \caption{A comparison of the memory consumed by the systems R4, Sparkwave and Etalis throughout a run processing 1.1 million triples from a single input stream, given sliding windows of 10 seconds over the input stream.}
+    %   \labelfig{all-systems-memory-consumption}
+    % \end{figure}
   \end{nestedsection}
